### 2022-06-29 minos/transitions/housing/housing_mice.Rmd
add mwb framework. need to get ols model working now.

---
title: "Housing mice Notebook"
output: html_notebook
---

This notebook demonstrates how multiple imputation is applied to the household variables for application in MINOS transition probability models. It goes into visualisation of missingness structure for the six household variables and what variables are influential in them being missing. A MICE pipeline is constructed demostrating any differences between complete case and imputed demographics. First load in US data and some packages..

```{r}
source("utils.R") # for data formatting functions and other useful globals. 
require(VIM) # for missingness visualisation.
require(mice) # for mice multiple imputation.
require(lattice) # for mice imputation functions.
require(ordinal) # for clm models.
require(tidyr) # for data frame formatting.
require(viridis) # for python's nice colour palettes
require(ggplot2) # for confusion matrix plots
#require(dplyr) # matrix manipulation
require(stringr)# for fancy wrapped text in confusion matrix axis labels
require(shadowtext) #^

# load data from 2011 and format it via utils.R functions. 
data_source<- "data/corrected_US/"
years <- c("2011")
file_names <- get_US_file_names(data_source, years)
data <- get_US_data(file_names)
data <- format_US_housing_data(data)

n_heating_missing <- nrow(data) - sum(complete.cases(data[,c("heating")]))
n_five_missing <- nrow(data) - sum(complete.cases(data[,c("fridge_freezer", "washing_machine", "tumble_dryer", "dishwasher", "microwave")]))
n_six_missing <- nrow(data) - sum(complete.cases(data[,c("fridge_freezer", "washing_machine", "tumble_dryer", "dishwasher", "microwave", "heating")]))
household_vars = c("fridge_freezer",
                   "washing_machine",
                   "tumble_dryer",
                   "dishwasher",
                   "microwave",
                   "heating")
```

Data description demonstrates the six SIPHER7 household variables as well as other variables used in imputation that may indicate missingness. To more clearly demonstrate missingness structure first use aggr plots from VIM.

```{r}
a <- aggr(data, cex.axis=.5, oma=c(5,2,2,5),numbers=T, cex.numbers=0.5) # cex and oma adjustments so all labels show up.

b <- aggr(data[, household_vars], numbers=T, cex.axis=.5, oma=c(5,2,2,5))
```

There are 40197 out of 134598 (29.9%) rows missing at least one of the six household variables. The vast majority 38757 (28.8%) are missing solely heating values. 1440 (1.07%) are missing 5/6 composite variables simultaneously with 621 (0.46%) missing all 6.

Since 5 household variables are simultaneous missing it is worth compositing them together and imputing them as one variable https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3074241/pdf/MPR-20-40.pdf. Heating will be imputed separately and then the two variables will be merged into one final SIPHER7 value.

The next part of the imputation strategy is which variables to include in analysis. It is suggested to have an imputation dataset larger than any dataset used in transition probability calibration ensuring any variable that may be indicative of missingness. VIM and MICE packages have good visualisation tools to represent this. These tools quantify distributional differences of variables between those missing houshold values and those not.

```{r}
is_missing_heating <- is.na(data$heating)
is_missing_five <- is.na(data$fridge_freezer)
histogram(~age|is_missing_heating, data)
histogram(~age|is_missing_five, data)
histogram(~SF.12|is_missing_heating, data)
histogram(~SF.12|is_missing_five, data)
```

```{r}
# needs better scaling for low % missing data
# scaling works with ylim change but ticks needs relabelling. 
#spineMiss(data[, c("job_sec", "heating")], col = c(blue, orange), ylim=c(0, 0.2))
#spineMiss(data[, c("job_sec", "fridge_freezer")], col = c(blue, orange), )

missing_barchart <- function(data, v1, v2){
  print(v1)
  print(v2)
  missing_table <- table(data[,v1], is.na(data[,v2]))
  # not guaranteed to be on these rows. needs forcing to correct labels somehow. 
  missing_percent <- missing_table[,2]/(missing_table[,1]+missing_table[,2])
  barchart(missing_percent, ylab = v1)
}
missing_barchart(data, "ethnicity", "heating")
```

```{r}
missing_barchart(data, "ethnicity", "heating")
missing_barchart(data, "depression_change", "heating")
missing_barchart(data, "education_state", "heating")
missing_barchart(data, "labour_state", "heating")
missing_barchart(data, "sex", "heating")
missing_barchart(data, "job_sec", "heating")

```

Can apply these plots to all variables used in imputation of heating and five others respectively.

```{r}
imp_columns <- c("five_household", "heating", "labour_state", "education_state",
                 "sex", "job_sec", "ethnicity", "age", "SF.12", "hh_netinc")
house_mice <- mice(data = data[, imp_columns],
                 m = 10, maxit = 10, seed = 500,
                 remove.collinear=T)
```

```{r}
print(summary(complete(house_mice)[,c("five_household","heating")]))
plot(house_mice)
```

```{r}
data2<- complete(house_mice)
data2 <- tidyr::drop_na(data2)
# format final housing composite. If they have no appliances assign 0.
# any up to 5 assign 1.
# all assign 2.
data2$housing <- rowSums(data2[, c("five_household", "heating")])
data2[which(data2$housing <= 0), "housing"] <- 0
data2[which(data2$housing <= 5 & data2$housing > 1), "housing"] <- 1
data2[which(data2$housing >= 6), "housing"] <- 2

data2$housing <- factor(data2$housing, levels = c(0,1,2))
n <- nrow(data2)
set.seed(88)
Inx <- sample(n, 0.8*n)
train<-data2[Inx,]
test<-data2[-Inx,]
y <- test$housing
test <- subset(test, select= - housing)

formula <- "housing ~ factor(sex) + 
                      age + 
                      SF.12 + 
                      factor(labour_state) +
                      factor(job_sec)"
housing.clmm <- clm(housing ~ (factor(sex) + 
                                factor(education_state) + 
                                age + 
                                SF.12 + 
                                factor(labour_state) +
                                factor(job_sec) +
                                factor(ethnicity) + hh_netinc)**2, 
                                data= train[complete.cases(train),],
                                link = "logit",
                                threshold = "flexible",
                                Hess=T)

print(summary(housing.clmm))
prs<- 1 - logLik(housing.clmm)/logLik(clm2(housing ~ 1,data= train))
print(prs)
```

```{r}
# remove housing column or predict function doesnt give full multinom probabilities.

preds <- predict(housing.clmm, test)$fit
markers <-as.numeric(as.character(factor(y, labels = c(0, 10, 25))))
markers <- as.integer(markers)
colours <-cividis(3)
v_colours <- factor(y, labels = cividis(3))

table(y, predict(housing.clmm, test, type="class")$fit)

test2<-test
test2$housing <- y
#par(mfrow(c(2,1)))
plot(predict(housing.clmm, test2, type="linear.predictor")$eta2, ylim=c(-10, 10)
, col=v_colours, pch=markers)
plot(predict(housing.clmm, test2, type="linear.predictor")$eta2, ylim=c(-100005, -99995)
, col=v_colours, pch=markers)
legend("topright", legend=c("None", "Some", "All"), col = colours, pch =c(0, 10, 25))

w1 <- which(y == 0)
plot(preds[w1, 1], col =  colours[1], pch=0)#dev.off()

w2 <- which(y == 1)
plot(preds[w2, 2], col =  colours[2], pch=10)#dev.off()

w3 <- which(y == 2)
plot(preds[w3, 3], col =  colours[3], pch=25)#dev.off()

```

```{r}
confusion.matrix<- table(y, predict(housing.clmm, test, type="class")$fit)
group_pops<-rowSums(confusion.matrix)
confusion.frame <- as.data.frame(confusion.matrix/group_pops) # convert to data frame
confusion.frame$Freq <-round(confusion.frame$Freq, 3) # round to 2dp
row_names <- paste0(rownames(confusion.matrix), " [")
row_names <- paste0(row_names,  as.character(group_pops))
row_names <- paste0(row_names, "]")
row_names <- str_wrap(row_names, width = 12)
col_names <- str_wrap(colnames(confusion.matrix), 10)

confusion.plot<- ggplot(confusion.frame, aes(Var2, rev(y), fill= Freq)) +
  scale_fill_viridis_c("% of state")+
  geom_tile() + 
  geom_shadowtext(aes(label=Freq)) +
  #scale_fill_gradient(high="#87ceeb", low="#e34234") +
  labs(y = "True State [population in state]",x = "Predicted State (no units)") + 
  scale_y_discrete(labels=rev(row_names)) + # label axes
  scale_x_discrete(labels=col_names)
print(confusion.plot)
```