### 2022-12-15 docsrc/Coefficients/summarise_transitions_2.Rmd
attempting a workaround for ugly notebooks using Rmd>tex>rst conversion

---
title: "MINOS Transition Models"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Setup

## Tobacco

Tobacco consumption is measured in the usual number of cigarettes smoked per day, and is taken directly from a US variable ([ncigs](https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/variable/ncigs)).

We estimated a zero inflated poisson model, with the formula:

> ncigs \~ age + sex + SF_12 + labour_state + job_sec + ethnicity + hh_income \| ethnicity + labour_state + SF_12

```{r}
require(tidyverse)
require(pscl)

workingDir <- "../.." #TODO stored in docsrc/Coefficients for now. move to data/transitions and change rst filepaths accordingly.
knitr::opts_knit$set(root.dir = workingDir)
trans.path <- paste0(workingDir,'/data/transitions/test/')
tobacco <- readRDS(paste0(trans.path, 'tobacco_zip_2018_2019.rds'))
summary(tobacco)
```