### 2022-12-15 docsrc/Coefficients/summarise_transitions_2.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={MINOS Transition Models},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{MINOS Transition Models}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\hypertarget{setup}{%
\section{Setup}\label{setup}}

\hypertarget{tobacco}{%
\subsection{Tobacco}\label{tobacco}}

Tobacco consumption is measured in the usual number of cigarettes smoked
per day, and is taken directly from a US variable
(\href{https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/variable/ncigs}{ncigs}).

We estimated a zero inflated poisson model, with the formula:

\begin{quote}
ncigs \textasciitilde{} age + sex + SF\_12 + labour\_state + job\_sec +
ethnicity + hh\_income \textbar{} ethnicity + labour\_state + SF\_12
\end{quote}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{require}\NormalTok{(tidyverse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: tidyverse
\end{verbatim}

\begin{verbatim}
## -- Attaching packages --------------------------------------- tidyverse 1.3.2 --
## v ggplot2 3.3.6      v purrr   0.3.4 
## v tibble  3.1.8      v dplyr   1.0.10
## v tidyr   1.2.0      v stringr 1.4.0 
## v readr   2.1.2      v forcats 0.5.1 
## -- Conflicts ------------------------------------------ tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{require}\NormalTok{(pscl)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Loading required package: pscl
## Classes and Methods for R developed in the
## Political Science Computational Laboratory
## Department of Political Science
## Stanford University
## Simon Jackman
## hurdle and zeroinfl functions by Achim Zeileis
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{workingDir }\OtherTok{\textless{}{-}} \StringTok{"../.."} \CommentTok{\#TODO stored in docsrc/Coefficients for now. move to data/transitions and change rst filepaths accordingly.}
\NormalTok{knitr}\SpecialCharTok{::}\NormalTok{opts\_knit}\SpecialCharTok{$}\FunctionTok{set}\NormalTok{(}\AttributeTok{root.dir =}\NormalTok{ workingDir)}
\NormalTok{trans.path }\OtherTok{\textless{}{-}} \FunctionTok{paste0}\NormalTok{(workingDir,}\StringTok{\textquotesingle{}/data/transitions/test/\textquotesingle{}}\NormalTok{)}
\NormalTok{tobacco }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(trans.path, }\StringTok{\textquotesingle{}tobacco\_zip\_2018\_2019.rds\textquotesingle{}}\NormalTok{))}
\FunctionTok{summary}\NormalTok{(tobacco)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## zeroinfl(formula = y ~ factor(sex) + scale(age) + scale(SF_12) + factor(labour_state) | 
##     relevel(factor(ethnicity), ref = "WBI") + scale(age), data = data, 
##     dist = "pois")
## 
## Pearson residuals:
##      Min       1Q   Median       3Q      Max 
## -1.75024 -0.76329 -0.05563  0.58683 17.88004 
## 
## Count model coefficients (poisson with log link):
##                                      Estimate Std. Error z value Pr(>|z|)    
## (Intercept)                          1.158187   0.024071  48.116  < 2e-16 ***
## factor(sex)Male                      0.106336   0.025077   4.240 2.23e-05 ***
## scale(age)                           0.107561   0.017073   6.300 2.97e-10 ***
## scale(SF_12)                        -0.029366   0.013324  -2.204 0.027518 *  
## factor(labour_state)Family Care      0.207191   0.057175   3.624 0.000290 ***
## factor(labour_state)Maternity Leave -0.302883   0.501065  -0.604 0.545525    
## factor(labour_state)PT Employed      0.039547   0.044232   0.894 0.371273    
## factor(labour_state)Retired         -0.028738   0.042805  -0.671 0.501989    
## factor(labour_state)Self-employed    0.006792   0.047900   0.142 0.887239    
## factor(labour_state)Sick/Disabled    0.107874   0.044940   2.400 0.016377 *  
## factor(labour_state)Student         -0.391335   0.101514  -3.855 0.000116 ***
## factor(labour_state)Unemployed       0.052690   0.047488   1.110 0.267196    
## 
## Zero-inflation model coefficients (binomial with logit link):
##                                             Estimate Std. Error z value
## (Intercept)                                 -1.74690    0.06856 -25.480
## relevel(factor(ethnicity), ref = "WBI")BAN  -0.14975    0.51269  -0.292
## relevel(factor(ethnicity), ref = "WBI")BLA  -0.65053    0.86881  -0.749
## relevel(factor(ethnicity), ref = "WBI")BLC   0.31791    0.40897   0.777
## relevel(factor(ethnicity), ref = "WBI")CHI   1.51817    0.70312   2.159
## relevel(factor(ethnicity), ref = "WBI")IND   0.52296    0.36411   1.436
## relevel(factor(ethnicity), ref = "WBI")MIX   0.20122    0.35581   0.566
## relevel(factor(ethnicity), ref = "WBI")OAS   0.93407    0.60401   1.546
## relevel(factor(ethnicity), ref = "WBI")OBL -13.01639  538.48954  -0.024
## relevel(factor(ethnicity), ref = "WBI")OTH  -0.19411    0.99293  -0.195
## relevel(factor(ethnicity), ref = "WBI")PAK   0.03483    0.37297   0.093
## relevel(factor(ethnicity), ref = "WBI")WHO  -0.17955    0.31947  -0.562
## scale(age)                                  -0.31921    0.06479  -4.927
##                                            Pr(>|z|)    
## (Intercept)                                 < 2e-16 ***
## relevel(factor(ethnicity), ref = "WBI")BAN   0.7702    
## relevel(factor(ethnicity), ref = "WBI")BLA   0.4540    
## relevel(factor(ethnicity), ref = "WBI")BLC   0.4370    
## relevel(factor(ethnicity), ref = "WBI")CHI   0.0308 *  
## relevel(factor(ethnicity), ref = "WBI")IND   0.1509    
## relevel(factor(ethnicity), ref = "WBI")MIX   0.5717    
## relevel(factor(ethnicity), ref = "WBI")OAS   0.1220    
## relevel(factor(ethnicity), ref = "WBI")OBL   0.9807    
## relevel(factor(ethnicity), ref = "WBI")OTH   0.8450    
## relevel(factor(ethnicity), ref = "WBI")PAK   0.9256    
## relevel(factor(ethnicity), ref = "WBI")WHO   0.5741    
## scale(age)                                 8.35e-07 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
## 
## Number of iterations in BFGS optimization: 30 
## Log-likelihood: -5657 on 25 Df
\end{verbatim}

\end{document}