### 2022-12-19 docsrc/documentation/notebooks/household_income.Rmd
automate rendering of R notebooks into sphinx. Rmd references as well using pandoc.

---
title: "Household_Income"
output: html_document
date: "2022-12-16"
bibliography: references.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Household Income

## Data

## Methods

## Results