### 2022-12-19 docsrc/documentation/notebooks/summarise_transitions_2.Rmd

---
title: "MINOS Transition Models"
output: html_document
bibliography: 
  - "refs.bib"
---

# Tobacco Coefficients

Tobacco consumption is measured in the usual number of cigarettes smoked per day, and is taken directly from a US variable ([ncigs](https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/variable/ncigs)).

Test references [@1987:nelson].

We estimated a zero inflated poisson model, with the formula:

> ncigs \~ age + sex + SF_12 + labour_state + job_sec + ethnicity + hh_income \| ethnicity + labour_state + SF_12

```{r tobacco_output}
options(width=120)
require(tidyverse)
require(pscl)

workingDir <- "../../.." #TODO stored in docsrc/Coefficients for now. move to data/transitions and change rst filepaths accordingly.
knitr::opts_knit$set(root.dir = workingDir)
trans.path <- paste0(workingDir,'/data/transitions/test/')
tobacco <- readRDS(paste0(trans.path, 'tobacco_zip_2018_2019.rds'))
summary(tobacco)
```

References
----------