### 2022-12-19 docsrc/documentation/notebooks/tobacco.Rmd

---
title: "Tobacco Transition Module"
output: html_document
bibliography: 
  - "refs.bib"  # switch bibliography position. depending on where the file is being run. stick to this one when compiiling the website using the makefile
  #- "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging.
---

# Tobacco

Number of cigarettes consumed is an indicator of XXXX. test reference [@1987:nelson].

```{r setup, include=FALSE}
source('sphinx_notebook_utils.R')
```

## Methods

  What methods are used? Justification due to output data type. explanation of model output.

## Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

## Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r tobacco_output, echo=FALSE}
options(width=120)
zip_output("../../../data/transitions/test/tobacco_zip_2018_2019.rds")
```

## References