### 2022-12-21 docsrc/documentation/notebooks/neighbourhood.Rmd

---
title: "Neighbourhood"
output: html_document
bibliography: 
  #- "refs.bib"  # switch bibliography position. depending on where the file is being run. stick to this one when compiling the website using the makefile.
  - "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging.
---

## Neighbourhood

Introductory fluff. Why do we need this module? test reference [@1987:nelson].

```{r setup, include=FALSE}
source('sphinx_notebook_utils.R')
```

### Methods

  What methods are used? Justification due to output data type. explanation of model output.

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r housing_output, echo=FALSE}
options(width=120)
clm_output("../../../data/transitions/test/neighbourhood_clm_2014_2017.rds") #e.g. ../../../data/transitions/test/housing_clm_2018_2019.rds
```

### References