### 2023-01-06 docsrc/documentation/notebooks/loneliness.Rmd
update naming in some transition modules. improve website some more

---
title: "Loneliness"
output: html_document
bibliography: 
  - "refs.bib"  # bib reference depending on where the file is being run. Use this one when compiling the website using the makefile.
  #- "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging. Make sure to switch back when done..
---

## Loneliness

Introductory fluff. Why do we need this module? test reference [@1987:nelson].

```{r setup, include=FALSE}
source('sphinx_notebook_utils.R')
```

### Methods

  What methods are used? Justification due to output data type. explanation of model output.

```{r loneliness_data}
discrete_barplot("loneliness")
```

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r housing_output, echo=FALSE}
options(width=120)
clm_barplot("../../../data/transitions/test/loneliness_clm_2018_2019.rds", "loneliness_next") 
clm_output("../../../data/transitions/test/loneliness_clm_2018_2019.rds") #e.g. ../../../data/transitions/test/housing_clm_2018_2019.rds
```

### References