### 2023-01-17 docsrc/documentation/notebooks/mental_well_being.Rmd

---
title: "Mental Well-Being"
output: html_document
bibliography: 
  - "refs.bib" # switch bibliography position. depending on where the file is being run. stick to this one when compiiling the website using the makefile
  #- "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging.
---

## Mental Well-Being

Introductory fluff. Why do we need this module? test reference [@1987:nelson].

```{r setup, include=FALSE}
source('sphinx_notebook_utils.R')
```

### Methods

  What methods are used? Justification due to output data type. explanation of model output.

```{r neighbourhood_barchart, echo=FALSE}
continuous_density("SF_12")
```
### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r housing_output, echo=FALSE}
options(width=120)
ols_histogram("../../../data/transitions/test/sf12_ols_2018_2019.rds", "y")
ols_output("../../../data/transitions/test/sf12_ols_2018_2019.rds")
```

### References