### 2023-02-20 docsrc/documentation/notebooks/loneliness.Rmd

---
title: "Loneliness"
output: html_document
bibliography: 
  - "docsrc/refs.bib"  # bib reference depending on where the file is being run. Use this one when compiling the website using the makefile.
  #- "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging. Make sure to switch back when done..
---

## Loneliness

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir= "../minos")
```

Introductory fluff. Why do we need this module? test reference [@1987:nelson].

```{r load_data, include=FALSE}
source('docsrc/documentation/notebooks/sphinx_notebook_utils.R')
```

### Methods

  What methods are used? Justification due to output data type. explanation of model output.

```{r loneliness_data}
discrete_barplot("data/transitions/loneliness/clm/loneliness_2018_2019.rds", 'next_loneliness')
```

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r housing_output, echo=FALSE}
options(width=120)
clm_barplot("data/transitions/loneliness/clm/loneliness_2018_2019.rds", "next_loneliness") 
clm_output("data/transitions/loneliness/clm/loneliness_2018_2019.rds")
```

### References