### 2023-02-21 docsrc/documentation/notebooks/labour.Rmd

---
title: "Labour"
output: html_document
bibliography: 
  #- "docsrc/refs.bib"  # switch bibliography position. depending on where the file is being run. stick to this one when compiling the website using the makefile.
  - "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging.
---

```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir= "../minos")
```

```{r load_data, include=FALSE}
source('docsrc/documentation/notebooks/sphinx_notebook_utils.R')
```

## Labour

Labour state is a measure of what an individual does. There are 8 distinctive categories including employment, unemployment, and retired. 
The encodings of these states can be found [here]](https://leeds-mrg.github.io/Minos/documentation/data_tables.html).

```{r labour_barchart, echo=FALSE}
discrete_barplot("labour_state")
```

### Methods

Labour state is a complex categorical data type. Single layer neural network is a simple way to estimate this state. Use multinom function from R's nnet package. Formula for weights included given as.

$$labour\_state\_next = sex + ethnicity + age + education\_state + SF\_12 + housing\_quality + labour\_state + job\_sec + hh\_income + alcohol\_spending$$

### Data

Variables used in this model. Encodings for discrete variables found in data tables.

- sex. Biological sex male/female.
- ethnicity. Ethnicity e.g. white british. XXXX cite.
- age in years. XXXX cite.
- education. Highest qualification attained. XXXX cite 
- sf12. Mental well-being score. XXXX cite
- housing quality. Number of household appliances. Ordinal 1-3. XXXX cite
- labour_state. Previous labour state. XXXX cite. Probably remove this. dominates prediction..
- nssec. Socioeconomic code of job. Indicates job quality with 1-9 ordinal variable. XXXX cite
- household income. Monthly disposable income of individuals household. XXXX cite.
- alcohol spending. How much is spent on alcohol. XXXX cite. remove.

### Results

  - hard to determine goodness of fit. 
  - use confusion matrix to estimate quality of fit. 
  - employed/retired well predicted. unemployed/student volatile socially and expectedly hard to predict. 
  - some deterministic replacement needed for categories like student that have specific time frames. e.g. three years for a degree.

```{r labour_output, echo=FALSE}
options(width=120)
#nnet_output("data/transitions/labour/nnet/labour_nnet_2018_2019.rds")
nnet_confusion_matrix("data/transitions/labour/nnet/labour_nnet_2018_2019.rds", mode='multinom', 'labour_state_next')
```

### References