### 2023-02-27 docsrc/documentation/notebooks/neighbourhood.Rmd

---
title: "Neighbourhood"
output: html_document
bibliography: 
  - "docsrc/refs.bib"  # switch bibliography position. depending on where the file is being run. stick to this one when compiling the website using the makefile.
  #- "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging.
---

```{r load_data, include=FALSE}
knitr::opts_knit$set(root.dir= "../minos")
```

## Neighbourhood

Neighbourhood quality is an indicator of well-being suggested by the SIPHER-7 XXXX ref and others?.

```{r neighbourhood_setup, include=FALSE}
source('docsrc/documentation/notebooks/sphinx_notebook_utils.R')
```

### Methods

```{r neighbourhood_barchart, echo=FALSE}
discrete_barplot("data/transitions/neighbourhood_safety/clm/neighbourhood_safety_2014_2017.rds", 'next_neighbourhood_safety')
```

  Neighbourhood data are presented as a likert scale from 1-6 indicating the number of harmful activities that happen in a persons
  neighbourhood for individual health. These include robberies and vandalism with the full encoding available in the data tables.

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r housing_output, echo=FALSE}
options(width=120)
clm_barplot("data/transitions/neighbourhood_safety/clm/neighbourhood_safety_2014_2017.rds", "next_neighbourhood_safety") 
clm_output("data/transitions/neighbourhood_safety/clm/neighbourhood_safety_2014_2017.rds")  
```

### References