### 2023-03-01 minos/validation/mwb_pathways.Rmd
Added WIP notebook for looking at the mental wellbeing pathway variables in the outputs

---
title: "Mental Wellbeing Pathways"
output: html_notebook
---

# Setup

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(purrr)

workingDir <- "/home/luke/Documents/WORK/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

Source utils for some functions.
```{r}
source('R_Scripts/utils.r')
```

## Data

```{r}
output.path <- 'Minos/output/default_config/'
baseline.path <- paste0(output.path, 'baseline')
base.path <- get_runtime_subdirectory(baseline.path)

# read in all the output data files
base.files <- list.files(base.path,
                         pattern = '[0-9]{4}.csv',
                         full.names = TRUE)
data <- do.call(rbind, lapply(base.files, read.csv))
```

# Analysis

## ncigs

DONE: Fixed problem where the rescale factor was still being applied in prediction but not in the estimation of the model.
DONE: Fixed line in the r_utils prediction method where the non-smokers were being predicted as smokers.
TODO: Fix the input data issue where lots of these vars are being reported as missing still (negative numbers). This should be in the run script somewhere.

```{r}
ncigs <- data %>% 
  select(pidp, time, ncigs)

# calculate and plot mean over time
ncigs.byyr <- data %>% group_by(time) %>%
  summarise(ncigs = mean(ncigs))

ggplot(data = ncigs.byyr, mapping = aes(x = time, y = ncigs)) +
  geom_line()
```

```{r}
# Lets overlay some histograms at 3 time points - 2018 (before model), 2020 (pathways kick in?), 2030 (long term)
ncigs.2018 <- ncigs %>% filter(time == 2018)
ncigs.2020 <- ncigs %>% filter(time == 2020)
ncigs.2030 <- ncigs %>% filter(time == 2030)

ncigs.joined <- rbind(ncigs.2018, ncigs.2020, ncigs.2030)
ncigs.joined$time <- as.factor(ncigs.joined$time)

ggplot(data = ncigs.joined, mapping = aes(x = ncigs, group = time, fill = time, color = time)) +
  geom_histogram(position='dodge')
```

```{r}
# Lets also look at the number of smokers vs non-smokers
smokers.count <- ncigs %>% 
  group_by(time) %>%
  summarise(non.smoker = sum(ncigs == 0),
            smoker = sum(ncigs > 0)) %>%
  pivot_longer(cols = non.smoker:smoker,
               names_to = 'smoking_status',
               values_to = 'count')

ggplot(data = smokers.count, mapping = aes(x = time, y = count, group = smoking_status, color = smoking_status)) +
  geom_line()
```

## Housing Quality

```{r}
# plot counts of each state over time (will only change with the replenishing but good to see)
housing.state <- data %>% 
  group_by(time, housing_quality) %>%
  count()

# set column types
housing.state$housing_quality <- as.factor(housing.state$housing_quality)

ggplot(data = housing.state, mapping = aes(x = time, y = n, group = housing_quality, color = housing_quality)) +
  geom_line()
```

## Neighbourhood Safety

```{r}
# plot counts of each state over time (will only change with the replenishing but good to see)
neighbour.state <- data %>% 
  group_by(time, neighbourhood_safety) %>%
  count()

# set column types
neighbour.state$neighbourhood_safety <- as.factor(neighbour.state$neighbourhood_safety)

ggplot(data = neighbour.state, mapping = aes(x = time, y = n, group = neighbourhood_safety, color = neighbourhood_safety)) +
  geom_line()
```

## Loneliness

```{r}
# plot counts of each state over time (will only change with the replenishing but good to see)
loneliness.state <- data %>% 
  group_by(time, loneliness) %>%
  count()

# set column types
loneliness.state$loneliness <- as.factor(loneliness.state$loneliness)

ggplot(data = loneliness.state, mapping = aes(x = time, y = n, group = loneliness, color = loneliness)) +
  geom_line()
```

```{r}

```