### 2023-03-01 minos/validation/interventions.Rmd
Added plots assessing the how the pathways react to the energy downlift intervention

---
title: "Investigating Interventions"
output: html_notebook
---

# Setup

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(purrr)
require(ordinal)
require(nnet)
require(pscl)

workingDir <- "/home/luke/Documents/WORK/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

Source utils for some functions.
```{r}
source('Minos/minos/validation/utils.r')
```

## Data

```{r}
out.path <- 'Minos/output/default_config/'
#out.path <- 'Minos/output/scotland_mode/'

base <- read_output(out.path, 'baseline')
allchild <- read_output(out.path, 'hhIncomeChildUplift')
poverty <- read_output(out.path, 'hhIncomePovertyLineChildUplift')
wage <- read_output(out.path, 'livingWageIntervention')
energy <- read_output(out.path, 'energyDownlift')

```

## Transition Models

```{r, include=FALSE}
trans.path <- 'Minos/data/transitions/'

educ.mod <- readRDS(paste0(trans.path, 'education_state/nnet/education_state_2018_2019.rds'))
sf_12.mod <- readRDS(paste0(trans.path, 'SF_12/ols/SF_12_2018_2019.rds'))
house.mod <- readRDS(paste0(trans.path, 'housing_quality/clm/housing_quality_2018_2019.rds'))
income.mod <- readRDS(paste0(trans.path, 'hh_income/ols/hh_income_2018_2019.rds'))
nut.mod <- readRDS(paste0(trans.path, 'nutrition_quality/ols/nutrition_quality_2018_2019.rds'))
```

# Energy Downlift

## Household Income

```{r}
income.base <- base %>% 
  select(pidp, time, hh_income)

income.energy <- energy %>% 
  select(pidp, time, hh_income, income_boosted)

# first rename and cbind
income.b <- rename(income.base, base = hh_income)
income.e <- rename(income.energy, energy = hh_income)

income.merged <- merge(income.b, income.e, by=c('pidp', 'time'))

income.merged <- pivot_longer(data = income.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'hh_income')

income.merged.byyr <- income.merged %>%
  group_by(time, scenario) %>%
  summarise(hh_income = mean(hh_income))

ggplot(data = income.merged.byyr, mapping = aes(x = time, y = hh_income, color=scenario)) +
  geom_line()
```

## SF_12

```{r}
sf12.base <- base %>% 
  select(pidp, time, SF_12)

sf12.energy <- energy %>% 
  select(pidp, time, SF_12, income_boosted)

# first rename and cbind
sf12.b <- rename(sf12.base, base = SF_12)
sf12.e <- rename(sf12.energy, energy = SF_12)

sf12.merged <- merge(sf12.b, sf12.e, by=c('pidp', 'time'))

sf12.merged <- pivot_longer(data = sf12.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'SF_12')

sf12.merged.byyr <- sf12.merged %>%
  group_by(time, scenario) %>%
  summarise(SF_12 = mean(SF_12))

ggplot(data = sf12.merged.byyr, mapping = aes(x = time, y = SF_12, color=scenario)) +
  geom_line()
```

## Pathways

### Housing Quality

```{r}
housing.base <- base %>% 
  select(pidp, time, housing_quality)

housing.energy <- energy %>% 
  select(pidp, time, housing_quality, income_boosted)

# first rename and cbind
housing.b <- rename(housing.base, base = housing_quality)
housing.e <- rename(housing.energy, energy = housing_quality)

housing.merged <- merge(housing.b, housing.e, by=c('pidp', 'time'))

housing.merged <- pivot_longer(data = housing.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'housing')

housing.merged.byyr <- housing.merged %>%
  group_by(time, scenario) %>%
  summarise(housing = mean(housing))

ggplot(data = housing.merged.byyr, mapping = aes(x = time, y = housing, color=scenario)) +
  geom_line()
```

### Neighbourhood Safety

```{r}
neighbourhood.base <- base %>% 
  select(pidp, time, neighbourhood_safety)

neighbourhood.energy <- energy %>% 
  select(pidp, time, neighbourhood_safety, income_boosted)

# first rename and cbind
neighbourhood.b <- rename(neighbourhood.base, base = neighbourhood_safety)
neighbourhood.e <- rename(neighbourhood.energy, energy = neighbourhood_safety)

neighbourhood.merged <- merge(neighbourhood.b, neighbourhood.e, by=c('pidp', 'time'))

neighbourhood.merged <- pivot_longer(data = neighbourhood.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'neighbourhood')

neighbourhood.merged.byyr <- neighbourhood.merged %>%
  group_by(time, scenario) %>%
  summarise(neighbourhood = mean(neighbourhood))

ggplot(data = neighbourhood.merged.byyr, mapping = aes(x = time, y = neighbourhood, color=scenario)) +
  geom_line()
```

### Tobacco Use

```{r}
ncigs.base <- base %>% 
  select(pidp, time, ncigs)

ncigs.energy <- energy %>% 
  select(pidp, time, ncigs, income_boosted)

# first rename and cbind
ncigs.b <- rename(ncigs.base, base = ncigs)
ncigs.e <- rename(ncigs.energy, energy = ncigs)

ncigs.merged <- merge(ncigs.b, ncigs.e, by=c('pidp', 'time'))

ncigs.merged <- pivot_longer(data = ncigs.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'ncigs')

ncigs.merged.byyr <- ncigs.merged %>%
  group_by(time, scenario) %>%
  summarise(ncigs = mean(ncigs))

ggplot(data = ncigs.merged.byyr, mapping = aes(x = time, y = ncigs, color=scenario)) +
  geom_line()
```

### Nutrition Quality

```{r}
nut.base <- base %>% 
  select(pidp, time, nutrition_quality)

nut.energy <- energy %>% 
  select(pidp, time, nutrition_quality, income_boosted)

# first rename and cbind
nut.b <- rename(nut.base, base = nutrition_quality)
nut.e <- rename(nut.energy, energy = nutrition_quality)

nut.merged <- merge(nut.b, nut.e, by=c('pidp', 'time'))

nut.merged <- pivot_longer(data = nut.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'nutrition')

nut.merged.byyr <- nut.merged %>%
  group_by(time, scenario) %>%
  summarise(nutrition = mean(nutrition))

ggplot(data = nut.merged.byyr, mapping = aes(x = time, y = nutrition, color=scenario)) +
  geom_line()
```

### Loneliness (NOT CURRENTLY IN SF12 MODEL)

```{r}
lnly.base <- base %>% 
  select(pidp, time, loneliness)

lnly.energy <- energy %>% 
  select(pidp, time, loneliness, income_boosted)

# first rename and cbind
lnly.b <- rename(lnly.base, base = loneliness)
lnly.e <- rename(lnly.energy, energy = loneliness)

lnly.merged <- merge(lnly.b, lnly.e, by=c('pidp', 'time'))

lnly.merged <- pivot_longer(data = lnly.merged,
                              cols = base:energy,
                              names_to = 'scenario',
                              values_to = 'loneliness')

lnly.merged.byyr <- lnly.merged %>%
  group_by(time, scenario) %>%
  summarise(loneliness = mean(loneliness))

ggplot(data = lnly.merged.byyr, mapping = aes(x = time, y = loneliness, color=scenario)) +
  geom_line()
```

## Boost Amount

```{r}
income.energy <- energy %>% 
  select(pidp, time, hh_income, income_boosted, boost_amount)

income.energy.byyr <- income.energy %>%
  group_by(time) %>%
  summarise(boost_amount = mean(boost_amount))

ggplot(data = income.energy.byyr, mapping = aes(x = time, y = boost_amount)) +
  geom_line()
```

## Who Boosted?

```{r}
boosted.byyr <- income.energy %>% 
  group_by(time) %>%
  count(income_boosted)

ggplot(data = filter(boosted.byyr, income_boosted == 'True'), mapping = aes(x = time, y = n)) +
  geom_line()
```

# Living Wage

## Household Income

```{r}
income.base <- base %>% 
  select(pidp, time, hh_income)

income.wage <- wage %>% 
  select(pidp, time, hh_income, income_boosted)

# first rename and cbind
income.b <- rename(income.base, base = hh_income)
income.w <- rename(income.wage, wage = hh_income)

income.merged <- merge(income.b, income.w, by=c('pidp', 'time'))

income.merged <- pivot_longer(data = income.merged,
                              cols = base:wage,
                              names_to = 'scenario',
                              values_to = 'hh_income')

income.merged.byyr <- income.merged %>%
  group_by(time, scenario) %>%
  summarise(hh_income = mean(hh_income))

ggplot(data = income.merged.byyr, mapping = aes(x = time, y = hh_income, color=scenario)) +
  geom_line()
```

## Boost Amount

```{r}
income.wage <- wage %>% 
  select(pidp, time, hh_income, income_boosted, boost_amount)

income.wage.byyr <- income.wage %>%
  group_by(time) %>%
  summarise(boost_amount = mean(boost_amount))

ggplot(data = income.wage.byyr, mapping = aes(x = time, y = boost_amount)) +
  geom_line()
```

## Who Boosted?

```{r}
boosted.byyr <- income.wage %>% 
  group_by(time) %>%
  count(income_boosted)

ggplot(data = filter(boosted.byyr, income_boosted == 'True'), mapping = aes(x = time, y = n)) +
  geom_line()

ggplot(data = boosted.byyr, mapping = aes(x = time, y = n, color = income_boosted)) +
  geom_line()
```

# All Child Uplift

## Household Income

```{r}
income.base <- base %>% 
  select(pidp, time, hh_income)

income.child <- allchild %>% 
  select(pidp, time, hh_income, income_boosted)

# first rename and cbind
income.base <- rename(income.base, base = hh_income)
income.child <- rename(income.child, allchild = hh_income)

income.merged <- merge(income.base, income.child, by=c('pidp', 'time'))

income.merged <- pivot_longer(data = income.merged,
                              cols = base:allchild,
                              names_to = 'scenario',
                              values_to = 'hh_income')

income.merged.byyr <- income.merged %>%
  group_by(time, scenario) %>%
  summarise(hh_income = mean(hh_income))

ggplot(data = income.merged.byyr, mapping = aes(x = time, y = hh_income, color=scenario)) +
  geom_line()
```

## Boost Amount

```{r}
income.child <- allchild %>% 
  select(pidp, time, hh_income, income_boosted, boost_amount)

income.child.byyr <- income.child %>%
  group_by(time) %>%
  summarise(boost_amount = mean(boost_amount))

ggplot(data = income.child.byyr, mapping = aes(x = time, y = boost_amount)) +
  geom_line()
```

## Who Boosted?

```{r}
boosted.byyr <- income.child %>% 
  group_by(time) %>%
  count(income_boosted)

ggplot(data = filter(boosted.byyr, income_boosted == 'True'), mapping = aes(x = time, y = n)) +
  geom_line()

ggplot(data = boosted.byyr, mapping = aes(x = time, y = n, color = income_boosted)) +
  geom_line()
```

# END

```{r}

```