### 2023-03-08 minos/validation/cross_validation.Rmd
Created functions for processing data (combining multiple sources, pivoting, and summarising), producing plots to compare means, and producing box plots. Also have some code for running t-tests

---
title: "Cross Validation Analysis"
output: html_notebook
---

This notebook will compare the results of the cross-validation simulation run with real data from Understanding Society, so that we can evaluate the transition models.

### Method

After generation of the final datasets (final_US), the respondents were separated into 2 roughly equal halves based on their pidp. One half was labelled as the transition dataset, and the other for simulation. Transition models were then fit using the transition dataset, with the other half then simulated with these models.

The results of the simulation will now be compared to the simulation dataset to see how closely our models mimick reality.

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)

workingDir <- "/home/luke/Documents/WORK/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

Source utils for some functions.

```{r}
source('Minos/minos/validation/utils.r')
```

## Data

```{r}
# cross validation simulation output
out.path <- 'Minos/output/cross_validation/'
cv <- read_output(out.path, 'baseline')

# cv simulation pop (not used for transitions)
cv.simul.files <- list.files('Minos/data/final_US/cross_validation/simulation/',
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw <- do.call(rbind, lapply(cv.simul.files, read.csv))
```

# Analysis

First things first, we need to make sure that we are comparing the exact same people from each dataset. Therefore, we need a step here to check the pidps from both, and remove anybody not in both datasets.

```{r}
cv.pidps <- cv %>% select(pidp) %>% unique()
raw.pidps <- raw %>% select(pidp) %>% unique()

both.pidps <- intersect(cv.pidps, raw.pidps)

cv <- cv %>% filter(pidp %in% both.pidps$pidp)
raw <- raw %>% filter(pidp %in% both.pidps$pidp)
```

## hh_income

### Mean

```{r}
income.pivoted <- combine_and_pivot_long(cv, 'simulated', raw, 'raw', 'hh_income')

income.merged.byyr <- income.pivoted %>%
  group_by(time, scenario) %>%
  summarise(hh_income = mean(hh_income))

ggplot(data = income.merged.byyr, mapping = aes(x = time, y = hh_income, color=scenario)) +
  geom_line()
```

### T-test

```{r}
qqnorm(cv$hh_income)
qqnorm(raw$hh_income)

# Check variance of both
cv.inc.var <- var(cv$hh_income)
raw.inc.var <- var(raw$hh_income, na.rm = TRUE)

# if ratio of largest variance to smallest is less than 4 then assume variance is equal for t.test
(raw.inc.var / cv.inc.var) < 4

test.t <- t.test(cv$hh_income, raw$hh_income, var.equal = TRUE)

for(year in unique(cv$time)) {
  print(paste('T tests for year', year))
  # filter the year
  tmp.cv <- cv %>% filter(time == year)
  tmp.raw <- raw %>% filter(time == year)

  # check variance and test for equal variance
  var.tmp.cv <- var(tmp.cv$hh_income)
  var.tmp.raw <- var(tmp.raw$hh_income, na.rm = TRUE)

  if(var.tmp.raw > var.tmp.cv) {
    equal.var <- (var.tmp.cv / var.tmp.raw) < 4
  } else {
    equal.var <- (var.tmp.raw / var.tmp.cv) < 4
  }

  print(t.test(tmp.cv$hh_income, tmp.raw$hh_income, var.equal = equal.var))
}
```

```{r}
compare_boxplot_long(income.pivoted,
                    var = 'hh_income',
                    scen1 = 'raw',
                    scen2 = 'simulation',
                    subset.max = 5000,
                    subset.min = -5000)
```

## SF12

```{r}
SF_12.cv <- cv %>% 
  select(pidp, time, SF_12)

SF_12.raw <- raw %>%
  select(pidp, time, SF_12)

# first rename and cbind
SF_12.cv <- rename(SF_12.cv, simulation = SF_12)
SF_12.raw <- rename(SF_12.raw, raw = SF_12)

SF_12.merged <- merge(SF_12.cv, SF_12.raw, by=c('pidp', 'time'))

SF_12.merged <- pivot_longer(data = SF_12.merged,
                              cols = simulation:raw,
                              names_to = 'scenario',
                              values_to = 'SF_12')

SF_12.merged.byyr <- SF_12.merged %>%
  group_by(time, scenario) %>%
  summarise(SF_12 = mean(SF_12))

ggplot(data = SF_12.merged.byyr, mapping = aes(x = time, y = SF_12, color=scenario)) +
  geom_line()
```

## Housing Quality

```{r}

```