### 2023-05-04 minos/outcomes/equivalent_income_out.Rmd
Created an RNotebook for visualising outputs from equivalent income scenarios, including comparisons of equivalent income between interventions and baseline for both the full population and treated population

---
title: "Equivalent Income Outcome Visualisation"
output: html_notebook
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)

workingDir <- "/home/luke/Documents/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

```{r, include=FALSE}
source('Minos/minos/validation/utils.r')
```

# Data

```{r}
# set output path to SIPHER 7 experiment out
out.path <- 'Minos/output/SIPHER7/'
# read all files into dataframe (util function handles runtime subdirectory etc.)
base.dat <- read_output(out.path, 'baseline')
allchild.dat <- read_output(out.path, 'hhIncomeChildUplift')
pov.dat <- read_output(out.path, 'hhIncomePovertyLineChildUplift')
wage.dat <- read_output(out.path, 'livingWageIntervention')
energy.dat <- read_output(out.path, 'energyDownlift')
```

# Baseline

```{r}
base.subset <- base.dat %>%
  select(pidp, time, equivalent_income) %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income))

base.subset$scenario <- 'Baseline'

ggplot(data = base.subset, aes(x = time, y = equivalent_income)) +
  geom_line()
```

# Comparisons

## Function

```{r}
equivalent_income_comparison_fullpop <- function(base, int, int.name, save=FALSE, save.path=NULL) {
  base.subset <- base %>%
    select(pidp, time, equivalent_income, alive) %>%
    filter(alive != 'dead') %>%
    group_by(time) %>%
    summarise(equivalent_income = mean(equivalent_income))
  int.subset <- int %>%
    select(pidp, time, equivalent_income, alive) %>%
    filter(alive != 'dead') %>%
    group_by(time) %>%
    summarise(equivalent_income = mean(equivalent_income))

  base.subset$scenario <- 'Baseline'
  int.subset$scenario <- int.name

  combined <- rbind(base.subset, int.subset)

  p <- ggplot(combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_line() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Full Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)

  if(save) {
    if(is.null(save.path)) {
      stop('ERROR: save.path cannot be missing if save=TRUE')
    }
    file.name <- paste0('EI_comp_', int.name, '_full.png')
    ggsave(filename = file.name,
           plot = p,
           path = save.path)
  }
}
```

## Plots

```{r}
equivalent_income_comparison_fullpop(base.dat, allchild.dat, 'allChild', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_fullpop(base.dat, pov.dat, 'povertyChild', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_fullpop(base.dat, wage.dat, 'LivingWage', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_fullpop(base.dat, energy.dat, 'EnergyDownlift', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
```

## Combined

```{r}
base.subset <- base.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income))
base.subset$scenario <- 'Baseline'

allchild.subset <- allchild.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income))
allchild.subset$scenario <- 'AllChild'

pov.subset <- pov.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income))
pov.subset$scenario <- 'PovertyChild'

wage.subset <- wage.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income))
wage.subset$scenario <- 'LivingWage'

energy.subset <- energy.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income))
energy.subset$scenario <- 'EnergyDownlift'

all.combined <- rbind(base.subset, allchild.subset, pov.subset, wage.subset, energy.subset)
all.combined$scenario <- factor(all.combined$scenario, levels = c('Baseline', 'AllChild', 'PovertyChild', 'LivingWage', 'EnergyDownlift'))

ggplot(data = all.combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line() +
  labs(title = 'Equivalent Income Comparison', subtitle = 'Whole Population') +
  xlab('Year') +
  ylab('Equivalent Income (£)')
```

# Treatment on the Treated

## Function

```{r}
equivalent_income_comparison_treated <- function(base, int, int.name, save=FALSE, save.path=NULL) {
  base.sub <- base.dat %>%
    select(pidp, time, equivalent_income, alive) %>%
    filter(alive != 'dead') %>%
    select(-alive) %>%
    rename(baseline = equivalent_income)
  int.sub <- int %>%
    select(pidp, time, equivalent_income, income_boosted, alive) %>%
    filter(alive != 'dead') %>%
    select(-alive) %>%
    set_names(c('pidp', 'time', int.name, 'income_boosted'))

  merged <- merge(base.sub, int.sub, by = c('pidp', 'time'))

  boosted.pidps <- int.sub %>%
    filter(income_boosted == 'True') %>%
    select(pidp)

  boosted <- merged %>%
    filter(pidp %in% boosted.pidps$pidp) %>%
    filter(income_boosted == 'True' | time == 2020) %>%
    select(-income_boosted) %>%
    pivot_longer(cols = baseline:.data[[int.name]],
                 names_to = 'scenario',
                 values_to = 'equivalent_income') %>%
    group_by(time, scenario) %>%
    summarise(equivalent_income = mean(equivalent_income))

  p <- ggplot(data = boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_line() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Treated Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)

  if(save) {
    if(is.null(save.path)) {
      stop('ERROR: save.path cannot be missing if save=TRUE')
    }
    file.name <- paste0('EI_comp_', int.name, '_treated.png')
    ggsave(filename = file.name,
           plot = p,
           path = save.path)
  }
}
```

## AllChild

```{r}
equivalent_income_comparison_treated(base.dat, allchild.dat, 'allChild', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_treated(base.dat, pov.dat, 'povertyChild', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_treated(base.dat, wage.dat, 'LivingWage', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_treated(base.dat, energy.dat, 'EnergyDownlift', save=TRUE, save.path = '/home/luke/Documents/MINOS/Minos/plots/Equivalent_Income/')
```

# Remove Income Effect

```{r}
# THink about this...
# We can use the income_boosted and boost_amount variables to remove the income boost from people and visualise that

base.sub2 <- base.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  select(-alive) %>%
  rename(baseline = equivalent_income)

allchild.sub2 <- allchild.dat %>%
  select(pidp, time, equivalent_income, income_boosted, boost_amount, alive) %>%
  filter(alive != 'dead') %>%
  select(-alive) %>%
  rename(allchild = equivalent_income)

allchild.merged <- merge(base.sub2, allchild.sub2, by = c('pidp', 'time'))

boosted.pidps <- allchild.sub2 %>%
  filter(income_boosted == 'True') %>%
  select(pidp)

allchild.boosted <- allchild.merged %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  filter(income_boosted == 'True' | time == 2020) %>%
  mutate(adjusted_EI = allchild - boost_amount) %>%
  select(-allchild, -income_boosted, -boost_amount) %>%
  group_by(time) %>%
  summarise(baseline = mean(baseline),
            allchild_adj = mean(adjusted_EI)) %>%
  pivot_longer(cols = baseline:allchild_adj,
               names_to = 'scenario',
               values_to = 'equivalent_income')

ggplot(data = allchild.boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line()

```

NOTE: Can we 'remove' the income effect?
Use the boost amount and income_boosted variables to sort of 'undo' the intervention effect and then we are only visualising the difference from pathway changes and not just the income effect.

ALSO: Difference plots.