### 2023-05-18 minos/testing/rate_of_change_testing.Rmd
Rate of change models in a good working state for both income and SF_12_MCS now, success! Also added some testing plots to testing/rate_of_change_testing.Rmd

---
title: "Testing Rate of Change Transition models"
output: html_notebook
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)

workingDir <- "/home/luke/Documents/WORK/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

```{r, include=FALSE}
source('Minos/minos/utils_datain.R')
source('Minos/minos/validation/utils.r')
```

## Data

```{r, include=FALSE}
# Read raw datafiles in
raw.files <- list.files('/home/luke/Documents/WORK/MINOS/Minos/data/final_US/', pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

out.path <- 'Minos/output/default_config/'
base.dat <- read_singular_local_out(out.path, 'baseline')
```

The purpose of this notebook will be to test rate of change models for use in Minos. Early attempts at the rate of change models have been relatively successful, seeing vast improvements in the regression to the mean issue that has been a problem for a while. However, even with the new rate of change models we are seeing some regression to the mean (or something that looks like it at least), with those earning the highest amounts seeing massive decreases and those earning very little or negative values seeing larger increases. What I think we need to do, is to investigate the actual rate of change across different income groups, as well as fitting these models here and seeing what a prediction looks like where we can actually investigate it. We can also plot some handovers of the rate of change from raw US to baseline outputs.

# Raw Data Analysis

## hh_income

First stop here will be to test the rate of change in raw understanding Society data across varying income levels.

```{r}
# difference between rows by group is more complicated than it sounds
# see https://stackoverflow.com/questions/14846547/calculate-difference-between-values-in-consecutive-rows-by-group for info
# appears dplyr::lag function could be the solution

# Can we just remove the first year? Nobody has a difference for first year obvs

raw <- raw.dat %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time)) %>%
  filter(time != 2009)

print(mean(raw$diff.hh_income, na.rm=TRUE))
print(sd(raw$diff.hh_income, na.rm=TRUE))

raw.sum <- raw %>%
  group_by(time) %>%
  summarise(diff.mean = mean(diff.hh_income, na.rm=TRUE),
            diff.sd = sd(diff.hh_income, na.rm=TRUE))

#raw.sum$time <- as.factor(raw.sum$time)

ggplot(data = raw.sum, aes(x = time, y = diff.mean)) +
  geom_line()

ggplot(data = raw.sum, aes(x = time, y = diff.sd)) +
  geom_line()

# Why wont this work with time as a factor?! Annoying!

rm(raw.sum)
```

So that is the mean rate of change and sd for the whole population. Now to do the same by income quintile.

QUESTION: Should I separate people into quintiles based on their average hh_income over all years? Or is it ok to stick with quintiles based on each year?
We are fitting one yearly models still (which is what need to be investi), so the rate of change per quintile per year is still valid I think.

### Yearly Quintile Assignment

```{r}
#raw.quintile <- raw %>%
#  mutate(rank = rank(hh_income, ties.method = "first"),
#         income.quint = cut(hh_income, breaks=quantile(rank, prob=c(0:5)/5, na.rm=TRUE), labels = c(1, 2, 3, 4, 5), include_lowest = TRUE))

# This is a pain in the arse
raw.quintile <- raw
raw.quintile$quintile <- cut(raw.quintile$hh_income, breaks=quantile(raw.quintile$hh_income, prob=c(0:5)/5, na.rm=TRUE), labels = c(1, 2, 3, 4, 5), include_lowest = TRUE)
# FINALLY

raw.quintile.sum <- raw.quintile %>%
  group_by(quintile) %>%
  summarise(diff.mean = mean(diff.hh_income, na.rm=TRUE),
            diff.sd = sd(diff.hh_income, na.rm=TRUE)) %>%
  filter(!is.na(quintile))

ggplot(data = raw.quintile.sum, aes(x = quintile, y = diff.mean, group=quintile, fill = quintile)) +
  geom_col()

ggplot(data = raw.quintile.sum, aes(x = quintile, y = diff.sd, group=quintile, fill = quintile)) +
  geom_col()

rm(raw.quintile.sum)
```

```{r}
raw.quintile.sum.time <- raw.quintile %>%
  group_by(quintile, time) %>%
  summarise(diff.mean = mean(diff.hh_income, na.rm=TRUE),
            diff.sd = sd(diff.hh_income, na.rm=TRUE)) %>%
  filter(!is.na(quintile))

ggplot(data = raw.quintile.sum.time, aes(x = time, y = diff.mean, group=quintile, color = quintile)) +
  geom_line()

ggplot(data = raw.quintile.sum.time, aes(x = time, y = diff.sd, group=quintile, color = quintile)) +
  geom_line()

rm(raw.quintile.sum.time)
```

### Whole Period Quintile Assignment

NOW CATEGORISE BY HH_INCOME QUINTILE ACROSS THE WHOLE TIME PERIOD.

```{r}
# Group people into income quintiles based on their mean income across the whole period
raw.time.quint <- raw %>%
  group_by(pidp) %>%
  summarise(hh_income = mean(hh_income, na.rm=TRUE)) %>%
  mutate(income.quint = cut(hh_income, breaks = quantile(x = hh_income, prob = c(0:5)/5), labels = c(1:5), include.lowest = TRUE)) %>%
  select(-hh_income)

# Add the income quintiles to the original raw data
raw.quint2 <- merge(raw, raw.time.quint, by='pidp')

# Now sort the data by time 
raw.quint2 <- raw.quint2 %>%
  select(pidp, time, age, sex, education_state, hh_income, diff.hh_income, income.quint) %>%
  group_by(pidp) %>%
  arrange(time, .by_group = TRUE)

#raw.quint2[order(raw.quint2$pidp, -raw.quint2$time), ]

raw.quintile.sum2 <- raw.quint2 %>%
  group_by(income.quint) %>%
  summarise(diff.mean = mean(diff.hh_income, na.rm = TRUE),
            diff.sd = sd(diff.hh_income, na.rm=TRUE))

#raw.quintile.sum2 <- raw.quint2 %>%
#  group_by(pidp) %>%
#  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time)) %>%
#  group_by(income.quint) %>%
#  summarise(diff.mean = mean(diff.hh_income, na.rm=TRUE),
#            diff.sd = sd(diff.hh_income, na.rm=TRUE)) %>%
#  filter(!is.na(income.quint))

ggplot(data = raw.quintile.sum2, aes(x = income.quint, y = diff.mean, group=income.quint, fill = income.quint)) +
  geom_col()

ggplot(data = raw.quintile.sum2, aes(x = income.quint, y = diff.sd, group=income.quint, fill = income.quint)) +
  geom_col()

rm(raw.time.quint, raw.quint2, raw.quintile.sum2)
```

## SF12

```{r}

diff_by_whole_period_quintile <- function(raw, var) {
  # first calculate diffs 
  raw <- raw %>%
    filter(!is.na(.data[[var]])) %>%  # remove missing
    filter(.data[[var]] >= 0) %>%  # remove negative missing codes
    group_by(pidp) %>%
    mutate(diff.var = .data[[var]] - lag(.data[[var]], order_by = time)) %>%
    filter(time != 2009) # remove 2009 as we can't have a diff in the first year

  #return(raw)

  raw.time.quint <- raw %>%
    group_by(pidp) %>%
    summarise(var.mean = mean(.data[[var]], na.rm = TRUE)) %>%
    mutate(var.quint = cut(var.mean, breaks = quantile(x = var.mean, prob = c(0:5)/5), labels = c(1:5), include.lowest = TRUE)) %>%
    select(-var.mean)

  #return(raw.time.quint)

  raw.quint <- merge(raw, raw.time.quint, by = 'pidp')

  raw.quint <- raw.quint %>%
    select(pidp, time, age, sex, education_state, hh_income, diff.var, var.quint) %>%
    group_by(pidp) %>%
    arrange(time, .by_group = TRUE)

  raw.quint.sum <- raw.quint %>%
    group_by(var.quint) %>%
    summarise(diff.mean = mean(diff.var, na.rm = TRUE),
              diff.sd = sd(diff.var, na.rm = TRUE),
              min = min(diff.var, na.rm = TRUE),
              max = max(diff.var, na.rm = TRUE))

  p1 <-ggplot(data = raw.quint.sum, aes(x = var.quint, y = diff.mean, group=var.quint, fill = var.quint)) +
    geom_col()

  p2 <- ggplot(data = raw.quint.sum, aes(x = var.quint, y = diff.sd, group=var.quint, fill = var.quint)) +
    geom_col()

  print(p1)
  print(p2)

  return(raw.quint.sum)
}

SF_12.stats <- diff_by_whole_period_quintile(raw.dat, 'SF_12')
```

```{r}
hh_income.stats <- diff_by_whole_period_quintile(raw.dat, 'hh_income')
```

```{r}
diff_by_year1_quintile <- function(raw, var) {
  # first calculate diffs 
  raw <- raw %>%
    filter(!is.na(.data[[var]])) %>%  # remove missing
    filter(.data[[var]] >= 0) %>%  # remove negative missing codes
    group_by(pidp) %>%
    mutate(diff.var = .data[[var]] - lag(.data[[var]], order_by = time))

  #return(raw)

  raw.time.quint <- raw %>%
    filter(time == 2009) %>%
    mutate(varx = .data[[var]],
           var.quint = cut(varx, breaks = unique(quantile(x = varx, prob = c(0:5)/5)), include.lowest = TRUE))# %>% , labels = c(1:5)
    #select(-var.mean)

  return(raw.time.quint)

  raw.quint <- merge(raw, raw.time.quint, by = 'pidp')

  raw.quint <- raw.quint %>%
    select(pidp, time, age, sex, education_state, hh_income, diff.var, var.quint) %>%
    group_by(pidp) %>%
    arrange(time, .by_group = TRUE)

  raw.quint.sum <- raw.quint %>%
    group_by(var.quint) %>%
    summarise(diff.mean = mean(diff.var, na.rm = TRUE),
              diff.sd = sd(diff.var, na.rm = TRUE),
              min = min(diff.var, na.rm = TRUE),
              max = max(diff.var, na.rm = TRUE))

  p1 <-ggplot(data = raw.quint.sum, aes(x = var.quint, y = diff.mean, group=var.quint, fill = var.quint)) +
    geom_col()

  p2 <- ggplot(data = raw.quint.sum, aes(x = var.quint, y = diff.sd, group=var.quint, fill = var.quint)) +
    geom_col()

  print(p1)
  print(p2)

  return(raw.quint.sum)
}

#hh_income.stats.yr1 <- diff_by_year1_quintile(raw.dat, 'hh_income')
```

```{r}
#test <- hh_income.stats.yr1

#test2 <- test %>%
#  filter(time == 2009)

#print(sum(is.na(test2$hh_income)))
#print(sum(test2$hh_income == -9))

#print(min(test2$hh_income))
#print(sum(test2$hh_income == 0))

#ggplot(data = test2, aes(x = hh_income)) +
#  geom_histogram()
```

# Comparing SF12_MCS Diff

```{r}
cv.path <- 'Minos/data/final_US/cross_validation/simulation/'
cv.outpath <- 'Minos/output/cross_validation/'

raw.files <- list.files(cv.path, pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.cv <- do.call(rbind, lapply(raw.files, read.csv))

base.cv <- read_singular_local_out(cv.outpath, 'baseline')

## MODELS
mod.path <- 'Minos/data/transitions/cross_validation/'
sf12.mod <- readRDS(paste0(mod.path, 'SF_12/ols2/SF_12_2017_2018.rds'))
income.mod <- readRDS(paste0(mod.path, 'hh_income/ols2/hh_income_2019_2020.rds'))

```

## Model Summaries

```{r}
summary(sf12.mod)
```
```{r}
summary(income.mod)
```

```{r}
miss.values <- c(-9, -8, -7, -3, -2, -1, -10)
raw <- raw.cv %>%
  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
  filter(time != 2009)

base <- base.cv %>%
  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
  filter(time != 2014)
```

```{r}
raw.sub <- raw %>%
  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
  mutate(source = 'raw')
base.sub <- base %>%
  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
  mutate(source = 'output')

combined <- rbind(raw.sub, base.sub)

combined.sum <- combined %>%
  group_by(time, source) %>%
  summarise(hh_income = mean(diff.hh_income, na.rm = TRUE),
            SF12_MCS = mean(diff.SF12_MCS, na.rm = TRUE))

ggplot(combined.sum, aes(x = time, y = hh_income, group = source, color = source)) +
  geom_line()

ggplot(combined.sum, aes(x = time, y = SF12_MCS, group = source, color = source)) +
  geom_line()
```

```{r}
test <- base.cv %>%
  group_by(time) %>%
  summarise(mean.income = mean(hh_income, na.rm = TRUE),
            mean.SF12 = mean(SF_12, na.rm = TRUE))

ggplot(test, aes(x = time, y = mean.income)) +
  geom_line()

ggplot(test, aes(x = time, y = mean.SF12)) +
  geom_line()
```

```{r}

```