### 2023-05-20 minos/outcomes/lineplot_notebook.Rmd
putting lineplots into one file and tidying a lot

---
title: "Untitled"
output: html_document
date: "2023-05-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
```

```{python}
# packages need installing in R using py_install("package"). honestly this seems not worth it..
from minos.outcomes.make_lineplots import main as lineplot_main
# define test parameters and run.
directories = "baseline,livingWageIntervention"
tags = "Baseline,Living Wage Intervention"
subset_function_strings = "who_below_living_wage,who_boosted"
prefix="baseline_living_wage_"
lineplot_main(directories, tags, subset_function_strings, prefix, mode='default_config', ref="Baseline", v="SF_12", method='nanmean')
```