### 2023-05-24 minos/outcomes/lineplot_notebook.Rmd
getting notebook working for all interventions and replenishment columns for default and glasgow_scaled configs

---
title: "R MINOS Lineplot notebooks noice"
output: html_document
date: "2023-05-20"
---

# MINOS Lineplot Outputs

IMPORTANT NOTE TO RUN PYTHON YOU NEED TO SET THE PYTHON INTERPRETER 
IN TOOLS>GLOBAL_OPTIIONS>PYTHON>CONDA_INTERPRETERS.

Can just use minos_conda or create a miniconda env using reticulate package.

```{r setup, include=FALSE}
library(reticulate)
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../..')
knitr::knit_engines$set(python = reticulate::eng_python)
```

# Baseline vs All Child Uplift

```{python all_child_uplift_SF_12, fig.path="plots/"}
# packages need installing in R using py_install("package"). honestly this seems not worth it..
from minos.outcomes.make_lineplots import main as lineplot_main
# define test parameters and run.
directories = "baseline,hhIncomeIntervention"
tags = "Baseline,All Child Uplift"
subset_function_strings = "who_kids,who_boosted"
prefix="baseline_all_child_uplift"
config_mode = "default_config"
lineplot_main(directories, tags, subset_function_strings, prefix, mode=config_mode, ref="Baseline", v="SF_12", method='nanmean')
```

```{r }
knitr::include_graphics("../../plots/baseline_all_child_uplift_SF_12_aggs_by_year.png", error=FALSE)
```

# Baseline vs Poverty Child Uplift

```{python poverty_line_child_uplift_SF_12, fig.path="plots/"}
# packages need installing in R using py_install("package"). honestly this seems not worth it..
from minos.outcomes.make_lineplots import main as lineplot_main
# define test parameters and run.
directories = "baseline,hhIncomePovertyLineChildUplift"
tags = "Baseline,Poverty Line Child Uplift"
subset_function_strings = "who_below_poverty_line_and_kids,who_boosted"
prefix="baseline_poverty_child_uplift"
config_mode = "default_config"
lineplot_main(directories, tags, subset_function_strings, prefix, mode=config_mode, ref="Baseline", v="SF_12", method='nanmean')
```

```{r }
knitr::include_graphics("../../plots/baseline_poverty_child_uplift_SF_12_aggs_by_year.png", error=FALSE)
```

# Baseline vs Living Wage

```{python living_wage_SF_12, fig.path="plots/"}
# packages need installing in R using py_install("package"). honestly this seems not worth it..
from minos.outcomes.make_lineplots import main as lineplot_main
# define test parameters and run.
directories = "baseline,livingWageIntervention"
tags = "Baseline,Living Wage Intervention"
subset_function_strings = "who_below_living_wage,who_boosted"
prefix="baseline_living_wage"
config_mode = "default_config"
lineplot_main(directories, tags, subset_function_strings, prefix, mode=config_mode, ref="Baseline", v="SF_12", method='nanmean')
```

```{r }
knitr::include_graphics("../../plots/baseline_living_wage_SF_12_aggs_by_year.png", error=FALSE)
```

# Baseline vs Energy Downlift

```{python energy_downlift_SF_12, fig.path="plots/"}
# packages need installing in R using py_install("package"). honestly this seems not worth it..
from minos.outcomes.make_lineplots import main as lineplot_main
# define test parameters and run.
directories = "baseline,energyDownlift"
tags = "Baseline,Energy Downlift"
subset_function_strings = "who_uses_energy,who_boosted"
prefix="baseline_energy_downlift"
config_mode = "default_config"
lineplot_main(directories, tags, subset_function_strings, prefix, mode=config_mode, ref="Baseline", v="SF_12", method='nanmean')
```

```{r }
knitr::include_graphics("../../plots/baseline_energy_downlift_SF_12_aggs_by_year.png", error=FALSE)
```

# All Five combined

```{python all_five_SF_12, fig.path="plots/"}
# packages need installing in R using py_install("package"). honestly this seems not worth it..
from minos.outcomes.make_lineplots import main as lineplot_main
# define test parameters and run.
directories = "baseline,hhIncomeIntervention,hhIncomePovertyLineChildUplift,livingWageIntervention,energyDownlift"
tags = "Baseline,All Child Uplift,Poverty Line Child Uplift,Living Wage Intervention,Energy Downlift"
subset_function_strings = "who_alive,who_boosted,who_boosted,who_boosted,who_boosted"
prefix="all_five_combined"
config_mode = "default_config"
lineplot_main(directories, tags, subset_function_strings, prefix, mode=config_mode, ref="Baseline", v="SF_12", method='nanmean')
```

```{r }
knitr::include_graphics("../../plots/all_five_combined_SF_12_aggs_by_year.png", error=FALSE)
```