### 2023-06-12 minos/validation/handovers_S7.Rmd
Added a separate handovers script for SIPHER 7 vars, and modified the cv ordinal plots function to allow for handovers as they are useful in combination with other handovers

---
title: "Handovers - SIPHER 7 Equivalent Income"
output: html_notebook
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)

workingDir <- "/home/luke/Documents/WORK/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

Source utils for some functions.

```{r}
source('Minos/minos/utils_datain.R')
source('Minos/minos/utils_validation_vis.R')
```

## Data

```{r}
# Read raw datafiles in
raw.files <- list.files('/home/luke/Documents/WORK/MINOS/Minos/data/final_US/', pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

out.path <- 'Minos/output/SIPHER7/'
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
```

## Constants

```{r}
start.year <- 2020

save.path <- '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/SIPHER7/Handovers/'
```

# INPUTS

## Income

```{r}
# first figure out how to plot final_US
# start with hh_income
raw.income <- raw.dat %>% 
  dplyr::select(pidp, time, hh_income, weight) %>%
  group_by(time) %>%
  summarise(income = mean(hh_income, na.rm=TRUE)) %>%
  mutate(source = 'final_US')

base.income <- base.dat %>%
  dplyr::select(pidp, time, hh_income, weight) %>%
  group_by(time) %>%
  summarise(income = mean(hh_income)) %>%
  mutate(source = 'baseline_output')

# merge before plot
income <- rbind(raw.income, base.income)

# Now plot
ggplot(data = income, mapping = aes(x = time, y = income, group = source, colour = source)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Household Income', subtitle = 'Full Sample') +
  xlab('Year') +
  ylab('Income')

## Try a version where final_US is limited to only those present from wave 1 
# onwards because the sample refreshments are messing with the plot
raw.income.wave1 <- raw.dat$pidp[raw.dat$time == 2009]

raw.income2 <- raw.dat %>% 
  dplyr::select(pidp, time, hh_income) %>%
  filter(pidp %in% raw.income.wave1) %>%
  group_by(time) %>%
  summarise(income = mean(hh_income, na.rm = TRUE)) %>%
  mutate(source = 'final_US')

base.income2 <- base.dat %>%
  dplyr::select(pidp, time, hh_income) %>%
  filter(pidp %in% raw.income.wave1) %>%
  group_by(time) %>%
  summarise(income = mean(hh_income, na.rm = TRUE)) %>%
  mutate(source = 'baseline_output')

income2 <- rbind(raw.income2, base.income2)

# Now plot
ggplot(data = income2, mapping = aes(x = time, y = income, group = source, colour = source)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Household Income') +
  xlab('Year') +
  ylab('Income')

ggsave(filename = 'hh_income_wav1.png',
       plot = last_plot(),
       path = save.path)

raw.income.check <- raw.dat %>% 
  dplyr::select(pidp, time, hh_income) %>%
  filter(pidp %in% raw.income.wave1)

base.income.check <- base.dat %>%
  dplyr::select(pidp, time, hh_income) %>%
  filter(pidp %in% raw.income.wave1)

rm(raw.income, base.income, income, raw.income.wave1, raw.income2, base.income2, 
   raw.income.check, base.income.check)
```

### Spaghetti

```{r}
raw.inc <- raw.dat %>%
  select(pidp, age, time, hh_income)
base.inc <- base.dat %>%
  select(pidp, age, time, hh_income)

income.spag <- rbind(raw.inc, base.inc)

spaghetti_plot(income.spag, 'hh_income',
               save = TRUE,
               save.path = save.path)

rm(raw.inc, base.inc, income.spag)
```

# OUTCOMES

## Equivalent Income

```{r}
# first figure out how to plot final_US
# start with hh_income
raw.ei <- raw.dat %>% 
  dplyr::select(pidp, time, equivalent_income) %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income, na.rm = TRUE)) %>%
  mutate(source = 'final_US')

base.ei <- base.dat %>%
  dplyr::select(pidp, time, equivalent_income) %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income, na.rm = TRUE)) %>%
  mutate(source = 'baseline_output')

# merge before plot
ei <- rbind(raw.ei, base.ei)
# drop -9 values (missing)
ei <- ei %>%
  filter(equivalent_income != -9)

# Now plot
ggplot(data = ei, mapping = aes(x = time, y = equivalent_income, group = source, colour = source)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Equivalent Income', subtitle = 'Full Sample') +
  xlab('Year') +
  ylab('EI')

## Try a version where final_US is limited to only those present from wave 1 
# onwards because the sample refreshments are messing with the plot
raw.ei.wave1 <- raw.dat$pidp[raw.dat$time == 2009]

raw.ei.2 <- raw.dat %>% 
  dplyr::select(pidp, time, equivalent_income) %>%
  filter(pidp %in% raw.ei.wave1) %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income, na.rm = TRUE)) %>%
  mutate(source = 'final_US')

base.ei.2 <- base.dat %>%
  dplyr::select(pidp, time, equivalent_income) %>%
  filter(pidp %in% raw.ei.wave1) %>%
  group_by(time) %>%
  summarise(equivalent_income = mean(equivalent_income, na.rm = TRUE)) %>%
  mutate(source = 'baseline_output')

ei.2 <- rbind(raw.ei.2, base.ei.2)
# drop -9 values (missing)
ei.2 <- ei.2 %>%
  filter(equivalent_income != -9)

# Now plot
ggplot(data = ei.2, mapping = aes(x = time, y = equivalent_income, group = source, colour = source)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Equivalent Income') +
  xlab('Year') +
  ylab('EI')

ggsave(filename = 'EI_wav1.png',
       plot = last_plot(),
       path = save.path)

rm(raw.ei, base.ei, ei, raw.ei.wave1, raw.ei.2, base.ei.2, ei.2)
```

### Spaghetti

```{r}
raw.ei <- raw.dat %>%
  select(pidp, age, time, equivalent_income)
base.ei <- base.dat %>%
  select(pidp, age, time, equivalent_income)

ei.spag <- rbind(raw.ei, base.ei)

spaghetti_plot(ei.spag, 'equivalent_income',
               save = TRUE,
               save.path = save.path)

rm(raw.ei, base.ei, ei.spag)
```

# PATHWAYS

## Housing Quality

```{r}
# first figure out how to plot final_US
# start with hh_income
raw <- raw.dat %>% 
  dplyr::select(pidp, time, S7_housing_quality, weight) %>%
  group_by(time) %>%
  summarise(housing_quality = mean(S7_housing_quality, na.rm=TRUE)) %>%
  mutate(source = 'final_US')

base <- base.dat %>%
  dplyr::select(pidp, time, S7_housing_quality, weight) %>%
  group_by(time) %>%
  summarise(housing_quality = mean(S7_housing_quality)) %>%
  mutate(source = 'baseline_output')

# merge before plot
housing_quality <- rbind(raw, base)

# Now plot
ggplot(data = housing_quality, mapping = aes(x = time, y = housing_quality, group = source, colour = source)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Housing Quality', subtitle = 'Full Sample') +
  xlab('Year') +
  ylab('Income')
```

TODO: Make these ordinal plots as handovers (might just need to change cv to baseline?)

```{r}
hous.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'S7_housing_quality')

cv_ordinal_plots(pivoted.df = hous.pivoted, 
                 var = 'housing_quality',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

### Spaghetti

```{r}
raw.s <- raw.dat %>%
  select(pidp, age, time, S7_housing_quality)
base.s <- base.dat %>%
  select(pidp, age, time, S7_housing_quality)

spag <- rbind(raw.s, base.s)

spaghetti_plot(spag, 'S7_housing_quality',
               save = TRUE,
               save.path = save.path)

rm(raw.s, base.s, spag)
```

```{r}

```