### 2023-06-26 minos/validation/cross_validation.Rmd
Adapted and improved the densigram plots for cross-validation. Also added lots of observed vs predicted plots to the cross_validation notebook

---
title: "Cross Validation Analysis"
output: html_notebook
---

This notebook will compare the results of the cross-validation simulation run with real data from Understanding Society, so that we can evaluate the transition models.

# Method

After generation of the final datasets (final_US), the respondents were separated into 2 roughly equal halves based on their pidp. One half was labelled as the transition dataset, and the other for simulation. Transition models were then fit using the transition dataset, with the other half then simulated with these models.

The results of the simulation will now be compared to the simulation dataset to see how closely our models mimick reality.

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(here)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
#source(here::here('minos', 'utils_outcome_vis.R'))
```

## Data

```{r, include=FALSE}
# cross validation simulation output
#out.path <- 'Minos/output/cross_validation/'
out.path <- here::here('output', 'cross_validation_default/')
cv.dat <- read_singular_local_out(out.path, 'baseline')

# cv simulation pop (not used for transitions)
cv.simul.files <- list.files(here::here('data', 'final_US', 'cross_validation', 'simulation'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw.dat <- do.call(rbind, lapply(cv.simul.files, read.csv))
```

# Analysis

First things first, we need to make sure that we are comparing the exact same people from each dataset. Therefore, we need a step here to check the pidps from both, and remove anybody not in both datasets. We can also filter the years out of the raw data here, as we only care about years that have been simulated for comparison.

```{r}
# pidp filter
cv.pidps <- cv.dat %>% select(pidp) %>% unique()
raw.pidps <- raw.dat %>% select(pidp) %>% unique()
both.pidps <- intersect(cv.pidps, raw.pidps)

# year filter
cv.years <- cv.dat %>% select(time) %>% unique()
cv.years <- cv.years[['time']]

cv <- cv.dat %>% filter(pidp %in% both.pidps$pidp)
raw <- raw.dat %>% filter(pidp %in% both.pidps$pidp, 
                      time %in% cv.years)

rm(cv.pidps, raw.pidps, both.pidps, cv.years)
```

## hh_income

```{r}
income.pivoted <- combine_and_pivot_long(df1 = cv, 
                                         df1.name = 'simulated', 
                                         df2 = raw, 
                                         df2.name = 'raw', 
                                         var = 'hh_income')

plot_mean_comparison(pivoted.df = income.pivoted, 
                     var = 'hh_income', 
                     group.var = 'scenario',
                     save = TRUE,
                     save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')

compare_boxplot_long(pivoted.df = income.pivoted,
                     var = 'hh_income',
                     scen1 = 'raw',
                     scen2 = 'simulation')

compare_boxplot_long(pivoted.df = income.pivoted,
                     var = 'hh_income',
                     scen1 = 'raw',
                     scen2 = 'simulation',
                     subset.max = 5000,
                     subset.min = -5000)
```

### Snapshots

Going to produce some comparison plots for snapshots throughout the simulation.

NOTE: LOG SCALE!

```{r}
snap.timelist <- c(2014, 2016, 2018, 2020)

cv.snap <- cv %>%
  select(pidp, time, hh_income) %>%
  filter(time %in% snap.timelist) %>%
  rename(predicted = hh_income)

raw.snap <- raw %>%
  select(pidp, time, hh_income) %>%
  filter(time %in% snap.timelist) %>%
  rename(observed = hh_income)

snap <- merge(cv.snap, raw.snap,
                   by = c('pidp', 'time'))

ggplot(snap, aes(x = observed, y = predicted)) +
  geom_point() +
  geom_abline(slope=1, intercept=0) +
  facet_wrap(~time) +
  labs(title = 'hh_income observed vs predicted')

rm(snap.timelist, cv.snap, raw.snap, snap)
```

```{r}
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2014)
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2016)
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2018)
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2020)
```

### T-tests

```{r}
income.t.tests <- two_sample_t_test(cv, 'cv', raw, 'raw', 'hh_income')
income.t.tests
```

```{r}
yearly_box_plots(pivoted.df = income.pivoted,
                 var = 'hh_income',
                 scen1 = 'raw',
                 scen2 = 'simulation',
                 subset.max = 5000,
                 subset.min = -5000)
```

## SF12

```{r}
sf12.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'SF_12')

plot_mean_comparison(pivoted.df = sf12.pivoted, 
                     var = 'SF_12', 
                     group.var = 'scenario',
                     save = TRUE,
                     save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')

compare_boxplot_long(pivoted.df = sf12.pivoted,
                     var = 'SF_12',
                     scen1 = 'raw',
                     scen2 = 'simulation')
```

### Snapshots

```{r}
observed <- raw
predicted <- cv
var <- 'SF_12'
target.year <- 2020

# get just one year
o.end <- observed %>% filter(time == target.year, .data[[var]] != -8.0)
p.end <- predicted %>% filter(time == target.year, .data[[var]] != -8.0)
# get just the SF12 columns and rename for later
o.end <- o.end %>% select(pidp, .data[[var]]) %>% rename(observed = var)
p.end <- p.end %>% select(pidp, .data[[var]]) %>% rename(predicted = var)

# combine before we plot
combined <- merge(o.end, p.end, by = 'pidp')

p <- ggplot(data = combined, aes(x = observed, y = predicted)) +
  geom_point(alpha = 0.6, size=0.1) +
  geom_smooth() +
  geom_abline(intercept = 0, scale=1) +
  theme(legend.position = c(0.15, 0.9)) +
  labs(title = paste0(var, ' - ', target.year))
p

p1 <- ggMarginal(p, type='densigram', xparams = list(position = 'dodge'), yparams=list(position = 'dodge'))
p1
```

```{r}
snap.timelist <- c(2014, 2016, 2018, 2020)

cv.snap <- cv %>%
  select(pidp, time, SF_12) %>%
  filter(time %in% snap.timelist) %>%
  rename(predicted = SF_12)

raw.snap <- raw %>%
  select(pidp, time, SF_12) %>%
  filter(time %in% snap.timelist) %>%
  rename(observed = SF_12)

snap <- merge(cv.snap, raw.snap,
                   by = c('pidp', 'time'))
snap$time <- as.factor(snap$time)

ggplot(filter(snap, time == 2014), aes(x = observed, y = predicted)) +
  geom_point() +
  geom_abline(slope=1, intercept=0) +
  stat_ellipse(color='green') +
  labs(title = 'SF_12 : 2014')

ggplot(filter(snap, time == 2016), aes(x = observed, y = predicted)) +
  geom_point() +
  geom_abline(slope=1, intercept=0) +
  stat_ellipse(color='green') +
  labs(title = 'SF_12 : 2016')

ggplot(filter(snap, time == 2018), aes(x = observed, y = predicted)) +
  geom_point() +
  geom_abline(slope=1, intercept=0) +
  labs(title = 'SF_12 : 2018')

ggplot(filter(snap, time == 2020), aes(x = observed, y = predicted)) +
  geom_point() +
  geom_abline(slope=1, intercept=0) +
  labs(title = 'SF_12 : 2020')

ggplot(snap, aes(x = observed, y = predicted, group = time, color = time, shape = time)) +
  geom_point(size = 1) +
  geom_abline(slope=1, intercept=0) +
  labs(title = 'SF_12')

rm(snap.timelist, cv.snap, raw.snap, snap)
```

```{r}
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2014)
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2016)
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2018)
marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2020)
```

### T-Tests

```{r}
sf12.t.tests <- two_sample_t_test(cv, 'cv', raw, 'raw', 'hh_income')
sf12.t.tests
```

```{r}
yearly_box_plots(pivoted.df = sf12.pivoted,
                 var = 'SF_12',
                 scen1 = 'raw',
                 scen2 = 'simulation')
```

## Nutrition Quality

```{r}
nut.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'nutrition_quality')

plot_mean_comparison(pivoted.df = nut.pivoted, 
                     var = 'nutrition_quality', 
                     group.var = 'scenario',
                     save = TRUE,
                     save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')

compare_boxplot_long(pivoted.df = nut.pivoted,
                     var = 'nutrition_quality',
                     scen1 = 'raw',
                     scen2 = 'simulation')
```

```{r}
yearly_box_plots(pivoted.df = nut.pivoted,
                 var = 'nutrition_quality',
                 scen1 = 'raw',
                 scen2 = 'simulation')
```

## Tobacco Use

```{r}
tobacco.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'ncigs')

plot_mean_comparison(pivoted.df = tobacco.pivoted, 
                     var = 'ncigs', 
                     group.var = 'scenario',
                     save = TRUE,
                     save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')

compare_boxplot_long(pivoted.df = tobacco.pivoted,
                     var = 'ncigs',
                     scen1 = 'raw',
                     scen2 = 'simulation')

compare_boxplot_long(pivoted.df = tobacco.pivoted,
                     var = 'ncigs',
                     scen1 = 'raw',
                     scen2 = 'simulation',
                     subset.max = 100,
                     subset.min = 0)
```

```{r}
yearly_box_plots(pivoted.df = tobacco.pivoted,
                 var = 'ncigs',
                 scen1 = 'raw',
                 scen2 = 'simulation',
                 subset.max = 100,
                 subset.min = 0)
```

## Housing Quality

```{r}
hous.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'housing_quality')

cv_ordinal_plots(pivoted.df = hous.pivoted, 
                 var = 'housing_quality',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

## Neighbourhood Safety

```{r}
nhsafe.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'neighbourhood_safety')

cv_ordinal_plots(pivoted.df = nhsafe.pivoted, 
                 var = 'neighbourhood_safety',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

## Loneliness

```{r}
lnly.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'loneliness')

cv_ordinal_plots(pivoted.df = lnly.pivoted, 
                 var = 'loneliness',
                 save = TRUE,
                 save.path = '/home/luke/Documents/WORK/MINOS/VALIDATION_PLOTS/Cross-Validation/')
```

```{r}

```