### 2023-07-05 minos/validation/cross_validation_S7.Rmd

---
title: "Cross-Validation Default"
output: html_notebook
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(ggExtra)
require(here)
require(scales)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
#source(here::here('minos', 'utils_outcome_vis.R'))
```

## Data

```{r, include=FALSE}
# cross validation simulation output
#out.path <- 'Minos/output/cross_validation/'
out.path <- here::here('output', 'cv')
cv1 <- read_singular_local_out(here::here(out.path, 'S7_1/'), 'baseline', drop.dead = TRUE)
cv2 <- read_singular_local_out(here::here(out.path, 'S7_2/'), 'baseline', drop.dead = TRUE)
cv3 <- read_singular_local_out(here::here(out.path, 'S7_3/'), 'baseline', drop.dead = TRUE)
cv4 <- read_singular_local_out(here::here(out.path, 'S7_4/'), 'baseline', drop.dead = TRUE)
cv5 <- read_singular_local_out(here::here(out.path, 'S7_5/'), 'baseline', drop.dead = TRUE)

cvlist <- list(cv1, cv2, cv3, cv4, cv5)
cv <- do.call(rbind, cvlist)

# cv simulation batches
cv.simul.files1 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch1'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw1 <- do.call(rbind, lapply(cv.simul.files1, read.csv))
cv.simul.files2 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch2'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw2 <- do.call(rbind, lapply(cv.simul.files2, read.csv))
cv.simul.files3 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch3'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw3 <- do.call(rbind, lapply(cv.simul.files3, read.csv))
cv.simul.files4 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch4'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw4 <- do.call(rbind, lapply(cv.simul.files4, read.csv))
cv.simul.files5 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch5'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw5 <- do.call(rbind, lapply(cv.simul.files5, read.csv))

rawlist <- list(raw1, raw2, raw3, raw4, raw5)
raw.dat <- do.call(rbind, rawlist)
```

First things first, we need to make sure that we are comparing the exact same people from each dataset. Therefore, we need a step here to check the pidps from both, and remove anybody not in both datasets. We can also filter the years out of the raw data here, as we only care about years that have been simulated for comparison.

```{r}
# pidp filter
#cv.pidps <- cv %>% select(pidp) %>% unique()
#raw.pidps <- raw %>% select(pidp) %>% unique()
#both.pidps <- intersect(cv.pidps, raw.pidps)

# year filter
cv.years <- cv %>% select(time) %>% unique()
cv.years <- cv.years[['time']]

#cv <- cv.dat %>% filter(pidp %in% both.pidps$pidp)
raw <- raw.dat %>% filter(time %in% cv.years)

#rm(cv.pidps, raw.pidps, both.pidps, cv.years)
rm(cv.years)
```

For all analyses, results will be calculated on each output separately and pooled. For example, the mean hh_income over time will be calculated for each output run seperately, and the mean of these 5 means is what will be plotted. This will be plotted against the entire raw sample.

# hh_income

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'hh_income')
multi_year_boxplots(raw, cv, 'hh_income')
snapshot_OP_plots(raw, cv, 'hh_income', target.years = c(2014, 2016, 2018, 2020))
q_q_comparison(raw, cv, 'hh_income')
```

```{r}
inc.14 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2014)
print(inc.14)
```
```{r}
inc.16 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2016)
print(inc.16)
```
```{r}
inc.18 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2018)
print(inc.18)
```
```{r}
inc.20 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2020)
print(inc.20)
rm(inc.14, inc.16, inc.18, inc.20)
```

# Equivalent Income

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'equivalent_income')
multi_year_boxplots(raw, cv, 'equivalent_income')
snapshot_OP_plots(raw, cv, 'equivalent_income', target.years = c(2014, 2017, 2020))
q_q_comparison(raw, cv, 'equivalent_income')
```
```{r}
ei.14 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'equivalent_income', 
                                 target.year = 2014)
print(ei.14)
```
```{r}
ei.17 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'equivalent_income', 
                                 target.year = 2017)
print(ei.17)
```
```{r}
ei.20 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'equivalent_income', 
                                 target.year = 2020)
print(ei.20)
rm(ei.14, ei.17, ei.20)
```

# S7 Housing Quality

```{r}
hous.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'S7_housing_quality')

cv_ordinal_plots(pivoted.df = hous.pivoted, 
                 var = 'S7_housing_quality',
                 save = FALSE)
rm(hous.pivoted)
```

# S7 Neighbourhood Safety

```{r}
nhsafe.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'S7_neighbourhood_safety')

cv_ordinal_plots(pivoted.df = nhsafe.pivoted, 
                 var = 'S7_neighbourhood_safety',
                 save = FALSE)
rm(nhsafe.pivoted)
```

# S7 Physical Health

```{r}
physhealth.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'S7_physical_health')

cv_ordinal_plots(pivoted.df = physhealth.pivoted, 
                 var = 'S7_physical_health',
                 save = FALSE)
rm(physhealth.pivoted)
```

# S7 Mental Health

```{r}
menhealth.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'S7_mental_health')

cv_ordinal_plots(pivoted.df = menhealth.pivoted, 
                 var = 'S7_mental_health',
                 save = FALSE)
rm(menhealth.pivoted)
```

# S7 Labour State

```{r}
labstate.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'S7_labour_state')

cv_ordinal_plots(pivoted.df = labstate.pivoted, 
                 var = 'S7_labour_state',
                 save = FALSE)
rm(labstate.pivoted)
```

# Loneliness

```{r}
lon.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'loneliness')

cv_ordinal_plots(pivoted.df = lon.pivoted, 
                 var = 'loneliness',
                 save = FALSE)
rm(lon.pivoted)
```

# Education

```{r}
educ.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'education_state')

cv_ordinal_plots(pivoted.df = educ.pivoted, 
                 var = 'education_state',
                 save = FALSE)
rm(educ.pivoted)
```

```{r}

```