### 2023-07-11 minos/testing/auditc.Rmd
Added alcohol to the physical wellbeing model and module, and updated handovers and cross_validation with more plots

---
title: "AUDITC Testing"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

Source utils for some functions.

```{r}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

## Data

```{r}
# Read raw datafiles in
comp.files <- list.files(here::here('data', 'composite_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
comp.dat <- do.call(rbind, lapply(comp.files, read.csv))

final.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
final.dat <- do.call(rbind, lapply(final.files, read.csv))

cv.files <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch1'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
cv.dat <- do.call(rbind, lapply(final.files, read.csv))
```

# Analysis

```{r}
comp <- comp.dat %>%
  dplyr::select(pidp, time, auditc) %>%
  dplyr::filter(time >= 2015) %>%
  dplyr::filter(time != 2016, time != 2018) %>%
  group_by(time, auditc) %>%
  count()

comp$auditc <- as.factor(comp$auditc)

ggplot(comp, aes(x = time, y = n, group = auditc, color = auditc)) +
  geom_line() +
  labs(title = 'AUDITC - Corrected Data') +
  ylab('Counts')
```

```{r}
final <- final.dat %>%
  dplyr::select(pidp, time, auditc) %>%
  dplyr::filter(time >= 2015) %>%
  dplyr::filter(time != 2016, time != 2018) %>%
  group_by(time, auditc) %>%
  count()

final$auditc <- as.factor(final$auditc)

ggplot(final, aes(x = time, y = n, group = auditc, color = auditc)) +
  geom_line() +
  labs(title = 'AUDITC - Final Data') +
  ylab('Counts')
```

```{r}
cv <- cv.dat %>%
  dplyr::select(pidp, time, auditc) %>%
  dplyr::filter(time >= 2015) %>%
  dplyr::filter(time != 2016, time != 2018) %>%
  group_by(time, auditc) %>%
  count()

cv$auditc <- as.factor(cv$auditc)

ggplot(cv, aes(x = time, y = n, group = auditc, color = auditc)) +
  geom_line() +
  labs(title = 'AUDITC - Cross-Validation Batch 1') +
  ylab('Counts')
```

```{r}

```