### 2023-07-13 docsrc/documentation/notebooks/tobacco.Rmd
Started a notebook for physical activity documentation

---
title: "Tobacco"
output:
  pdf_document: default
  html_document: default
bibliography: docsrc/refs.bib
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir= "../minos")
```
## Tobacco

Number of cigarettes consumed is an indicator of several mental illnesses including anxiety [@lawrence2010anxiety].

```{r load_data, include=FALSE}
source('docsrc/documentation/notebooks/sphinx_notebook_utils.R')
```

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

```{r tobacco_data}
counts_density("data/transitions/ncigs/zip/ncigs_2018_2019.rds", "y")
```

### Methods

The number of zero inflated values is higher than expected for a count distribution such as a poisson distribution. This inflation occurs naturally as a large proportion (over 50\%) of the population do not smoke. There are two sources of cigarette consumption that can be modelled using zero inflated models. In this case a zero-inflated poisson (ZIP) is used. Two models are fitted simulatenously. One is a logistic regression that estimates whether a person smokes cigarettes or not. This provides a simple probability of smoking or not. The second is a poisson counts model estimating the number of cigarettes consumed.

### Data

Two set of variables are needed for the logistic and poisson parts of the ZIP model respectively.

Variables that predict how much a person smokes.

- age. persons age. generally older people and very young smoke. 
- SF_12. wellbeing estimates number of cigarettes smoked.
- labour_state. whether a person is employed or not.
- ethnicity. certain ethnicities more likely to smoke cigarettes. 
- education_state. highest qualification.
- job_sec job quality
- hh_income household income
- ncigs previous number consumed.

Variables that predict whether a person smokes

- ethnicity. certain ethnicities more likely to smoke cigarettes. 
- labour_state. whether a person is employed or not.
- age
- SF_12. wellbeing estimates number of cigarettes smoked.
- ncigs previous number consumed.

### Results

Almost all coefficients significant. Particularly previous consumption of cigarettes. Good estimation of the number of non-smokers in the population at around 55\%. Counts of smoking are underdispersed and fail to estimate consumption over 20 cigarettes.

```{r tobacco_output, echo=FALSE}
options(width=120)
zip_output("data/transitions/ncigs/zip/ncigs_2018_2019.rds")
zip_density("data/transitions/ncigs/zip/ncigs_2018_2019.rds", 'y')
```

### References