### 2023-07-24 minos/testing/imputation_testing.Rmd
Another version of the imputation step. This time, we "seed" variables where information is completely missing within a year (copy another wave onto and remove half of values). This is to give the imputater something to work with as complete missing within a year was causing big problems

---
title: "Imputation Assessment"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  out.width = 3500
)
# 3000
# '100%'
rm(workingDir)
```

Source utils for some functions.

```{r}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

## Data

```{r}
# Read raw datafiles in
imp.files <- list.files(here::here('data', 'stock'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
imp.dat <- do.call(rbind, lapply(imp.files, read.csv))

final.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
final.dat <- do.call(rbind, lapply(final.files, read.csv))

complete.files <- list.files(here::here('data', 'complete_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
complete.dat <- do.call(rbind, lapply(complete.files, read.csv))
```

# Functions

```{r}
compare_continuous <- function(final, complete, imputed, var) {

  final.var <- final %>% 
    dplyr::select(pidp, time, all_of(var), weight) %>%
    filter(time != 2009) %>%
    filter(!.data[[var]] %in% miss.values) %>%
    drop_na() %>%
    group_by(time) %>%
    summarise(mean = weighted.mean(x = .data[[var]],
                                   w = weight,
                                   na.rm=TRUE)) %>%
    mutate(source = 'final')

  complete.var <- complete %>% 
    dplyr::select(pidp, time, all_of(var), weight) %>%
    filter(time != 2009) %>%
    filter(!.data[[var]] %in% miss.values) %>%
    drop_na() %>%
    group_by(time) %>%
    summarise(mean = weighted.mean(x = .data[[var]],
                                   w = weight,
                                   na.rm=TRUE)) %>%
    mutate(source = 'complete')

  imputed.var <- imputed %>% 
    dplyr::select(pidp, time, all_of(var), weight) %>%
    filter(time != 2009) %>%
    filter(!.data[[var]] %in% miss.values) %>%
    drop_na() %>%
    group_by(time) %>%
    summarise(mean = weighted.mean(x = .data[[var]],
                                   w = weight,
                                   na.rm=TRUE)) %>%
    mutate(source = 'imputed')

  merged <- rbind(final.var, complete.var)
  merged <- rbind(merged, imputed.var)

  # Now plot
  p1 <- ggplot(data = merged, mapping = aes(x = time, y = mean, group = source, colour = source)) +
    geom_line() +
    labs(title = var, subtitle = 'Full Sample') +
    xlab('Year') +
    ylab(var)

  print(p1)
}

compare_ordinal <- function(final, complete, imputed, var) {

  final.var <- final %>% 
    dplyr::select(pidp, time, all_of(var)) %>%
    filter(.data[[var]] >= 0) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    mutate(source = 'final')

  complete.var <- complete %>% 
    dplyr::select(pidp, time, all_of(var)) %>%
    filter(.data[[var]] >= 0) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    mutate(source = 'complete')

  imputed.var <- imputed %>% 
    dplyr::select(pidp, time, all_of(var)) %>%
    filter(.data[[var]] >= 0) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    mutate(source = 'imputed')

  merged <- rbind(final.var, complete.var)
  merged <- rbind(merged, imputed.var)
  merged[[var]] <- as.factor(merged[[var]])
  merged$time = as.factor(merged$time)

  # Now plot
  p1 <- ggplot(data = merged, mapping = aes(x = .data[[var]], y = n, group = source, fill = source)) +
    geom_col(position='dodge') +
    labs(title = var, subtitle = 'Counts by Level') +
    xlab('Year') +
    ylab('Count')

  p2 <- ggplot(data = merged, mapping = aes(x = time, y = n, group = .data[[var]], colour = .data[[var]])) +
    geom_line() +
    geom_point() +
    facet_grid(.~source) +
    theme(legend.position="none") +
    labs(title = var, subtitle = 'Counts by Level') +
    xlab('Year') +
    ylab('Count')

  print(p1)
  print(p2)

  var.norm <- merged %>%
    group_by(time, source) %>%
    mutate(total = sum(n)) %>%
    mutate(prct = (n / total))

  p3 <- ggplot(data = var.norm, mapping = aes(x = time, y = prct, fill=.data[[var]], position='dodge')) +
    geom_bar(stat = 'identity') +
    facet_grid(.~source) +
    theme(legend.position="none") +
    labs(title = var) +
    xlab('Year') +
    ylab('Proportion')
  print(p3)
}

multi_year_boxplots <- function(final, complete, imputed, var) {
  final.var <- final %>%
    dplyr::select(pidp, time, all_of(var)) %>%
    filter(!.data[[var]] %in% miss.values)
  final.var$source <- 'final'

  complete.var <- complete %>%
    dplyr::select(pidp, time, all_of(var)) %>%
    filter(!.data[[var]] %in% miss.values)
  complete.var$source <- 'complete'

  imputed.var <- imputed %>%
    dplyr::select(pidp, time, all_of(var)) %>%
    filter(!.data[[var]] %in% miss.values)
  imputed.var$source <- 'imputed'

  combined <- rbind(final.var, complete.var)
  combined <- rbind(combined, imputed.var)
  combined$time <- as.factor(combined$time)
  combined <- drop_na(combined)
  combined <- filter(combined, .data[[var]] != -9)

  if (var %in% c('hh_income', 'equivalent_income')) {
    combined <- filter(combined, .data[[var]] < quantile(.data[[var]], 0.99), .data[[var]] > quantile(.data[[var]], 0.01))
  } else if (var == 'ncigs') {
    #combined <- filter(combined, .data[[var]] < quantile(.data[[var]], 0.99))
    combined <- filter(combined, .data[[var]] < quantile(.data[[var]], 0.99), !.data[[var]] == 0)
  }

  ggplot(data = combined, aes(x = time, y = .data[[var]], group = interaction(time, source), color = source)) +
    geom_boxplot(notch=TRUE) +
    labs(title = paste0(var, ': Yearly box plots'))
}

```

# Float Variables

## hh_income

```{r}
compare_continuous(final.dat, complete.dat, imp.dat, 'hh_income')
multi_year_boxplots(final.dat, complete.dat, imp.dat, 'hh_income')
```

## SF_12

```{r}
compare_continuous(final.dat, complete.dat, imp.dat, 'SF_12')
multi_year_boxplots(final.dat, complete.dat, imp.dat, 'SF_12')
```

# Str Variables

## Region

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'region')
```

## Ethnicity

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'ethnicity')
```

## Loneliness

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'loneliness')
```

## Sex

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'sex')
```

## Housing Quality

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'housing_quality')
```

## S7 Housing Quality

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'S7_housing_quality')
```

## Neighbourhood Safety

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'neighbourhood_safety')
```

## S7 Neighbourhood Safety

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'S7_neighbourhood_safety')
```

## S7 Labour State

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'S7_labour_state')
```

## Marital Status

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'marital_status')
```

# Int Variables

## Heating

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'heating')
```

## Job Sec

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'job_sec')
```

## ncigs

```{r}
compare_continuous(final.dat, complete.dat, imp.dat, 'ncigs')
multi_year_boxplots(final.dat, complete.dat, imp.dat, 'ncigs')
```

## Financial Situation

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'financial_situation')
```

## Education State

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'education_state')
```

## S7 Physical Health

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'S7_physical_health')
```

## S7 Mental Health

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'S7_mental_health')
```

## Housing Tenure

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'housing_tenure')
```

## Nutrition Quality

```{r}
compare_continuous(final.dat, complete.dat, imp.dat, 'nutrition_quality')
multi_year_boxplots(final.dat, complete.dat, imp.dat, 'nutrition_quality')
```

## Smoker

```{r}
compare_ordinal(final.dat, complete.dat, imp.dat, 'smoker')
```

```{r}

```