### 2023-08-14 minos/outcomes/equivalent_income_out.Rmd
Lots of additions to equivalent income validation and outcomes scripts, and tweaks to the pipeline

---
title: "Equivalent Income Outcome Visualisation"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(here)
require(dplyr)
require(tidyr)
require(purrr)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

# Data

```{r}
# set output path to SIPHER 7 experiment out
out.path <- here::here('output', 'SIPHER7/')
# read all files into dataframe (util function handles runtime subdirectory etc.)
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
allchild.dat <- read_singular_local_out(out.path, 'hhIncomeChildUplift', drop.dead = TRUE)
pov.dat <- read_singular_local_out(out.path, 'hhIncomePovertyLineChildUplift', drop.dead = TRUE)
wage.dat <- read_singular_local_out(out.path, 'livingWageIntervention', drop.dead = TRUE)
energy.dat <- read_singular_local_out(out.path, 'energyDownlift', drop.dead = TRUE)
```

## Remove 2020 Data Point

```{r}
base.dat <- filter(base.dat, time > 2020)
allchild.dat <- filter(allchild.dat, time > 2020)
pov.dat <- filter(pov.dat, time > 2020)
wage.dat <- filter(wage.dat, time > 2020)
energy.dat <- filter(energy.dat, time > 2020)
```

## BATCH DATA

```{r}
S7.var.list <- c('hh_income', 'equivalent_income', 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness')
out.path.batch <- here::here('output', 'SIPHER7_batch/')

base.batch <- read_batch_out_all_years(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=TRUE)
wage.batch <- read_batch_out_all_years(out.path.batch, 'livingWageIntervention', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=TRUE)
energy.batch <- read_batch_out_all_years(out.path.batch, 'energyDownlift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=TRUE)
```

# Baseline

```{r}
base.subset <- base.dat %>%
  select(pidp, time, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))

base.subset$scenario <- 'Baseline'

ggplot(data = base.subset, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line()
```

## Batch

```{r}
base.bat.sub <- base.batch %>%
  select(pidp, time, run_id, equivalent_income, weight) %>%
  group_by(run_id, time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm=TRUE))

ggplot(data = base.bat.sub, aes(x = time, y = equivalent_income, group=run_id, color = run_id)) +
  geom_line()

ggplot(data = base.bat.sub, aes(x = time, y = equivalent_income)) +
  geom_smooth()

base.bat.sub2 <- base.bat.sub %>%
  group_by(time) %>%
  summarise(var = mean(equivalent_income),
         min = min(equivalent_income),
         max = max(equivalent_income))

ggplot(data = base.bat.sub2, aes(x = time, y = var)) +
  geom_line() +
  geom_ribbon(aes(ymin = min, ymax = max), 
                alpha = 0.1,
                linetype = 'dashed')
```

# Comparisons

## Function

```{r}
equivalent_income_comparison_fullpop <- function(base, int, int.name, save=FALSE, save.path=NULL) {
  base.subset <- base %>%
    select(pidp, time, equivalent_income, alive, weight) %>%
    filter(alive != 'dead') %>%
    group_by(time) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
  int.subset <- int %>%
    select(pidp, time, equivalent_income, alive, weight) %>%
    filter(alive != 'dead') %>%
    group_by(time) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))

  base.subset$scenario <- 'Baseline'
  int.subset$scenario <- int.name

  combined <- rbind(base.subset, int.subset)

  p <- ggplot(combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_line() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Full Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)

  if(save) {
    if(is.null(save.path)) {
      stop('ERROR: save.path cannot be missing if save=TRUE')
    }
    file.name <- paste0('EI_comp_', int.name, '_full.png')
    ggsave(filename = file.name,
           plot = p,
           path = save.path)
  }
}

equivalent_income_comparison_fullpop_batch <- function(base.bat, int.bat, int.name, save=FALSE, save.path=NULL) {
  base.subset <- base.bat %>%
    select(pidp, time, run_id, equivalent_income, weight) %>%
    group_by(time, run_id) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
  int.subset <- int.bat %>%
    select(pidp, time, run_id, equivalent_income, weight) %>%
    group_by(time, run_id) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))

  base.subset$scenario <- 'Baseline'
  int.subset$scenario <- int.name

  combined <- rbind(base.subset, int.subset)

  p <- ggplot(combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_smooth() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = '10 Runs') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)
}
```

## Plots

```{r}
equivalent_income_comparison_fullpop(base.dat, allchild.dat, 'allChild', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_fullpop(base.dat, pov.dat, 'povertyChild', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_fullpop(base.dat, wage.dat, 'LivingWage', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_fullpop(base.dat, energy.dat, 'EnergyDownlift', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')

equivalent_income_comparison_fullpop_batch(base.batch, wage.batch, 'LivingWage')
equivalent_income_comparison_fullpop_batch(base.batch, energy.batch, 'EnergyCrisis')
```

## Combined

```{r}
base.subset <- base.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
base.subset$scenario <- 'Baseline'

allchild.subset <- allchild.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
allchild.subset$scenario <- 'AllChild'

pov.subset <- pov.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
pov.subset$scenario <- 'PovertyChild'

wage.subset <- wage.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
wage.subset$scenario <- 'LivingWage'

energy.subset <- energy.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
energy.subset$scenario <- 'EnergyDownlift'

all.combined <- rbind(base.subset, allchild.subset, pov.subset, wage.subset, energy.subset)
all.combined$scenario <- factor(all.combined$scenario, levels = c('Baseline', 'AllChild', 'PovertyChild', 'LivingWage', 'EnergyDownlift'))

ggplot(data = all.combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line() +
  labs(title = 'Equivalent Income Comparison', subtitle = 'Whole Population') +
  xlab('Year') +
  ylab('Equivalent Income (£)')
```

# Treated Population

## Function

```{r}
equivalent_income_comparison_treated <- function(base, int, int.name, save=FALSE, save.path=NULL) {
  base.sub <- base %>%
    select(pidp, time, equivalent_income, alive) %>%
    filter(alive != 'dead') %>%
    select(-alive) %>%
    rename(baseline = equivalent_income)
  int.sub <- int %>%
    select(pidp, time, equivalent_income, income_boosted, alive, weight) %>%
    filter(alive != 'dead') %>%
    select(-alive) %>%
    set_names(c('pidp', 'time', int.name, 'income_boosted', 'weight'))

  merged <- merge(base.sub, int.sub, by = c('pidp', 'time'))

  boosted.pidps <- int.sub %>%
    filter(income_boosted == 'True') %>%
    select(pidp)

  boosted <- merged %>%
    filter(pidp %in% boosted.pidps$pidp) %>%
    filter(income_boosted == 'True' | time == 2020) %>%
    select(-income_boosted) %>%
    pivot_longer(cols = baseline:.data[[int.name]],
                 names_to = 'scenario',
                 values_to = 'equivalent_income') %>%
    group_by(time, scenario) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w=weight))

  p <- ggplot(data = boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_line() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Treated Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)

  if(save) {
    if(is.null(save.path)) {
      stop('ERROR: save.path cannot be missing if save=TRUE')
    }
    file.name <- paste0('EI_comp_', int.name, '_treated.png')
    ggsave(filename = file.name,
           plot = p,
           path = save.path)
  }
}

equivalent_income_comparison_treated_batch <- function(base.bat, int.bat, int.name, save=FALSE, save.path=NULL) {
  base.sub <- base.bat %>%
    select(pidp, time, run_id, equivalent_income) %>%
    rename(baseline = equivalent_income)
  int.sub <- int.bat %>%
    select(pidp, time, run_id, equivalent_income, income_boosted, weight) %>%
    set_names(c('pidp', 'time', 'run_id', int.name, 'income_boosted', 'weight'))

  merged <- merge(base.sub, int.sub, by = c('pidp', 'time', 'run_id'))

  boosted.pidps <- int.sub %>%
    filter(income_boosted == 'True') %>%
    select(pidp)

  boosted <- merged %>%
    filter(pidp %in% boosted.pidps$pidp) %>%
    filter(income_boosted == 'True' | time == 2020) %>%
    select(-income_boosted) %>%
    pivot_longer(cols = baseline:.data[[int.name]],
                 names_to = 'scenario',
                 values_to = 'equivalent_income') %>%
    group_by(time, scenario, run_id) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w=weight))

  p <- ggplot(data = boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_smooth() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Treated Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)
}
```

## Plots

```{r}
equivalent_income_comparison_treated(base.dat, allchild.dat, 'allChild', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_treated(base.dat, pov.dat, 'povertyChild', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_treated(base.dat, wage.dat, 'LivingWage', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')
equivalent_income_comparison_treated(base.dat, energy.dat, 'EnergyDownlift', save=TRUE, save.path = '/home/luke/Documents/WORK/MINOS/Minos/plots/Equivalent_Income/')

equivalent_income_comparison_treated_batch(base.batch, wage.batch, 'LivingWage')
equivalent_income_comparison_treated_batch(base.batch, energy.batch, 'EnergyDownlift')
```

# Remove Income Effect

```{r, include=FALSE}
# THink about this...
# We can use the income_boosted and boost_amount variables to remove the income boost from people and visualise that

base.sub2 <- base.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  select(-alive) %>%
  rename(baseline = equivalent_income)

allchild.sub2 <- allchild.dat %>%
  select(pidp, time, equivalent_income, income_boosted, boost_amount, alive) %>%
  filter(alive != 'dead') %>%
  select(-alive) %>%
  rename(allchild = equivalent_income)

allchild.merged <- merge(base.sub2, allchild.sub2, by = c('pidp', 'time'))

boosted.pidps <- allchild.sub2 %>%
  filter(income_boosted == 'True') %>%
  select(pidp)

allchild.boosted <- allchild.merged %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  filter(income_boosted == 'True' | time == 2020) %>%
  mutate(adjusted_EI = allchild - boost_amount) %>%
  select(-allchild, -income_boosted, -boost_amount) %>%
  group_by(time) %>%
  summarise(baseline = mean(baseline),
            allchild_adj = mean(adjusted_EI)) %>%
  pivot_longer(cols = baseline:allchild_adj,
               names_to = 'scenario',
               values_to = 'equivalent_income')
```

```{r}
ggplot(data = allchild.boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line()
```

NOTE: Can we 'remove' the income effect?
Use the boost amount and income_boosted variables to sort of 'undo' the intervention effect and then we are only visualising the difference from pathway changes and not just the income effect.

ALSO: Difference plots.

# DETAILED ANALYSIS

## Living Wage

For this, we're going to look at the difference in equivalent income and try to explain it by looking at the pathway variables. What we want to see is that as income is uplifted, the various pathways show a positive shift and that in part explains the difference in equivalent income.

First visualise both income and equivalent income on one graph.

### Income and Equivalent Income

```{r}
base.inc <- base.dat %>%
  select(pidp, time, hh_income, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
            equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Baseline')

int.inc <- wage.dat %>%
  select(pidp, time, hh_income, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
            equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Living_Wage')

combined <- rbind(base.inc, int.inc)
combined <- combined %>%
  pivot_longer(cols=hh_income:equivalent_income,
               names_to = 'income_type',
               values_to = 'value')

ggplot(combined, aes(x = time, y = value, group = source, color = source)) +
  geom_line() +
  facet_wrap( ~ income_type, nrow=2, scales='free')
```

## Energy Crisis

### Housing Quality

```{r}
S7_ordinal_pathway_vis <- function(base.dat, int.dat, int.name, var) {
  base.hous <- base.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = 'Baseline')

  inc.hous <- int.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = int.name)

  combined <- rbind(base.hous, inc.hous)
  combined[[var]] <- as.factor(combined[[var]])

  p1 <- ggplot(combined, aes(x = time, y = prop, group = interaction(.data[[var]], source), color = .data[[var]], linetype = source)) +
    geom_line() +
    geom_point()

  print(p1)

  base.hous2 <- base.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(Baseline = prop)

  int.hous2 <- int.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(int = prop)

  combined2 <- merge(base.hous2, int.hous2, by = c('time', var))

  combined2[[var]] <- as.factor(combined2[[var]])

  combined2$diff <- combined2$int - combined2$Baseline

  p2 <- ggplot(combined2, aes(x = time, y = diff, group = .data[[var]], color = .data[[var]])) +
    geom_line() +
    geom_point()

  print(p2)
}

S7_ordinal_pathway_vis_batch <- function(base.bat, int.bat, int.name, var) {
  base.hous <- base.bat %>%
    select(pidp, time, run_id, all_of(var)) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = 'Baseline')

  inc.hous <- int.bat %>%
    select(pidp, time, run_id, all_of(var)) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = int.name)

  combined <- rbind(base.hous, inc.hous)
  combined[[var]] <- as.factor(combined[[var]])

  p1 <- ggplot(combined, aes(x = time, y = prop, group = interaction(.data[[var]], source), color = .data[[var]], linetype = source)) +
    geom_smooth()

  print(p1)

  base.hous2 <- base.bat %>%
    select(pidp, time, run_id, all_of(var)) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(Baseline = prop)

  int.hous2 <- int.bat %>%
    select(pidp, time, run_id, all_of(var)) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(int = prop)

  combined2 <- merge(base.hous2, int.hous2, by = c('time', 'run_id', var))

  combined2[[var]] <- as.factor(combined2[[var]])

  combined2$diff <- combined2$int - combined2$Baseline

  p2 <- ggplot(combined2, aes(x = time, y = diff, group = .data[[var]], color = .data[[var]]), fill = .data[[var]]) +
    geom_smooth() +
    labs(title = paste0(int.name, ': Change from Baseline')) +
    xlab('Year') +
    ylab('Change')

  print(p2)
}

```

### Housing Quality

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_housing_quality')
S7_ordinal_pathway_vis_batch(base.batch, energy.batch, 'EnergyCrisis', 'S7_housing_quality')
```

### Neighbourhood Safety

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_neighbourhood_safety')
S7_ordinal_pathway_vis_batch(base.batch, energy.batch, 'EnergyCrisis', 'S7_neighbourhood_safety')
```

### Mental Wellbeing

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_mental_health')
S7_ordinal_pathway_vis_batch(base.batch, energy.batch, 'EnergyCrisis', 'S7_mental_health')
```

### Physical Health

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_physical_health')
S7_ordinal_pathway_vis_batch(base.batch, energy.batch, 'EnergyCrisis', 'S7_physical_health')
```

### Loneliness

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'loneliness')
S7_ordinal_pathway_vis_batch(base.batch, energy.batch, 'EnergyCrisis', 'loneliness')
```

### Labour State

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_labour_state')
S7_ordinal_pathway_vis_batch(base.batch, energy.batch, 'EnergyCrisis', 'S7_labour_state')
```

```{r}
boosted.pidps <- energy.dat %>%
  filter(income_boosted == 'True') %>%
  select(pidp)

base.hous <- base.dat %>%
  select(pidp, time, S7_housing_quality) %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  group_by(time, S7_housing_quality) %>%
  count() %>%
  group_by(time) %>%
  mutate(total = sum(n)) %>%
  mutate(prop = (n / total) * 100) %>%
  select(-n, -total) %>%
  rename(Baseline = prop)

int.hous <- energy.dat %>%
  select(pidp, time, S7_housing_quality, income_boosted) %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  group_by(time, S7_housing_quality) %>%
  count() %>%
  group_by(time) %>%
  mutate(total = sum(n)) %>%
  mutate(prop = (n / total) * 100) %>%
  select(-n, -total) %>%
  rename(EnergyCrisis = prop)

combined <- merge(base.hous, int.hous, by = c('time', 'S7_housing_quality'))

combined$diff <- combined$EnergyCrisis - combined$Baseline

ggplot(combined, aes(x = time, y = diff, group = S7_housing_quality, color = S7_housing_quality)) +
  geom_line() +
  geom_point()
```

```{r}
## just look at means in case the means are causing some problems
base.dat.2 <- base.dat
base.dat.2$S7_housing_quality2[base.dat.2$S7_housing_quality == 'No to all'] <- 1
base.dat.2$S7_housing_quality2[base.dat.2$S7_housing_quality == 'Yes to some'] <- 2
base.dat.2$S7_housing_quality2[base.dat.2$S7_housing_quality == 'Yes to all'] <- 3
wage.dat.2 <- energy.dat
wage.dat.2$S7_housing_quality2[wage.dat.2$S7_housing_quality == 'No to all'] <- 1
wage.dat.2$S7_housing_quality2[wage.dat.2$S7_housing_quality == 'Yes to some'] <- 2
wage.dat.2$S7_housing_quality2[wage.dat.2$S7_housing_quality == 'Yes to all'] <- 3

base.hous.mean <- base.dat.2 %>%
  select(pidp, time, S7_housing_quality2, weight) %>%
  group_by(time) %>%
  summarise(mean_housing = weighted.mean(S7_housing_quality2, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Baseline')
wage.hous.mean <- wage.dat.2 %>%
  select(pidp, time, S7_housing_quality2, weight) %>%
  group_by(time) %>%
  summarise(mean_housing = weighted.mean(S7_housing_quality2, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'EnergyCrisis')

combined <- rbind(base.hous.mean, wage.hous.mean)

ggplot(combined, aes(x = time, y = mean_housing, group = source, color = source)) +
  geom_line() +
  geom_point()
```

```{r}

df <- combine_and_pivot_long(base.dat, 'Baseline', wage.dat, 'LivingWage', c('S7_housing_quality', 'income_boosted'))

outcomes.prep <- function(base.dat, int.dat, int.name, var) {
  boosted.pidps <- int.dat %>%
    filter(income_boosted == 'True') %>%
    select(pidp, time)

  df1 <- base.dat %>%
    select(pidp, time, all_of(var), weight) %>%

}

combine_and_pivot_long <- function(df1, df1.name, df2, df2.name, var) {
  # get only the columns we want a rename
  df1 <- df1 %>%
    select('pidp', 'time', all_of(var)) %>%
    set_names(c('pidp', 'time', df1.name))
  df2 <- df2 %>%
    select('pidp', 'time', all_of(var)) %>%
    set_names(c('pidp', 'time', df2.name))
  # merge on pidp and time
  merged <- merge(df1, df2, by = c('pidp', 'time'), all=TRUE)
  pivoted <- pivot_longer(data = merged,
                          cols = df1.name:df2.name,
                          names_to = 'scenario',
                          values_to = var)
  return(pivoted)
}

```