### 2023-08-24 minos/outcomes/paper_1_notebook.Rmd
getting R plots from paper working on new branch

---
title: "Paper 1 Notebook"
output: html_notebook
---

```{r "preamble"}
source("minos/transitions/utils.R")
source("minos/transitions/transition_model_functions.R")
source("minos/transitions/paper_1_plots.R")
library(dplyr)
library(texreg)
library(lme4)
library(glmnet)
library(mice)
library(sjPlot)
library(visreg)
library(ggplot2)
library(ggpattern)
library(VIM)
```

```{r "raw_data_processing"}

dataDir <- "data/raw_US/"
filelist <- list.files(dataDir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[21:23] # get files for 2011-2013.
raw.data <- do.call(rbind, lapply(filelist, read.csv))

model_formula_string <- "SF_12 ~ factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state_raw) + 
                  relevel(factor(job_sec), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_netinc)"
model_formula <- as.formula(model_formula_string)
variables <- all.vars(model_formula)
variables[1] <- "SF_12"

raw.data <- raw.data[, append(variables, c("time", 'pidp', 'weight'))]

raw.data <- raw.data %>%
        group_by(pidp) %>%
        #mutate(diff = .data[[dependent]] - lag(.data[[dependent]], order_by = time)) %>%
        mutate(last = lag(.data[["SF_12"]], order_by = time)) %>%
        rename_with(.fn = ~paste0('SF_12_', .), .cols = last)  # add the dependent as prefix to the calculated diff

raw.data <- replace.missing(raw.data)
raw.data <- drop_na(raw.data)
```

```{r "OLS"}

sf12.lm <- lm(model_formula_string,
              data= raw.data)
a<- summary(sf12.lm)
print(a)
print(a$adj.r.squared)

save.path <- "minos/outcomes/paper1_plots/"
create.if.not.exists(save.path)
texreg(sf12.lm, dcolumn=T, booktabs=T, file=paste0(save.path,'ols_coefficients.txt'), custom.model.names=c("OLS"), single.row=T)
save(sf12.lm, file = paste0(save.path,"baseline_OLS.RData"))

write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_OLS_coefficients.txt')
  tex_label <- paste0("table: baseline_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("Baseline SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(sf12.lm, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}

forest_plot_lm(sf12.lm, paste0(save.path, "presentation_forest.pdf"))
qq_plot(resid(sf12.lm), paste0(save.path, "presentation_qq.pdf"))
residual_density_plot(res=resid(sf12.lm), file_name=paste0(save.path, "presentation_residual_density.pdf"), guide="normal")

squareRootRes <- sqrt(abs(scale(resid(sf12.lm))))
fitted_residuals <- as.data.frame(cbind(fitted(sf12.lm), squareRootRes))
colnames(fitted_residuals) <- c("fitted", "sqrt_residuals")
fitted_residual_plot(fitted_residuals, paste0(save.path, "presentation_fitted_residual_plot.pdf"))

#visreg(sf12.lm, "age", by="sex", overlay=T)

pdf(paste0(save.path, "total_missingness_structure1.pdf"))
aggr(subset(raw.data[,reg_columns], select=-c(y)), sortVars=T,  oma=c(8,4,4,4), numbers=T, cex.axis=1.0, col = c(blue, orange), prop=F, combined=T, cex.numbers=0.5)
dev.off()

```

```{r final_data_processing}
# new formula and get data from final

dataDir <- "data/composite_US/"
filelist <- list.files(dataDir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[3:5] # get files for 2011-2013.
final.data <- do.call(rbind, lapply(filelist, read.csv))

# scale(SF_12_last) +
model_formula_string <- "SF_12 ~
                  factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  scale(I(age**2)) +
                  factor(education_state) + 
                  factor(S7_labour_state) + 
                  relevel(factor(job_sec), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income) + 
                  scale(I(hh_income**2)) +
                  (1+time|pidp)"

final.data <- final.data %>%
        group_by(pidp) %>%
        #mutate(diff = .data[[dependent]] - lag(.data[[dependent]], order_by = time)) %>%
        mutate(last = lag(.data[["SF_12"]], order_by = time)) %>%
        rename_with(.fn = ~paste0('SF_12_', .), .cols = last)  # add the dependent as prefix to the calculated diff

glmm_model_formula <- as.formula(model_formula_string)
variables <- all.vars(glmm_model_formula)
#variables[2] <- "SF_12"

final.data <- final.data[, append(variables, c("time", 'pidp', 'weight'))]

final.data <- replace.missing(final.data)
final.data <- drop_na(final.data)

```

```{r MICE}

dataDir <- "data/composite_US/"
filelist <- list.files(dataDir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[3:5] # get files for 2011-2013.
mice.data <- do.call(rbind, lapply(filelist, read.csv))
mice.data <-replace.missing(mice.data)

model_formula_string <- "SF_12 ~ factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(S7_labour_state) + 
                  relevel(factor(job_sec), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income)"

# set up MICE and cache data
# plots of missingness pre and post raw US
# figures of convergence and difference in distribution.
cached <- F # if you just want the plots used cached data. MICE can take forever..
imp_columns <- c("SF_12",
                 'sex',
                 "job_sec",
                 "education_state",
                 "S7_labour_state", 
                 "region", 
                 "ethnicity", 
                 "age", 
                 "hh_income")

method <-   c('pmm',
            'pmm',#'polr',
            'pmm',
            'pmm', #'polr',
            'pmm',
            'pmm',
            'pmm',
            'pmm',
            'pmm' )

# structure plots for adding in auxliary variables

pdf(paste0(save.path, "total_missingness_structure2.pdf"))
aggr(subset(mice.data, select=-c(y)), sortVars=T,  oma=c(8,4,4,4), numbers=T, cex.axis=1.0, col = c(blue, orange), prop=F, combined=T, cex.numbers=0.5)
dev.off()

cached<-T
if (cached){
  load('papers/phd1/data/mice_set.RData')
  load('papers/phd1/data/mice_data.RData')
} else{
  pidps <- mice.data$pidp
  save(pidps, file = paste0(save.path, 'mice_pidps.RData'))
  mice.weights <- mice.data$weight
  save(mice.weights, file = paste0(save.path, 'mice_weights.RData'))

  mice_set <- mice(data = mice.data[,imp_columns], method=method,
                   m = 20, maxit = 10,
                   #m = 2, maxit = 2,
                   remove.collinear=T)
  mice.data <- complete(mice_set)

  save(mice_set, file = paste0(save.path,'mice_set.RData'))
  save(mice.data, file = paste0(save.path,'mice_data.RData'))
}

#pdf("papers/phd1/plots/ols_mice_convergence.pdf")
#print(plot(mice_set, y = SF_12 + y ~ .it | .ms))
#dev.off()
pdf(paste0(paste0(save.path,"ols_mice_convergence.pdf")))
print(plot(mice_set, SF_12 = SF_12 ~ .it | .ms))
dev.off()

mice.sf12.lm <- with(mice_set, lm(as.formula(model_formula_string)))
final.pool<- pool(mice.sf12.lm)

pooled_lm <- mice.sf12.lm$analyses[[1]]
pool.sum <- summary(final.pool)
pooled_lm$coefficients = pool.sum$estimate
mice.preds<- predict(pooled_lm)

pool.r.squared(final.pool)

# mice objects dont go in forest plot annoyingly. move coefficients over to a
# dummy lm that plot_models can read.
mice.lm.object <- summary(lm(model_formula_string, data = complete(mice_set)))
mice.lm.object$coefficients[,1] <- pool.sum$estimate
mice.lm.object$coefficients[,2] <- pool.sum$std.error
mice.lm.object$coefficients[,3] <- pool.sum$statistic
mice.lm.object$coefficients[,4] <- pool.sum$p.value
save(mice.lm.object, file = paste0(paste0(save.path,'mice_lm_object.RData')))

#texreg(mice.lm.object, dcolumn=T, booktabs=T, file='../plots/mice_ols_coefficients.txt', title="SF12 MICE OLS Coefficients", custom.model.names=c("SF12 OLS"), single.row=T, include.aic=T)

res <- as.data.frame(predict(pooled_lm))
real <- as.data.frame(data$SF_12)
res$type <- c("Predicted")
real$type <- c("Real")
colnames(res) <- c("SF12", 'type')
colnames(real) <- c("SF12", 'type')
hist.data <- rbind(res,real)
hist.data <- tidyr::drop_na(hist.data)

res <- as.data.frame(resid(pooled_lm))
residual_density_plot(res, paste0(save.path,'mice_ols_densities.pdf'), guide='normal')

forest_plot(mice.lm.object, paste0(save.path,'mice_SF12_forest.pdf'))

write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_MICE_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(pooled_lm, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}
```

```{r cross validated nowcasting}
# split data into five
# transition models
# compare coefficients in plots.

```

```{r lasso}
# just import from old paper
lasso.data <- tidyr::drop_na(final.data)
X <- model.matrix(as.formula(model_formula_string), data=lasso.data)
y <- lasso.data$SF_12
weight <- lasso.data$weight
cv.lasso <- cv.glmnet(X, y, family = "gaussian", weights = weight)
lasso <- glmnet(X, y, alpha = 1,  family = "gaussian", weights = weight)

# plots for this.

  pdf(paste0(save.path, 'ols_lasso_2013.pdf'))
  plot(lasso, xvar = "lambda", label = T) 
  abline(v = log(cv.lasso$lambda.min), col = "blue",lty=2)
  abline(v = log(cv.lasso$lambda.1se), col = "red", lty=3)
  legend('bottomright', legend=c("Minimum Standard Error", "First Standard Error"), col=c('blue', 'red'), lty=c(2, 3))
  dev.off()

  pdf(paste0(save.path, 'ols_cv_lasso_2013.pdf'))
  plot(cv.lasso)
  dev.off()

  print(coef(cv.lasso, s = "lambda.min"))
  print(coef(cv.lasso, s = "lambda.1se"))
```

```{r projection of weights}
# get from old paper as well? interesting idea on macro micro
sf12.lm.weighted <- lm(formula, weights = weights, data = final.data)

# plots as with SF12

```

```{r handovers}
# boxplots for handovers
#.estimating from 2014-2020.
# get data for 2013- 2020
# run 2013 pop through transition probability model a few times. 
# profit. 
# run it through model a few times
# compare with real data in boxplots and ridgelines.

dataDir <- "data/raw_US/"
filelist <- list.files(dataDir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[23:30] # get files for 2011-2013.
raw.data <- do.call(rbind, lapply(filelist, read.csv))
raw.data <- replace.missing(raw.data)
raw.data <- raw.data[, c('pidp', 'time', variables)]
raw.data <- drop_na(raw.data)
```

```{r handovers}
start.data <- raw.data[which(raw.data$time==2013),]
rolling.data <- start.data[, c("pidp", "SF_12", "time")]
handover.data <- rolling.data
for (i in 1:7){
  rolling.data$SF_12 <- predict(sf12.lm, start.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data)
}

handover_boxplots(raw.data, handover.data, "SF_12")
density_ridges(raw.data, "SF_12", save=TRUE, save.path=save.path, filename.tag = "raw_ridges")
density_ridges(handover.data, "SF_12", save=TRUE, save.path=save.path, filename.tag = "handover_ridges")

```

## HETEROGENEITY ###############################################################

```{r "glmm"}
glmm.data <- final.data
# yoink transition model function pre processing
reflect <- T
yeo_johnson <- T
include_weights <- T
depend <- "SF_12"
dependent_data <- glmm.data$SF_12

if (reflect) {
    max_value <- nanmax(glmm.data[[depend]])
    glmm.data[, c(depend)] <- max_value - glmm.data[, c(depend)] 
  }
if (yeo_johnson)
{
  yj <- yeojohnson(glmm.data[,c(depend)])
  glmm.data[, c(depend)] <- predict(yj)
}

min_value <- nanmin(glmm.data[[depend]])
glmm.data[[depend]] <- glmm.data[[depend]] - min_value + 0.001

if(include_weights) {
  sf12.glmm.model <- glmer(glmm_model_formula,  
                 nAGQ=0, # fast but inaccurate optimiser. nAGQ=1 takes forever..
                 family=Gamma(link='log'), # gamma family with log link.
                 weights=weight, 
                 data = glmm.data)
} else {
  sf12.glmm.model <- glmer(model_formula, 
                 nAGQ=0, 
                 family=Gamma(link='log'),
                 data = glmm.data)
}

save(sf12.glmm.model, file = paste0(save.path,"sf12_glmm.RData"))

write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF_12_GLMM_coefficients.txt')
  tex_label <- paste0("table: GLMM_SF_12_coefficients") 
  tex_headers <- c(texreg::names2latex("SF_12 GLMM"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "glmm"))    
  texreg(glmm.sf12.lm, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}

glmm.fitted <- (max_value - predict(yj, predict(sf12.glmm.model, type='response') + min_value, inverse=T))
glmm.resid <- scale(dependent_data - glmm.fitted)
forest_plot_glmm(sf12.glmm.model, paste0(save.path, "glmm_forest.pdf"))
qq_plot(glmm.resid, paste0(save.path, "glmm_qq.pdf"))
residual_density_plot(res=glmm.resid , file_name=paste0(save.path, "glmm_residual_density.pdf"), guide="normal")

squareRootRes <- sqrt(abs(scale(glmm.resid)))
fitted_residuals <- as.data.frame(cbind(glmm.fitted, squareRootRes))
colnames(fitted_residuals) <- c("fitted", "sqrt_residuals")
fitted_residual_plot(fitted_residuals, paste0(save.path, "glmm_fitted_residual_plot.pdf"))

```