### 2023-08-31 minos/outcomes/equivalent_income_out.Rmd
Additions to testing scripts for equivalent income interventions, and expansion of the local batch logic

---
title: "Equivalent Income Outcome Visualisation"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(here)
require(dplyr)
require(tidyr)
require(purrr)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

# Data

```{r}
# set output path to SIPHER 7 experiment out
out.path <- here::here('output', 'SIPHER7/')
# read all files into dataframe (util function handles runtime subdirectory etc.)
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
allchild.dat <- read_singular_local_out(out.path, 'hhIncomeChildUplift', drop.dead = TRUE)
pov.dat <- read_singular_local_out(out.path, 'hhIncomePovertyLineChildUplift', drop.dead = TRUE)
wage.dat <- read_singular_local_out(out.path, 'livingWageIntervention', drop.dead = TRUE)
energy.dat <- read_singular_local_out(out.path, 'energyDownlift', drop.dead = TRUE)
```

## Remove 2020 Data Point

```{r}
base.dat <- filter(base.dat, time > 2020)
allchild.dat <- filter(allchild.dat, time > 2020)
pov.dat <- filter(pov.dat, time > 2020)
wage.dat <- filter(wage.dat, time > 2020)
energy.dat <- filter(energy.dat, time > 2020)
```

# Baseline

```{r}
base.subset <- base.dat %>%
  select(pidp, time, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))

base.subset$scenario <- 'Baseline'

ggplot(data = base.subset, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line()
```

# Comparisons

## Function

```{r}
equivalent_income_comparison_fullpop <- function(base, int, int.name, save=FALSE, save.path=NULL) {
  base.subset <- base %>%
    select(pidp, time, equivalent_income, alive, weight) %>%
    filter(alive != 'dead') %>%
    group_by(time) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
  int.subset <- int %>%
    select(pidp, time, equivalent_income, alive, weight) %>%
    filter(alive != 'dead') %>%
    group_by(time) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))

  base.subset$scenario <- 'Baseline'
  int.subset$scenario <- int.name

  combined <- rbind(base.subset, int.subset)

  p <- ggplot(combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_line() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Full Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)

  if(save) {
    if(is.null(save.path)) {
      stop('ERROR: save.path cannot be missing if save=TRUE')
    }
    file.name <- paste0('EI_comp_', int.name, '_full.png')
    ggsave(filename = file.name,
           plot = p,
           path = save.path)
  }
}
```

## Plots

```{r}
equivalent_income_comparison_fullpop(base.dat, allchild.dat, 'allChild')
equivalent_income_comparison_fullpop(base.dat, pov.dat, 'povertyChild')
equivalent_income_comparison_fullpop(base.dat, wage.dat, 'LivingWage')
equivalent_income_comparison_fullpop(base.dat, energy.dat, 'EnergyDownlift')
```

## Combined

```{r}
base.subset <- base.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
base.subset$scenario <- 'Baseline'

allchild.subset <- allchild.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
allchild.subset$scenario <- 'AllChild'

pov.subset <- pov.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
pov.subset$scenario <- 'PovertyChild'

wage.subset <- wage.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
wage.subset$scenario <- 'LivingWage'

energy.subset <- energy.dat %>%
  select(pidp, time, equivalent_income, alive, weight) %>%
  filter(alive != 'dead') %>%
  group_by(time) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
energy.subset$scenario <- 'EnergyDownlift'

all.combined <- rbind(base.subset, allchild.subset, pov.subset, wage.subset, energy.subset)
all.combined$scenario <- factor(all.combined$scenario, levels = c('Baseline', 'AllChild', 'PovertyChild', 'LivingWage', 'EnergyDownlift'))

ggplot(data = all.combined, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line() +
  labs(title = 'Equivalent Income Comparison', subtitle = 'Whole Population') +
  xlab('Year') +
  ylab('Equivalent Income (£)')
```

# Treated Population

## Function

```{r}
equivalent_income_comparison_treated <- function(base, int, int.name, save=FALSE, save.path=NULL) {
  base.sub <- base %>%
    select(pidp, time, equivalent_income, alive) %>%
    filter(alive != 'dead') %>%
    select(-alive) %>%
    rename(baseline = equivalent_income)
  int.sub <- int %>%
    select(pidp, time, equivalent_income, income_boosted, alive, weight) %>%
    filter(alive != 'dead') %>%
    select(-alive) %>%
    set_names(c('pidp', 'time', int.name, 'income_boosted', 'weight'))

  merged <- merge(base.sub, int.sub, by = c('pidp', 'time'))

  boosted.pidps <- int.sub %>%
    filter(income_boosted == 'True') %>%
    select(pidp)

  boosted <- merged %>%
    filter(pidp %in% boosted.pidps$pidp) %>%
    filter(income_boosted == 'True' | time == 2020) %>%
    select(-income_boosted) %>%
    pivot_longer(cols = baseline:.data[[int.name]],
                 names_to = 'scenario',
                 values_to = 'equivalent_income') %>%
    group_by(time, scenario) %>%
    summarise(equivalent_income = weighted.mean(equivalent_income, w=weight))

  p <- ggplot(data = boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
    geom_line() +
    labs(title = paste0('Equivalent Income: ', int.name), subtitle = 'Treated Population') +
    xlab('Year') +
    ylab('Equivalent Income (£)')

  print(p)

  if(save) {
    if(is.null(save.path)) {
      stop('ERROR: save.path cannot be missing if save=TRUE')
    }
    file.name <- paste0('EI_comp_', int.name, '_treated.png')
    ggsave(filename = file.name,
           plot = p,
           path = save.path)
  }
}
```

## Plots

```{r}
equivalent_income_comparison_treated(base.dat, allchild.dat, 'allChild')
equivalent_income_comparison_treated(base.dat, pov.dat, 'povertyChild')
equivalent_income_comparison_treated(base.dat, wage.dat, 'LivingWage')
equivalent_income_comparison_treated(base.dat, energy.dat, 'EnergyDownlift')
```

# Remove Income Effect

```{r, include=FALSE}
# THink about this...
# We can use the income_boosted and boost_amount variables to remove the income boost from people and visualise that

base.sub2 <- base.dat %>%
  select(pidp, time, equivalent_income, alive) %>%
  filter(alive != 'dead') %>%
  select(-alive) %>%
  rename(baseline = equivalent_income)

allchild.sub2 <- allchild.dat %>%
  select(pidp, time, equivalent_income, income_boosted, boost_amount, alive) %>%
  filter(alive != 'dead') %>%
  select(-alive) %>%
  rename(allchild = equivalent_income)

allchild.merged <- merge(base.sub2, allchild.sub2, by = c('pidp', 'time'))

boosted.pidps <- allchild.sub2 %>%
  filter(income_boosted == 'True') %>%
  select(pidp)

allchild.boosted <- allchild.merged %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  filter(income_boosted == 'True' | time == 2020) %>%
  mutate(adjusted_EI = allchild - boost_amount) %>%
  select(-allchild, -income_boosted, -boost_amount) %>%
  group_by(time) %>%
  summarise(baseline = mean(baseline),
            allchild_adj = mean(adjusted_EI)) %>%
  pivot_longer(cols = baseline:allchild_adj,
               names_to = 'scenario',
               values_to = 'equivalent_income')
```

```{r}
ggplot(data = allchild.boosted, aes(x = time, y = equivalent_income, group = scenario, color = scenario)) +
  geom_line()
```

NOTE: Can we 'remove' the income effect?
Use the boost amount and income_boosted variables to sort of 'undo' the intervention effect and then we are only visualising the difference from pathway changes and not just the income effect.

ALSO: Difference plots.

# DETAILED ANALYSIS

## Living Wage

For this, we're going to look at the difference in equivalent income and try to explain it by looking at the pathway variables. What we want to see is that as income is uplifted, the various pathways show a positive shift and that in part explains the difference in equivalent income.

First visualise both income and equivalent income on one graph.

### Income and Equivalent Income

```{r}
base.inc <- base.dat %>%
  select(pidp, time, hh_income, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
            equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Baseline')

int.inc <- wage.dat %>%
  select(pidp, time, hh_income, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
            equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Living_Wage')

combined <- rbind(base.inc, int.inc)
combined <- combined %>%
  pivot_longer(cols=hh_income:equivalent_income,
               names_to = 'income_type',
               values_to = 'value')

ggplot(combined, aes(x = time, y = value, group = source, color = source)) +
  geom_line() +
  facet_wrap( ~ income_type, nrow=2, scales='free')
```

### CHANGE IN INCOME / EQUIVALENT INCOME

```{r}
base.inc <- base.dat %>%
  select(pidp, time, hh_income, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
            equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Baseline')

int.inc <- energy.dat %>%
  select(pidp, time, hh_income, equivalent_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
            equivalent_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'EnergyCrisis')

combined <- rbind(base.inc, int.inc)
combined <- combined %>%
  pivot_longer(cols=hh_income:equivalent_income,
               names_to = 'income_type',
               values_to = 'value')

ggplot(combined, aes(x = time, y = value, group = source, color = source)) +
  geom_line() +
  facet_wrap( ~ income_type, nrow=2, scales='free')
```

How about on the same plot to make it more clear?

```{r}
income_equivinc_change <- function(base.dat, int.dat) {
  base.inc <- base.dat %>%
    select(pidp, time, hh_income, equivalent_income, weight) %>%
    group_by(time) %>%
    summarise(base_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              base_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  int.inc <- int.dat %>%
    select(pidp, time, hh_income, equivalent_income, weight) %>%
    group_by(time) %>%
    summarise(energy_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              energy_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  combined <- merge(base.inc, int.inc, by = c('time'))

  combined$diff_income <- combined$energy_income - combined$base_income
  combined$diff_equiv_income <- combined$energy_equiv_income - combined$base_equiv_income

  combined <- combined %>%
    pivot_longer(cols = diff_income:diff_equiv_income,
                 names_prefix = 'diff_',
                 names_to = 'variable',
                 values_to = 'diff')

  ggplot(combined, aes(x = time, y = diff, group = variable, color = variable)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype = 'dashed')
}

income_equivinc_change(base.dat, energy.dat)
income_equivinc_change(base.dat, wage.dat)
```

## Energy Crisis

```{r}
S7_ordinal_pathway_vis <- function(base.dat, int.dat, int.name, var) {
  base.hous <- base.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = 'Baseline')

  inc.hous <- int.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = int.name)

  combined <- rbind(base.hous, inc.hous)
  combined[[var]] <- as.factor(combined[[var]])

  p1 <- ggplot(combined, aes(x = time, y = prop, group = interaction(.data[[var]], source), color = .data[[var]], linetype = source)) +
    geom_line() +
    geom_point()

  print(p1)

  base.hous2 <- base.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(Baseline = prop)

  int.hous2 <- int.dat %>%
    select(pidp, time, all_of(var)) %>%
    group_by(time, .data[[var]]) %>%
    count() %>%
    group_by(time) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(int = prop)

  combined2 <- merge(base.hous2, int.hous2, by = c('time', var))

  combined2[[var]] <- as.factor(combined2[[var]])

  combined2$diff <- combined2$int - combined2$Baseline

  p2 <- ggplot(combined2, aes(x = time, y = diff, group = .data[[var]], color = .data[[var]])) +
    geom_line() +
    geom_point()

  print(p2)
}
```

### Housing Quality

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_housing_quality')
```

### Neighbourhood Safety

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_neighbourhood_safety')
```

### Mental Wellbeing

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_mental_health')
```

### Physical Health

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_physical_health')
```

### Loneliness

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'loneliness')
```

### Labour State

```{r}
S7_ordinal_pathway_vis(base.dat, energy.dat, 'EnergyCrisis', 'S7_labour_state')
```

```{r}
boosted.pidps <- energy.dat %>%
  filter(income_boosted == 'True') %>%
  select(pidp)

base.hous <- base.dat %>%
  select(pidp, time, S7_housing_quality) %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  group_by(time, S7_housing_quality) %>%
  count() %>%
  group_by(time) %>%
  mutate(total = sum(n)) %>%
  mutate(prop = (n / total) * 100) %>%
  select(-n, -total) %>%
  rename(Baseline = prop)

int.hous <- energy.dat %>%
  select(pidp, time, S7_housing_quality, income_boosted) %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  group_by(time, S7_housing_quality) %>%
  count() %>%
  group_by(time) %>%
  mutate(total = sum(n)) %>%
  mutate(prop = (n / total) * 100) %>%
  select(-n, -total) %>%
  rename(EnergyCrisis = prop)

combined <- merge(base.hous, int.hous, by = c('time', 'S7_housing_quality'))

combined$diff <- combined$EnergyCrisis - combined$Baseline

ggplot(combined, aes(x = time, y = diff, group = S7_housing_quality, color = S7_housing_quality)) +
  geom_line() +
  geom_point()
```

```{r}
## just look at means in case the means are causing some problems
base.dat.2 <- base.dat
base.dat.2$S7_housing_quality2[base.dat.2$S7_housing_quality == 'No to all'] <- 1
base.dat.2$S7_housing_quality2[base.dat.2$S7_housing_quality == 'Yes to some'] <- 2
base.dat.2$S7_housing_quality2[base.dat.2$S7_housing_quality == 'Yes to all'] <- 3
wage.dat.2 <- energy.dat
wage.dat.2$S7_housing_quality2[wage.dat.2$S7_housing_quality == 'No to all'] <- 1
wage.dat.2$S7_housing_quality2[wage.dat.2$S7_housing_quality == 'Yes to some'] <- 2
wage.dat.2$S7_housing_quality2[wage.dat.2$S7_housing_quality == 'Yes to all'] <- 3

base.hous.mean <- base.dat.2 %>%
  select(pidp, time, S7_housing_quality2, weight) %>%
  group_by(time) %>%
  summarise(mean_housing = weighted.mean(S7_housing_quality2, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'Baseline')
wage.hous.mean <- wage.dat.2 %>%
  select(pidp, time, S7_housing_quality2, weight) %>%
  group_by(time) %>%
  summarise(mean_housing = weighted.mean(S7_housing_quality2, w = weight, na.rm = TRUE)) %>%
  mutate(source = 'EnergyCrisis')

combined <- rbind(base.hous.mean, wage.hous.mean)

ggplot(combined, aes(x = time, y = mean_housing, group = source, color = source)) +
  geom_line() +
  geom_point()
```

```{r}

# df <- combine_and_pivot_long(base.dat, 'Baseline', wage.dat, 'LivingWage', c('S7_housing_quality', 'income_boosted'))
# 
# outcomes.prep <- function(base.dat, int.dat, int.name, var) {
#   boosted.pidps <- int.dat %>%
#     filter(income_boosted == 'True') %>%
#     select(pidp, time)
#   
#   df1 <- base.dat %>%
#     select(pidp, time, all_of(var), weight) %>%
#     
# }
# 
# 
# 
# 
# combine_and_pivot_long <- function(df1, df1.name, df2, df2.name, var) {
#   # get only the columns we want a rename
#   df1 <- df1 %>%
#     select('pidp', 'time', all_of(var)) %>%
#     set_names(c('pidp', 'time', df1.name))
#   df2 <- df2 %>%
#     select('pidp', 'time', all_of(var)) %>%
#     set_names(c('pidp', 'time', df2.name))
#   # merge on pidp and time
#   merged <- merge(df1, df2, by = c('pidp', 'time'), all=TRUE)
#   pivoted <- pivot_longer(data = merged,
#                           cols = df1.name:df2.name,
#                           names_to = 'scenario',
#                           values_to = var)
#   return(pivoted)
# }

```