### 2023-09-17 minos/outcomes/EI_PresPlots_Refined.Rmd
Lots of testing on S7 component models. Also added a script to check the proportional odds assumption for deciding between clm and nnet. Still working on a function to add to the transition estimation scripts to produce a file with each model

---
title: "Equivalent Income Presentation Plots - FINAL"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(here)
require(dplyr)
require(tidyr)
require(purrr)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

# Data

## BATCH DATA

```{r}
S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
                 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
                 'ethnicity', 'age', 'region', 'job_sec', 'education_state', 'nkids_ind', 'housing_tenure', 'urban')
out.path.batch <- here::here('output', 'SIPHER7_batch/')

base.batch <- read_batch_out_all_years(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=TRUE)
int.batch <- read_batch_out_all_years(out.path.batch, 'energyDownlift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=TRUE)

# out.path.arc2 <- here::here('output', 'batch')
# 
# #target.pattern <- paste0('aggregated_run_id_', year, '.csv')
# base.list <- list.files(path = paste0(out.path.arc2, '/baseline'),
#                             full.names = TRUE)
# base.batch <- do.call(rbind, lapply(base.list, read.csv))
# 
# energy.list <- list.files(path = paste0(out.path.arc2, '/energy'),
#                             full.names = TRUE)
# int.batch <- do.call(rbind, lapply(energy.list, read.csv))

int.name <- 'Energy Crisis'
```

## Calculate Age Group

```{r}
calc_age_group <- function(batch.dat) {
  batch.dat <- batch.dat %>%
    mutate(age.group = cut(age, 
                            breaks = c(15, 30, 45, 60, 75, 120),
                            labels = c('16-30', '30-45', '45-60', '60-75', '75+')))
  return(batch.dat)
}

base.batch <- calc_age_group(base.batch)
int.batch <- calc_age_group(int.batch)
```

## Calculate Income Quintile

```{r}
calc_income_quintile <- function(batch.dat) {
  batch.dat <- batch.dat %>%
    mutate(income.quint = cut(batch.dat$hh_income, breaks=c(quantile(batch.dat$hh_income, probs = seq(0, 1, by = 0.20))), 
                          labels=c(1,2,3,4,5), include.lowest=TRUE))
}

base.batch <- calc_income_quintile(base.batch)
int.batch <- calc_income_quintile(int.batch)
```

## FUNCTIONS

```{r}
income_equivinc_change_batch <- function(base.batch, int.batch, int.name) {
  base.inc <- base.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, weight) %>%
    group_by(time, run_id) %>%
    summarise(base_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              base_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  int.inc <- int.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, weight) %>%
    group_by(time, run_id) %>%
    summarise(energy_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              energy_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  combined <- merge(base.inc, int.inc, by = c('time', 'run_id'))

  combined$diff_income <- combined$energy_income - combined$base_income
  combined$diff_equiv_income <- combined$energy_equiv_income - combined$base_equiv_income

  combined <- combined %>%
    pivot_longer(cols = diff_income:diff_equiv_income,
                 names_prefix = 'diff_',
                 names_to = 'variable',
                 values_to = 'diff')

  combined2 <- combined %>%
    group_by(time, variable) %>%
    summarise(mean.diff = mean(diff))

  ggplot(combined, aes(x = time, y = diff, group = variable, color = variable)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = 'Change in Income and Equivalent Income', 
         subtitle = int.name)
}

equivinc_change_batch_subpop <- function(base.batch, int.batch, int.name, group.var) {
  base.inc <- base.batch %>%
    select(pidp, time, run_id, equivalent_income, weight, all_of(group.var)) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(base_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  int.inc <- int.batch %>%
    select(pidp, time, run_id, equivalent_income, weight, all_of(group.var)) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(int_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  combined <- merge(base.inc, int.inc, by = c('time', 'run_id', group.var))
  combined[[group.var]] <- as.factor(combined[[group.var]])

  combined$diff_equiv_income <- combined$int_equiv_income - combined$base_equiv_income

  combined <- combined %>%
    group_by(time, .data[[group.var]]) %>%
    summarise(mean_diff_equiv_income = mean(diff_equiv_income))

  ggplot(combined, aes(x = time, y = mean_diff_equiv_income, group = .data[[group.var]], color = .data[[group.var]])) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0('Equivalent Income change by ', group.var), subtitle = int.name) +
    xlab('Year') +
    ylab('Equivalent Income (£s)')
}

inc_change_batch_subpop <- function(base.batch, int.batch, int.name, group.var) {
  base.inc <- base.batch %>%
    select(pidp, time, run_id, hh_income, weight, all_of(group.var)) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(base_hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE))

  int.inc <- int.batch %>%
    select(pidp, time, run_id, hh_income, weight, all_of(group.var)) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(int_hh_income = weighted.mean(hh_income, w = weight, na.rm = TRUE))

  combined <- merge(base.inc, int.inc, by = c('time', 'run_id', group.var))
  combined[[group.var]] <- as.factor(combined[[group.var]])

  combined$diff_hh_income <- combined$int_hh_income - combined$base_hh_income

  combined <- combined %>%
    group_by(time, .data[[group.var]]) %>%
    summarise(mean_diff_hh_income = mean(diff_hh_income))

  ggplot(combined, aes(x = time, y = mean_diff_hh_income, group = .data[[group.var]], color = .data[[group.var]])) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0('Household Income change by ', group.var), subtitle = int.name) +
    xlab('Year') +
    ylab('Household Income (£s)')
}

income_equivinc_percent_change_batch <- function(base.batch, int.batch, int.name, group.var) {
  base.inc <- base.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, all_of(group.var), weight) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(base_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              base_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  int.inc <- int.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, all_of(group.var), weight) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(int_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              int_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

  combined <- merge(base.inc, int.inc, by = c('time', 'run_id', group.var))

  combined$diff_income <- combined$int_income - combined$base_income
  combined$diff_equiv_income <- combined$int_equiv_income - combined$base_equiv_income

  combined <- combined %>%
    group_by(time, .data[[group.var]]) %>%
    summarise(mean_income = mean(base_income),
              mean_equiv_income = mean(base_equiv_income),
              mean_diff_income = mean(diff_income),
              mean_diff_equiv_income = mean(diff_equiv_income)) %>%
    mutate(percent_change_income = (mean_diff_income / mean_income) * 100,
           percent_change_equiv_income = (mean_diff_equiv_income / mean_equiv_income) * 100) %>%
    select(-mean_income, -mean_equiv_income, -mean_diff_income, -mean_diff_equiv_income) %>%
    pivot_longer(cols = percent_change_income:percent_change_equiv_income,
                 names_prefix = 'percent_change_',
                 names_to = 'variable',
                 values_to = 'change')

  ggplot(combined, aes(x = time, y = change, group = variable, color = variable)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    facet_wrap(~ income.quint, nrow = 1) +
    labs(title = 'Percent change: Income vs Equivalent Income',
         subtitle = int.name) +
    xlab('Year') +
    ylab('Percent Change')
}

S7_pathway_change_byGroup <- function(base.batch, int.batch, group.var, var) {
  base.sub <- base.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           base_prop = n / total) %>%
    select(-n, -total)

  int.sub <- int.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           int_prop = n / total) %>%
    select(-n, -total)

  combined <- merge(base.sub, int.sub, by = c('time', 'run_id', group.var, var))
  combined$prop_diff <- combined$int_prop - combined$base_prop

  combined[[var]] <- as.factor(combined[[var]])

  ggplot(data = combined, aes(x = time, y = prop_diff, group = .data[[var]], color = .data[[var]])) +
    geom_smooth() +
    facet_wrap(~ .data[[group.var]], nrow=1) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
}

S7_cumulative_pathwayChange_byGroup <- function(base.batch, int.batch, group.var, var) {
  base.sub <- base.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           base_prop = n / total) %>%
    select(-n, -total)

  int.sub <- int.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           int_prop = n / total) %>%
    select(-n, -total)

  combined <- merge(base.sub, int.sub, by = c('time', 'run_id', group.var, var))
  combined$prop_diff <- combined$int_prop - combined$base_prop

  combined[[var]] <- as.factor(combined[[var]])

  combined <- combined %>%
    group_by(run_id, .data[[group.var]], .data[[var]]) %>%
    summarise(cumulative_diff = sum(prop_diff)) %>%
    group_by(.data[[group.var]], .data[[var]]) %>%
    summarise(cumulative_diff = mean(cumulative_diff))

  ggplot(data = combined, aes(x = .data[[var]], y = cumulative_diff, fill = .data[[var]])) +
    geom_col() +
    facet_wrap(~ .data[[group.var]], nrow=1) +
    labs(title = paste0('Cumulative Difference in ', var, ' by ', group.var)) +
    ylab('Cumulative Difference (%)') +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = 'bottom')
}

simplified_component_change_BOXPLOT <- function(base.batch, int.batch, group.var, var.list, print = FALSE) {

  # it would be lovely if this was completely generic, but for some things that
  # is more hassle than its worth. Handling string vars here is one of these 
  # things

  base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                          levels = c('No to all',
                                                     'Yes to some',
                                                     'Yes to all'),
                                          labels = c(1, 2, 3)))
  int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                          levels = c('No to all',
                                                     'Yes to some',
                                                     'Yes to all'),
                                          labels = c(1, 2, 3)))
  base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                               levels = c('Often',
                                                          'Some of the time',
                                                          'Hardly ever'),
                                               labels = c(1, 2, 3)))
  int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                               levels = c('Often',
                                                          'Some of the time',
                                                          'Hardly ever'),
                                               labels = c(1, 2, 3)))
  base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                  levels = c('Often',
                                             'Sometimes',
                                             'Never'),
                                  labels = c(1, 2, 3)))
  int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                  levels = c('Often',
                                             'Sometimes',
                                             'Never'),
                                  labels = c(1, 2, 3)))

  base.sub <- base.batch %>%
    select(pidp, time, run_id, all_of(group.var), all_of(var.list)) %>%
    rename_with(.fn = ~paste0('Baseline_', .), .cols = all_of(var.list))

  int.sub <- int.batch %>%
    select(pidp, time, run_id, all_of(var.list)) %>%
    rename_with(.fn = ~paste0('Intervention_', .), .cols = all_of(var.list))

  merged <- merge(base.sub, int.sub,
                  by=c('pidp', 'time', 'run_id'),
                  all = FALSE)

  # https://stackoverflow.com/questions/49694353/dplyr-mutating-multiple-columns-by-prefix-and-suffix

  merged2 <- map(var.list, ~ merged %>%
                   mutate(!!paste0("Change_", .x) := !!as.name(paste0("Intervention_", .x)) - !!as.name(paste0("Baseline_", .x)))) %>%
    reduce(left_join) %>%
    select(-starts_with('Baseline_'), -starts_with('Intervention_')) %>%
    pivot_longer(cols = paste0('Change_', first(var.list)):paste0('Change_', last(var.list)),
                 names_to = 'Variable',
                 names_prefix = 'Change_',
                 values_to = 'Change') %>%
    mutate(effect = case_when(Change > 0 ~ "Positive",
                              Change < 0 ~ "Negative",
                              Change == 0 ~ "No_Change")) %>%
    select(-Change) %>%
    group_by(time, run_id, .data[[group.var]], Variable, effect) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]], Variable) %>%
    mutate(total = sum(n),
           effect_prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    pivot_wider(names_from = 'effect',
                values_from = 'effect_prop') %>%
    mutate(change = Positive - Negative) %>%
    select(-Positive, -Negative, -No_Change) %>%
    group_by(run_id, .data[[group.var]], Variable) %>%
    summarise(sum.change = sum(change))

  ggplot(merged2, aes(x = Variable, y = sum.change, group = Variable, colour = Variable, fill = Variable)) +
    geom_boxplot(notch = FALSE) +
    facet_wrap(~ .data[[group.var]], nrow = 1) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0("Change in component variables by ", group.var)) +
    ylab("Net Change (%)") +
    theme(legend.position = 'bottom',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())

}
```

# Income and Equivalent Income

## Absolute Change

```{r}
income_equivinc_change_batch(base.batch, int.batch, int.name)
```

## Percent Change

```{r}

```

# EquivInc split by Income Quintile

```{r}
inc_change_batch_subpop(base.batch, int.batch, int.name, group.var = 'income.quint')
equivinc_change_batch_subpop(base.batch, int.batch, int.name, group.var = 'income.quint')
```

## Percent Change

```{r}
income_equivinc_percent_change_batch(base.batch, int.batch, int.name, group.var = 'income.quint')
```

# Pathway Vars by Income Quintile

```{r}
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness')
```

## Cumulative Pathway Change

### Income Quintile

```{r}
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness')
```

### Age Group

```{r}
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_housing_quality')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_neighbourhood_safety')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_physical_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_mental_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'loneliness')
```

# Simplified Cumulative Change

```{r}
varlist <- c('S7_housing_quality', 'S7_neighbourhood_safety', 'loneliness',
             'S7_physical_health', 'S7_mental_health')

simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var.list = varlist)
```

```{r}

```