### 2023-09-17 minos/testing/proportional_odds_assumption_test.Rmd

---
title: "R Notebook"
output: html_notebook
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(here)
require(dplyr)
require(tidyr)
require(purrr)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

Source utils for some functions.

```{r}
# source(here::here('minos', 'utils_datain.R'))
# source(here::here('minos', 'utils_validation_vis.R'))
# source(here::here('minos', 'validation', 'utils.r'))
```

## Data

```{r}
# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

#out.path <- 'Minos/output/SIPHER7/'
#out.path <- here::here('output', 'SIPHER7/')
#base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)

# raw.dat <- raw.dat %>%
#   #filter(loneliness != '-9')
#   #filter(S7_housing_quality != '-9')
#   filter(S7_labour_state != '-9')
```

# TEST

```{r}
require(car)
require(nnet)
#library(xlsx)
require(MASS)
#options(warn=1)
options(digits = 3)

var.name <- 'job_sec'

raw.dat <- raw.dat %>%
  filter(.data[[var.name]] != '-9')

file.name <- paste0('PROP_ODDS_TESTING/Prop_odds_', var.name, '.txt')

#
#Trial <- read.xlsx("Trial.xls", "Sheet 1")
# Set up an out file structure
sink(file.name)
# Trial$Outcome is assessed on a six point scale 0-5
# loneliness
#schtyp_M_M.f <- factor(raw.dat$loneliness, labels = c('Never', 'Sometimes', 'Often'))
#schtyp_M_M.f <- factor(raw.dat$S7_housing_quality, labels = c('No to all', 'Yes to some', 'Yes to all'))
#schtyp_M_M.f <- factor(raw.dat$S7_labour_state, labels = c('FT Employed', 'PT Employed', 'Job Seeking', 'Family Care', 'Not Working', 'FT Education'))
#schtyp_M_M.f <- factor(raw.dat$S7_labour_state)
schtyp_M_M.f <- factor(raw.dat[[var.name]])
#
cat("Multinomial logistic regression \n")
# Assign takes on a value of 1 (Treatment) or 0 (Control) 
mod.multinom <-multinom(schtyp_M_M.f~sex, data = raw.dat)
print(summary(mod.multinom, cor=F, Wald=T))
x1<-logLik(mod.multinom)
cat("Degrees of freedom Multinomial logistic regression \n")
print(df_of_multinom_model <- attributes(x1)$df)
cat("Proportional odds logistic regression\n")
mod.polr <- polr(schtyp_M_M.f ~ sex, data=raw.dat)
print(summary(mod.polr))
x2<-logLik(mod.polr)
cat("Degrees of freedom Proportional Odds Logistic Regression \n")
print(df_of_polr_model <- attributes(x2)$df)

cat("Answering the question: Is proportional odds model assumption violated\n")
cat("P value for difference in AIC between POLR and Multinomial Logit model\n")
# abs since the values could be negative. That is negative difference of degrees of freedom would produce p=NaN
print(1-pchisq(abs(mod.polr$deviance-mod.multinom$deviance),   abs(df_of_multinom_model-df_of_polr_model)))
sink()
```

```{r}

prop_odds_assumption_test <- function(test.dat, test.var, form.right) {
  # NOTE
  # The following code came from this stack overflow which itself came from a book:
  # https://stackoverflow.com/questions/37016215/testing-the-proportional-odds-assumption-in-r

  require(car)
  require(nnet)
  require(MASS)
  options(warn=1)
  options(digits = 3)

  # Set up an out file structure
  sink(paste0('Testing_adequacy_of_Prop_odds_', test.var, '.txt'))
  # set up response factor
  schtyp_M_M.f <- factor(test.dat[[test.var]])
  # set up formula object for multinom
  form <- as.formula(paste0("schtyp_M_M.f ~ ", form.right))
  print(form)
  #
  cat("Multinomial logistic regression \n")
  # Test against the formula from file
  mod.multinom <-multinom(formula = form, data = test.dat)
  print(summary(mod.multinom, cor=F, Wald=T))
  x1<-logLik(mod.multinom)
  cat("Degrees of freedom Multinomial logistic regression \n")
  print(df_of_multinom_model <- attributes(x1)$df)
  cat("Proportional odds logistic regression\n")
  mod.polr <- polr(formula = form, data = test.dat)
  print(summary(mod.polr))
  x2<-logLik(mod.polr)
  cat("Degrees of freedom Proportional Odds Logistic Regression \n")
  print(df_of_polr_model <- attributes(x2)$df)

  cat("Answering the question: Is proportional odds model assumption violated\n")
  cat("P value for difference in AIC between POLR and Multinomial Logit model\n")
  # abs since the values could be negative. That is negative difference of degrees of freedom would produce p=NaN
  print(1-pchisq(abs(mod.polr$deviance-mod.multinom$deviance),   abs(df_of_multinom_model-df_of_polr_model)))
  sink()
}
```

```{r}
raw.dat <- raw.dat %>%
  filter(loneliness != '-9')

form.rightside <- "factor(loneliness) + scale(age) + factor(sex) + relevel(factor(education_state), ref = '3') + relevel(factor(job_sec), ref = '3') + scale(hh_income) + relevel(factor(hh_comp), ref = '3') + relevel(factor(marital_status), ref = 'Partnered') + relevel(factor(ethnicity), ref = 'WBI')"

form.rightside <- 'sex'

#prop_odds_assumption_test(test.dat = raw.dat, test.var = 'loneliness', form.right = form.rightside)

test.dat <- raw.dat
test.var <- 'loneliness'
form.right <- form.rightside

require(car)
require(nnet)
require(MASS)
options(warn=1)
options(digits = 3)

# Set up an out file structure
sink(paste0('Testing_adequacy_of_Prop_odds_', test.var, '.txt'))
# set up response factor
schtyp_M_M.f <- factor(test.dat[[test.var]])
# set up formula object for multinom
form <- as.formula(paste0("schtyp_M_M.f ~ ", form.right))
print(form)
#
cat("Multinomial logistic regression \n")
# Test against the formula from file
mod.multinom <-multinom(formula = form, data = test.dat)
print(summary(mod.multinom, cor=F, Wald=T))
x1<-logLik(mod.multinom)
cat("Degrees of freedom Multinomial logistic regression \n")
print(df_of_multinom_model <- attributes(x1)$df)
cat("Proportional odds logistic regression\n")
mod.polr <- polr(formula = form, data = test.dat)
print(summary(mod.polr))
x2<-logLik(mod.polr)
cat("Degrees of freedom Proportional Odds Logistic Regression \n")
print(df_of_polr_model <- attributes(x2)$df)

cat("Answering the question: Is proportional odds model assumption violated\n")
cat("P value for difference in AIC between POLR and Multinomial Logit model\n")
# abs since the values could be negative. That is negative difference of degrees of freedom would produce p=NaN
print(1-pchisq(abs(mod.polr$deviance-mod.multinom$deviance),   abs(df_of_multinom_model-df_of_polr_model)))
sink()
```

```{r}

```