### 2023-10-30 minos/testing/QALY_comparison_energyCrises.Rmd
Implemented utility score and QALY calculations, using the Energy Crises interventions as a test comparison

---
title: "QALY Comparisons - Energy Crisis Interventions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
```

## Data

```{r}
# read in QALY files for baseline and interventions
out.path <- here::here('output', 'default_config')

base.path <- get_latest_runtime_subdirectory(here::here(out.path, 'baseline'))
support.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownlift'))
noSupport.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownliftNoSupport'))

base <- read.csv(paste0(base.path, 'qalys.csv'))
support <- read.csv(paste0(support.path, 'qalys.csv'))
noSupport <- read.csv(paste0(noSupport.path, 'qalys.csv'))
```

# QALY comparison

```{r}
# combine the dataframes
combined <- rbind(base, support, noSupport)

ggplot(data = combined, aes(x = year, y = QALYs, group = intervention, colour = intervention)) +
  geom_line()
```

## Area Under the Curve Comparison

```{r}
#base.auc <- sum(diff(base$time)*rollmean(base$QALYs, 2))
base.auc <- data.frame(intervention = 'baseline', AUC = auc(x=base$year, y=base$QALYs, type = 'spline'))
support.auc <- data.frame(intervention = 'Support', AUC = auc(x=support$year, y=support$QALYs, type = 'spline'))
noSupport.auc <- data.frame(intervention = 'NoSupport', AUC = auc(x=noSupport$year, y=noSupport$QALYs, type = 'spline'))

combined.auc <- rbind(base.auc, support.auc, noSupport.auc)
```

```{r}
print(combined.auc)

ggplot(data = combined.auc, aes(x = intervention, y = AUC, group = intervention, fill = intervention)) +
  geom_col()
```

```{r}
# Another version showing difference from baseline for better comparison
combined.auc2 <- combined.auc %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'AUC') %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = Support - baseline,
         change_noSupport = NoSupport - baseline) %>%
  pivot_longer(cols = change_baseline:change_noSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'AUC_difference')

print(combined.auc2)

ggplot(combined.auc2, aes(x = intervention, y = AUC_difference, group = intervention, fill = intervention)) +
  geom_col()
```