### 2023-10-31 minos/testing/QALY_comparison_energyCrises.Rmd
Added cost effectiveness calculations to QALY comparison energyCrises testing script. Added Make target to run QALY calculation for living wage intervention. Added MESS package to environment.yml

---
title: "QALY Comparisons - Energy Crisis Interventions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
```

## Data

```{r}
# # read in QALY files for baseline and interventions
out.path <- here::here('output', 'default_config')

base.path <- get_latest_runtime_subdirectory(here::here(out.path, 'baseline'))
support.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownlift'))
noSupport.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownliftNoSupport'))

base <- read.csv(paste0(base.path, 'qalys.csv'))
support <- read.csv(paste0(support.path, 'qalys.csv'))
noSupport <- read.csv(paste0(noSupport.path, 'qalys.csv'))
```

# QALY comparison

```{r}
# combine the dataframes
combined <- rbind(base, support, noSupport)

ggplot(data = combined, aes(x = year, y = QALYs, group = intervention, colour = intervention)) +
  geom_smooth()
```

## Groupby and Mean

```{r}
base <- base %>%
  group_by(year) %>%
  summarise(QALYs = mean(QALYs),
            boost_amount = mean(total_boost))
support <- support %>%
  group_by(year) %>%
  summarise(QALYs = mean(QALYs),
            boost_amount = mean(total_boost))
noSupport <- noSupport %>%
  group_by(year) %>%
  summarise(QALYs = mean(QALYs),
            boost_amount = mean(total_boost))
```

## Area Under the Curve Comparison

```{r}
#base.auc <- sum(diff(base$time)*rollmean(base$QALYs, 2))
base.auc <- data.frame(intervention = 'baseline', AUC = auc(x=base$year, y=base$QALYs, type = 'spline'))
support.auc <- data.frame(intervention = 'Support', AUC = auc(x=support$year, y=support$QALYs, type = 'spline'))
noSupport.auc <- data.frame(intervention = 'NoSupport', AUC = auc(x=noSupport$year, y=noSupport$QALYs, type = 'spline'))

combined.auc <- rbind(base.auc, support.auc, noSupport.auc)
```

```{r}
print(combined.auc)

ggplot(data = combined.auc, aes(x = intervention, y = AUC, group = intervention, fill = intervention)) +
  geom_col()
```

```{r}
# Another version showing difference from baseline for better comparison
combined.auc2 <- combined.auc %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('AUC')) %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = Support - baseline,
         change_noSupport = NoSupport - baseline) %>%
  pivot_longer(cols = change_baseline:change_noSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'AUC_difference')

print(combined.auc2)

ggplot(combined.auc2, aes(x = intervention, y = AUC_difference, group = intervention, fill = intervention)) +
  geom_col()
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}

combined.cost <- combined %>%
  select(-X, -pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  mutate(QALY_value = QALYs * 60000)

ggplot(combined.cost, aes(x = year, y = QALY_value, group = intervention, colour = intervention)) +
  geom_line()

combined.cost.change <- combined.cost %>%
  select(-total_boost, -QALYs) %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'QALY_value') %>%
  mutate(change_Support = energyDownlift - baseline,
         change_NoSupport = energyDownliftNoSupport - baseline) %>%
  select(-baseline, -energyDownlift, -energyDownliftNoSupport) %>%
  pivot_longer(cols = -year,
               names_prefix = 'change_',
               names_to = 'intervention',
               values_to = 'value_change')

ggplot(combined.cost.change, aes(x = year, y = value_change, group = intervention, colour = intervention)) +
  geom_line()

  # select(-QALYs) %>%
  # pivot_wider(names_from = 'intervention',
  #             values_from = c('total_boost', 'QALY_value'),
  #             names_sep = '000') %>%
  # mutate(QALY_value_change000energyDownlift = QALY_value000energyDownlift - QALY_value000baseline,
  #        QALY_value_change000energyDownliftNoSupport = QALY_value000energyDownliftNoSupport - QALY_value000baseline) %>%
  # select(-QALY_value000baseline, -QALY_value000energyDownlift, -QALY_value000energyDownliftNoSupport, -total_boost000baseline) %>%
  # pivot_longer(cols = total_boost000energyDownlift:QALY_value_change000energyDownliftNoSupport,
  #              names_sep = '000',
  #              names_to = c('.value', 'intervention'),
  #              names_repair = 'unique') #%>%
  # pivot_longer(cols = QALY_value_change_Support:QALY_value_change_NoSupport,
  #              names_prefix = 'QALY_value_change_',
  #              names_to = 'intervention2',
  #              values_to = 'QALY_value_change')

# combined.cost <- combined %>%
#   select(-X, -pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
#   pivot_wider(names_from = 'intervention',
#               values_from = c('mean_boost', 'QALYs')) %>%
#   mutate(change_energyDownlift = QALYs_energyDownlift - QALYs_baseline,
#          change_energyDownliftNoSupport = QALYs_energyDownliftNoSupport - QALYs_baseline) %>%
#   mutate(costPerQALY_energyDownlift = change_energyDownlift / mean_boost_energyDownlift,
#          costPerQALY_energyDownliftNoSupport = change_energyDownliftNoSupport / mean_boost_energyDownliftNoSupport) %>%
#   pivot_longer(cols = costPerQALY_energyDownlift:costPerQALY_energyDownliftNoSupport,
#                names_prefix = 'costPerQALY_',
#                names_to = 'intervention',
#                values_to = 'costPerQALY') %>%
#   select(year, intervention, costPerQALY)

# options(scipen=999)
# 
# ggplot(data = combined.cost, aes(x = year, y = costPerQALY, group = intervention, colour = intervention)) +
#   geom_line()
```

```{r}
total.combined.cost <- combined.cost %>%
  group_by(intervention) %>%
  summarise(total_value_change = sum(value_change))

ggplot(data = total.combined.cost, aes(x = intervention, y = total_value_change, group = intervention, fill = intervention)) +
  geom_col()
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).

```{r}
# ICER = (C1 - C0) / (E1 - E0)
# where
# C1 = cost of intervention
# E1 = effect of intervention
# C0 = cost of control group
# E0 = effect of control group

ICER <- combined.cost %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('total_boost', 'QALYs', 'QALY_value')) %>%
  mutate(ICER_Support = (total_boost_energyDownlift - total_boost_baseline) / (QALYs_energyDownlift - QALYs_baseline),
         ICER_NoSupport = (total_boost_energyDownliftNoSupport - total_boost_baseline) / (QALYs_energyDownliftNoSupport - QALYs_baseline)) %>%
  select(year, ICER_Support, ICER_NoSupport) %>%
  pivot_longer(cols = -year,
               names_prefix = 'ICER_',
               names_to = 'intervention',
               values_to = 'ICER')

ggplot(ICER, aes(x = year, y = ICER, group = intervention, colour = intervention)) +
  geom_line()

ICER.final <- ICER %>%
  group_by(intervention) %>%
  summarise(ICER = mean(ICER))

ggplot(ICER.final, aes(x = intervention, y = ICER, group = intervention, fill = intervention, colour = intervention)) +
  geom_col()
```