### 2023-11-01 minos/testing/QALY_comparison_energyCrises.Rmd
Added run_id to QALY_calculation outputs for plotting confidence. Improved script for comparing energy crises interventions, and added new script for QALY calculations on living wage intervention. Also fixed function to get runtime subdirectory path in utils

---
title: "QALY Comparisons - Energy Crisis Interventions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
```

## Data

```{r}
# # read in QALY files for baseline and interventions
out.path <- here::here('output', 'default_config')

base.path <- get_latest_runtime_subdirectory(here::here(out.path, 'baseline'))
support.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownlift'))
noSupport.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownliftNoSupport'))

base <- read.csv(paste0(base.path, 'qalys.csv'))
support <- read.csv(paste0(support.path, 'qalys.csv'))
noSupport <- read.csv(paste0(noSupport.path, 'qalys.csv'))

# out.path <- '/home/luke/Documents/WORK/MINOS/PLOTS_FROM_ARC/QALY_TESTING/'
# 
# base <- read.csv(paste0(out.path, 'qaly_baseline.csv'))
# support <- read.csv(paste0(out.path, 'qalys_support.csv'))
# noSupport <- read.csv(paste0(out.path, 'qalys_noSupport.csv'))
```

# QALY comparison

```{r}
# combine the dataframes
combined <- rbind(base, support, noSupport)

ggplot(data = combined, aes(x = year, y = QALYs, group = intervention, colour = intervention)) +
  geom_smooth()
```

## Groupby and Mean

```{r}
base <- base %>%
  group_by(year) %>%
  summarise(QALYs = mean(QALYs),
            boost_amount = mean(total_boost))
support <- support %>%
  group_by(year) %>%
  summarise(QALYs = mean(QALYs),
            boost_amount = mean(total_boost))
noSupport <- noSupport %>%
  group_by(year) %>%
  summarise(QALYs = mean(QALYs),
            boost_amount = mean(total_boost))
```

## Area Under the Curve Comparison

```{r}
#base.auc <- sum(diff(base$time)*rollmean(base$QALYs, 2))
base.auc <- data.frame(intervention = 'baseline', AUC = auc(x=base$year, y=base$QALYs, type = 'spline'))
support.auc <- data.frame(intervention = 'Support', AUC = auc(x=support$year, y=support$QALYs, type = 'spline'))
noSupport.auc <- data.frame(intervention = 'NoSupport', AUC = auc(x=noSupport$year, y=noSupport$QALYs, type = 'spline'))

combined.auc <- rbind(base.auc, support.auc, noSupport.auc)
```

```{r}
print(combined.auc)

ggplot(data = combined.auc, aes(x = intervention, y = AUC, group = intervention, fill = intervention)) +
  geom_col()
```

```{r}
# Another version showing difference from baseline for better comparison
combined.auc2 <- combined.auc %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('AUC')) %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = Support - baseline,
         change_noSupport = NoSupport - baseline) %>%
  pivot_longer(cols = change_baseline:change_noSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'AUC_difference')

print(combined.auc2)

ggplot(combined.auc2, aes(x = intervention, y = AUC_difference, group = intervention, fill = intervention)) +
  geom_col()
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}

combined.cost <- combined %>%
  select(-X, -pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  mutate(QALY_value = QALYs * 60000)

ggplot(combined.cost, aes(x = year, y = QALY_value, group = intervention, colour = intervention)) +
  geom_smooth()

combined.cost.change <- combined.cost %>%
  select(-total_boost, -QALYs) %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'QALY_value') #%>%
  # mutate(change_Support = energyDownlift - baseline,
  #        change_NoSupport = energyDownliftNoSupport - baseline) %>%
  # select(-baseline, -energyDownlift, -energyDownliftNoSupport) %>%
  # pivot_longer(cols = -year,
  #              names_prefix = 'change_',
  #              names_to = 'intervention',
  #              values_to = 'value_change')

ggplot(combined.cost.change, aes(x = year, y = value_change, group = intervention, colour = intervention)) +
  geom_smooth()

  # select(-QALYs) %>%
  # pivot_wider(names_from = 'intervention',
  #             values_from = c('total_boost', 'QALY_value'),
  #             names_sep = '000') %>%
  # mutate(QALY_value_change000energyDownlift = QALY_value000energyDownlift - QALY_value000baseline,
  #        QALY_value_change000energyDownliftNoSupport = QALY_value000energyDownliftNoSupport - QALY_value000baseline) %>%
  # select(-QALY_value000baseline, -QALY_value000energyDownlift, -QALY_value000energyDownliftNoSupport, -total_boost000baseline) %>%
  # pivot_longer(cols = total_boost000energyDownlift:QALY_value_change000energyDownliftNoSupport,
  #              names_sep = '000',
  #              names_to = c('.value', 'intervention'),
  #              names_repair = 'unique') #%>%
  # pivot_longer(cols = QALY_value_change_Support:QALY_value_change_NoSupport,
  #              names_prefix = 'QALY_value_change_',
  #              names_to = 'intervention2',
  #              values_to = 'QALY_value_change')

# combined.cost <- combined %>%
#   select(-X, -pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
#   pivot_wider(names_from = 'intervention',
#               values_from = c('mean_boost', 'QALYs')) %>%
#   mutate(change_energyDownlift = QALYs_energyDownlift - QALYs_baseline,
#          change_energyDownliftNoSupport = QALYs_energyDownliftNoSupport - QALYs_baseline) %>%
#   mutate(costPerQALY_energyDownlift = change_energyDownlift / mean_boost_energyDownlift,
#          costPerQALY_energyDownliftNoSupport = change_energyDownliftNoSupport / mean_boost_energyDownliftNoSupport) %>%
#   pivot_longer(cols = costPerQALY_energyDownlift:costPerQALY_energyDownliftNoSupport,
#                names_prefix = 'costPerQALY_',
#                names_to = 'intervention',
#                values_to = 'costPerQALY') %>%
#   select(year, intervention, costPerQALY)

# options(scipen=999)
# 
# ggplot(data = combined.cost, aes(x = year, y = costPerQALY, group = intervention, colour = intervention)) +
#   geom_line()
```

```{r}
combined.cost.mean <- combined.cost %>%
  group_by(year, intervention) %>%
  summarise(across(everything(), mean))

total.combined.cost <- combined.cost.change %>%
  group_by(intervention) %>%
  summarise(total_value_change = sum(value_change))

ggplot(data = total.combined.cost, aes(x = intervention, y = total_value_change, group = intervention, fill = intervention)) +
  geom_col()
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).

```{r}
# ICER = (C1 - C0) / (E1 - E0)
# where
# C1 = cost of intervention
# E1 = effect of intervention
# C0 = cost of control group
# E0 = effect of control group

ICER_test2 <- combined.cost.mean %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('total_boost', 'QALYs', 'QALY_value')) %>%
  mutate(ICER_Support = (total_boost_energyDownlift - total_boost_baseline) / (QALYs_energyDownlift - QALYs_baseline),
         ICER_NoSupport = (total_boost_energyDownliftNoSupport - total_boost_baseline) / (QALYs_energyDownliftNoSupport - QALYs_baseline)) %>%
select(year, ICER_Support, ICER_NoSupport) %>%
  pivot_longer(cols = -year,
               names_prefix = 'ICER_',
               names_to = 'intervention',
               values_to = 'ICER')

ggplot(ICER, aes(x = year, y = ICER, group = intervention, colour = intervention)) +
  geom_line()

ICER.final <- ICER %>%
  group_by(intervention) %>%
  summarise(ICER = mean(ICER))

ggplot(ICER.final, aes(x = intervention, y = ICER, group = intervention, fill = intervention, colour = intervention)) +
  geom_col()

ICER_test3 <- combined.cost.mean %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('total_boost', 'QALYs', 'QALY_value')) %>%
  mutate(C_Support = total_boost_energyDownlift - total_boost_baseline,
         E_Support = QALYs_energyDownlift - QALYs_baseline,
         C_NoSupport = total_boost_energyDownliftNoSupport - total_boost_baseline,
         E_NoSupport = QALYs_energyDownliftNoSupport - QALYs_baseline) %>%
  select(year, C_Support, E_Support, C_NoSupport, E_NoSupport) %>%
  mutate(ICER_Support = C_Support / E_Support,
         ICER_NoSupport = C_NoSupport / E_NoSupport)# %>%
  #select(year, ICER_Support, ICER_NoSupport)# %>%
#   pivot_longer(cols = -year,
#                names_prefix = 'ICER_',
#                names_to = 'intervention',
#                values_to = 'ICER')

```

```{r}

```