### 2023-11-01 minos/testing/QALY_comparison_energyCrises.Rmd
Added standard error bars, see how this works before moving to 95% confidence intervals

---
title: "QALY Comparisons - Energy Crisis Interventions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)
require(scales)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  options(scipen = 999)
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
```

## Data

```{r}
# # read in QALY files for baseline and interventions
out.path <- here::here('output', 'default_config')

base.path <- get_latest_runtime_subdirectory(here::here(out.path, 'baseline'))
support.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownlift'))
noSupport.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownliftNoSupport'))

base <- read.csv(paste0(base.path, 'qalys.csv'))
support <- read.csv(paste0(support.path, 'qalys.csv'))
noSupport <- read.csv(paste0(noSupport.path, 'qalys.csv'))

# out.path <- '/home/luke/Documents/WORK/MINOS/PLOTS_FROM_ARC/QALY_TESTING/'
# 
# base <- read.csv(paste0(out.path, 'qaly_baseline.csv'))
# support <- read.csv(paste0(out.path, 'qalys_support.csv'))
# noSupport <- read.csv(paste0(out.path, 'qalys_noSupport.csv'))
```

# QALY comparison

```{r}
# combine the dataframes
combined <- rbind(base, support, noSupport)

ggplot(data = combined, aes(x = year, y = QALYs, group = intervention, colour = intervention)) +
  geom_smooth()
```

## Area Under the Curve Comparison

```{r}
base.auc <- base %>%
  group_by(run_id) %>%
  summarise(AUC = auc(x=year, y = QALYs, type = 'spline')) %>%
  mutate(intervention = 'baseline')

support.auc <- support %>%
  group_by(run_id) %>%
  summarise(AUC = auc(x=year, y = QALYs, type = 'spline')) %>%
  mutate(intervention = 'Support')

noSupport.auc <- noSupport %>%
  group_by(run_id) %>%
  summarise(AUC = auc(x=year, y = QALYs, type = 'spline')) %>%
  mutate(intervention = 'NoSupport')

combined.auc <- rbind(base.auc, support.auc, noSupport.auc)
```

```{r}
print(combined.auc)

combined.auc.plot <- combined.auc %>%
  group_by(intervention) %>%
  summarise(sd = sd(AUC),
            se = sd / sqrt(n()),
            AUC = mean(AUC))

ggplot(data = combined.auc.plot, aes(x = intervention, y = AUC, group = intervention, fill = intervention)) +
  geom_col() +
  geom_errorbar(aes(ymin = AUC - se, ymax = AUC + se, width = 0.4, group = intervention)) +
  labs(title = 'Total AUC comparison', subtitle = 'Baseline vs Energy Crisis Interventions') +
  xlab('Intervention')
```

```{r}
# Another version showing difference from baseline for better comparison
combined.auc2 <- combined.auc %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('AUC')) %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = Support - baseline,
         change_noSupport = NoSupport - baseline) %>%
  pivot_longer(cols = change_baseline:change_noSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'AUC_difference') %>%
  filter(intervention != 'baseline')

print(combined.auc2)

combined.auc2.plot <- combined.auc2 %>%
  group_by(intervention) %>%
  summarise(se = sd(AUC_difference) / sqrt(n()),
            AUC_difference = mean(AUC_difference))

ggplot(combined.auc2.plot, aes(x = intervention, y = AUC_difference, group = intervention, fill = intervention)) +
  geom_col() +
  geom_errorbar(aes(ymin = AUC_difference - se, ymax = AUC_difference + se, width = 0.4, group = intervention)) +
  labs(title = 'Change in AUC', subtitle = 'Baseline vs Energy Crisis Interventions') +
  xlab('Intervention')
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}
combined.cost <- combined %>%
  select(-pop_size, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  mutate(QALY_value = QALYs * 60000)

ggplot(combined.cost, aes(x = year, y = QALY_value, group = intervention, colour = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma)

combined.cost.change <- combined.cost %>%
  select(-total_boost, -QALYs) %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'QALY_value') %>%
  mutate(change_baseline = baseline - baseline,
         change_Support = energyDownlift - baseline,
         change_NoSupport = energyDownliftNoSupport - baseline) %>%
  select(-baseline, -energyDownlift, -energyDownliftNoSupport) %>%
  pivot_longer(cols = change_baseline:change_NoSupport,
               names_to = 'intervention',
               names_prefix = 'change_',
               values_to = 'QALY_value_change') %>%
  filter(intervention != 'baseline')

ggplot(combined.cost.change, aes(x = year, y = QALY_value_change, group = intervention, colour = intervention, fill = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma)
```

```{r}
total.combined.cost <- combined.cost.change %>%
  group_by(run_id, intervention) %>%
  summarise(TotalChange = sum(QALY_value_change)) %>%
  group_by(intervention) %>%
  summarise(se = sd(TotalChange) / sqrt(n()),
            TotalChange = mean(TotalChange))

ggplot(total.combined.cost, aes(x = intervention, y = TotalChange, group = intervention, fill = intervention)) +
  geom_col() +
  geom_errorbar(aes(ymin = TotalChange - se, ymax = TotalChange + se, width = 0.4, group = intervention)) +
  scale_y_continuous(label = comma) +
  labs(title = 'Total Change in QALY Value', subtitle = 'Baseline vs Energy Crisis Interventions') +
  xlab('Intervention')
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).

```{r}
combined.cost.mean <- combined.cost %>%
  group_by(run_id, year, intervention) %>%
  summarise(across(everything(), mean))

# ICER = (C1 - C0) / (E1 - E0)
# where
# C1 = cost of intervention
# E1 = effect of intervention
# C0 = cost of control group
# E0 = effect of control group

ICER <- combined.cost.mean %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('total_boost', 'QALYs', 'QALY_value')) %>%
  mutate(ICER_Support = (total_boost_energyDownlift - total_boost_baseline) / (QALYs_energyDownlift - QALYs_baseline),
         ICER_NoSupport = (total_boost_energyDownliftNoSupport - total_boost_baseline) / (QALYs_energyDownliftNoSupport - QALYs_baseline)) %>%
  select(run_id, year, ICER_Support, ICER_NoSupport) %>%
  pivot_longer(cols = -c(run_id, year),
               names_prefix = 'ICER_',
               names_to = 'intervention',
               values_to = 'ICER')

ggplot(ICER, aes(x = year, y = ICER, group = intervention, fill = intervention, colour = intervention)) +
  geom_smooth()

ICER.final <- ICER %>%
  group_by(run_id, intervention) %>%
  summarise(ICER = mean(ICER)) %>%
  group_by(intervention) %>%
  summarise(se = sd(ICER) / sqrt(n()),
            ICER = mean(ICER))

ggplot(ICER.final, aes(x = intervention, y = ICER, group = intervention, fill = intervention)) +
  geom_col() +
  geom_errorbar(aes(ymin = ICER - se, ymax = ICER + se, width = 0.4, group = intervention)) +
  scale_y_continuous(label = comma) +
  labs(title = 'ICER', subtitle = 'Living Wage Intervention') +
  xlab('Intervention')
```

```{r}

```