### 2023-11-14 minos/testing/QALY_comparison_energyCrises_diff.Rmd
Attempting to fix weird bug where PCS doesnt transition. Think this will fix it. Updated some paths in config/uk_scaled.yaml, updated some functions in minos/utils_qaly.R and some tweaks to the QALY comparison scripts

---
title: "QALY Comparisons - Energy Crisis Interventions - NoSupport Baseline"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)
require(scales)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  options(scipen = 999)
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_qaly.R'))
```

## Data

```{r}
# # read in QALY files for baseline and interventions
out.path <- here::here('output', 'default_config')
#out.path <- here::here('output', 'glasgow_scaled')
#out.path <- here::here('output', 'SF12_uk_scaled')

base.path <- get_latest_runtime_subdirectory(here::here(out.path, 'baseline'))
support.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownlift'))
noSupport.path <- get_latest_runtime_subdirectory(here::here(out.path, 'energyDownliftNoSupport'))

base <- read.csv(paste0(base.path, 'qalys.csv'))
support <- read.csv(paste0(support.path, 'qalys.csv'))
noSupport <- read.csv(paste0(noSupport.path, 'qalys.csv'))

# out.path <- '/home/luke/Documents/WORK/MINOS/PLOTS_FROM_ARC/QALY_TESTING/'
# 
# base <- read.csv(paste0(out.path, 'qalys_baseline.csv'))
# support <- read.csv(paste0(out.path, 'qalys_support.csv'))
# noSupport <- read.csv(paste0(out.path, 'qalys_nosupport.csv'))

# combine the dataframes
combined <- rbind(support, noSupport)
```

# Intervention Information

Just going to plot some things here like how many people affected by an intervention and what the mean / total boost amounts look like.

```{r}
ggplot(combined, aes(x = year, y = pop_boosted, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Number Individuals Receiving Intervention')

combined.diff <- combined %>%
  select(run_id, year, intervention, total_boost) %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'total_boost') %>%
  mutate(diff.boost = energyDownlift - energyDownliftNoSupport)

ggplot(combined.diff, aes(x = year, y = diff.boost)) +
  geom_smooth() +
  labs(title = 'Total Boost amount by year') +
  scale_y_continuous(label = comma)

ggplot(combined, aes(x = year, y = alive_ratio, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Ratio of living to dead in output dataframe')

## incidence of mortality
deaths <- combined %>%
  select(run_id, intervention, year, alive_pop, dead_pop) %>%
  group_by(run_id, intervention) %>%
  mutate(total_pop = alive_pop + dead_pop,
         deaths_this_year = dead_pop - lag(dead_pop)) %>%
  group_by(run_id, year, intervention) %>%
  summarise(death_incidence = deaths_this_year / total_pop)

ggplot(deaths, aes(x = year, y = death_incidence, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Incidence of Mortality Comparison')
```

# QALY comparison

```{r}
QALY_comparison(base = noSupport,
                base.name = 'energyDownliftNoSupport',
                int = support,
                int.name = 'energyDownlift')
```

## SF12 Change

```{r}
sf12.plots(base = noSupport,
           base.name = 'energyDownliftNoSupport',
           int = support,
           int.name = 'energyDownlift')
```

```{r}
# base <- noSupport
# base.name <- 'energyDownliftNoSupport'
# int <- support
# int.name <- 'energyDownlift'
# 
# combined <- rbind(base, int)
# 
# ## now SF12 plots for comparison
# p1 <- ggplot(combined, aes(x = year, y = SF_12_MCS, group = intervention, colour = intervention, fill = intervention)) +
#   geom_smooth()
# p2 <- ggplot(combined, aes(x = year, y = SF_12_PCS, group = intervention, colour = intervention, fill = intervention)) +
#   geom_smooth()
# 
# print(p1)
# print(p2)
# 
# combined.SF12 <- combined %>%
#   select(run_id, year, intervention, SF_12_MCS, SF_12_PCS) %>%
#   pivot_wider(names_from = 'intervention',
#               values_from = c('SF_12_MCS', 'SF_12_PCS')) %>%
#   mutate(MCS_diff = .data[[paste0('SF_12_MCS_', int.name)]] - .data[[paste0('SF_12_MCS_', base.name)]],
#          PCS_diff = .data[[paste0('SF_12_PCS_', int.name)]] - .data[[paste0('SF_12_PCS_', base.name)]]) %>%
#   select(run_id, year, MCS_diff, PCS_diff)
# 
# p3 <- ggplot(combined.SF12, aes(x = year, y = MCS_diff)) +
#   geom_smooth() +
#   geom_hline(yintercept = 0, linetype = 'dashed') +
#   labs(title = 'Change in SF_12_MCS', subtitle = paste0(int.name, ' vs ', base.name)) +
#   xlab('Year') +
#   ylab('Change in MCS')
# 
# p4 <- ggplot(combined.SF12, aes(x = year, y = PCS_diff)) +
#   geom_smooth() +
#   geom_hline(yintercept = 0, linetype = 'dashed') +
#   labs(title = 'Change in SF_12_PCS', subtitle = paste0(int.name, ' vs ', base.name)) +
#   xlab('Year') +
#   ylab('Change in MCS')
# 
# print(p3)
# print(p4)

```

## Area Under the Curve

```{r}
auc.plots(base = noSupport, 
          base.name = 'NoSupport',
          intervention = support,
          int.name = 'Support')
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}
cost.per.qaly(base = noSupport,
              base.name = 'energyDownliftNoSupport',
              int = support,
              int.name = 'energyDownlift',
              QALY_value = 60000,  # QALY value == £60,000
              int.label <- 'EPCG')
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).

```{r}
ICER(base = noSupport,
     base.name = 'energyDownliftNoSupport',
     int = support,
     int.name = 'energyDownlift',
     QALY_value = 60000,
     int.label = 'EPCG')
```

```{r}
#ICER <- function(base, base.name, int, int.name, QALY_value) {

base <- noSupport
base.name <- 'energyDownliftNoSupport'
int <- support
int.name <- 'energyDownlift'
QALY_value <- 60000

combined <- rbind(base, int)

combined.cost.mean <- combined %>%
  #select(-alive_pop, -SF_12_MCS, -SF_12_PCS, -utility) %>%
  select(run_id, year, total_boost, intervention, QALYs) %>%
  mutate(QALY_value = QALYs * QALY_value) %>%
  group_by(run_id, year, intervention) %>%
  summarise(across(everything(), mean))

# ICER = (C1 - C0) / (E1 - E0)
# where
# C1 = cost of intervention
# E1 = effect of intervention
# C0 = cost of control group
# E0 = effect of control group

# prepare var names
c1 <- paste('total_boost', int.name, sep='_')
c0 <- paste('total_boost', base.name, sep='_')
e1 <- paste('QALYs', int.name, sep='_')
e0 <- paste('QALYs', base.name, sep='_')

ICER <- combined.cost.mean %>%
  pivot_wider(names_from = 'intervention',
              values_from = c('total_boost', 'QALYs', 'QALY_value')) %>%
  mutate(ICER = (.data[[c1]] - .data[[c0]]) / (.data[[e1]] - .data[[e0]])) %>%
  select(run_id, year, ICER) %>%
  mutate(intervention = int.name)

pc1 <- quantile(ICER$ICER, .01)
print(pc1)
print(min(ICER$ICER))
pc99 <- quantile(ICER$ICER, .99)
print(pc99)
print(max(ICER$ICER))

ICER <- filter(ICER, ICER < pc99, ICER > pc1)
print(min(ICER$ICER))
print(max(ICER$ICER))

p1 <- ggplot(ICER, aes(x = year, y = ICER, group = intervention, fill = intervention, colour = intervention)) +
  geom_smooth() +
  scale_y_continuous(label = comma) +
  labs(title = 'ICER', subtitle = paste0(int.name, ' vs ', base.name))
print(p1)

ICER.final <- ICER %>%
  group_by(run_id, intervention) %>%
  summarise(ICER = mean(ICER)) %>%
  group_by(intervention) %>%
  summarise(margin = (qt(0.975, df=n() - 1) * sd(ICER)) / sqrt(n()),
            ICER = mean(ICER))

p2 <- ggplot(ICER.final, aes(x = intervention, y = ICER, group = intervention, fill = intervention)) +
  geom_col() +
  geom_errorbar(aes(ymin = ICER - margin, ymax = ICER + margin, width = 0.4, group = intervention)) +
  scale_y_continuous(label = comma) +
  labs(title = 'ICER', subtitle = paste0(int.name, ' vs ', base.name)) +
  xlab('Intervention')
print(p2)

print(paste0('ICER == ', ICER.final$ICER, ' +- ', ICER.final$margin))
```

```{r}

```