### 2023-11-14 minos/testing/QALY_comparison_livingWage.Rmd

---
title: "QALY Comparison - Living Wage Intervention"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)
require(scales)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  options(scipen=999)
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_qaly.R'))
```

## Data

```{r}
# # read in QALY files for baseline and interventions
#out.path <- here::here('output', 'default_config')
#out.path <- here::here('output', 'glasgow_scaled')
out.path <- here::here('output', 'SF12_uk_scaled')

base.path <- get_latest_runtime_subdirectory(here::here(out.path, 'baseline'))
wage.path <- get_latest_runtime_subdirectory(here::here(out.path, 'livingWageIntervention'))

base <- read.csv(paste0(base.path, 'qalys.csv'))
wage <- read.csv(paste0(wage.path, 'qalys.csv'))

# out.path <- '/home/luke/Documents/WORK/MINOS/PLOTS_FROM_ARC/QALY_TESTING/'
# 
# base <- read.csv(paste0(out.path, 'qalys_baseline.csv'))
# wage <- read.csv(paste0(out.path, 'qalys_livwage.csv'))

# combine the dataframes
combined <- rbind(base, wage)
```

```{r}
diff <- combined %>%
  pivot_longer(cols = -c(run_id, year, intervention, total_pop_size, alive_pop, dead_pop),
               names_to = 'measure') %>%
  group_by(run_id, year, measure) %>%
  summarise(diff = value[intervention == 'livingWageIntervention'] - value[intervention == 'baseline']) %>%
  pivot_wider(names_from = 'measure',
              values_from = 'diff')

wage.short <- wage %>%
  select(run_id, year, total_pop_size, alive_pop, dead_pop)

combined.diff <- merge(wage.short, diff, by=c('run_id', 'year'))

qalys_3sd <- 3 * sd(combined.diff$QALYs)
combined.culled <- filter(combined.diff, 
                          QALYs < QALYs + qalys_3sd,
                          QALYs > QALYs - qalys_3sd)

  # pivot_wider(id_cols = c(run_id, year),
  #             names_from = 'intervention',
  #             values_from = -c(run_id, year)) %>%
  # mutate(across(ends_with('_livingWageIntervention'), .names = "{.col}_diff") - across(ends_with('_baseline')))
```

# Intervention Information

Just going to plot some things here like how many people affected by an intervention and what the mean / total boost amounts look like.

```{r}

test <- combined %>%
  filter(intervention != 'baseline')

ggplot(test, aes(x = year, y = pop_boosted, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Number Individuals Receiving Intervention')

ggplot(filter(combined, intervention != 'baseline'), aes(x = year, y = total_boost, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Total Boost amount by year')

# ggplot(combined, aes(x = year, y = alive_ratio, group = intervention, colour = intervention)) +
#   geom_smooth() +
#   labs(title = 'Ratio of living to dead in output dataframe')

## incidence of mortality
deaths <- combined %>%
  select(run_id, intervention, year, alive_pop, dead_pop) %>%
  group_by(run_id, intervention) %>%
  mutate(total_pop = alive_pop + dead_pop,
         deaths_this_year = dead_pop - lag(dead_pop)) %>%
  group_by(run_id, year, intervention) %>%
  summarise(death_incidence = deaths_this_year / total_pop)

ggplot(deaths, aes(x = year, y = death_incidence, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Incidence of Mortality Comparison')
```

# QALY comparison

```{r}
QALY_comparison(base = base,
                base.name = 'baseline',
                int = wage,
                int.name = 'livingWageIntervention')
```

## SF12 Change

```{r}
sf12.plots(base = base,
           base.name = 'baseline',
           int = wage,
           int.name = 'livingWageIntervention')
```

## Area Under the Curve Comparison

```{r}
auc.plots(base = base, 
          base.name = 'baseline',
          intervention = wage,
          int.name = 'LivingWage')
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}
cost.per.qaly(base = base,
              base.name = 'baseline',
              int = wage,
              int.name = 'livingWageIntervention',
              QALY_value = 60000,
              int.label <- 'LivingWage')  # QALY value == £60,000
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).There is also a hard limit of £100,000, where any intervention with an ICER above this value is never acceptable.

```{r}
ICER(base = base,
     base.name = 'baseline',
     int = wage,
     int.name = 'livingWageIntervention',
     QALY_value = 60000,
     int.label = 'LivingWage')
```

# Total Cost of Intervention

```{r}
intervention.cost(base = baseline,
                  base.name = 'baseline',
                  int = livWage,
                  int.name = 'livingWageIntervention',
                  int.label = 'Living Wage')
```

```{r}

```