### 2023-11-17 minos/validation/cross_validation_default.Rmd
a lot of bug fixing and inclusion in (cross validated) handovers

---
title: "Cross-Validation Default"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(ggExtra)
require(here)
require(scales)
require(dplyr)
require(tidyr)
require(purrr)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
#source(here::here('minos', 'utils_outcome_vis.R'))
```

## Data

```{r, include=FALSE}
# cross validation simulation output
#out.path <- 'Minos/output/cross_validation/'
out.path <- here::here('output', 'cv')
cv1 <- read_singular_local_out(here::here(out.path, 'default1/'), 'baseline', drop.dead = TRUE)
cv2 <- read_singular_local_out(here::here(out.path, 'default2/'), 'baseline', drop.dead = TRUE)
cv3 <- read_singular_local_out(here::here(out.path, 'default3/'), 'baseline', drop.dead = TRUE)
cv4 <- read_singular_local_out(here::here(out.path, 'default4/'), 'baseline', drop.dead = TRUE)
cv5 <- read_singular_local_out(here::here(out.path, 'default5/'), 'baseline', drop.dead = TRUE)

cvlist <- list(cv1, cv2, cv3, cv4, cv5)
cv <- do.call(rbind, cvlist)

# cv simulation batches
cv.simul.files1 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch1'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw1 <- do.call(rbind, lapply(cv.simul.files1, read.csv))
cv.simul.files2 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch2'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw2 <- do.call(rbind, lapply(cv.simul.files2, read.csv))
cv.simul.files3 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch3'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw3 <- do.call(rbind, lapply(cv.simul.files3, read.csv))
cv.simul.files4 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch4'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw4 <- do.call(rbind, lapply(cv.simul.files4, read.csv))
cv.simul.files5 <- list.files(here::here('data', 'final_US', 'cross_validation', 'batch5'),
                             pattern = '[0-9]{4}_US_cohort.csv',
                             full.names = TRUE)
raw5 <- do.call(rbind, lapply(cv.simul.files5, read.csv))

rawlist <- list(raw1, raw2, raw3, raw4, raw5)
raw.dat <- do.call(rbind, rawlist)
```

First things first, we need to make sure that we are comparing the exact same people from each dataset. Therefore, we need a step here to check the pidps from both, and remove anybody not in both datasets. We can also filter the years out of the raw data here, as we only care about years that have been simulated for comparison.

```{r, include=FALSE}
# pidp filter
#cv.pidps <- cv %>% select(pidp) %>% unique()
#raw.pidps <- raw %>% select(pidp) %>% unique()
#both.pidps <- intersect(cv.pidps, raw.pidps)

# year filter
cv.years <- cv %>% select(time) %>% unique()
cv.years <- cv.years[['time']]

#cv <- cv.dat %>% filter(pidp %in% both.pidps$pidp)
# getting only people who were in 2014 population. 
#who_2014_pidps <- raw.dat %>% filter(time %in% c(2014) & hh_income<10000) %>% select(pidp) %>% unique()
raw <- raw.dat %>% filter(time %in% cv.years)
#raw <- raw %>% filter (pidp %in% who_2014_pidps$pidp)

#rm(cv.pidps, raw.pidps, both.pidps, cv.years)
rm(cv.years)
```

For all analyses, results will be calculated on each output separately and pooled. For example, the mean hh_income over time will be calculated for each output run seperately, and the mean of these 5 means is what will be plotted. This will be plotted against the entire raw sample.

# hh_income

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'hh_income')
multi_year_boxplots(raw, cv, 'hh_income')
snapshot_OP_plots(raw, cv, 'hh_income', target.years = c(2014, 2016, 2018, 2020))
q_q_comparison(raw, cv, 'hh_income')
```
```{r}
inc.14 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2014)
print(inc.14)
```
```{r}
inc.16 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2016)
print(inc.16)
```
```{r}
inc.18 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2018)
print(inc.18)
```
```{r}
inc.20 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'hh_income', 
                                 target.year = 2020)
print(inc.20)
rm(inc.14, inc.16, inc.18, inc.20)
```

# SF_12

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'SF_12')
multi_year_boxplots(raw, cv, 'SF_12')
snapshot_OP_plots(raw, cv, 'SF_12', target.years = c(2014, 2016, 2018, 2020))
q_q_comparison(raw, cv, 'SF_12')
```
```{r}
sf12.14 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2014)
print(sf12.14)
```
```{r}
sf12.16 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2016)
print(sf12.16)
```
```{r}
sf12.18 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2018)
print(sf12.18)
```
```{r}
sf12.20 <- marg_dist_densigram_plot_oneyear(observed = raw, 
                                 predicted = cv, 
                                 var = 'SF_12', 
                                 target.year = 2020)
print(sf12.20)
rm(sf12.14, sf12.16, sf12.18, sf12.20)
```

# Housing Quality

```{r}
hous.pivoted <- combine_and_pivot_long(df1 = cv, 
                                       df1.name = 'simulated', 
                                       df2 = raw, 
                                       df2.name = 'raw', 
                                       var = 'housing_quality')

cv_ordinal_plots(pivoted.df = hous.pivoted, 
                 var = 'housing_quality',
                 save = FALSE)
rm(hous.pivoted)
```

# Neighbourhood Safety

```{r}
nhsafe.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'neighbourhood_safety')

cv_ordinal_plots(pivoted.df = nhsafe.pivoted, 
                 var = 'neighbourhood_safety',
                 save = FALSE)
rm(nhsafe.pivoted)
```

# Loneliness

```{r}
lnly.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'loneliness')

cv_ordinal_plots(pivoted.df = lnly.pivoted, 
                 var = 'loneliness',
                 save = FALSE)
rm(lnly.pivoted)
```

# Nutrition Quality

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'nutrition_quality')
multi_year_boxplots(raw, cv, 'nutrition_quality')
q_q_comparison(raw, cv, 'nutrition_quality')
```

# Tobacco Use

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'ncigs')
multi_year_boxplots(raw, cv, 'ncigs')
q_q_comparison(raw, cv, 'ncigs')
```

# Financial Situation

```{r}
fsit.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'financial_situation')

cv_ordinal_plots(pivoted.df = fsit.pivoted, 
                 var = 'financial_situation',
                 save = FALSE)
rm(fsit.pivoted)
```

# Heating

```{r}
heat.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'heating')

cv_ordinal_plots(pivoted.df = heat.pivoted, 
                 var = 'heating',
                 save = FALSE)
rm(heat.pivoted)
```

# SECONDARY VARS

## Marital Status

```{r}
mstat.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'marital_status')

cv_ordinal_plots(pivoted.df = mstat.pivoted, 
                 var = 'marital_status',
                 save = FALSE)
rm(mstat.pivoted)
```

## hhsize

```{r}
cv.mean.plots(cv1, cv2, cv3, cv4, cv5, raw, 'hhsize')
multi_year_boxplots(raw, cv, 'hhsize')
q_q_comparison(raw, cv, 'hhsize')
```

## housing tenure

```{r}
tenure.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'housing_tenure')

cv_ordinal_plots(pivoted.df = tenure.pivoted, 
                 var = 'housing_tenure',
                 save = FALSE)
rm(tenure.pivoted)
```

## Ethnicity

```{r}
ethnic.pivoted <- combine_and_pivot_long(df1 = cv, 
                                          df1.name = 'simulated', 
                                          df2 = raw, 
                                          df2.name = 'raw', 
                                          var = 'ethnicity')

cv_ordinal_plots(pivoted.df = ethnic.pivoted, 
                 var = 'ethnicity',
                 save = FALSE)

ethnic.piv.nowhite <- filter(ethnic.pivoted, ethnicity != 'WBI')

cv_ordinal_plots(pivoted.df = ethnic.piv.nowhite, 
                 var = 'ethnicity',
                 save = FALSE)

rm(ethnic.pivoted, ethnic.piv.nowhite)
```

## Parity

```{r}
cv.parity <- cv %>%
  filter(sex == "Female")

cv1.parity <- cv1 %>%
  filter(sex == "Female")

cv2.parity <- cv2 %>%
  filter(sex == "Female")

cv3.parity <- cv3 %>%
  filter(sex == "Female")

cv4.parity <- cv4 %>%
  filter(sex == "Female")

cv5.parity <- cv5 %>%
  filter(sex == "Female")

raw.parity <- raw %>%
  filter(sex == "Female")

cv.mean.plots(cv1.parity, cv2.parity, cv3.parity, cv4.parity, cv5.parity, raw.parity, 'nkids_ind')
multi_year_boxplots(raw.parity, cv.parity, 'nkids_ind')
q_q_comparison(raw.parity, cv.parity, 'nkids_ind')
```
```{r}
parity.pivoted <- combine_and_pivot_long(df1 = cv.parity,
                                         df1.name = 'simulated',
                                         df2 = raw.parity,
                                         df2.name = 'raw',
                                         var = 'nkids_ind')

cv_ordinal_plots(pivoted.df = parity.pivoted, 
                 var = 'nkids_ind',
                 save = FALSE)

rm(parity.pivoted)
```

## Children Ageing

Children ageing is now calculated in MINOS.

```{r}

format_child_age_data <- function(dataset) {
  formatted.age.data <- dataset %>%
                        dplyr::select(hidp, time, child_ages, pidp) %>%
                        filter(time >= 2014) %>%
                        group_by(pidp, time) %>%
                        summarise(child_ages = max(child_ages))

  split.ages <- strsplit(formatted.age.data$child_ages, "_")

  child.ages.expanded <- data.frame(time = rep(formatted.age.data$time, sapply(split.ages, length)), 
                                    #hidp= rep(formatted.age.data$hidp, sapply(split.ages, length)), 
                                    pidp= rep(formatted.age.data$pidp, sapply(split.ages, length)), 
                                    expanded_child_ages = as.numeric(unlist(split.ages)))

  #child.ages.expanded <- child.ages.expanded %>%
  #  group_by(pidp, time) %>%
  #  summarise(expanded_child_ages = median(expanded_child_ages))

  return (child.ages.expanded)
}

find_mode <- function(x) {
  u <- unique(x)
  tab <- tabulate(match(x, u))
  u[tab == max(tab)]
}

cv.child.ages <- format_child_age_data(cv)
cv1.child.ages <- format_child_age_data(cv1)
cv2.child.ages <- format_child_age_data(cv2)
cv3.child.ages <- format_child_age_data(cv3)
cv4.child.ages <- format_child_age_data(cv4)
cv5.child.ages <- format_child_age_data(cv5)
raw.child.ages <- format_child_age_data(raw.dat)

cv.mean.plots(cv1.child.ages, cv2.child.ages, cv3.child.ages, cv4.child.ages, cv5.child.ages, raw.child.ages, 'expanded_child_ages')
multi_year_boxplots(raw.child.ages, cv.child.ages, 'expanded_child_ages')
q_q_comparison(raw.child.ages, cv.child.ages, 'expanded_child_ages')

```

```{r} 
#nkids

format_nkids_data <- function(dataset) {
  dataset %>%
  group_by(time, hidp) %>%
  summarise(nkids = max(nkids)) %>%
  #group_by(time) %>%
  #summarise(nkids_by_year = sum(nkids_counts))
  return (dataset)
}

cv.nkids <- format_nkids_data(cv)
cv1.nkids <- format_nkids_data(cv1)
cv2.nkids <- format_nkids_data(cv2)
cv3.nkids <- format_nkids_data(cv3)
cv4.nkids <- format_nkids_data(cv4)
cv5.nkids <- format_nkids_data(cv5)
raw.nkids <- format_nkids_data(raw.dat)

cv.mean.plots(cv1.nkids, cv2.nkids, cv3.nkids, cv4.nkids, cv5.nkids, raw.nkids, 'nkids')
multi_year_boxplots(raw.nkids, cv.nkids, 'nkids')
q_q_comparison(raw.nkids, cv.nkids, 'nkids')
```