### 2023-11-29 minos/transitions/paper_1_notebook.Rmd

---
title: "Paper 1 Notebook"
output: html_notebook
---

```{r "preamble"}
source("minos/transitions/utils.R")
source("minos/transitions/transition_model_functions.R")
library(dplyr)
```

```{r "OLS"}

model_formula_string <- "SF_12 ~ factor(sex) + 
                  relevel(factor(data$ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state_raw) + 
                  relevel(factor(data$job_sec), ref=1) +
                  relevel(factor(data$region), ref='London') +
                  scale(hh_netinc)"
model_formula <- as.formula(model_formula_string)

dataDir <- "data/raw_US/"
filelist <- list.files(dataDir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[21:23] # get files for 2011-2013.
raw.data <- do.call(rbind, lapply(filelist, read.csv))

raw.data <- raw.data[, append(all.vars(model_formula), c("time", 'pidp', 'weight'))]

raw.data <- raw.data %>%
        group_by(pidp) %>%
        #mutate(diff = .data[[dependent]] - lag(.data[[dependent]], order_by = time)) %>%
        mutate(last = lag(.data[["SF_12"]], order_by = time)) %>%
        rename_with(.fn = ~paste0('SF_12_', .), .cols = last)  # add the dependent as prefix to the calculated diff

```