### 2023-12-07 minos/validation/scot_input_pop_validation.Rmd
forgot a couple of things

---
title: "Scotpop validation notebook"
output:
  html_document:
    df_print: paged
---

This notebook is for validation of the scottish population for child poverty interventions. It contains descriptive statistics for the population and validated against provided child poverty data spatially and nationally.

```{r}
library(tidyr)
library(dplyr)
library(sf)
library(sp)
library(tibble) # converting spatial objects to data frames.
library(broom)
library(here)
library(ggplot2)

source( here::here("minos", "outcomes", "minos_SF12_maps.R"))
```

```{r}
scotland.data <- read.csv("data/scaled_scotland_US/2020_US_cohort.csv")
#scotland.data <- drop_na(scotland.data)
uk.geojson <- st_read("persistent_data/spatial_data/UK_super_outputs.geojson")
uk.geojson <- geojson_to_tibble(uk.geojson)

scotland.data.zones <- read.csv("persistent_data/spatial_data/scotland_data_zones.csv")
glasgow.data.zones <- read.csv("persistent_data/spatial_data/glasgow_data_zones.csv")
edinburgh.data.zones <- read.csv("persistent_data/spatial_data/edinburgh_data_zones.csv")

scotland.geojson <- uk.geojson[which(uk.geojson$ZoneID %in% (scotland.data.zones$LSOA11CD)), ]
glasgow.geojson <- uk.geojson[which(uk.geojson$ZoneID %in% (glasgow.data.zones)$LSOA11CD), ]
edinburgh.geojson <- uk.geojson[which(uk.geojson$ZoneID %in% (edinburgh.data.zones)$LSOA11CD), ]

scotland.LADs.geojson <-  st_read("persistent_data/spatial_data/scotland_LADs.json")
scotland.LADs.geojson <- geojson_to_tibble(scotland.LADs.geojson)
scotland.LADs.geojson[6, "LAD13NM"] <- "Na h-Eileanan Siar" # conflicting name. Think this has been updated since 2013?

```

Description of SF12 MCS

```{r}
hist(scotland.data$SF_12, breaks=100)
print(summary(scotland.data$SF_12))
print(sd(scotland.data$SF_12))
```

```{r}
#https://www.gov.scot/publications/tackling-child-poverty-delivery-plan-forecasting-child-poverty-scotland/pages/4/
glasgow.poverty.stats <- read.csv("persistent_data/spatial_data/Child-Poverty-Glasgow-City-2021-22.csv")
#glasgow.poverty.stats$X..of.children.in.poverty <- glasgow.poverty.stats$X..of.children.in.poverty %% 100
# getting households in relative poverty
scotland.relative_poverty <- scotland.data %>%
                                filter(hh_income < 0.7*median(hh_income))

glasgow.relative_poverty.map <- scotland.relative_poverty %>%
                                  filter(ZoneID %in% (glasgow.data.zones)) %>%
                                  group_by(ZoneID) %>%
                                  summarize(sf12_mean = mean(SF_12, na.rm=T))

glasgow.relative_poverty.map.data <- merge(glasgow.geojson, glasgow.relative_poverty.map, by ='ZoneID')

glasgow.relative.poverty.map <- ggplot(data = glasgow.relative_poverty.map.data , aes(geometry = geometry, fill=.data[["sf12_mean"]])) +
  #geom_sf() + # black outlines on borders for clarity.
  geom_sf(color=NA) + # black outlines on borders for clarity.
  scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
  labs(fill="sf12_mean")  + # colour bar label.
  theme(aspect.ratio=9/16) + # change plot to widescreen dimensions.
  xlab("Longitude") + # axis labels
  ylab("Latitude")
plot(glasgow.relative.poverty.map)
```

```{r}

glasgow.percentage.relative.poverty.data <- scotland.data %>%
                                  filter(ZoneID %in% (glasgow.data.zones)) %>%
                                  mutate(in_relative_poverty = hh_income < 0.75*median(hh_income)) %>%
                                  group_by(hidp) %>%
                                  summarize(ZoneID= max(ZoneID), nkids = max(nkids), in_relative_poverty = max(in_relative_poverty)) %>%
                                  group_by(ZoneID) %>%
                                  summarize(relative.poverty.freq = 100 * sum(nkids*in_relative_poverty)/sum(nkids))

glasgow.percentage.relative.poverty.map.data <- merge(glasgow.geojson, glasgow.percentage.relative.poverty.data, by ='ZoneID')

glasgow.percentage.relative.poverty.map <- ggplot(data = glasgow.percentage.relative.poverty.map.data  , aes(geometry = geometry, fill=.data[["relative.poverty.freq"]])) +
  #geom_sf() + # black outlines on borders for clarity.
  geom_sf(color=NA) + # black outlines on borders for clarity.
  scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
  labs(fill="Percentage Poverty")  + # colour bar label.
  theme(aspect.ratio=9/16) + # change plot to widescreen dimensions.
  xlab("Longitude") + # axis labels
  ylab("Latitude")
plot(glasgow.percentage.relative.poverty.map)

glasgow.percentage.relative.poverty.diff <- merge(glasgow.percentage.relative.poverty.map.data, glasgow.poverty.stats, by.x = "ZoneID", by.y="DataZone")
glasgow.percentage.relative.poverty.diff$diff <- glasgow.percentage.relative.poverty.diff$X..of.children.in.poverty - glasgow.percentage.relative.poverty.diff$relative.poverty.freq

# bluer indicates the synth pop underestimates percentage poverty. orange indicates overestimation. 
minos_diff_map(glasgow.percentage.relative.poverty.diff, "plots/poverty_diff_map.pdf", "PP", F)
print("MAE:")
print(mean(abs(glasgow.percentage.relative.poverty.diff$diff)))
hist(abs(glasgow.percentage.relative.poverty.diff$diff))
```

```{r}
#https://www.healthscotland.scot/media/2607/child-poverty-scales-and-trends.pdf
scotland.lad.poverty.data <- scotland.data[, c("hh_income", "nkids", "hidp", "ZoneID")]
scotland.lad.poverty.data <- merge(scotland.lad.poverty.data, scotland.geojson, by = "ZoneID")
scotland.lad.poverty.data <- scotland.lad.poverty.data[, c("hh_income", "nkids", "hidp", "LAName")]

scotland.lad.percentage.relative.poverty.data <- scotland.lad.poverty.data %>%
                                  mutate(in_relative_poverty = hh_income < 0.6*median(hh_income)) %>%
                                  group_by(hidp) %>%
                                  summarize(LAName= max(LAName), nkids = max(nkids), in_relative_poverty = max(in_relative_poverty)) %>%
                                  group_by(LAName) %>%
                                  summarize(relative.poverty.freq = 100 * sum(nkids*in_relative_poverty)/sum(nkids))

scotland.lad.percentage.relative.poverty.map.data  <- merge(scotland.lad.percentage.relative.poverty.data, scotland.LADs.geojson, by.x = "LAName", by.y="LAD13NM")

scotland.lad.percentage.relative.poverty.map <- ggplot(data = scotland.lad.percentage.relative.poverty.map.data  , aes(geometry = geometry, fill=.data[["relative.poverty.freq"]])) +
  #geom_sf() + # black outlines on borders for clarity.
  geom_sf(color=NA) + # black outlines on borders for clarity.
  scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
  labs(fill="Percentage Poverty")  + # colour bar label.
  xlab("Longitude") + # axis labels
  ylab("Latitude")
plot(scotland.lad.percentage.relative.poverty.map)
print(mean(scotland.lad.percentage.relative.poverty.map.data$relative.poverty.freq))

# download from here https://www.jrf.org.uk/data/uk-child-poverty-rates-local-authority.
rowntree.percentage.child.poverty <- readxl::read_xlsx("persistent_data/spatial_data/6_chld_pov_rate_by_la.xlsx")
rowntree.percentage.child.poverty <- rowntree.percentage.child.poverty[which(rowntree.percentage.child.poverty$`Title:` %in% scotland.LADs.geojson$LAD13NM), ]
rowntree.percentage.child.poverty$percentage_poverty <- as.numeric(rowntree.percentage.child.poverty$`Child poverty rate estimate by local authority`) * 100

scotland.lad.percentage.relative.poverty.diff <- merge(scotland.lad.percentage.relative.poverty.map.data, rowntree.percentage.child.poverty, by.x = "LAName", by.y="Title:")
scotland.lad.percentage.relative.poverty.diff $diff <- scotland.lad.percentage.relative.poverty.diff$percentage_poverty - scotland.lad.percentage.relative.poverty.diff$relative.poverty.freq

# bluer indicates the synth pop underestimates percentage poverty. orange indicates overestimation. 
minos_diff_map(scotland.lad.percentage.relative.poverty.diff, "plots/poverty_diff_map.pdf", "PP", F)
print("MAE:")
print(mean(abs(scotland.lad.percentage.relative.poverty.diff$diff)))
hist(abs(scotland.lad.percentage.relative.poverty.diff$diff), breaks=100)

```

Description of Household Income

```{r}

```

Description of Household (Absolute/Relative) Poverty.

```{r}

```

Description of Material Deprivation and Low Income

```{r}

```

Description of Persistent poverty? Not sure this is feasible with one year of synth data.

```{r}

```