### 2023-12-11 minos/validation/replenishing.Rmd

---
title: "Assessing Replenishing Population"
output: html_notebook
---

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(purrr)

workingDir <- "/home/luke/Documents/WORK/MINOS/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

Source utils for some functions.
```{r}
source('Minos/minos/validation/utils.r')
```

```{r}
# Replenishing data
repl <- read.csv('Minos/data/replenishing/replenishing_pop_2015-2070.csv')

# scotland replenishing
#scot.repl <- read.csv('Minos/data/replenishing/scotland/replenishing_pop_2015-2070.csv')
```

## EDUCATION

```{r}
# plot counts of each state over time (will only change with the replenishing but good to see)
educ.state <- repl %>% 
  group_by(time, education_state) %>%
  count()

# set column types
educ.state$education_state <- as.factor(educ.state$education_state)

ggplot(data = educ.state, mapping = aes(x = education_state, y = n, fill=education_state)) +
  geom_col(position='dodge')
```

## ETHNICITY

```{r}
# plot counts of each state over time (will only change with the replenishing but good to see)
ethnic.state <- repl %>% 
  group_by(time, ethnicity) %>%
  count()

# set column types
ethnic.state$ethnicity <- as.factor(ethnic.state$ethnicity)

ggplot(data = ethnic.state, mapping = aes(x = ethnicity, y = n, fill=ethnicity)) +
  geom_col(position='dodge')

ggplot(data = filter(ethnic.state, ethnicity != 'WBI'), mapping = aes(x = ethnicity, y = n, fill=ethnicity)) +
  geom_col(position='dodge')
```