### 2024-01-06 minos/testing/clustering_EI.Rmd
Clustering script and tweaks to some utils

---
params:
    out_path: 'SIPHER7_batch/' # default arguments. specify others in command line.
    intervention: 'livingWageIntervention'
    type: 'batch'
title: "`r paste(params$file_path, ' ', params$intervention, ' ', params$type, ' CLUSTER ANALYSIS')`"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(cluster)
require(foreach)
require(doParallel)
require(clustMixType)
require(ggplot2)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
)
rm(workingDir)
```

Source utils for some functions.

```{r}
source(here::here('minos', 'transitions', 'utils.R'))
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
source(here::here('minos', 'utils_S7.R'))
```

## Data

```{r}
## Parameters
out.path.batch <- here::here('output', params$out_path)
intervention <- params$intervention
out.type <- params$type

# base.dat <- read_batch_out_all_years(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=FALSE)
# wage.dat <- read_batch_out_all_years(out.path.batch, 'livingWageIntervention', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)
# energy.dat <- read_batch_out_all_years(out.path.batch, 'energyDownlift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)
# child.dat <- read_batch_out_all_years(out.path.batch, 'hhIncomeChildUplift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)
# pov.dat <- read_batch_out_all_years(out.path.batch, 'hhIncomePovertyLineChildUplift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)

if (out.type == 'batch') {
  S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
                 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
                 'ethnicity', 'age', 'sex', 'region', 'job_sec', 'education_state', 'nkids_ind', 'hourly_wage')

  base.dat <- read_batch_out_summarise(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=FALSE)
  int.dat <- read_batch_out_summarise(out.path.batch, paste0(intervention, '/'), start.year = 2021, var.list = c('income_boosted', 'boost_amount', S7.var.list))
} else if (out.type == 'single') {
  base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
  int.dat <- read_singular_local_out(out.path, intervention, drop.dead = TRUE)
} else {
  print('You have supplied an incorrect output type. The options are batch or
        single.')
}
```

```{r}

large_df <- base.dat

# Function to identify variable types
variable_type <- function(x) {
  if (is.numeric(x) && !is.integer(x)) {
    return("continuous")  # float values
  } else if (is.numeric(x) && is.integer(x) && length(unique(x)) > 15) {
    return("continuous")  # int but lots of unique values
  } else if (is.numeric(x) && is.integer(x) && length(unique(x)) < 15) {
    return("ordinal")
  } else if (is.character(x)) {
    return("nominal")
  } else {
    return("other")
  }
}

# Now identify columns and calculate summary values
# Define a list of columns to group by
group_cols <- c("pidp", "hidp", "time")

# Identify variable types
var_types <- sapply(large_df, variable_type)

# Separate variables by type
continuous_vars <- names(var_types[var_types == "continuous"])
ordinal_vars <- names(var_types[var_types == "ordinal"])
nominal_vars <- names(var_types[var_types == "nominal"])

# Calculate mean for continuous variables and mode for ordinal and nominal variables
result_df <- large_df %>%
  group_by(across(all_of(group_cols))) %>%
  summarize(
    across(all_of(setdiff(continuous_vars, group_cols)), mean, na.rm = TRUE),
    across(all_of(ordinal_vars), ~ mode(.x)),
    across(all_of(nominal_vars), ~ mode(.x))
  )
```

Some data preparation.

```{r}
calc_age_group <- function(batch.dat) {
  batch.dat <- batch.dat %>%
    mutate(age.group = cut(age, 
                            breaks = c(15, 30, 45, 60, 75, 120),
                            labels = c('16-30', '30-45', '45-60', '60-75', '75+')))
  return(batch.dat)
}

calc_income_quintile <- function(batch.dat) {
  batch.dat <- batch.dat %>%
    mutate(income.quint = cut(batch.dat$hh_income, breaks=c(quantile(batch.dat$hh_income, probs = seq(0, 1, by = 0.20))), 
                          labels=c(1,2,3,4,5), include.lowest=TRUE))
}

base.dat <- calc_age_group(base.dat)
int.dat <- calc_age_group(int.dat)

base.dat <- calc_income_quintile(base.dat)
int.dat <- calc_income_quintile(int.dat)
```

# CLUSTERING

Going to start with the k-prototypes algorithm as this is a clustering algorithm that can handle ordinal and continuous input variables.

Also need to be clear about the structure of this clustering. I am going to cluster individuals in a single year (start wit 2030) based on the change in each component variable (i.e. S7_mental_health) and the change in Equivalent Income. I can then investigate the clusters and look at things like:
- age range
- other demographics characteristics (sex, ethnicity, region, education_state)
- Change in components (is one group dominated by people who improved housing_quality for example?)
- Change in Equivalent Income

```{r}
# # first add scenario column to each
# base <- base.dat %>%
#   mutate(scenario = 'baseline',
#          boost_amount = 0,
#          income_boosted = 0)
# int <- int.dat %>%
#   mutate(scenario = 'LivingWage')
```

```{r}
# so need a function to generate a dataframe with all those additional metrics for the clustering

calculate_clustering_metrics <- function(base.dat, int.dat, int.name, target.year) {

  # first add scenario column to each
  base.dat <- base.dat %>%
    mutate(scenario = 'baseline',
           boost_amount = 0,
           income_boosted = 0)
  int.dat <- int.dat %>%
    mutate(scenario = int.name)

  miss.cols <- setdiff(names(int.dat), names(base.dat))
  print('Missing Columns!')
  print(miss.cols)

  # now bind together before subsetting
  combined <- rbind(base.dat, int.dat)

  # drop the original datafiles now for better memory handling with batch
  rm(base.dat, int.dat)

  combined <- combined %>%
    select(pidp, hidp, time, scenario, age, age.group, sex, ethnicity, region, education_state, hh_income, equivalent_income, S7_housing_quality, S7_neighbourhood_safety, S7_physical_health, S7_mental_health, S7_labour_state, loneliness, job_sec, hourly_wage)

  combined$S7_housing_quality <- as.numeric(factor(combined$S7_housing_quality,
                                          levels = c('No to all',
                                                     'Yes to some',
                                                     'Yes to all'),
                                          labels = c(1, 2, 3)))

  combined$S7_neighbourhood_safety <- as.numeric(factor(combined$S7_neighbourhood_safety,
                                               levels = c('Often',
                                                          'Some of the time',
                                                          'Hardly ever'),
                                               labels = c(1, 2, 3)))

  combined$loneliness <- as.numeric(factor(combined$loneliness,
                                  levels = c(1, 2, 3),
                                  labels = c(3, 2, 1)))

  # Now calculate metrics. We need:
  # - change in Equivalent Income
  # - change in components
  #
  # This is all change from start.year (usually 2021) to target.year (start with 2031)
  combined <- combined %>%
    filter(time %in% c(target.year)) %>%
    pivot_wider(
      names_from = 'scenario',
      values_from = c('age', 'age.group', 'sex', 'ethnicity', 'region', 'education_state', 'hh_income', 'equivalent_income', 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness', 'job_sec', 'hourly_wage')
    )

  return(combined)

  combined <- combined[complete.cases(combined), ]
  #return(combined)

  ############# Now calculate the difference between intervention and baseline in target.year ###########
  # List of variables to ignore (no difference calculation)
  ignore_vars <- c('pidp', 'hidp', 'time')
  rename_vars <- c("age", "sex", "ethnicity", "region", "age.group", "S7_labour_state", "education_state")

  for (var in rename_vars) {
    combined <- combined %>%
      select(-contains(paste0(var, '_LivingWage'))) %>%
      rename(!!var := .data[[paste0(var, '_baseline')]])
  }

  # Get the names of all variables in the 'combined' dataframe
  all_vars <- names(combined)

  # Find the variables that are in 'all_vars' but not in 'ignore_vars' or 'rename_vars'
  remaining_vars <- setdiff(all_vars, c(ignore_vars, rename_vars))
  # drop the scenarios suffixes
  remaining_vars <- lapply(remaining_vars, function(x) modified_string <- sub("_[^_]*$", "", x))
  remaining_vars <- unique(remaining_vars)

  for (var in remaining_vars) {
    base.var <- paste0(var, '_baseline')
    int.var <- paste0(var, '_', int.name)

    combined <- combined %>%
      mutate(!!var := !!sym(int.var) - !!sym(base.var)) %>%
      select(-base.var, -int.var)
  }

  # Final step is to pivot long to get the run_id column back
  combined <- combined %>%
    pivot_longer(age_baseline_1)

  return(combined)
}

# First calculate clustering metrics
combined <- calculate_clustering_metrics(base.dat, int.dat, int.name = intervention, target.year = 2031)
# Now remove original data for better memory handling
#rm(base.dat, int.dat)
```

```{r}
run_clustering <- function(combined) {

  # Identify categorical columns
  categorical_cols <- c('sex', 'ethnicity', 'region', 'education_state', 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness', 'job_sec')

  # Convert categorical variables to factors
  combined[, categorical_cols] <- lapply(combined[, categorical_cols], as.factor)

  cluster_model <- kproto(x = combined,
                    k = 4, 
                    num_rounds = 20, 
                    verbose = FALSE, 
                    cat.vars = which(names(combined) %in% categorical_cols))

  return(cluster_model)
}

clust <- run_clustering(combined)
```

## Plotting

```{r}
# Now analyze the clustering output
table(clust$cluster)

cluster_profiles <- aggregate(combined, by = list(clust$cluster), FUN = function(x) {
  if (is.numeric(x)) mean(x) else names(sort(table(x), decreasing = TRUE)[1])
})

plot.vars <- c('age', 'region', 'education_state', 'hh_income', 'equivalent_income', 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness', 'job_sec', 'hourly_wage')

for (var in plot.vars) {
  p1 <-ggplot(combined, aes(x = factor(clust$cluster), y = .data[[var]])) +
    geom_boxplot() +
    labs(x = "Cluster", y = var) +
    theme_minimal()
  print(p1)
}
```

# Plotting Change in Components

```{r}
varlist <- c('S7_neighbourhood_safety', 'loneliness',
             'S7_physical_health', 'S7_mental_health')

sccb_all_vars_BOXPLOT(base.dat, int.dat, group.var = 'income.quint', varlist = varlist)
```

# Testing Number of Clusters

```{r}
# # so need a function to generate a dataframe with all those additional metrics for the clustering
# base.data <- base.dat
# int.dat <- wage.dat
# int.name <- 'livingWage'
# 
# # first add scenario column to each
# base <- base.data %>%
#   mutate(scenarios = 'baseline',
#          boost_amount = 0,
#          income_boosted = 0)
# int <- int.dat %>%
#   mutate(scenarios = int.name)
# 
# # now bind together before subsetting
# combined <- rbind(base, int)
# combined <- combined %>%
#   select(pidp, hidp, time, age, sex, ethnicity, region, education_state, hh_income, equivalent_income, S7_housing_quality, S7_neighbourhood_safety, S7_physical_health, S7_mental_health, S7_labour_state, loneliness, job_sec, nkids_ind, hourly_wage)
# 
# # Identify categorical columns
# categorical_cols <- c('sex', 'ethnicity', 'region', 'education_state', 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness', 'job_sec')
# 
# # Convert categorical variables to factors
# combined[, categorical_cols] <- lapply(combined[, categorical_cols], as.factor)
# 
# # Choose a range of values for 'k' (number of clusters)
# k_values <- 2:8 # You can adjust the range
# 
# # Initialize vectors to store evaluation metrics
# wss <- vector('numeric', length(k_values))  # Within-cluster sum of squares
# silhouette_scores <- vector('numeric', length(k_values))
# 
# # # Register parallel cores (adjust the number as needed)
# # num_cores <- 8
# # cl <- makeCluster(num_cores)
# # registerDoParallel(cl)
# # 
# # # Create a function to compute wss and silhouette scores for a given k
# # compute_metrics <- function(k, data) {
# #   kp_model <- kproto(data, k = k, num_rounds = 20, verbose = FALSE, cat.vars = which(names(data) %in% categorical_cols))
# #   wss <- kp_model$tot.withinss
# #   silhouette_score <- silhouette(kp_model$cluster, dist(data))$avg.width
# #   return(c(k, wss, silhouette_score))
# # }
# 
# # Iterate through different values of 'k'
# for (i in 1:length(k_values)) {
#   # Fit the k-prototypes model
#   kp_model <- kproto(combined, k = k_values[i], num_rounds = 20, verbose = FALSE, cat.vars = which(names(combined) %in% categorical_columns))
#   # Compute within-cluster sum of squares
#   wss[i] <- kp_model$tot.withinss
#   # Compute silhouette score
#   # Silhouette score measures the quality of clustering (higher is better)
#   #silhouette_scores[i] <- silhouette(kp_model$cluster, dist(combined))$avg.width
# }
# 
# # Plot the elbow curve
# #par(mfrow = c(1, 2))
# plot(k_values, wss, type = "b", pch = 19, xlab = "Number of Clusters (k)", ylab = "Within-Cluster Sum of Squares (WSS)", main = "Elbow Method")
# #plot(k_values, silhouette_scores, type = "b", pch = 19, xlab = "Number of Clusters (k)", ylab = "Silhouette Score", main = "Silhouette Score Method")
# 
# # Use the elbow or silhouette method to select the optimal number of clusters
# # Elbow method: Look for an 'elbow point' where the WSS starts to level off
# # Silhouette method: Look for the highest silhouette score
```