### 2024-01-09 minos/testing/S7_comp_change_child.Rmd
Didnt render properly on arc, moving to explicitly save the image we want

---
title: "COMPONENT ANALYSIS"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(here)
require(dplyr)
require(tidyr)
require(purrr)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r, include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

```{r}
# out.path.batch <- here::here('output', params$out_path)
# intervention <- params$intervention

out.path.batch <- here::here('output', 'SIPHER7')
intervention <- 'hhIncomeChildUplift'

S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
                 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
                 'ethnicity', 'age', 'region', 'job_sec', 'education_state', 'nkids_ind', 'housing_tenure', 'urban')
out.path.batch <- here::here('output', 'SIPHER7/')

base.batch <- read_batch_out_all_years(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=FALSE)
int.batch <- read_batch_out_all_years(out.path.batch, intervention, start.year = 2021, var.list = c('income_boosted', 'boost_amount', S7.var.list), verbose=FALSE)

# add income_boosted and boost_amount to baseline
base.batch$income_boosted <- 0
base.batch$boost_amount <- 0
```

```{r}
# # select only top 10% of equivalent income increases
# base.sub <- base.batch %>%
#   select(pidp, time, equivalent_income) %>%
#   rename(baseline = equivalent_income)
# 
# int.sub <- int.batch %>%
#   select(pidp, time, equivalent_income) %>%
#   rename(intervention = equivalent_income)
# 
# merged <- merge(x = base.sub,
#                 y = int.sub,
#                 by = c('pidp', 'time'))
# 
# rm(base.sub, int.sub)
# 
# top10p <- merged %>%
#   mutate(diff = intervention - baseline) %>%
#   group_by(pidp) %>%
#   summarise(sum_diff = sum(diff)) %>%
#   filter(sum_diff > quantile(sum_diff, .9))
# 
# rm(merged)
# 
# print(paste0('Top 10% of improvers: n = ', nrow(top10p)))
# 
# base.batch <- base.batch %>%
#   filter(pidp %in% unique(top10p$pidp))
# int.batch <- int.batch %>%
#   filter(pidp %in% unique(top10p$pidp))
# 
# rm(top10p)
```

```{r}
simplified_component_change_BOXPLOT <- function(base.batch, int.batch, var, print = FALSE) {

  # it would be lovely if this was completely generic, but for some things that
  # is more hassle than its worth. Handling string vars here is one of these
  # things

  if (var == 'S7_housing_quality') {
    base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
    int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
  } else if (var == 'S7_neighbourhood_safety') {
    base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
    int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
  } else if (var == 'loneliness') {
    base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                    levels = c(1, 2, 3),
                                    labels = c(3, 2, 1)))
    int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                    levels = c(1, 2, 3),
                                    labels = c(3, 2, 1)))
  }

  base.sub <- base.batch %>%
    select(pidp, time, run_id, all_of(var)) %>%
    rename(Baseline_var = .data[[var]])

  int.sub <- int.batch %>%
    select(pidp, time, run_id, all_of(var)) %>%
    rename(Intervention_var = .data[[var]])

  merged <- merge(base.sub, int.sub,
                  by=c('pidp', 'time', 'run_id'),
                  all = FALSE)

  merged2 <- merged %>%
    mutate(varChange = Intervention_var - Baseline_var,
           effect = case_when(varChange > 0 ~ "Positive",
                              varChange < 0 ~ "Negative",
                              varChange == 0 ~ "No_Change")) %>%
    select(-Baseline_var, -Intervention_var, -varChange) %>%
    group_by(time, run_id, effect) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n),
           effect_prop = n / total) %>%
    select(-n, -total) %>%
    pivot_wider(names_from = 'effect',
                values_from = 'effect_prop') %>%
    mutate(change = Positive - Negative) %>%
    select(-Positive, -Negative, -No_Change) %>%
    group_by(run_id) %>%
    summarise(sum.change = sum(change))

  p1 <- ggplot(data = merged2, aes(y = sum.change)) +
    geom_boxplot(notch = FALSE) +
    labs(title = paste0('Change in ', var)) +
    ylab('% Change')

  if(print) {
    print(p1)
  }

  return(merged2)
}
```

```{r}
varlist <- c('S7_neighbourhood_safety', 'loneliness',
             'S7_physical_health', 'S7_mental_health')

sccb_all_vars_BOXPLOT <- function(base.batch, int.batch, varlist) {
  df <- simplified_component_change_BOXPLOT(base.batch, int.batch, var = 'S7_housing_quality')
  colnames(df) <- c('run_id', 'S7_housing_quality')
  for (S7_var in varlist) {
    df2 <- simplified_component_change_BOXPLOT(base.batch, int.batch, var = S7_var)
    colnames(df2) <- c('run_id', S7_var)
    df <- merge(df, df2, by = c('run_id'))
  }
  rm(df2)

  # finally pivot_longer for ggplot and facet_wrap to take over
  df2 <- df %>%
    pivot_longer(cols = S7_housing_quality:S7_mental_health,
                 names_to = 'Variable',
                 values_to = 'Change')

  p1 <- ggplot(df2, aes(x = Variable, y = Change, group = Variable, colour = Variable, fill = Variable)) +
    geom_boxplot(notch = FALSE) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    theme(legend.position = 'bottom',
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  print(p1)

  ggsave(filename = paste0('sccb_boxplot_', intervention, '.png'),
         plot = p1,
         path = 'plots/')
}

sccb_all_vars_BOXPLOT(base.batch, int.batch, varlist = varlist)

```