### 2024-03-04 minos/outcomes/QALY_comparison2.Rmd
Third time lucky. Switched some paths and constants back from testing to normal operation

---
params:
  experiment: 'default_config/'
  base: 'baseline'
  intervention: '50All'
title: "QALY Comparisons - `r params$experiment` - `r params$intervention` - `r params$baseline`"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)
require(MESS)
require(scales)

#workingDir <- "/home/luke/Documents/WORK/MINOS/PLOTS_FROM_ARC/QALY_SCP_28Feb/QALY_analysis_SCP_test/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/PLOTS_FROM_ARC/tmp/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  options(scipen = 999)
)
rm(workingDir)
```

```{r}
#source_dir <- '/home/luke/Documents/WORK/MINOS/Minos/minos'
#datain <- here::here(source_dir, 'utils_datain.R')

source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_qaly.R'))
```

## Parameters

```{r}
experiment <- params$experiment
baseline <- params$base
intervention <- params$intervention
# 
# experiment <- 'default_config'
# baseline <- 'baseline'
# intervention <- '100All'
```

## Data

```{r}
# read in QALY files for baseline and interventions
out.path <- here::here('output', experiment)

base.path <- get_latest_runtime_subdirectory(here::here(out.path, baseline))
int.path <- get_latest_runtime_subdirectory(here::here(out.path, intervention))

base <- read.csv(paste0(base.path, 'qalys.csv'))
int <- read.csv(paste0(int.path, 'qalys.csv'))

# combine the dataframes
combined <- rbind(base, int)
```

TESTING TEMPORARY TEMPORARY TEMPORARY TEMPORARY TEMPORARY TEMPORARY TEMPORARY 
```{r}
# base <- read.csv('qalys_base.csv')
# int <- read.csv('qalys_25All.csv')
# 
# combined <- rbind(base, int)
# 
# baseline <- 'baseline'
# intervention <- '25All'
```
TESTING TEMPORARY TEMPORARY TEMPORARY TEMPORARY TEMPORARY TEMPORARY TEMPORARY

# Intervention Information

Just going to plot some things here like how many people affected by an intervention and what the mean / total boost amounts look like.

```{r}
combined.pop_boosted <- combined %>%
  group_by(year, run_id, intervention) %>%
  summarise(pop_boosted = sum(pop_boosted))

ggplot(combined.pop_boosted, aes(x = year, y = pop_boosted, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Number Individuals Receiving Intervention')

combined.diff <- combined %>%
  select(run_id, year, intervention, total_boost) %>%
  pivot_wider(names_from = 'intervention',
              values_from = 'total_boost') %>%
  mutate(diff.boost = .data[[intervention]] - .data[[baseline]])

ggplot(combined.diff, aes(x = year, y = diff.boost)) +
  geom_smooth() +
  labs(title = 'Total Boost amount by year') +
  scale_y_continuous(label = comma)

ggplot(combined, aes(x = year, y = alive_ratio, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Ratio of living to dead in output dataframe')

## incidence of mortality
deaths <- combined %>%
  select(run_id, intervention, year, alive_pop, dead_pop) %>%
  group_by(run_id, intervention) %>%
  mutate(total_pop = alive_pop + dead_pop,
         deaths_this_year = dead_pop - lag(dead_pop)) %>%
  group_by(run_id, year, intervention) %>%
  summarise(death_incidence = deaths_this_year / total_pop)

ggplot(deaths, aes(x = year, y = death_incidence, group = intervention, colour = intervention)) +
  geom_smooth() +
  labs(title = 'Incidence of Mortality Comparison')
```

# QALY comparison

```{r}
QALY_comparison(base = base,
                base.name = baseline,
                int = int,
                int.name = intervention)
```

## SF12 Change

```{r}
sf12.plots(base = base,
           base.name = baseline,
           int = int,
           int.name = intervention)
```

## Area Under the Curve

```{r}
auc.plots(base = base, 
          base.name = baseline,
          intervention = int,
          int.name = intervention)
```

# Cost per QALY

The [UK Government](https://www.gov.uk/government/publications/valuation-of-risks-to-life-and-health-monetary-value-of-a-life-year-voly/a-scoping-study-on-the-valuation-of-risks-to-life-and-health-the-monetary-value-of-a-life-year-voly) has set the value of 1 QALY to be equal to $60,000. This is the number we will use for our calculations.

```{r}
cost.per.qaly(base = base,
              base.name = baseline,
              int = int,
              int.name = intervention,
              QALY_value = 70000,  # QALY value == £60,000
              int.label <- intervention)
```

# Incremental Cost-Effectiveness Ratio

From [NICE guidelines](https://www.nice.org.uk/process/pmg6/chapter/assessing-cost-effectiveness) about the cost-effectiveness of health interventions (7.3), and works with a statistic called the Incremental Cost-Effectiveness Ratio (ICER). The formula for calculating the ICER can be found [here](https://en.wikipedia.org/wiki/Incremental_cost-effectiveness_ratio), but in short, it represents the average incremental cost associated with 1 additional unit of the measurement of effect. In our case, the unit of measurement of effect is a QALY.

From the NICE guidelines, there is a general consensus that anything below £20,000 is cost effective, and anything above £30,000 is not. Between £20,000 and £30,000 is a grey area, and depends on the field and other factors to decide (i.e. if a rare or important condition is tackled by an intervention then the cost-effective threshold is higher).

```{r}
ICER(base = base,
     base.name = baseline,
     int = int,
     int.name = intervention,
     QALY_value = 70000,
     int.label = intervention)
```

# Total Cost of Intervention

```{r}
intervention.cost(base = base,
                  base.name = baseline,
                  int = int,
                  int.name = intervention,
                  int.label = intervention)
```

```{r}

```