### 2024-03-13 docsrc/documentation/notebooks/physical_activity.Rmd

---
title: "Physical Activity"
output: html_document
bibliography: 
  - "docsrc/refs.bib"  # switch bibliography position. depending on where the file is being run. stick to this one when compiiling the website using the makefile
  #- "../../refs.bib" # use this one if knitting directly in Rstudio. Easier for debugging.
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir= "../minos")
```

## Physical Activity

Whether a person is physically active or not has strong links to both mental and physical wellbeing, as well as other variables included in the MINOS pathways.

```{r load_data, include=FALSE}
source('docsrc/documentation/notebooks/sphinx_notebook_utils.R')
```

### Methods

What methods are used? Justification due to output data type. explanation of model output.

```{r housing_barchart, echo=FALSE}
discrete_barplot("data/transitions/active/logit/active_2018_2019.rds", 'next_active')
```

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

  What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r active_output, echo=FALSE}
options(width=120)
clm_barplot("data/transitions/active/logit/active_2018_2019.rds", "next_active") 
clm_output("data/transitions/active/logit/active_2018_2019.rds")
```

### References