### 2024-03-13 minos/validation/scotland_synthpop_validation.Rmd
getting basic validation notebook working

---
title: "Synthpop Validation Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r, include=FALSE}
# include flag mutes package load messages for tidiness
library(here)
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'utils_validation_map_vis.R'))
library(dplyr)
```

```{r}
raw_data <- read.csv("../../data/final_US/2020_US_cohort.csv")
raw_data$group <- "Raw Data"
scotland_synth_data <- read.csv("../../data/scaled_scotland_US/2020_US_cohort.csv")
scotland_synth_data$group <- "Scotland Synthetic Data"

#lsoa_to_LA_map <- read.csv(here::here("persistent_data", "EWS_LSOA_2011_LA_2022.csv"))
# merge onto two datasets to get LA level information
```

```{r}
lsoas <- get_subsetted_lsoa_geojson_data("scotland")
las <- get_subsetted_la_geojson_data("scotland")
```

```{r}
density_comparison(raw_data, scotland_synth_data, "SF_12")
density_comparison(raw_data, scotland_synth_data, "hh_income")
```

```{r}
bars_comparison(raw_data, scotland_synth_data, "housing_quality")
bars_comparison(raw_data, scotland_synth_data, "neighbourhood_safety")
```

```{r}
map_la_and_lsoa_level_variable(scotland_synth_data, "SF_12")
```