### 2024-03-21 minos/testing/intervention_pathway_testing.Rmd

---
title: "Testing Interventions - QALY Focus"
output: html_notebook
---

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(knitr)
require(here)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r}
source(here::here('minos', 'utils_datain.R'))
```

```{r}
output.dir <- here::here('output', 'default_config/')

var.list <- c('hh_income', 'age', 'sex', 'ethnicity', 'region', 'education_state', 'nkids', 'hhsize', 
              'housing_quality', 'neighbourhood_safety', 'loneliness', 'ncigs', 'nutrition_quality', 
              'active', 'auditc', 
              'S7_labour_state', 
              'SF_12_MCS', 'SF_12_PCS')

base <- read_batch_out_all_years(out.path = output.dir,
                                 scenario = 'baseline',
                                 var.list = var.list)
all50 <- read_batch_out_all_years(out.path = output.dir,
                                 scenario = '50All',
                                 var.list = var.list)
UC50 <- read_batch_out_all_years(out.path = output.dir,
                                 scenario = '50UniversalCredit',
                                 var.list = var.list)

base$scen <- 'baseline'
all50$scen <- 'all50'
UC50$scen <- 'UC50'

data <- rbind(base, all50, UC50)

```

# Functions

## Ordinal Vars

```{r}
pathway_plots_ordinal <- function(data, var, levels = c(1,2,3)) {
  df <- data %>% 
    group_by(time, scen, .data[[var]]) %>%
    summarise(n = n()) %>%
    group_by(time, scen) %>%
    mutate(prop = (n / sum(n)) * 100)

  for (level in levels) {
    p <- ggplot(dplyr::filter(df, .data[[var]] == level), aes(x = time, y = prop, group = scen, fill = scen, color = scen)) +
            geom_line() +
            labs(title = var, subtitle = level)
    print(p)
  }
}

pathway_plots_continuous <- function(data, var) {
  df <- data %>%
    group_by(time, scen, .data[[var]]) %>%
    summarise(mean_var = mean(.data[[var]]))

  p1 <- ggplot(data, aes(x = time, y = .data[[var]], group = scen, fill = scen, color = scen)) +
    geom_smooth() +
    labs(title = var, subtitle = 'Loess Smoothed')

  p2 <- ggplot(df, aes(x = time, y = mean_var, group = scen, fill = scen, color = scen)) +
    geom_smooth() +
    labs(title = var, subtitle = 'Mean')

  print(p1)
  print(p2)
}
```

# INCOME CHANGE DUE TO INTERVENTION

```{r}
ggplot(data, aes(x = time, y = hh_income, group = scen, fill = scen, color = scen)) +
  geom_smooth() +
  labs(title = 'Household Income Comparison') +
  xlab('Year') +
  ylab('Mean Household Income')
```

# PATHWAYS CHANGE DUE TO INTERVENTION

## Housing Quality

```{r}
pathway_plots_ordinal(data, var = 'housing_quality', levels = c('Low', 'Medium', 'High'))
```

## Neighbourhood Safety

```{r}
pathway_plots_ordinal(data, var = 'neighbourhood_safety')
```

## Loneliness

```{r}
pathway_plots_ordinal(data, var = 'loneliness')
```

## Tobacco

```{r}
pathway_plots_continuous(data, var = 'ncigs')
```

## Tobacco

```{r}
pathway_plots_continuous(data, var = 'nutrition_quality')
```

## Active

```{r}
pathway_plots_ordinal(data, var = 'active', levels = c(0,1))
```

## AUDITC

```{r}
pathway_plots_ordinal(data, var = 'auditc', levels = c('Non-drinker', 'Low Risk', 'Increased Risk', 'High Risk'))
```

# OUTCOMES

## MCS

```{r}
pathway_plots_continuous(data, var = 'SF_12_MCS')
```

## PCS

```{r}
pathway_plots_continuous(data, var = 'SF_12_PCS')
```

```{r}

```