### 2024-03-27 minos/validation/housing_handovers.Rmd
getting sensible looking handovers for net income, ncars, and ncigs

---
params:
    file_path: '' # default arguments. specify others in command line.
    intervention: ''
title: "Housing Default Handovers"
output:
  #output_file: "`r paste('minos/validation/', params$file_path, '_', params$intervention, '_handovers.html') `"
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide

---

This notebook shows the handover plots for all pathways in the MINOS microsimulation model.

Handover plots are not strictly statistical validation, but more a type of 'sanity check' for transitioned variables. A handover plot shows the 'handover' from input data to predicted outcomes, and allows us to see at a glance if the predicted outputs are following the same trends seen in the input data.

# SETUP

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)
require(ggridges) # for densities over time plots. 
require(knitr)
require(here)
require(dplyr)
require(tidyr)
require(purrr)
require(gghighlight)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

Source utils for some functions.

```{r}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

## Data

```{r}

# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)[1:12]
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

out.path <- here::here('output', 'energy_default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
#base.dat <- read_singular_local_out(out.path, 'goodHeatingDummy', drop.dead = TRUE)
```

## Constants

```{r}
save.path <- here::here("plots", "handovers", 'default')
create.if.not.exists(save.path)
start.year <- 2020
# make sure we only overlap on one year for now or else tomfoolery happens. (the plots look bad).
# we may want this overlap in the future? probably not.
raw.dat <- raw.dat[which(raw.dat$time <= start.year), ]
base.dat <- base.dat[which(base.dat$time>=start.year), ]

shall.we.save <- FALSE
```

# PATHWAYS

## Income

```{r}
# first figure out how to plot final_US
# start with hh_income

handover_boxplots(raw.dat, base.dat, 'hh_income')
handover_lineplots(raw.dat, base.dat, "hh_income")
```

### Spaghetti

```{r}
raw.inc <- raw.dat %>%
  select(pidp, age, time, hh_income)
base.inc <- base.dat %>%
  select(pidp, age, time, hh_income)

income.spag <- rbind(raw.inc, base.inc)

density_ridges(income.spag, 'hh_income',
               save=TRUE, 
               save.path=save.path)

pidp_sample <- sample(unique(income.spag$pidp), size=nrow(raw.inc)/10, replace=FALSE)
income.spag.sample <- income.spag[which(income.spag$pidp %in% pidp_sample), ]
spaghetti_plot(income.spag.sample, 'hh_income',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(income.spag.sample, 'hh_income',
               save = shall.we.save,
               save.path = save.path)

rm(raw.inc, base.inc, income.spag)
```

## Net Income & Rent & Mortgage

```{r yearly_energy}
# get required variables. 
raw.inc <- raw.dat %>%
  select(pidp, age, time, net_hh_income, hh_rent, hh_mortgage, council_tax, oecd_equiv)
base.inc <- base.dat %>%
  select(pidp, age, time, net_hh_income, hh_rent, hh_mortgage, council_tax, oecd_equiv)

net.income.spag <- rbind(raw.inc, base.inc)
net.income.spag <- net.income.spag[which(net.income.spag$net_hh_income<50000), ]
density_ridges(net.income.spag, 'net_hh_income',
               save=TRUE, 
               save.path=save.path)
handover_boxplots(raw.dat, base.dat, 'net_hh_income')

pidp_sample <- sample(unique(net.income.spag$pidp), size=nrow(raw.inc)/10, replace=FALSE)
net.income.spag.sample <- net.income.spag[which(net.income.spag$pidp %in% pidp_sample), ]
spaghetti_plot(net.income.spag.sample, 'net_hh_income',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(net.income.spag.sample, 'net_hh_income',
               save = shall.we.save,
               save.path = save.path)

# net household income. 
handover_boxplots(raw.inc, base.inc, 'net_hh_income')
handover_lineplots(raw.inc, base.inc, "net_hh_income")
inc.spag<-rbind(raw.inc, base.inc)
density_ridges(inc.spag, 'net_hh_income',
               save=TRUE, 
               save.path=save.path)

# rent
handover_boxplots(raw.inc[which(raw.inc$hh_rent>1), ], base.inc[which(base.inc$hh_rent>1), ], 'hh_rent')
handover_lineplots(raw.inc[which(raw.inc$hh_rent>1), ], base.inc[which(base.inc$hh_rent>1), ], "hh_rent")
density_ridges(inc.spag[which(inc.spag$hh_rent>1), ], 'hh_rent',
               save=TRUE, 
               save.path=save.path)
spaghetti_plot(net.income.spag.sample, 'hh_rent',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(net.income.spag.sample, 'hh_rent',
               save = shall.we.save,
               save.path = save.path)
# mortgage
handover_boxplots(raw.inc[which(raw.inc$hh_mortgage>1 & raw.inc$hh_mortgage<5000), ], base.inc[which(base.inc$hh_mortgage>1 & base.inc$hh_mortgage<5000), ], 'hh_mortgage')
handover_lineplots(raw.inc[which(raw.inc$hh_mortgage>1), ], base.inc[which(base.inc$hh_mortgage>1), ], "hh_mortgage")
density_ridges(inc.spag[which(inc.spag$hh_mortgage>1 & inc.spag$hh_mortgage<5000), ], 'hh_mortgage',
               save=TRUE, 
               save.path=save.path)
spaghetti_plot(net.income.spag.sample, 'hh_mortgage',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(net.income.spag.sample, "hh_mortgage",
               save = shall.we.save,
               save.path = save.path)
#council tax
handover_boxplots(raw.inc[which(raw.inc$council_tax>1), ], base.inc[which(base.inc$council_tax>1), ], 'council_tax')
handover_lineplots(raw.inc[which(raw.inc$council_tax>1), ], base.inc[which(base.inc$council_tax>1), ], "council_tax")
density_ridges(inc.spag[which(inc.spag$council_tax>1), ], 'council_tax',
               save=TRUE, 
               save.path=save.path)
spaghetti_plot(net.income.spag.sample, 'council_tax',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(net.income.spag.sample, "council_tax",
               save = shall.we.save,
               save.path = save.path)

# oecd equivalence scale. 
#TODO not fully transitioned yet for leaving the nest.
handover_boxplots(raw.inc, base.inc, 'oecd_equiv')
handover_lineplots(raw.inc, base.inc, 'oecd_equiv')
density_ridges(inc.spag, 'oecd_equiv',
               save=TRUE, 
               save.path=save.path)

# tidy up.
rm(raw.inc, base.inc, inc.spag)
```

## Yearly energy

```{r net_income}
# get required variables. 
raw.energy <- raw.dat %>%
  select(pidp, age, time, yearly_energy, yearly_gas, yearly_electric)
base.energy <- base.dat %>%
  select(pidp, age, time, yearly_energy, yearly_gas, yearly_electric)

# gas expenditure
handover_boxplots(raw.energy, base.energy, 'yearly_gas')
handover_lineplots(raw.energy, base.energy, "yearly_gas")
#inc.spag<-rbind(raw.inc, base.inc)
#density_ridges(inc.spag, 'yearly_gas',
#               save=TRUE, 
#               save.path=save.path)

# yearly electric
handover_boxplots(raw.energy, base.energy, 'yearly_electric')
handover_lineplots(raw.energy, base.energy, "yearly_electric")
#density_ridges(inc.spag[which(inc.spag$hh_rent>1), ], 'hh_rent',
#               save=TRUE, 
#               save.path=save.path)

# yearly energy
handover_boxplots(raw.energy, base.energy, 'yearly_energy')
handover_lineplots(raw.energy, base.energy, "yearly_energy")
#density_ridges(inc.spag[which(inc.spag$hh_mortgage>1), ], 'hh_mortgage',
#               save=TRUE, 
#               save.path=save.path)

# tidy up.
rm(raw.energy, base.energy)#, inc.spag)
```

## SF12

```{r}
# sf12 boxplot handovers
handover_boxplots(raw.dat, base.dat, 'SF_12')
handover_lineplots(raw.dat, base.dat, "SF_12")
# 
# raw.sf12 <- raw.dat %>% 
#   dplyr::select(pidp, time, SF_12) %>%
#   group_by(time) %>%
#   summarise(sf12 = mean(SF_12, na.rm = TRUE)) %>%
#   mutate(source = 'final_US')
# 
# base.sf12 <- base.dat %>%
#   dplyr::select(pidp, time, SF_12) %>%
#   group_by(time) %>%
#   summarise(sf12 = mean(SF_12, na.rm = TRUE)) %>%
#   mutate(source = 'baseline_output')
# 
# # merge before plot
# sf12 <- rbind(raw.sf12, base.sf12)
# 
# # Now plot
# ggplot(data = sf12, mapping = aes(x = time, y = sf12, group = source, colour = source)) +
#   geom_smooth() +
#   #geom_line() +
#   geom_vline(xintercept=start.year, linetype='dotted') +
#   labs(title = 'Mental Wellbeing (SF12)', subtitle = 'Full Sample') +
#   xlab('Year') +
#   ylab('SF12')
# 
# ## Try a version where final_US is limited to only those present from wave 1 
# # onwards because the sample refreshments are messing with the plot
# raw.sf12.wave1 <- raw.dat$pidp[raw.dat$time == 2009]
# 
# raw.sf12.2 <- raw.dat %>% 
#   dplyr::select(pidp, time, SF_12) %>%
#   filter(pidp %in% raw.sf12.wave1) %>%
#   group_by(time) %>%
#   summarise(sf12 = mean(SF_12, na.rm = TRUE)) %>%
#   mutate(source = 'final_US')
# 
# base.sf12.2 <- base.dat %>%
#   dplyr::select(pidp, time, SF_12) %>%
#   filter(pidp %in% raw.sf12.wave1) %>%
#   group_by(time) %>%
#   summarise(sf12 = mean(SF_12, na.rm = TRUE)) %>%
#   mutate(source = 'baseline_output')
# 
# sf12.2 <- rbind(raw.sf12.2, base.sf12.2)
# 
# # Now plot
# ggplot(data = sf12.2, mapping = aes(x = time, y = sf12, group = source, colour = source)) +
#   geom_line() +
#   geom_vline(xintercept=start.year, linetype='dotted') +
#   labs(title = 'Mental Wellbeing (SF12 MCS)') +
#   xlab('Year') +
#   ylab('SF12')
# 
# if (shall.we.save) {
#   ggsave(filename = 'SF12_wav1.png',
#        plot = last_plot(),
#        path = save.path)
# }
# 
# rm(raw.sf12, base.sf12, sf12, raw.sf12.wave1, raw.sf12.2, base.sf12.2, sf12.2)
```

### Spaghetti

<!-- ```{r} -->
<!-- raw.sf12 <- raw.dat %>% -->
<!--   select(pidp, age, time, SF_12) -->
<!-- base.sf12 <- base.dat %>% -->
<!--   select(pidp, age, time, SF_12) -->

<!-- sf12.spag <- rbind(raw.sf12, base.sf12) -->

<!-- sf12.spag.sample <- sf12.spag[which(sf12.spag$pidp %in% pidp_sample), ] -->
<!-- spaghetti_plot(sf12.spag.sample, 'SF_12', -->
<!--                save = shall.we.save, -->
<!--                save.path = save.path) -->
<!-- spaghetti_highlight_max_plot(sf12.spag.sample, 'SF_12', -->
<!--                save = shall.we.save, -->
<!--                save.path = save.path) -->

<!-- density_ridges(sf12.spag, "SF_12",  -->
<!--                save = T,  -->
<!--                save.path = save.path) -->

<!-- rm(raw.sf12, base.sf12, sf12.spag) -->
<!-- ``` -->

<!-- ## Financial Situation -->

<!-- ```{r} -->
<!-- raw.financial_situation <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, financial_situation) %>% -->
<!--   group_by(time, financial_situation) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.financial_situation <- base.dat %>% -->
<!--   dplyr::select(pidp, time, financial_situation) %>% -->
<!--   group_by(time, financial_situation) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- financial_situation <- rbind(raw.financial_situation, base.financial_situation) -->
<!-- financial_situation$financial_situation <- as.factor(financial_situation$financial_situation) -->

<!-- # Now plot -->
<!-- ggplot(data = financial_situation, mapping = aes(x = time, y = n, group = financial_situation, colour = financial_situation)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Financial Situation', subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- financial_situation.norm <- financial_situation %>% -->
<!--   group_by(time) %>% -->
<!--   mutate(total = sum(n)) %>% -->
<!--   mutate(prct = (n / total)) -->

<!-- ggplot(data = financial_situation.norm, mapping = aes(x = time, y = prct, fill=financial_situation)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Financial situation') + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'financial_situation.png', -->
<!--        plot = last_plot(), -->
<!--        path = save.path) -->
<!-- } -->

<!-- rm(raw.financial_situation, base.financial_situation, financial_situation, financial_situation.norm) -->
<!-- ``` -->

<!-- ## Housing Quality -->

<!-- ```{r} -->
<!-- raw.housing <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, housing_quality) %>% -->
<!--   group_by(time, housing_quality) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.housing <- base.dat %>% -->
<!--   dplyr::select(pidp, time, housing_quality) %>% -->
<!--   group_by(time, housing_quality) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- housing <- rbind(raw.housing, base.housing) -->
<!-- housing$housing_quality <- as.factor(housing$housing_quality) -->

<!-- # Now plot -->
<!-- ggplot(data = housing, mapping = aes(x = time, y = n, group = housing_quality, colour = housing_quality)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Housing Quality', subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- housing.norm <- housing %>% -->
<!--   group_by(time) %>% -->
<!--   mutate(total = sum(n)) %>% -->
<!--   mutate(prct = (n / total)) -->

<!-- ggplot(data = housing.norm, mapping = aes(x = time, y = prct, fill=housing_quality)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Housing Quality') + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'housing_quality.png', -->
<!--        plot = last_plot(), -->
<!--        path = save.path) -->
<!-- } -->

<!-- rm(raw.housing, base.housing, housing, housing.norm) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- # raw.s <- raw.dat %>% -->
<!-- #   select(pidp, age, time, housing_quality) -->
<!-- # base.s <- base.dat %>% -->
<!-- #   select(pidp, age, time, housing_quality) -->
<!-- #  -->
<!-- # spag <- rbind(raw.s, base.s) -->
<!-- #  -->
<!-- # spaghetti_plot(spag, 'housing_quality', -->
<!-- #                save = shall.we.save, -->
<!-- #                save.path = save.path) -->
<!-- #  -->
<!-- # rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## Housing Tenure -->

<!-- ```{r} -->
<!-- raw.tenure <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, housing_tenure) %>% -->
<!--   group_by(time, housing_tenure) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.tenure <- base.dat %>% -->
<!--   dplyr::select(pidp, time, housing_tenure) %>% -->
<!--   group_by(time, housing_tenure) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- tenure <- rbind(raw.tenure, base.tenure) -->
<!-- tenure$housing_tenure <- as.factor(tenure$housing_tenure) -->

<!-- # Now plot -->
<!-- ggplot(data = tenure, mapping = aes(x = time, y = n, group = housing_tenure, colour = housing_tenure)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Housing Tenure', subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- tenure.norm <- tenure %>% -->
<!--   group_by(time) %>% -->
<!--   mutate(total = sum(n)) %>% -->
<!--   mutate(prct = (n / total)) -->

<!-- ggplot(data = tenure.norm, mapping = aes(x = time, y = prct, fill=housing_tenure)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Housing Tenure') + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'housing_tenure.png', -->
<!--        plot = last_plot(), -->
<!--        path = save.path) -->
<!-- } -->

<!-- rm(raw.tenure, base.tenure, tenure, tenure.norm) -->
<!-- ``` -->

<!-- ## Neighbourhood Safety -->

<!-- ```{r} -->
<!-- raw.neighbour <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, neighbourhood_safety) %>% -->
<!--   group_by(time, neighbourhood_safety) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.neighbour <- base.dat %>% -->
<!--   dplyr::select(pidp, time, neighbourhood_safety) %>% -->
<!--   group_by(time, neighbourhood_safety) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- neighbour <- rbind(raw.neighbour, base.neighbour) -->
<!-- neighbour$neighbourhood_safety <- as.factor(neighbour$neighbourhood_safety) -->

<!-- # Now plot -->
<!-- ggplot(data = neighbour, mapping = aes(x = time, y = n, group = neighbourhood_safety, colour = neighbourhood_safety)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Neighbourhood Safety', subtitle = 'Counts by level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- neighbour.norm <- neighbour %>% -->
<!--   group_by(time) %>% -->
<!--   filter(neighbourhood_safety != -9) %>% -->
<!--   mutate(total = sum(n)) %>% -->
<!--   mutate(prct = (n / total)) -->

<!-- ggplot(data = neighbour.norm, mapping = aes(x = time, y = prct, fill=neighbourhood_safety)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Neighbourhood Safety') + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'neighbourhood_safety.png', -->
<!--        plot = last_plot(), -->
<!--        path = save.path) -->
<!-- } -->

<!-- #rm(raw.neighbour, base.neighbour, neighbour, neighbour.norm) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- # raw.s <- raw.dat %>% -->
<!-- #   select(pidp, age, time, neighbourhood_safety) -->
<!-- # base.s <- base.dat %>% -->
<!-- #   select(pidp, age, time, neighbourhood_safety) -->
<!-- #  -->
<!-- # spag <- rbind(raw.s, base.s) -->
<!-- #  -->
<!-- # spaghetti_plot(spag, 'neighbourhood_safety', -->
<!-- #                save = shall.we.save, -->
<!-- #                save.path = save.path) -->
<!-- #  -->
<!-- # rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## Loneliness -->

<!-- ```{r} -->
<!-- raw.loneliness <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, loneliness) %>% -->
<!--   group_by(time, loneliness) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.loneliness <- base.dat %>% -->
<!--   dplyr::select(pidp, time, loneliness) %>% -->
<!--   group_by(time, loneliness) %>% -->
<!--   count() %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- loneliness <- rbind(raw.loneliness, base.loneliness) -->
<!-- loneliness$loneliness <- as.factor(loneliness$loneliness) -->

<!-- # Now plot -->
<!-- ggplot(data = loneliness, mapping = aes(x = time, y = n, group = loneliness, colour = loneliness)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Loneliness', subtitle = 'Counts by level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- loneliness.norm <- loneliness %>% -->
<!--   group_by(time) %>% -->
<!--   filter(loneliness != -9) %>% -->
<!--   mutate(total = sum(n)) %>% -->
<!--   mutate(prct = (n / total)) -->

<!-- ggplot(data = loneliness.norm, mapping = aes(x = time, y = prct, fill=loneliness)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Loneliness') + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'loneliness.png', -->
<!--        plot = last_plot(), -->
<!--        path = save.path) -->
<!-- } -->

<!-- rm(raw.loneliness, base.loneliness, loneliness, loneliness.norm) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- # raw.s <- raw.dat %>% -->
<!-- #   select(pidp, age, time, loneliness) -->
<!-- # base.s <- base.dat %>% -->
<!-- #   select(pidp, age, time, loneliness) -->
<!-- #  -->
<!-- # spag <- rbind(raw.s, base.s) -->
<!-- #  -->
<!-- # spaghetti_plot(spag, 'loneliness', -->
<!-- #                save = shall.we.save, -->
<!-- #                save.path = save.path) -->
<!-- #  -->
<!-- # rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## Nutrition Quality -->

<!-- TODO: Drop negative values ONLY from waves with no data i.e. not 7,9,11 -->

<!-- ```{r} -->
<!-- #  nutriton quality line/boxplots. -->
<!-- handover_boxplots(raw.dat, base.dat, 'nutrition_quality') -->
<!-- handover_lineplots(raw.dat, base.dat, 'nutrition_quality') -->

<!-- # Try a version where final_US is limited to only those present from wave 1 -->
<!-- # onwards because the sample refreshments are messing with the plot -->
<!-- # raw.wave1 <- raw.dat$pidp[raw.dat$time == 2009] -->
<!-- #  -->
<!-- # raw.nut2 <- raw.dat %>% -->
<!-- #   dplyr::select(pidp, time, nutrition_quality) %>% -->
<!-- #   filter(pidp %in% raw.wave1) %>% -->
<!-- #   filter(nutrition_quality >= 0) %>% -->
<!-- #   group_by(time) %>% -->
<!-- #   summarise(nutrition_quality = mean(nutrition_quality, na.rm = TRUE)) %>% -->
<!-- #   mutate(source = 'final_US') -->
<!-- #  -->
<!-- # base.nut2 <- base.dat %>% -->
<!-- #   dplyr::select(pidp, time, nutrition_quality) %>% -->
<!-- #   filter(pidp %in% raw.wave1) %>% -->
<!-- #   group_by(time) %>% -->
<!-- #   summarise(nutrition_quality = mean(nutrition_quality, na.rm = TRUE)) %>% -->
<!-- #   mutate(source = 'baseline_output') -->
<!-- #  -->
<!-- # nutrition_quality2 <- rbind(raw.nut2, base.nut2) -->
<!-- #  -->
<!-- # # Now plot -->
<!-- # ggplot(data = nutrition_quality2, mapping = aes(x = time, y = nutrition_quality, group = source, colour = source)) + -->
<!-- #   geom_line() + -->
<!-- #   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!-- #   labs(title = 'Nutrition Quality') + -->
<!-- #   xlab('Year') + -->
<!-- #   ylab('Nutrition Quality') -->
<!-- #  -->
<!-- # if (shall.we.save) { -->
<!-- #   ggsave(filename = 'nutrition_quality_wav1.png', -->
<!-- #        plot = last_plot(), -->
<!-- #        path = save.path) -->
<!-- # } -->
<!-- #  -->
<!-- # rm(raw.nut, base.nut, nutrition_quality, raw.wave1, raw.nut2, base.nut2, -->
<!-- #     nutrition_quality2) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- # raw.nut <- raw.dat %>% -->
<!-- #   select(pidp, age, time, nutrition_quality) -->
<!-- # base.nut <- base.dat %>% -->
<!-- #   select(pidp, age, time, nutrition_quality) -->
<!-- #  -->
<!-- # nut.spag <- rbind(raw.nut, base.nut) -->
<!-- #  -->
<!-- # spaghetti_plot(nut.spag, 'nutrition_quality', -->
<!-- #                save = shall.we.save, -->
<!-- #                save.path = save.path) -->
<!-- #  -->
<!-- #  -->
<!-- # density_ridges(nut.spag, "nutrition_quality",  -->
<!-- #                save = T,  -->
<!-- #                save.path = save.path) -->
<!-- #  -->
<!-- # rm(raw.nut, base.nut, nut.spag) -->
<!-- ``` -->

<!-- ## Physical Health -->

<!-- NOTE: At present, physical health is not predicted or transitioned. -->

<!-- ```{r} -->
<!-- # first figure out how to plot final_US -->
<!-- # start with hh_income -->
<!-- raw.phealth <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, phealth) %>% -->
<!--   filter(time > 2009) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(phealth = mean(phealth, na.rm = TRUE)) %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.phealth <- base.dat %>% -->
<!--   dplyr::select(pidp, time, phealth) %>% -->
<!--   filter(time > 2009) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(phealth = mean(phealth, na.rm = TRUE)) %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- phealth <- rbind(raw.phealth, base.phealth) -->

<!-- # Now plot -->
<!-- ggplot(data = phealth, mapping = aes(x = time, y = phealth, group = source, colour = source)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Physical Health', subtitle = 'Full Sample') + -->
<!--   xlab('Year') + -->
<!--   ylab('Average') -->

<!-- ## Try a version where final_US is limited to only those present from wave 1 -->
<!-- # onwards because the sample refreshments are messing with the plot -->
<!-- raw.wave1 <- raw.dat$pidp[raw.dat$time == 2009] -->

<!-- raw.phealth2 <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, phealth) %>% -->
<!--   filter(time > 2009) %>% -->
<!--   filter(pidp %in% raw.wave1) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(phealth = mean(phealth, na.rm = TRUE)) %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.phealth2 <- base.dat %>% -->
<!--   dplyr::select(pidp, time, phealth) %>% -->
<!--   filter(time > 2009) %>% -->
<!--   filter(pidp %in% raw.wave1) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(phealth = mean(phealth, na.rm = TRUE)) %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- phealth2 <- rbind(raw.phealth2, base.phealth2) -->

<!-- # Now plot -->
<!-- ggplot(data = phealth2, mapping = aes(x = time, y = phealth, group = source, colour = source)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Physical Health') + -->
<!--   xlab('Year') + -->
<!--   ylab('Average') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'physical_health_wav1.png', -->
<!--        plot = last_plot(), -->
<!--        path = save.path) -->
<!-- } -->

<!-- #raw.phealth.check <- raw.dat %>% -->
<!-- #  select(pidp, time, phealth) %>% -->
<!-- #  filter(time > 2009) %>% -->
<!-- #  filter(pidp %in% raw.wave1) -->

<!-- #base.phealth.check <- base.dat %>% -->
<!-- #  select(pidp, time, phealth) %>% -->
<!-- #  filter(time > 2009) %>% -->
<!-- #  filter(pidp %in% raw.wave1) -->

<!-- rm(raw.phealth, base.phealth, phealth, raw.wave1, raw.phealth2, base.phealth2, phealth2) -->
<!-- #rm(raw.phealth.check, base.phealth.check) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- # raw.s <- raw.dat %>% -->
<!-- #   select(pidp, age, time, phealth) -->
<!-- # base.s <- base.dat %>% -->
<!-- #   select(pidp, age, time, phealth) -->
<!-- #  -->
<!-- # spag <- rbind(raw.s, base.s) -->
<!-- #  -->
<!-- # spaghetti_plot(spag, 'phealth', -->
<!-- #              save = shall.we.save, -->
<!-- #              save.path = save.path) -->
<!-- #  -->
<!-- # rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## Education -->

<!-- ```{r} -->
<!-- prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'education_state') -->

<!-- var <- prepped.list$var -->
<!-- norm <- prepped.list$norm -->

<!-- tit <- 'Education State' -->

<!-- # Now plot -->
<!-- ggplot(data = var, mapping = aes(x = time, y = n, group = education_state, colour = education_state)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = tit, subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- ggplot(data = norm, mapping = aes(x = time, y = prct, fill=education_state)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = tit) + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'education_state.png', -->
<!--          plot = last_plot(), -->
<!--          path = save.path) -->
<!-- } -->

<!-- rm(prepped.list, var, norm, tit) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pivoted <- combine_and_pivot_long(df1 = base.dat, -->
<!--                                        df1.name = 'simulated', -->
<!--                                        df2 = raw.dat, -->
<!--                                        df2.name = 'raw', -->
<!--                                        var = 'education_state') -->

<!-- cv_ordinal_plots(pivoted.df = pivoted, -->
<!--                  var = 'education_state', -->
<!--                  save = shall.we.save, -->
<!--                  save.path = save.path) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- # raw.s <- raw.dat %>% -->
<!-- #   select(pidp, age, time, education_state) -->
<!-- # base.s <- base.dat %>% -->
<!-- #   select(pidp, age, time, education_state) -->
<!-- #  -->
<!-- # spag <- rbind(raw.s, base.s) -->
<!-- #  -->
<!-- # spaghetti_plot(spag, 'education_state', -->
<!-- #                save = shall.we.save, -->
<!-- #                save.path = save.path) -->
<!-- #  -->
<!-- # rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## nkids Household -->

<!-- ```{r} -->
<!-- handover_boxplots(raw.dat, base.dat, 'nkids') -->
<!-- handover_lineplots(raw.dat, base.dat, 'nkids') -->
<!-- ``` -->

<!-- ```{r} -->
<!-- raw.s <- raw.dat %>% -->
<!--   select(pidp, age, time, nkids) -->
<!-- base.s <- base.dat %>% -->
<!--   select(pidp, age, time, nkids) -->

<!-- spag <- rbind(raw.s, base.s) -->

<!-- spaghetti_plot(spag, 'nkids', -->
<!--                save = shall.we.save, -->
<!--                save.path = save.path) -->

<!-- rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## nkids Max -->

<!-- ```{r} -->
<!-- # first figure out how to plot final_US -->
<!-- # start with hh_income -->
<!-- raw.nkids_max <- raw.dat %>% -->
<!--   dplyr::select(pidp, time, nkids) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(max_nkids = max(nkids)) %>% -->
<!--   mutate(source = 'final_US') -->

<!-- base.nkids_max <- base.dat %>% -->
<!--   dplyr::select(pidp, time, nkids) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(max_nkids = max(nkids)) %>% -->
<!--   mutate(source = 'baseline_output') -->

<!-- # merge before plot -->
<!-- max_nkids <- rbind(raw.nkids_max, base.nkids_max) -->

<!-- # Now plot -->
<!-- ggplot(data = max_nkids, mapping = aes(x = time, y = max_nkids, group = source, colour = source)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'Max nkids', subtitle = 'Full Sample') + -->
<!--   xlab('Year') + -->
<!--   ylab('Average') -->
<!-- ``` -->

<!-- ## Parity -->

<!-- ```{r} -->
<!-- base.parity <- base.dat %>% -->
<!--   filter(sex == "Female") -->

<!-- raw.parity <- raw.dat %>% -->
<!--   filter(sex == "Female") -->

<!-- prepped.list <- validation_prep_ordinal(raw.parity, base.parity, 'nkids_ind') -->

<!-- var <- prepped.list$var -->
<!-- norm <- prepped.list$norm -->

<!-- tit <- 'Parity (total no. of children)' -->

<!-- # Now plot -->
<!-- ggplot(data = var, mapping = aes(x = time, y = n, group = nkids_ind, colour = nkids_ind)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = tit, subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- ggplot(data = norm, mapping = aes(x = time, y = prct, fill=nkids_ind)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = tit) + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'parity.png', -->
<!--          plot = last_plot(), -->
<!--          path = save.path) -->
<!-- } -->

<!-- rm(prepped.list, var, norm, tit) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- pivoted <- combine_and_pivot_long(df1 = base.parity, -->
<!--                                   df1.name = 'simulated',  -->
<!--                                   df2 = raw.parity,  -->
<!--                                   df2.name = 'raw',  -->
<!--                                   var = 'nkids_ind') -->

<!-- cv_ordinal_plots(pivoted.df = pivoted,  -->
<!--                  var = 'nkids_ind', -->
<!--                  save = shall.we.save, -->
<!--                  save.path = save.path) -->
<!-- ``` -->

<!-- ### Spaghetti -->

<!-- ```{r} -->
<!-- raw.s <- raw.parity %>% -->
<!--   select(pidp, age, time, nkids_ind) -->
<!-- base.s <- base.parity %>% -->
<!--   select(pidp, age, time, nkids_ind) -->

<!-- spag <- rbind(raw.s, base.s) -->

<!-- spaghetti_plot(spag, 'nkids_ind', -->
<!--                save = shall.we.save, -->
<!--                save.path = save.path) -->

<!-- rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ## Hourly Wage -->

<!-- ```{r} -->
<!-- handover_boxplots(raw.dat, base.dat, 'hourly_wage') -->
<!-- handover_lineplots(raw.dat, base.dat, 'hourly_wage') -->
<!-- ``` -->

<!-- ```{r} -->
<!-- raw.s <- raw.dat %>% -->
<!--   select(pidp, age, time, hourly_wage) -->
<!-- base.s <- base.dat %>% -->
<!--   select(pidp, age, time, hourly_wage) -->

<!-- spag <- rbind(raw.s, base.s) -->

<!-- spaghetti_plot(spag, 'hourly_wage', -->
<!--                save = shall.we.save, -->
<!--                save.path = save.path) -->

<!-- # now a version limited to 0-100 -->
<!-- spag <- spag %>% -->
<!--   filter(hourly_wage <= 300) -->
<!-- spaghetti_plot(spag, 'hourly_wage', -->
<!--                save = shall.we.save, -->
<!--                save.path = save.path) -->

<!-- rm(raw.s, base.s, spag) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # some basic statistics about hourly_wage variable -->
<!-- print("Some basic statistics on hourly_wage. These are all for 2020.") -->
<!-- raw.2020 <- raw.dat %>% filter(time == 2020) -->
<!-- print('All living people in 2020') -->
<!-- print(paste0("Number of people alive in 2020: ", nrow(raw.2020))) -->
<!-- # number individuals with an hourly_wage > 0 -->
<!-- print('People not in work') -->
<!-- print(paste0("Number of individuals not in paid work: ", sum(!raw.2020$S7_labour_state %in% c('FT Employed', 'PT Employed')))) -->
<!-- print(paste0("Number of individuals with hourly_wage == 0: ", sum(raw.2020$hourly_wage == 0))) -->

<!-- # Number in work -->
<!-- print('People in work') -->
<!-- print(paste0("Number of individuals in paid work: ", sum(raw.2020$S7_labour_state %in% c('FT Employed', 'PT Employed')))) -->
<!-- print(paste0("Number of individuals with hourly_wage > 0: ", sum(raw.2020$hourly_wage > 0))) -->

<!-- # Extreme hourly_wage values -->
<!-- print('Extreme hourly wage values: HIGH') -->
<!-- print(paste0("Number of individuals with hourly_wage > 30: ", sum(raw.2020$hourly_wage > 30))) -->
<!-- print(paste0("Number of individuals with hourly_wage > 50: ", sum(raw.2020$hourly_wage > 50))) -->
<!-- print(paste0("Number of individuals with hourly_wage > 100: ", sum(raw.2020$hourly_wage > 100))) -->
<!-- print(paste0("Number of individuals with hourly_wage > 500: ", sum(raw.2020$hourly_wage > 500))) -->
<!-- print(paste0("Number of individuals with hourly_wage > 1000: ", sum(raw.2020$hourly_wage > 1000))) -->
<!-- print(paste0("Number of individuals with hourly_wage > 2000: ", sum(raw.2020$hourly_wage > 2000))) -->

<!-- raw.2020.work <- raw.2020 %>% filter(hourly_wage > 0) -->
<!-- print('Extreme hourly wage values: LOW (non-zero)') -->
<!-- print(paste0("Number of individuals with hourly_wage < 20: ", sum(raw.2020.work$hourly_wage < 20))) -->
<!-- print(paste0("Number of individuals with hourly_wage < 10: ", sum(raw.2020.work$hourly_wage < 10))) -->
<!-- print(paste0("Number of individuals with hourly_wage < 5: ", sum(raw.2020.work$hourly_wage < 5))) -->
<!-- print(paste0("Number of individuals with hourly_wage < 1: ", sum(raw.2020.work$hourly_wage < 1))) -->
<!-- ``` -->

<!-- ## S7 Labour State -->

<!-- ```{r} -->
<!-- prepped.list <- validation_prep_ordinal(raw.dat, base.dat, 'S7_labour_state') -->

<!-- lab.state <- prepped.list$var -->
<!-- lab.state.norm <- prepped.list$norm -->

<!-- # Now plot -->
<!-- ggplot(data = lab.state, mapping = aes(x = time, y = n, group = S7_labour_state, colour = S7_labour_state)) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'S7 Labour State', subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- ggplot(data = lab.state.norm, mapping = aes(x = time, y = prct, fill=S7_labour_state)) + -->
<!--   geom_bar(stat = 'identity') + -->
<!--   geom_vline(xintercept=start.year, linetype='dotted') + -->
<!--   labs(title = 'S7 Labour State') + -->
<!--   xlab('Year') + -->
<!--   ylab('Proportion') -->

<!-- if (shall.we.save) { -->
<!--   ggsave(filename = 'S7_labour_state.png', -->
<!--          plot = last_plot(), -->
<!--          path = save.path) -->
<!-- } -->

<!-- rm(prepped.list, lab.state, lab.state.norm) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pivoted <- combine_and_pivot_long(df1 = base.dat,  -->
<!--                                        df1.name = 'simulated',  -->
<!--                                        df2 = raw.dat,  -->
<!--                                        df2.name = 'raw',  -->
<!--                                        var = 'S7_labour_state') -->

<!-- cv_ordinal_plots(pivoted.df = pivoted,  -->
<!--                  var = 'S7_labour_state', -->
<!--                  save = shall.we.save, -->
<!--                  save.path = save.path) -->
<!-- ``` -->

<!-- ## Child Ages -->

<!-- ```{r} -->
<!-- raw.child.ages <- raw.dat %>% -->
<!--   dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>% -->
<!--   filter(time > 2013 & child_ages != 'childless') %>% -->
<!--   group_by(time, hidp) %>% -->
<!--   drop_na() %>% -->
<!--   summarise(child_ages = find_mode(child_ages)) -->

<!-- r.s<- strsplit(raw.child.ages$child_ages, "_") -->

<!-- raw.child.ages.expanded <- data.frame(time = rep(raw.child.ages$time, sapply(r.s, length)), expanded_child_ages = unlist(r.s)) -->
<!-- raw.child.ages.expanded <- raw.child.ages.expanded %>% -->
<!--   group_by(time, expanded_child_ages) %>% -->
<!--   summarise(age_counts = n()) %>% -->
<!--   mutate(source = 'final_US') %>% -->
<!--   filter(expanded_child_ages <= 16) -->

<!-- base.child.ages <- base.dat %>% -->
<!--   dplyr::select(pidp, hidp, time, "child_ages", "nkids") %>% -->
<!--   filter(time > 2009) %>% -->
<!--   group_by(time, hidp) %>% -->
<!--   drop_na() %>% -->
<!--   summarise(child_ages = find_mode(child_ages)) -->

<!-- b.s<- strsplit(base.child.ages$child_ages, "_") -->

<!-- base.child.ages.expanded <- data.frame(time = rep(base.child.ages$time, sapply(b.s, length)), expanded_child_ages = unlist(b.s)) -->
<!-- base.child.ages.expanded <- base.child.ages.expanded %>% -->
<!--   group_by(time, expanded_child_ages) %>% -->
<!--   summarise(age_counts = n()) %>% -->
<!--   mutate(source = 'baseline') -->

<!-- combined <- rbind(raw.child.ages.expanded, base.child.ages.expanded) -->

<!-- ggplot(data = combined, mapping = aes(x = time, y = age_counts, group = expanded_child_ages, colour =  factor(as.numeric(expanded_child_ages)))) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept = start.year) + -->
<!--   labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') + -->
<!--   scale_colour_viridis_d() + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- ggplot(data = combined, mapping = aes(x = factor(as.numeric(expanded_child_ages)), y = age_counts, group = factor(time), colour = factor(time))) + -->
<!--   geom_line() + -->
<!--   geom_vline(xintercept = start.year) + -->
<!--   labs(title = "Child Age Counts Over Time", subtitle = 'Counts by Level') + -->
<!--   scale_colour_viridis_d() + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- base.child.ages.nkids <- base.dat %>% -->
<!--   group_by(time, hidp) %>% -->
<!--   summarise(nkids_counts = mean(nkids)) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(nkids_by_year = sum(nkids_counts)) %>% -->
<!--   mutate(source = 'baseline') -->

<!-- raw.child.ages.nkids <- raw.dat %>% -->
<!--   group_by(time, hidp) %>% -->
<!--   summarise(nkids_counts = mean(nkids)) %>% -->
<!--   group_by(time) %>% -->
<!--   summarise(nkids_by_year = sum(nkids_counts)) %>% -->
<!--   mutate(source = 'final_US') -->

<!-- combined.nkids <- rbind(base.child.ages.nkids, raw.child.ages.nkids) -->

<!-- ggplot(data = combined.nkids, mapping = aes(x = time, y = nkids_by_year, group = source, colour = source)) + -->
<!--   geom_line() + -->
<!--   labs(title = "nkids Count Over Time", subtitle = 'Counts by Level') + -->
<!--   xlab('Year') + -->
<!--   ylab('Count') -->

<!-- ``` -->

## Number of cars

```{r ncars}

raw.ncars <- raw.dat
base.ncars <- base.dat
raw.ncars[raw.ncars$ncars>=10, 'ncars'] <- "More than 9."
base.ncars[base.ncars$ncars>=10, 'ncars'] <- "More than 9."

raw.ncars <- raw.ncars %>%
  dplyr::select(pidp, time, ncars) %>%
  filter(ncars >= 0) %>%
  group_by(time, ncars) %>%
  count() %>%
  mutate(source = 'final_US')

base.ncars <- base.ncars %>%
  dplyr::select(pidp, time, ncars) %>%
  filter(ncars >= 0) %>%
  group_by(time, ncars) %>%
  count() %>%
  mutate(source = 'baseline_output')

# merge before plot
ncars <- rbind(raw.ncars, base.ncars)
ncars$ncars <- as.factor(ncars$ncars)

# Now plot
ggplot(data = ncars, mapping = aes(x = time, y = n, group = ncars, colour = ncars)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Number of Cars', subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ncars.norm <- ncars %>%
  group_by(time) %>%
  mutate(total = sum(n)) %>%
  mutate(prct = (n / total))

ggplot(data = ncars.norm, mapping = aes(x = time, y = prct, group = ncars, colour = ncars)) +
  geom_line() +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Number of Cars', subtitle = 'Counts by Level') +
  xlab('Year') +
  ylab('Count')

ggplot(data = ncars.norm, mapping = aes(x = time, y = prct, fill=ncars)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept=start.year, linetype='dotted') +
  labs(title = 'Number of Cars') +
  xlab('Year') +
  ylab('Proportion')

if (shall.we.save) {
  ggsave(filename = 'ncars.png',
       plot = last_plot(),
       path = save.path)
}

rm(raw.ncars, base.ncars, ncars, ncars.norm)
```

## ncigs

```{r}
handover_lineplots(raw.dat, base.dat, "ncigs")
raw.ncigs <- raw.dat %>%
  select(pidp, age, time, ncigs)
base.ncigs <- base.dat %>%
  select(pidp, age, time, ncigs)
ncigs.spag <- rbind(raw.ncigs, base.ncigs)
shall.we.save <- F
pidp_sample <- sample(unique(ncigs.spag$pidp), size=nrow(raw.dat)/10, replace=FALSE)
ncigs.spag.sample <- ncigs.spag[which(ncigs.spag$pidp %in% pidp_sample), ]
spaghetti_plot(ncigs.spag.sample, 'ncigs',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(ncigs.spag.sample, 'ncigs',
               save = shall.we.save,
               save.path = save.path)

zip_density_ridges(ncigs.spag, "ncigs",
               save = shall.we.save,
               save.path = save.path)
rm(raw.ncigs, base.ncigs, ncigs.spag)
```

## ncigs non-zero

```{r}

raw.ncigs <- raw.dat %>%
  filter(ncigs>0) %>%
  select(pidp, age, time, ncigs)
base.ncigs <- base.dat %>%
  filter(ncigs>0) %>%
  select(pidp, age, time, ncigs)

handover_boxplots(raw.ncigs, base.ncigs, 'ncigs')

ncigs.spag <- rbind(raw.ncigs, base.ncigs) # get all smokers.
shall.we.save <- F
pidp_sample <- sample(unique(ncigs.spag$pidp), size=nrow(raw.dat)/100, replace=FALSE)
ncigs.spag.sample <- ncigs.spag[which(ncigs.spag$pidp %in% pidp_sample), ]
spaghetti_plot(ncigs.spag.sample, 'ncigs',
               save = shall.we.save,
               save.path = save.path)
spaghetti_highlight_max_plot(ncigs.spag.sample, 'ncigs',
               save = shall.we.save,
               save.path = save.path)

zip_density_ridges(ncigs.spag, "ncigs", 
               save = shall.we.save, 
               save.path = save.path)
rm(raw.ncigs, base.ncigs, ncigs.spag)
```