### 2024-04-18 minos/testing/PCS_QALYs.Rmd
Added plots to handovers for visualising counts of extreme old age groups. Also started an R Notebook for the PCS QALY work

---
title: "Physical Component Score and QALYs"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# Introduction

```{r "setup-environment", include=False}
require(tidyverse)
require(ggplot2)
require(ggridges) # for densities over time plots. 
require(knitr)
require(here)
require(dplyr)
require(tidyr)
require(purrr)
require(gghighlight)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r "setup-utils", include=False}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
source(here::here('minos', 'utils_qaly.R'))
```

```{r "setup-parameters", include=False}
#experiment <- 'glasgow_scaled'
#baseline <- 'baseline'
#ints <- c('50All', '50UniversalCredit')
#start.year <- 2020
```

```{r "setup-data", include=False}
# Read raw datafiles in
raw.files <- list.files(here::here('data', 'final_US'), pattern='[0-9]{4}_US_cohort.csv', full.names = TRUE)
raw.dat <- do.call(rbind, lapply(raw.files, read.csv))

#out.path <- here::here('output', 'glasgow_scaled/')
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
```

# Tran

```{r}

```