### 2024-05-16 minos/outcomes/child_poverty_reduction_vis.Rmd
Tracking the ChildPovReduction plotting notebooks

---
title: "Child Poverty Interventions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

This document presents results from a host of interventions on child poverty, including visualisations of the change in SF_12_MCS compared with baseline.

```{r "setup", include=FALSE}

library(tidyverse)
library(ggplot2)
library(ggridges) # for densities over time plots. 
library(knitr)
library(here)
library(gghighlight)

#workingDir <- "/home/luke/Documents/WORK/MINOS/"
workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r "imports", include=FALSE}
source(here::here('minos', 'utils_datain.R'))
source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

```{r, include=FALSE}
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 'baseline', 
                                    drop.dead = TRUE,
                                    drop.zero.weight = TRUE)
int.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionRELATIVE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)
int2.dat <- read_singular_local_out(out.path, 'ChildPovertyReductionABSOLUTE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)

#base.dat <- read_agg_subset_batch_out(out.path, 'baseline', drop.dead = TRUE)
#int1.dat <- read_agg_subset_batch_out(out.path, 'ChildPovertyReduction', drop.dead = TRUE)
```

# Relative Poverty Reduction

The first intervention we will focus on is a reduction in the prevalence of children in relative poverty. In these interventions, we have a target level of child poverty to attain (10% of children in a household in poverty) by a specified target year, which is 2030 unless stated otherwise. To achieve this, we find households with a hh_income under the poverty line, and uplift a number of households each year to linearly achieve the target proportion by the target year.

## Threshold

Relative poverty is defined as the proportion of children living in households with income below 60% of the population median *in a given year*.

We can calculate the relative poverty threshold for our starting year of 2021:

```{r}
base.2021 <- base.dat %>%
  filter(time == 2021)

mean.hh_income <- weighted.mean(base.2021$hh_income, w = base.2021$weight)
rel.pov.threshold <- mean.hh_income * 0.6

print(paste0("Relative poverty threshold in 2020: ", rel.pov.threshold))
rm(base.2021, mean.hh_income, rel.pov.threshold)
```

We can also look at how this evolves over time in the baseline simulation:

```{r}
threshold <- base.dat %>%
  select(time, weight, hh_income) %>%
  group_by(time) %>%
  summarise(poverty_threshold = (weighted.mean(hh_income, w = weight) * 0.6))

ggplot(data = threshold, aes(x = time, y = poverty_threshold)) +
  geom_col() +
  geom_hline(yintercept = threshold$poverty_threshold[threshold$time == 2020], linetype='dashed') +
  labs(title = 'Relative poverty threshold over time') +
  xlab('Year') +
  ylab('Threshold')
```

The threshold for relative poverty decreases over time in our simulation (it's worth noting that monetary values in our simulation have been adjusted for inflation to 2020 £s). This suggests that our population is getting slightly less wealthy over time on average.

## Intervention

In this intervention, we will randomly select households to uplift in each year until we uplift the number of children required for that year.

### Boosted Sample

We can also look at how many people are receiving the intervention over time, and the total and mean amounts of money being added to the system. This should slowly increase to 2030 and then hold fairly steady.

```{r}

hh_boosted <- int.dat %>%
    select(hidp, time, income_boosted, boost_amount, nkids) %>%
    filter(income_boosted == 'True') %>%
    group_by(time) %>%
    summarise(n_boosted = length(unique(hidp)),
              total_boost = sum(boost_amount),
              kids_boosted = sum(nkids))

sample_information <- function(int.dat, target.year) {
  hh_boosted <- int.dat %>%
    select(hidp, time, income_boosted, boost_amount, nkids) %>%
    filter(income_boosted == 'True') %>%
    group_by(time) %>%
    summarise(n_boosted = length(unique(hidp)),
              total_boost = sum(boost_amount),
              kids_boosted = sum(nkids))

  p1 <- ggplot(hh_boosted, aes(x = time, y = n_boosted)) +
    geom_line() +
    geom_vline(xintercept = target.year, linetype = 'dashed') +
    labs(title = 'Number of households receiving intervention') +
    xlab('Year') +
    ylab('Number Boosted')

  p2 <- ggplot(hh_boosted, aes(x = time, y = kids_boosted)) +
    geom_line() +
    geom_vline(xintercept = target.year, linetype = 'dashed') +
    labs(title = 'Number of children in households receiving intervention') +
    xlab('Year') +
    ylab('Number Children')

  p3 <- ggplot(hh_boosted, aes(x = time, y = total_boost)) +
    geom_line() +
    geom_vline(xintercept = target.year, linetype = 'dashed') +
    labs(title = 'Total boost amount') +
    xlab('Year') +
    ylab('Boost Amount')

  mean.boost <- int.dat %>%
    select(time, income_boosted, boost_amount) %>%
    group_by(time, income_boosted) %>%
    summarise(mean_boost = mean(boost_amount)) %>%
    filter(income_boosted == 'True')

  p4 <- ggplot(mean.boost, aes(x = time, y = mean_boost)) +
    geom_line() +
    geom_vline(xintercept = target.year, linetype = 'dashed') +
    labs(title = 'Mean boost amount') +
    xlab('Year') +
    ylab('Boost Amount')

  print(p1)
  print(p2)
  print(p3)
  print(p4)
}

sample_information(int.dat, target.year = 2030)
```

Now lets look at SF_12_MCS comparisons.

### SF_12_MCS

SF_12_MCS stands for the Mental Component Score (MCS) from the 12-item Short Form (SF) Survey. See [this paper](https://link.springer.com/referenceworkentry/10.1007/978-94-007-0753-5_2698) for more information about SF_12.

#### Full Population

First lets look at the effect of the intervention on the entire sample.

```{r}
full.sample.change.SF12 <- function(base.dat, int.dat) {
  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.dat %>%
    select(time, weight, SF_12) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

  combined <- merge(x = base.all.SF12, y = int.all.SF12, by = 'time')
  combined$diff <- combined$int.SF12 - combined$base.SF12

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}

full.sample.change.SF12(base.dat, int.dat)
```

#### Families with children

Now focused only on families with children.

```{r}
all.families.change.SF12 <- function(base.dat, int.dat) {
  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids) %>%
    filter(nkids > 0) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.dat %>%
    select(time, weight, SF_12, nkids) %>%
    filter(nkids > 0) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

  combined <- merge(x = base.all.SF12, y = int.all.SF12, by = 'time')
  combined$diff <- combined$int.SF12 - combined$base.SF12

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}
all.families.change.SF12(base.dat, int.dat)
```

#### Families w/ children under poverty line

Now only households with children under the poverty line.

```{r}
poverty.families.change.SF12 <- function(base.dat, int.dat, threshold) {
  base.thresh <- merge(base.dat, threshold, by = 'time')
  int.thresh <- merge(int.dat, threshold, by = 'time')

  base.all.SF12 <- base.thresh %>%
    select(time, weight, SF_12, nkids, hh_income, poverty_threshold) %>%
    filter(nkids > 0) %>%
    filter(hh_income < poverty_threshold) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.thresh %>%
    select(time, weight, SF_12, nkids, hh_income, poverty_threshold) %>%
    filter(nkids > 0) %>%
    filter(hh_income < poverty_threshold) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

  combined <- merge(x = base.all.SF12, y = int.all.SF12, by = 'time')
  combined$diff <- combined$int.SF12 - combined$base.SF12

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}

poverty.families.change.SF12(base.dat, int.dat, threshold)
```

This plot is perhaps a bit disingenuous, as the closer we get to 2030 the fewer households are actually under the poverty line. We might achieve what we want better by looking only at the boosted population.

#### Treatment on the Treated

```{r}
treatment.on.treated.change.SF12 <- function(base.dat, int.dat) {
  boosted.byyear <- int.dat %>%
    filter(income_boosted == 'True')

  base.all.SF12 <- base.dat %>%
    select(time, weight, SF_12, nkids, pidp) %>%
    filter(pidp %in% boosted.byyear$pidp) %>%
    group_by(time) %>%
    summarise(base.SF12 = weighted.mean(SF_12, w = weight))

  int.all.SF12 <- int.dat %>%
    select(time, weight, SF_12, nkids, pidp) %>%
    filter(pidp %in% boosted.byyear$pidp) %>%
    group_by(time) %>%
    summarise(int.SF12 = weighted.mean(SF_12, w = weight))

  combined <- merge(x = base.all.SF12, y = int.all.SF12, by = 'time')
  combined$diff <- combined$int.SF12 - combined$base.SF12

  ggplot(combined, aes(x = time, y = diff)) +
    geom_line() +
    geom_hline(yintercept = 0, linetype='dashed') +
    labs(title = 'Relative change in SF12') +
    xlab('Year') +
    ylab('Difference')
}

treatment.on.treated.change.SF12(base.dat, int.dat)
```

## Absolute Poverty

### Boosted Sample

```{r}
sample_information(int2.dat, target.year = 2030)
```

### SF_12_MCS

#### Full Population

```{r}
full.sample.change.SF12(base.dat, int2.dat)
```

#### Families with children

```{r}
all.families.change.SF12(base.dat, int.dat)
```

#### Families w/ children under poverty line

```{r}
poverty.families.change.SF12(base.dat, int.dat, threshold)
```

#### Treatment on the Treated

```{r}
treatment.on.treated.change.SF12(base.dat, int.dat)
```

# Diagnostics

## Spaghetti

```{r}
base.spag <- base.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'baseline')
int.spag <- int.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'relative')
int2.spag <- int2.dat %>%
  select(pidp, age, time, hh_income) %>%
  mutate(scen = 'absolute')

# spaghetti_plot(base.spag, 'hh_income',
#                save = FALSE,
#                save.path = '')
# 
# spaghetti_plot(int.spag, 'hh_income',
#                save = FALSE,
#                save.path = '')
# 
# spaghetti_plot(int2.spag, 'hh_income',
#                save = FALSE,
#                save.path = '')

density_ridges(base.spag, 'hh_income',
               save=FALSE, 
               save.path='')

density_ridges(int.spag, 'hh_income',
               save=FALSE, 
               save.path='')

density_ridges(int2.spag, 'hh_income',
               save=FALSE, 
               save.path='')
```

```{r}

dens.1 <- rbind(base.spag, int.spag)
dens.2 <- rbind(base.spag, int2.spag)

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int.dat, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Relative Poverty Intervention')

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = int2.dat, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Absolute Poverty Intervention')

```

```{r}

```