### 2024-05-22 minos/outcomes/CPR_Intervention_Diagnostics.Rmd
Improved the nutrition model scaling of predicted values, and some more changes to the model_definitions default file. Also added some more print statements for reporting on relative child pov intervention and added to the visualisation scripts

---
title: "Diagnosing Child Poverty Reduction Interventions"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

# SETUP

This document presents results from a host of interventions on child poverty, including visualisations of the change in SF_12_MCS compared with baseline.

```{r "setup", include=FALSE}

library(tidyverse)
library(ggplot2)
library(ggridges) # for densities over time plots. 
library(knitr)
library(here)
library(gghighlight)

workingDir <- "/home/luke/Documents/MINOS/Minos/"
#workingDir <- normalizePath('../')
knitr::opts_knit$set(root.dir = workingDir)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
rm(workingDir)
```

```{r "imports", include=FALSE}
source(here::here('minos', 'utils_datain.R'))
#source(here::here('minos', 'utils_validation_vis.R'))
source(here::here('minos', 'validation', 'utils.r'))
```

```{r, include=FALSE}
out.path <- here::here('output', 'default_config/')
base.dat <- read_singular_local_out(out.path, 
                                    'baseline', 
                                    drop.dead = TRUE, 
                                    drop.zero.weight = TRUE)
rel.int <- read_singular_local_out(out.path, 
                                   'ChildPovertyReductionRELATIVE', 
                                   drop.dead = TRUE,
                                   drop.zero.weight = TRUE)
# rel.int.2 <- read_singular_local_out(out.path,
#                                      'ChildPovertyReductionRELATIVE_2',
#                                      drop.dead = TRUE,
#                                      drop.zero.weight = TRUE)
# abs.int <- read_singular_local_out(out.path, 
#                                    'ChildPovertyReductionABSOLUTE', 
#                                    drop.dead = TRUE,
#                                    drop.zero.weight = TRUE)
```

# Household Income

i.e. is the intervention working?

## All Pop

```{r}
base.inc <- base.dat %>%
  select(pidp, hidp, time, hh_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(x = hh_income,w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Baseline')

rel.inc <- rel.int %>%
  select(pidp, hidp, time, hh_income, weight) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(x = hh_income,w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Relative')

inc <- rbind(base.inc, rel.inc)

ggplot(inc, aes(x = time, y = hh_income, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(base.dat, aes(x = hh_income, y = time, group = time, color = 'blue', alpha = 0.3)) +
  geom_density_ridges() +
  geom_density_ridges(data = rel.int, aes(x = hh_income, y = time, group = time, color = 'red')) +
  theme(legend.position = 'none') +
  labs(title = 'Relative Poverty Reduction Intervention - Red')
```

## Households With Children

```{r}
base.inc <- base.dat %>%
  select(pidp, hidp, time, hh_income, age, nkids, weight) %>%
  mutate(nkids_any = (nkids > 0)) %>%
  filter(nkids_any == TRUE) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(x = hh_income, w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Baseline')

rel.inc <- rel.int %>%
  select(pidp, hidp, time, hh_income, age, nkids, weight) %>%
  mutate(nkids_any = (nkids > 0)) %>%
  filter(nkids_any = TRUE) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(x = hh_income, w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Relative')

inc <- rbind(base.inc, rel.inc)

ggplot(inc, aes(x = time, y = hh_income, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()
```

## Boosted Pop

```{r}
boosted.pidps <- rel.int %>%
  filter(income_boosted == 'True') %>%
  select(pidp)

base.inc <- base.dat %>%
  select(pidp, hidp, time, hh_income, SF_12, weight) %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(x = hh_income, w = weight, na.rm = TRUE),
            SF_12 = weighted.mean(x = SF_12, w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Baseline')

rel.inc <- rel.int %>%
  select(pidp, hidp, time, hh_income, SF_12, weight) %>%
  filter(pidp %in% boosted.pidps$pidp) %>%
  group_by(time) %>%
  summarise(hh_income = weighted.mean(x = hh_income, w = weight, na.rm = TRUE),
            SF_12 = weighted.mean(x = SF_12, w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Relative')

inc <- rbind(base.inc, rel.inc)

ggplot(inc, aes(x = time, y = hh_income, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = SF_12, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()
```

# TESTING THE BOOSTED POP

```{r}
boosted.pidps <- rel.int %>%
  filter(income_boosted == 'True') %>%
  select(pidp, time)

# Define the recoding mapping for housing_quality
recoding_mapping <- c("Low" = 1, "Medium" = 2, "High" = 3)  # Adjust as needed

base.inc <- base.dat %>%
  select(pidp, hidp, time, age, hh_income, SF_12, housing_quality, neighbourhood_safety, ncigs, nutrition_quality, loneliness, nkids, weight) %>%
  inner_join(boosted.pidps, by = c("pidp", "time")) %>%
  mutate_at(vars(housing_quality), ~recode(., !!!recoding_mapping)) %>%
  group_by(time) %>%
  summarise(boosted_pop = n(),
            age = mean(age, na.rm = TRUE),
            hh_income = weighted.mean(x = hh_income, w = weight, na.rm = TRUE),
            SF_12 = weighted.mean(x = SF_12, w = weight, na.rm = TRUE),
            housing_quality = weighted.mean(x = housing_quality, w = weight, na.rm = TRUE),
            neighbourhood_safety = weighted.mean(x = neighbourhood_safety, w = weight, na.rm = TRUE),
            ncigs = weighted.mean(x = ncigs, w = weight, na.rm = TRUE),
            nutrition_quality = weighted.mean(x = nutrition_quality, w = weight, na.rm = TRUE),
            loneliness = weighted.mean(x = loneliness, w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Baseline')

rel.inc <- rel.int %>%
  select(pidp, hidp, time, age, hh_income, SF_12, housing_quality, neighbourhood_safety, ncigs, nutrition_quality, loneliness, nkids, weight) %>%
  inner_join(boosted.pidps, by = c("pidp", "time")) %>%
  mutate_at(vars(housing_quality), ~recode(., !!!recoding_mapping)) %>%
  group_by(time) %>%
  summarise(boosted_pop = n(),
            age = mean(age, na.rm = TRUE),
            hh_income = weighted.mean(x = hh_income, w = weight, na.rm = TRUE),
            SF_12 = weighted.mean(x = SF_12, w = weight, na.rm = TRUE),
            housing_quality = weighted.mean(x = housing_quality, w = weight, na.rm = TRUE),
            neighbourhood_safety = weighted.mean(x = neighbourhood_safety, w = weight, na.rm = TRUE),
            ncigs = weighted.mean(x = ncigs, w = weight, na.rm = TRUE),
            nutrition_quality = weighted.mean(x = nutrition_quality, w = weight, na.rm = TRUE),
            loneliness = weighted.mean(x = loneliness, w = weight, na.rm = TRUE)) %>%
  mutate(scen = 'Relative')

base.treated <- base.dat %>%
  select(pidp, time, hh_income, SF_12, nutrition_quality, ncigs) %>%
  inner_join(boosted.pidps, by = c("pidp", "time")) %>%
  mutate(scen = 'Baseline')

rel.treated <- rel.int %>%
  select(pidp, time, hh_income, SF_12, nutrition_quality, ncigs) %>%
  inner_join(boosted.pidps, by = c("pidp", "time")) %>%
  mutate(scen = 'Intervention')

inc <- rbind(base.inc, rel.inc)
treated <- rbind(base.treated, rel.treated)
```

```{r}
ggplot(inc, aes(x = time, y = boosted_pop, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = age, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = hh_income, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(base.treated, aes(x = hh_income, y = time, group = time, alpha = 0.3), color = 'blue') +
  geom_density_ridges() +
  geom_density_ridges(data = rel.treated, aes(x = hh_income, y = time, group = time), color = 'red') +
  theme(legend.position = 'none') +
  labs(title = 'Relative Poverty Reduction Intervention - Red')

ggplot(inc, aes(x = time, y = SF_12, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = housing_quality, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = neighbourhood_safety, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = ncigs, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = nutrition_quality, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

ggplot(inc, aes(x = time, y = loneliness, group = scen, color = scen, fill = scen)) +
  geom_line() +
  geom_point()

```

```{r}
ggplot(treated, aes(x = time, y = hh_income, group = interaction(time, scen), fill = scen)) +
  geom_boxplot()

ggplot(treated, aes(x = time, y = SF_12, group = interaction(time, scen), fill = scen)) +
  geom_boxplot()

ggplot(treated, aes(x = time, y = nutrition_quality, group = interaction(time, scen), fill = scen)) +
  geom_boxplot()

ggplot(treated, aes(x = time, y = ncigs, group = interaction(time, scen), fill = scen)) +
  geom_boxplot()
```

```{r}
base.housing <- base.dat %>%
  dplyr::select(pidp, time, housing_quality) %>%
  group_by(time, housing_quality) %>%
  count() %>%
  mutate(scen = 'Baseline')

rel.housing <- rel.int %>%
  dplyr::select(pidp, time, housing_quality) %>%
  group_by(time, housing_quality) %>%
  count() %>%
  mutate(scen = 'Intervention')

housing <- rbind(base.housing, rel.housing)
housing$housing_quality <- as.factor(housing$housing_quality)

housing.norm <- housing %>%
  group_by(time, scen) %>%
  mutate(total = sum(n)) %>%
  mutate(prct = (n / total))

# Plot the data
ggplot(data = housing.norm, mapping = aes(x = factor(time), y = prct, fill = housing_quality)) +
  geom_bar(stat = 'identity', position = 'stack') +
  facet_wrap(~ scen, nrow = 2) +
  labs(title = 'Housing Quality Over Time', x = 'Year', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data = housing.norm, mapping = aes(x = factor(time), y = prct, fill = housing_quality)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.9)) +
  facet_wrap(~ scen, nrow = 2) +
  labs(title = 'Housing Quality Over Time', x = 'Year', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Plot the data
ggplot(data = housing.norm, mapping = aes(x = factor(time), y = prct, fill = housing_quality)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.8), aes(group = scen)) +
  labs(title = 'Housing Quality Over Time', x = 'Year', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ scen, scales = "free_y", ncol = 2)

ggplot(data = housing.norm, mapping = aes(x = interaction(time, scen), y = prct, fill = housing_quality)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(title = 'Housing Quality Over Time', x = 'Year and Scenario', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Create a combined variable for interspersed plotting
housing.norm2 <- housing.norm %>%
  mutate(year_scen = interaction(time, scen, sep = " - "))

# Plot the data
ggplot(data = housing.norm2, mapping = aes(x = year_scen, y = prct, fill = housing_quality)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(title = 'Housing Quality Over Time', x = 'Year and Scenario', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Prepare data for line plot
housing.norm3 <- housing %>%
  group_by(time, scen, housing_quality) %>%
  summarize(prct = sum(n) / sum(housing$n)) %>%
  ungroup()

# Line plot with faceting by scenario
ggplot(data = housing.norm3, mapping = aes(x = time, y = prct, color = housing_quality, group = housing_quality)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  facet_wrap(~ scen) +
  labs(title = 'Housing Quality Over Time by Scenario', x = 'Year', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Generate proportions
housing.norm5 <- housing %>%
  group_by(time, scen) %>%
  mutate(total = sum(n)) %>%
  mutate(prct = (n / total)) %>%
  ungroup()

# Create a combined variable for interspersed plotting
housing.norm5 <- housing.norm5 %>%
  mutate(year_scen = interaction(time, scen, sep = " - "))

# Plot the data with different color schemes
ggplot(data = housing.norm5, mapping = aes(x = factor(time), y = prct, fill = housing_quality, group = scen)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.9)) +
  facet_wrap(~ scen, ncol = 1) +
  labs(title = 'Housing Quality Over Time', x = 'Year', y = 'Proportion') +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

```