### 2024-08-01 minos/transitions/energy_price_forecasting.Rmd
saving scatterplot for gain vs cost

---
title: "Energy Price Forecasting."
output: 'html_document'
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
#library(reticulate)
#use_condaenv("minos_condagcc")  
# package preamble.
library(forecast)
library(ggfortify)
library(readxl)
library(tidyr)
library(here)
source(here::here("minos", "transitions", "utils.R"))
```

```{r}
process_input_data <- function(dataset, data.column, new.name){

  subvector <- dataset[, c(data.column)]
  subvector <- drop_na(subvector)
  colnames(subvector) <- c(new.name)
  return(subvector)
}
#energy_data <- readxl::read_xlsx("../data/energy_prices_simpler.xlsx")
## primary energy material sources. fuel and motor oil omitted due to no UKHLS data. 
#electricity_vector <- process_input_data(energy_data, "Real terms price indices: Electricity [Note 3]", 'electricity')
#gas_vector <- process_input_data(energy_data, "Real terms price indices: Gas [Note 3]", "gas")
#solid_fuel_vector <- process_input_data(energy_data, "Real terms price indices: Solid fuels [Note 3]", "solid_fuel")
#liquid_fuel_vector <- process_input_data(energy_data, "Real terms price indices: Liquid fuels [Note 3]", "liquid_fuel")

#https://www.gov.uk/government/statistical-data-sets/monthly-domestic-energy-price-stastics
energy_data <- readxl::read_xlsx(here::here("..", "data", "energy_prices_simpler_yearly.xlsx"))
# primary energy material sources. fuel and motor oil omitted due to no UKHLS data. 
electricity_vector <- process_input_data(energy_data, "Real price indices: Electricity [Note 3]", 'electricity')
gas_vector <- process_input_data(energy_data, "Real price indices: Gas [Note 3]", "gas")
solid_fuel_vector <- process_input_data(energy_data, "Real price indices: Solid fuels [Note 3]", "solid_fuel")
liquid_fuel_vector <- process_input_data(energy_data, "Real price indices: Liquid fuels [Note 3]", "liquid_fuel")
# aggregate of the above 4. 
#domestic_fuel_vector <- process_input_data(energy_data, "Real terms price indices: Domestic fuels\r\n[Note 1, 3]", "domestic_fuel")
plot(electricity_vector$electricity, type='l', xlab="Quarter (0 = Q1 1990", ylab=c("Real Electricity Price indices"), main="Electricity")
plot(gas_vector$gas, type='l', xlab="Quarter (0 = Q1 1990", ylab=c("Real Gas Price indices"), main="Gas")
plot(solid_fuel_vector$solid_fuel, type='l', xlab="Quarter (0 = Q1 1990", ylab=c("Real Solid Fuel Price indices"), main="Solid Fuel (Coal)")
plot(liquid_fuel_vector$liquid_fuel, type='l', xlab="Quarter (0 = Q1 1990", ylab=c("Real Liquid Fuel Price indices"), main="Liquid Fuel (Petroleum, Diesel, LPG, Heating Oil)")
#plot(domestic_fuel_vector$domestic_fuel, type='l', xlab="Quarter (0 = Q1 1990", ylab=c("Real Demostrict Fuel Price indices"), main="Domestic Fuel #(Heating Oil)")
```

```{r}
electric.series <- electricity_vector$electricity #+ rnorm(length(electric.series))
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(electric.series)
acf(diff(electric.series))
pacf(electric.series)
#arima.electric 
#m1 <- auto.arima(electric.series)
m1 <- arima(electric.series, order=c(7, 0, 1))
print(m1)
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m1))
autoplot(forecast(m1, 40))
checkresiduals(m1)
print(m1$aic)

pdf(here::here("plots", "electric_forecast.pdf"))
autoplot(forecast(m1, 15))
dev.off()
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
gas.series <- gas_vector$gas
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(gas.series)
acf(diff(gas.series))
pacf(gas.series)
#arima.electric 
m2 <- arima(gas.series, order=c(4, 0 ,1))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m2))
autoplot(forecast(m2, 15))
checkresiduals(m2)
print(m2$aic)

pdf(here::here("plots", "gas_forecast.pdf"))
autoplot(forecast(m2, 15))
dev.off()
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
solid.series <- solid_fuel_vector$solid_fuel
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (2, 1, 4) in ARIMA.
acf(solid.series)
acf(diff(solid.series))
pacf(solid.series)
#arima.electric 
m3 <- arima(solid.series, order=c(5, 1 ,5))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m3))
autoplot(forecast(m3, 15))
checkresiduals(m3)
print(m3$aic)
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals.

pdf(here::here("plots", "solid_forecast.pdf"))
autoplot(forecast(m3, 15))
dev.off()
```

```{r}
liquid.series <- liquid_fuel_vector$liquid_fuel
# ACF, difference 1 ACF and PACF suggests AR(2)  difference 1 and MA(4) structure (5, 0, 1) in ARIMA.
acf(liquid.series)
acf(diff(liquid.series))
pacf(liquid.series)
#arima.electric 
m4 <- arima(gas.series, order=c(5, 0 ,1))
#Box.test(residuals(m1), type="Ljung-Box")
plot(residuals(m4))
autoplot(forecast(m4,15))
checkresiduals(m4)
print(m4$aic)

pdf(here::here("plots", "oil_forecast.pdf"))
autoplot(forecast(m4, 15))
dev.off()
# residuals appear to be in nominal range. large LB p-value also support no autocorrelation between residuals. 
```

```{r}
# saving data to R objects for use in MINOS
create.if.not.exists(here::here("data", "transitions", "energy_prices"))
write.csv(electric.series, here::here("data", "transitions", "energy_prices", "solid_historic.csv"))
write.csv(gas.series, here::here("data", "transitions", "energy_prices", "gas_historic.csv"))
write.csv(solid.series, here::here("data", "transitions", "energy_prices", "solid_historic.csv"))
write.csv(liquid.series, here::here("data", "transitions", "energy_prices", "liquid_historic.csv"))
write.csv(forecast(m1, 15), here::here("data", "transitions", "energy_prices", "electric_arima.csv"))
write.csv(forecast(m2, 15), here::here("data", "transitions", "energy_prices", "gas_arima.csv"))
write.csv(forecast(m3, 15), here::here("data", "transitions", "energy_prices", "solid_arima.csv"))
write.csv(forecast(m4, 15), here::here("data", "transitions", "energy_prices", "liquid_arima.csv"))
```