### 2024-08-07 minos/outcomes/MHF_SCP_visualisations.Rmd
Improved plots, changed lots of labels and ensured consistent colour scheme across different plots for MHF results

---
title: "MHF SCP Experiment Visualisations"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r "setup", include=FALSE}

library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(stringr)
library(grid)

workingDir <- "/home/luke/Documents/WORK/MINOS/SCP_ARC_SUMMARIES/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

```{r "data", include=FALSE}
# whole pop to start
# base.whole <- read.csv(file = sprintf('%s/combined_summary_whole_pop.csv', 'baseline'))
# base.UC <- read.csv(file = sprintf('%s/combined_summary_UC.csv', 'baseline'))
```

```{r}
scen.colour.list <- c('baseline', '£10', '£20',
               '£25', '£30',
               '£35', '£40', '£45',
               '£50', '£60', '£70',
               '£80', '£90', '£100')

#colours <- display.brewer.all(length(scen.list.full))
#colours <- colorRampPalette( brewer.pal(11,"Spectral") )(14)
colours <- colorRampPalette( brewer.pal(9,"Set1") )(14)
my_palette <- set_names(colours, scen.colour.list)

scen.list.full <- c('baseline', '10UniversalCredit', '20UniversalCredit',
               '25UniversalCredit', '30UniversalCredit',
               '35UniversalCredit', '40UniversalCredit', '45UniversalCredit',
               '50UniversalCredit', '60UniversalCredit', '70UniversalCredit',
               '80UniversalCredit', '90UniversalCredit', '100UniversalCredit')

scen.list.50l <- c('baseline', '10UniversalCredit',
               '20UniversalCredit', '25UniversalCredit', '30UniversalCredit',
               '35UniversalCredit', '40UniversalCredit', '45UniversalCredit',
               '50UniversalCredit')

scen.list.10_50 <- c('baseline', '10UniversalCredit',
               '20UniversalCredit', '30UniversalCredit',
               '40UniversalCredit', '50UniversalCredit')

scen.list.60_100 <- c('baseline', '60UniversalCredit', '70UniversalCredit',
               '80UniversalCredit', '90UniversalCredit', '100UniversalCredit')

scen.list.25_50 <- c('baseline', '25UniversalCredit', '30UniversalCredit',
                    '35UniversalCredit', '40UniversalCredit', '45UniversalCredit',
                    '50UniversalCredit')

scen.list.10_100 <- c('baseline', '10UniversalCredit', '20UniversalCredit',
                      '30UniversalCredit', '40UniversalCredit', '50UniversalCredit', 
                      '60UniversalCredit', '70UniversalCredit', '80UniversalCredit', 
                      '90UniversalCredit', '100UniversalCredit')

scen.list.20_100 <- c('baseline', '20UniversalCredit', '40UniversalCredit',
                      '60UniversalCredit', '80UniversalCredit', '100UniversalCredit')
```

# Functions

```{r "datain", include=FALSE}
generate_df_list <- function(summary, scen.list) {
  df.list <- lapply(scen.list, function(scen) {
    df <- read.csv(file = sprintf('%s/combined_summary_%s.csv', scen, summary))
    df$scen <- scen
    return(df)
  })
  return(df.list)
}

```

```{r "cost_ratio_functions", include=FALSE}

SF_12_cost_ratio <- function(summary_type, scen_list, palette = my_palette) {

  data <- do.call(rbind, generate_df_list(summary_type, scen_list))

  data <- data %>%
    mutate_at("scen", str_replace, "UniversalCredit", "") %>%
    mutate(scen = if_else(scen == "baseline", scen, paste0("£", scen)))

  if (summary_type == "UC") {
    data <- data %>%
      filter(universal_credit == 1)
  }

  plot_data1 <- data %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"), 
                 names_to = c(".value", "scen"), 
                 names_sep = "\\.") %>%
    filter(scen != 'baseline')

  # Extract numeric part from 'scen' column then...
  scen_numeric <- as.numeric(gsub("[^0-9]", "", plot_data1$scen))
  # ...Reorder 'scen' column based on numeric values for plotting
  plot_data1$scen <- factor(plot_data1$scen, levels = unique(plot_data1$scen[order(scen_numeric)]))

  plot_data1_confint <- plot_data1 %>%
    group_by(year, scen) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data2 <- plot_data1 %>%
    mutate(SF12_to_cost_ratio = diff_SF_12 / mean_cost) %>%
    replace_na(list(SF12_to_cost_ratio = 0))

  plot_data2_confint <- plot_data2 %>%
    group_by(year, scen) %>%
    summarise(n = n(),
              mean_SF12_to_cost_ratio = mean(SF12_to_cost_ratio),
              margin = qt(0.975, df = n - 1) * (sd(SF12_to_cost_ratio) / sqrt(n)),
              lower = mean_SF12_to_cost_ratio - margin,
              upper = mean_SF12_to_cost_ratio + margin)

  # p1 <- ggplot(plot_data1, aes(x = year, y = diff_SF_12, group = scen, colour = scen, fill = scen)) +
  #   geom_smooth()
  p1 <- ggplot(plot_data1_confint, aes(x = year, y = mean_SF12_change, group = scen)) +
    geom_ribbon(aes(ymin = lower, ymax = upper, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = scen), linetype = 'dashed') +
    geom_point(aes(color = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    xlab('Year') +
    ylab('Mean SF-12 Change') +
    theme(
        legend.title = element_blank(),  # Remove the legend title
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(
        color = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        ),
        fill = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        )
    )

  # if (length(scen_list) > 5) {
  #   p1 <- ggplot(plot_data1_confint, aes(x = year, y = mean_SF12_change, group = scen)) +
  #     geom_ribbon(aes(ymin = lower, ymax = upper, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
  #     geom_line(aes(color = scen), linetype = 'dashed') +
  #     geom_point(aes(color = scen)) +
  #     scale_color_manual(values = palette) +
  #     scale_fill_manual(values = palette) +
  #     xlab('Year') +
  #     ylab('Mean SF-12 Change') +
  #     theme(legend.title = element_blank(),
  #           legend.position = c(0.2, 0.8),
  #           legend.text = element_text(size=5),
  #           legend.key.size = unit(1, 'cm')) +
  #     guides(fill=guide_legend(ncol=2))
  # }

  #          legend.background = element_rect(fill = "white")

  # p2 <- ggplot(plot_data2, aes(x = year, y = SF12_to_cost_ratio, group = scen, colour = scen, fill = scen)) +
  #   geom_smooth()

  p2 <- ggplot(plot_data2_confint, aes(x = year, y = mean_SF12_to_cost_ratio, group = scen)) +
    geom_ribbon(aes(ymin = lower, ymax = upper, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = scen), linetype = 'dashed') +
    geom_point(aes(color = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    xlab('Year') +
    ylab('Mean SF-12 Change') +
    theme(
        legend.title = element_blank(),  # Remove the legend title
        legend.position = c(0.2, 0.7),  # Position the legend inside the plot
        legend.text = element_text(size = 8),  # Increase the text size for better readability
        legend.key.size = unit(0.8, 'cm'),  # Reduce the key size to fit more entries
        legend.spacing.y = unit(0.1, 'cm')  # Add vertical space between legend items
    ) +
    guides(
        color = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        ),
        fill = guide_legend(
            ncol = ifelse(length(unique(plot_data1_confint$scen)) > 5, 2, 1)  # Use 2 columns if more than 5 groups
        )
    )

  # if (length(scen_list) > 5) {
  #   p2 <- ggplot(plot_data2_confint, aes(x = year, y = mean_SF12_to_cost_ratio, group = scen)) +
  #     geom_ribbon(aes(ymin = lower, ymax = upper, color = scen, fill = scen), linetype = 'dotted', alpha = 0.3) +
  #     geom_line(aes(color = scen), linetype = 'dashed') +
  #     geom_point(aes(color = scen)) +
  #     scale_color_manual(values = palette) +
  #     scale_fill_manual(values = palette) +
  #     xlab('Year') +
  #     ylab('Mean SF-12 Change') +
  #     theme(legend.title = element_blank(),
  #           legend.position = c(0.2, 0.8),
  #           legend.text = element_text(size=5),
  #           legend.key.size = unit(1, 'cm')) +
  #     guides(fill=guide_legend(ncol=2))
  # }

  print(p1)
  print(p2)
}

SF_12_cost_ratio_cumul <- function(summary_type, scen_list, palette = my_palette) {

  data <- do.call(rbind, generate_df_list(summary_type, scen_list))

  data <- data %>%
    mutate_at("scen", str_replace, "UniversalCredit", "") %>%
    mutate(scen = if_else(scen == "baseline", scen, paste0("£", scen)))

  if (summary_type == "UC") {
    data <- data %>%
      filter(universal_credit == 1)
  }

  plot_data1 <- data %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"), 
                 names_to = c(".value", "scen"), 
                 names_sep = "\\.") %>%
    filter(scen != 'baseline')

  # Extract numeric part from 'scen' column then...
  scen_numeric <- as.numeric(gsub("[^0-9]", "", plot_data1$scen))
  # ...Reorder 'scen' column based on numeric values for plotting
  plot_data1$scen <- factor(plot_data1$scen, levels = unique(plot_data1$scen[order(scen_numeric)]))

  # Now we need to calculate cumulative stuff for the ratio
  cumulative_SF12 <- plot_data1 %>%
    group_by(scen, run_id) %>%
    summarise(SF_12 = sum(diff_SF_12),
              cost = sum(mean_cost))

  cumulative_SF12_ratio <- cumulative_SF12 %>%
    mutate(SF12_to_cost_ratio = SF_12 / cost) %>%
    replace_na(list(SF12_to_cost_ratio = 0))

  cumulative_SF12_ratio_confint <- cumulative_SF12_ratio %>%
    group_by(scen) %>%
    summarise(n = n(),
              mean_SF12_to_cost_ratio = mean(SF12_to_cost_ratio),
              margin = qt(0.975, df = n - 1) * (sd(SF12_to_cost_ratio) / sqrt(n)),
              lower = mean_SF12_to_cost_ratio - margin,
              upper = mean_SF12_to_cost_ratio + margin)

  p1 <- ggplot(cumulative_SF12_ratio_confint, aes(x = scen, y = mean_SF12_to_cost_ratio)) +
    geom_col(aes(color = scen, fill = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    xlab('Intervention') +
    ylab('Ratio') +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
          legend.position = 'none')

  p2 <- ggplot(cumulative_SF12_ratio_confint, aes(x = scen, y = mean_SF12_to_cost_ratio)) +
    geom_col(aes(color = scen, fill = scen)) +
    scale_color_manual(values = palette) +
    scale_fill_manual(values = palette) +
    geom_errorbar(aes(ymin = lower, ymax = upper), position = 'dodge', width = 0.25) +
    xlab('Intervention') +
    ylab('Ratio') +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
          legend.position = 'none')

  print(p1)
  print(p2)
}
```

```{r "rel_abs_functions", include=FALSE}

compare_relative_absolute <- function(comparison_scen, kids = FALSE) {

  scen.list <- c('baseline', comparison_scen)

  if (kids) {
    UC.rel <- do.call(rbind, generate_df_list('UC_kids_rel_pov', scen.list))
    UC.abs <- do.call(rbind, generate_df_list('UC_kids_abs_pov', scen.list))
  } else {
    UC.rel <- do.call(rbind, generate_df_list('UC_rel_pov', scen.list))
    UC.abs <- do.call(rbind, generate_df_list('UC_abs_pov', scen.list))
  }

  UC.rel <- UC.rel %>%
    mutate_at("scen", str_replace, "UniversalCredit", "") %>%
    mutate(scen = if_else(scen == "baseline", scen, paste0("£", scen)))

  UC.abs <- UC.abs %>%
    mutate_at("scen", str_replace, "UniversalCredit", "") %>%
    mutate(scen = if_else(scen == "baseline", scen, paste0("£", scen)))

  rel.change <- UC.rel %>%
    filter(universal_credit == 1,
           relative_poverty == 1) %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    rename(RelativePoverty = diff_SF_12) %>%
    select(run_id, year, RelativePoverty)

  abs.change <- UC.abs %>%
    filter(universal_credit == 1,
           absolute_poverty == 1) %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    rename(AbsolutePoverty = diff_SF_12) %>%
    select(run_id, year, AbsolutePoverty)

  combined <- merge(rel.change, abs.change, by = c('run_id', 'year'))

  combined_confint <- combined %>%
    pivot_longer(cols = RelativePoverty:AbsolutePoverty,
                 names_to = 'subpop',
                 values_to = 'SF_12_change') %>%
      group_by(subpop, year) %>%
      summarise(n = n(),
                mean_SF12_change = mean(SF_12_change),
                margin = qt(0.975, df = n - 1) * (sd(SF_12_change) / sqrt(n)),
                lower = mean_SF12_change - margin,
                upper = mean_SF12_change + margin)

  comparison_scen <- gsub("[a-zA-Z ]", "", comparison_scen)
  comparison_scen <- paste0("£", comparison_scen)

  p1 <- ggplot(combined_confint, aes(x = year, y = mean_SF12_change, group = subpop, colour = subpop, fill = subpop)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = subpop), linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') + 
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('Mean SF-12 Change')

  p2 <- ggplot(combined_confint, aes(x = year, y = mean_SF12_change, group = subpop, colour = subpop, fill = subpop)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('Mean SF-12 Change')

  print(p1)
  print(p2)
}

compare_init_relative_absolute <- function(comparison_scen, kids = FALSE) {

  scen.list <- c('baseline', comparison_scen)

  if (kids) {
    UC.rel <- do.call(rbind, generate_df_list('UC_kids_init_rel_pov', scen.list))
    UC.abs <- do.call(rbind, generate_df_list('UC_kids_init_abs_pov', scen.list))
  } else {
    UC.rel <- do.call(rbind, generate_df_list('UC_init_rel_pov', scen.list))
    UC.abs <- do.call(rbind, generate_df_list('UC_init_abs_pov', scen.list))
  }

  UC.rel <- UC.rel %>%
    mutate_at("scen", str_replace, "UniversalCredit", "") %>%
    mutate(scen = if_else(scen == "baseline", scen, paste0("£", scen)))

  UC.abs <- UC.abs %>%
    mutate_at("scen", str_replace, "UniversalCredit", "") %>%
    mutate(scen = if_else(scen == "baseline", scen, paste0("£", scen)))

  rel.change <- UC.rel %>%
    filter(universal_credit == 1,
           init_relative_poverty == 1) %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    rename(RelativePoverty = diff_SF_12) %>%
    select(run_id, year, RelativePoverty)

  abs.change <- UC.abs %>%
    filter(universal_credit == 1,
           init_absolute_poverty == 1) %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    rename(AbsolutePoverty = diff_SF_12) %>%
    select(run_id, year, AbsolutePoverty)

  combined <- merge(rel.change, abs.change, by = c('run_id', 'year'))

  combined_confint <- combined %>%
    pivot_longer(cols = RelativePoverty:AbsolutePoverty,
                 names_to = 'subpop',
                 values_to = 'SF_12_change') %>%
      group_by(subpop, year) %>%
      summarise(n = n(),
                mean_SF12_change = mean(SF_12_change),
                margin = qt(0.975, df = n - 1) * (sd(SF_12_change) / sqrt(n)),
                lower = mean_SF12_change - margin,
                upper = mean_SF12_change + margin)

  comparison_scen <- gsub("[a-zA-Z ]", "", comparison_scen)
  comparison_scen <- paste0("£", comparison_scen)

  p1 <- ggplot(combined_confint, aes(x = year, y = mean_SF12_change, group = subpop, colour = subpop, fill = subpop)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = subpop), linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('Mean SF-12 Change')

  p2 <- ggplot(combined_confint, aes(x = year, y = mean_SF12_change, group = subpop, colour = subpop, fill = subpop)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('Mean SF-12 Change')

  print(p1)
  print(p2)
}

```

```{r "gender_function", include=FALSE}
compare_gender <- function(comparison_scen) {
  scen.list <- c('baseline', comparison_scen)
  UC.gender <- do.call(rbind, generate_df_list('UC_gender', scen.list))

  gender.change <- UC.gender %>%
    filter(universal_credit == 1) %>%
    select(run_id, sex, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(run_id, sex, year, diff_SF_12)

  gender.confint <- gender.change %>%
    group_by(sex, year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  comparison_scen <- gsub("[a-zA-Z ]", "", comparison_scen)
  comparison_scen <- paste0("£", comparison_scen)

  p1 <- ggplot(data = gender.confint, aes(x = year, y = mean_SF12_change, group = sex, colour = sex, fill = sex)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(aes(color = sex), linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
    scale_color_manual(values = c("#E69F00", "#56B4E9")) +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('SF-12 Change')

  p2 <- ggplot(data = gender.confint, aes(x = year, y = mean_SF12_change, group = sex, colour = sex, fill = sex)) +
    geom_smooth() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(subtitle = comparison_scen) +
    xlab('Year') +
    ylab('SF-12 Change')

  print(p1)
  #print(p2)
}
```

# Whole Population

```{r}
whole.pop.list <- generate_df_list('whole_pop', scen.list.full)
whole.pop <- do.call(rbind, whole.pop.list)
```

```{r}
whole.pop.sum <- whole.pop %>%
  group_by(scen, year) %>%
  summarise(hh_income = mean(hh_income),
            SF_12 = mean(SF_12))

ggplot(whole.pop.sum, aes(x = year, y = SF_12, group = scen, colour = scen, fill = scen)) +
  geom_line()

ggplot(whole.pop, aes(x = year, y = SF_12, group = scen, colour = scen, fill = scen)) +
  geom_smooth()

rm(whole.pop.sum, whole.pop.list, whole.pop)
```

# Families

```{r}
fam.list <- generate_df_list('families', scen.list.full)
families <- do.call(rbind, fam.list)
```

```{r}
fam.sum <- families %>%
  group_by(scen, year) %>%
  summarise(hh_income = mean(hh_income),
            SF_12 = mean(SF_12))

ggplot(fam.sum, aes(x = year, y = SF_12, group = scen, colour = scen, fill = scen)) +
  geom_line()

ggplot(families, aes(x = year, y = SF_12, group = scen, colour = scen, fill = scen)) +
  geom_smooth()

rm(fam.sum, fam.list, families)
```

```{r}
fam.2550 <- do.call(rbind, generate_df_list('families', scen.list.25_50))

ggplot(fam.2550, aes(x = year, y = SF_12, group = scen, colour = scen, fill = scen)) +
  geom_smooth()

rm(fam.2550)
```

# Improvement Ratios

```{r}
# test <- do.call(rbind, generate_df_list('families', scen.list.10_50))
# test2 <- do.call(rbind, generate_df_list('UC', scen.list.10_50))
# SF_12_cost_ratio(summary_type = 'families', scen_list = scen.list.10_50)
# SF_12_cost_ratio_cumul(summary_type = 'families', scen_list = scen.list.10_50)
```

## Cost Ratio: £25 - £50

### UC

```{r}
SF_12_cost_ratio(summary_type = 'UC', scen_list = scen.list.25_50)
SF_12_cost_ratio_cumul(summary_type = 'UC', scen_list = scen.list.25_50)
```

## Cost Ratio: £10 - £50

### UC

```{r}
SF_12_cost_ratio(summary_type = 'UC', scen_list = scen.list.10_50)
SF_12_cost_ratio_cumul(summary_type = 'UC', scen_list = scen.list.10_50)
```

## Cost Ratio: £60 - £100

### UC

```{r}
SF_12_cost_ratio(summary_type = 'UC', scen_list = scen.list.60_100)
SF_12_cost_ratio_cumul(summary_type = 'UC', scen_list = scen.list.60_100)
```

## Cost Ratio: £10 - £100

### UC

```{r}
SF_12_cost_ratio(summary_type = 'UC', scen_list = scen.list.10_100)
SF_12_cost_ratio_cumul(summary_type = 'UC', scen_list = scen.list.10_100)
```

# Compare Different Populations

## Relative and Absolute Poverty

### Whole Pop

```{r}
compare_relative_absolute(comparison_scen = '25UniversalCredit')
compare_relative_absolute(comparison_scen = '50UniversalCredit')
compare_relative_absolute(comparison_scen = '70UniversalCredit')
compare_relative_absolute(comparison_scen = '100UniversalCredit')
```

### Families with Children

```{r}
compare_relative_absolute(comparison_scen = '25UniversalCredit', kids=TRUE)
compare_relative_absolute(comparison_scen = '50UniversalCredit', kids=TRUE)
compare_relative_absolute(comparison_scen = '70UniversalCredit', kids=TRUE)
compare_relative_absolute(comparison_scen = '100UniversalCredit', kids=TRUE)
```

## INITIAL Relative and Absolute Poverty

### Whole Pop

```{r}
compare_init_relative_absolute(comparison_scen = '25UniversalCredit')
compare_init_relative_absolute(comparison_scen = '50UniversalCredit')
compare_init_relative_absolute(comparison_scen = '70UniversalCredit')
compare_init_relative_absolute(comparison_scen = '100UniversalCredit')
```

### Families with Children

```{r}
compare_init_relative_absolute(comparison_scen = '25UniversalCredit', kids=TRUE)
compare_init_relative_absolute(comparison_scen = '50UniversalCredit', kids=TRUE)
compare_init_relative_absolute(comparison_scen = '70UniversalCredit', kids=TRUE)
compare_init_relative_absolute(comparison_scen = '100UniversalCredit', kids=TRUE)
```

## Gender

```{r}
# scen.list.50 <- c('baseline', '50UniversalCredit')
# UC.gender <- do.call(rbind, generate_df_list('UC_gender', scen.list.50))
```

```{r}
compare_gender(comparison_scen = '25UniversalCredit')
compare_gender(comparison_scen = '50UniversalCredit')
compare_gender(comparison_scen = '70UniversalCredit')
compare_gender(comparison_scen = '100UniversalCredit')
```

```{r}

```

# OLD OR UNWANTED

## Cost Ratio: £10 - £50

### Whole Pop

```{r}
# SF_12_cost_ratio(summary_type = 'whole_pop', scen_list = scen.list.10_50)
# SF_12_cost_ratio_cumul(summary_type = 'whole_pop', scen_list = scen.list.10_50)
```

### Families

```{r}
# SF_12_cost_ratio(summary_type = 'families', scen_list = scen.list.10_50)
# SF_12_cost_ratio_cumul(summary_type = 'families', scen_list = scen.list.10_50)
```

## Cost Ratio: £60 - £100

### Whole Pop

```{r}
# SF_12_cost_ratio(summary_type = 'whole_pop', scen_list = scen.list.60_100)
# SF_12_cost_ratio_cumul(summary_type = 'whole_pop', scen_list = scen.list.60_100)
```

### Families

```{r}
# SF_12_cost_ratio(summary_type = 'families', scen_list = scen.list.60_100)
# SF_12_cost_ratio_cumul(summary_type = 'families', scen_list = scen.list.60_100)
```

## Cost Ratio: £10 - £100

### Whole Pop

```{r}
# SF_12_cost_ratio(summary_type = 'whole_pop', scen_list = scen.list.10_100)
# SF_12_cost_ratio_cumul(summary_type = 'whole_pop', scen_list = scen.list.10_100)
```

### Families

```{r}
# SF_12_cost_ratio(summary_type = 'families', scen_list = scen.list.10_100)
# SF_12_cost_ratio_cumul(summary_type = 'families', scen_list = scen.list.10_100)
```