### 2024-09-09 minos/outcomes/spatial_analysis.Rmd
modules fix for outgoins

---
title: "R Notebook"
output: html_notebook
---

spatial analysis notebook for chapter.

msoa/lsoa/la level maps for GBIS MCS/PCS/QALY.

regressions
ICER plots
disaggregation of good performing LSOAs/MSOAs.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.