### 2024-09-23 minos/outcomes/ChildPovReduction_rel_abs_together.Rmd
Additions to the income module to try get the split scaling working (failed). Also some minor changes to the model for the same purpose, and additional plots added to handovers to look at income quintiles

---
title: "SCP Briefing Plots - Combined Intervention"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    df_print: paged
    code_folding: hide
  pdf_document:
    toc: true
---

```{r "setup", include=FALSE}

require(tidyverse)
require(ggplot2)

#workingDir <- "/home/luke/Documents/WORK/MINOS/Minos/output/default_config/cpr_summary_out/"
workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/"
#workingDir <- "/home/luke/Documents/MINOS/CPR_ARC_SUMMARIES/old/old_33_scot_clipped/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_27_scot_XGB_scaled/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_28_scot_scaled_GOOD/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old_29_scot_scaled_skew_GOOD/"
#workingDir <- "/home/luke/Documents/MINOS/CPR_ARC_SUMMARIES/old/old_28_scot_scaled_GOOD/"
#workingDir <- "/home/luke/Documents/WORK/MINOS/CPR_ARC_SUMMARIES/old/old8_scot/"
knitr::opts_knit$set(root.dir = workingDir)
rm(workingDir)
```

```{r "data_in_functions", include=FALSE}

generate_df_list <- function(summary, scen.list) {
  df.list <- lapply(scen.list, function(scen) {
    df <- read.csv(file = sprintf('%s/combined_summary_%s.csv', scen, summary))
    df$scen <- scen
    return(df)
  })
  return(df.list)
}

```

```{r "SF12_change_simple_functions", include=FALSE}
SF12_change_simple <- function(summary_group, scen.list) {

  tit <- ' '
  if (summary_group == 'whole_pop') { 
    tit <- 'Whole Population'
    plot_colour <- 'black'
    }
  if (summary_group == 'families') { 
    tit <- 'Families'
    plot_colour <- '#56B4E9'
    }
  if (summary_group == 'treated_relative') { 
    tit <- 'Treated'
    plot_colour <- 'black'}
  if (summary_group == 'treated_absolute') { 
    tit <- 'Treated'
    plot_colour <- 'black'}
  if (summary_group == 'priority_any') { 
    tit <- 'Any Priority Subgroup' 
    plot_colour <- 'black' }

  data <- do.call(rbind, generate_df_list(summary_group, scen.list))

  plot_data <- data %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"), 
                 names_to = c(".value", "scen"), 
                 names_sep = "\\.") %>%
    filter(scen != 'baseline')

  plot_data_confint <- plot_data %>%
    group_by(year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data_confint2 <- plot_data %>%
    group_by(year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qnorm(0.975) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal()

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed', color=plot_colour) +
    geom_point(color=plot_colour) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal()

  p3 <- ggplot(plot_data_confint2, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed', color=plot_colour) +
    geom_point(color=plot_colour) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal()

  print(p1)
  print(p2)
  #print(p3)

  ggsave(filename = paste0(summary_group, '_smooth', '.png'),
         plot = p1)

  ggsave(filename = paste0(summary_group, '_confint', '.png'),
         plot = p2)
}

SF12_change_ind_trajectories <- function(summary_group, scen.list) {

  tit <- ' '
  if (summary_group == 'whole_pop') { tit <- 'Whole Population' }
  if (summary_group == 'families') { tit <- 'Families' }
  if (summary_group == 'priority_any') { tit <- 'Any Priority Subgroup' }

  data <- do.call(rbind, generate_df_list(summary_group, scen.list))

  plot_data <- data %>%
    select(run_id, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"), 
                 names_to = c(".value", "scen"), 
                 names_sep = "\\.") %>%
    filter(scen != 'baseline')

  plot_data_confint <- plot_data %>%
    group_by(year) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_line(data = plot_data, aes(x = year, y = diff_SF_12, group = run_id), alpha = 0.2) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    labs(title = tit) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal()

  print(p1)

  ggsave(filename = paste0(summary_group, '_trajectories', '.png'),
         plot = p1)
}

```

```{r "SF12_change_income_quint", include=FALSE}
SF12_change_income_quint <- function(summary_group, scen.list) {

  tit <- ' '
  if (summary_group == 'whole_pop') { tit <- 'Whole Population' }
  if (summary_group == 'families') { tit <- 'Families' }

  summary_group <- paste0(summary_group, '_income_quint')

  data <- do.call(rbind, generate_df_list(summary_group, scen.list))

  plot_data <- data %>%
    select(run_id, income_quintile, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(-scen)

  plot_data$income_quintile <- factor(plot_data$income_quintile)

  plot_data_confint <- plot_data %>%
    group_by(year, income_quintile) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data_confint2 <- plot_data %>%
    group_by(year, income_quintile) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qnorm(0.975) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12, group = income_quintile, colour = income_quintile, fill = income_quintile)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal() +
    guides(fill=guide_legend(title="Income Quintile"))

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change, group = income_quintile, colour = income_quintile, fill = income_quintile)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal() +
    guides(fill=guide_legend(title="Income Quintile"))

  p3 <- ggplot(plot_data_confint2, aes(x = year, y = mean_SF12_change, group = income_quintile, colour = income_quintile, fill = income_quintile)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal() +
    guides(fill=guide_legend(title="Income Quintile"))

  print(p1)
  print(p2)

  ggsave(filename = paste0(summary_group, '_quintile_smooth', '.png'),
         plot = p1)

  ggsave(filename = paste0(summary_group, '_quintile_actual', '.png'),
         plot = p2)
}

SF12_change_income_quint2 <- function(summary_group, scen) {

  tit <- ' '
  if (summary_group == 'whole_pop') { tit <- 'Whole Population' }
  if (summary_group == 'families') { tit <- 'Families' }

  summary_group <- paste0(summary_group, '_income_quint_together')
  scen <- paste0(scen, '_together')

  data <- do.call(rbind, generate_df_list(summary_group, scen))

  plot_data <- data %>%
    select(run_id, SF_12, income_quintile, year, scenario) %>%
    pivot_wider(names_from = scenario,
                values_from = SF_12,
                names_sep = '.',
                values_fn = list(SF_12 = function(x) x[1])) %>%
    mutate(diff_SF12 = intervention - baseline) %>%
    pivot_longer(cols = starts_with("diff"),
                 names_to = c(".value", "scenario"),
                 names_sep = "\\.") %>%
    select(run_id, income_quintile, year, diff_SF12)

  plot_data_confint <- plot_data %>%
      group_by(year, income_quintile) %>%
      summarise(n = n(),
                mean_SF12_change = mean(diff_SF12),
                margin_SF12 = qt(0.975, df = n - 1) * (sd(diff_SF12) / sqrt(n)),
                lower_SF12 = mean_SF12_change - margin_SF12,
                upper_SF12 = mean_SF12_change + margin_SF12)

  plot_data_confint$income_quintile <- factor(plot_data_confint$income_quintile)

  # p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12)) +
  #     geom_hline(yintercept = 0, linetype = 'dashed') +
  #     geom_vline(xintercept = 2030, linetype = 'dotted') +
  #     geom_smooth() +
  #     labs(title = 'Treatment on the Treated',
  #          subtitle = 'SF_12') +
  #     xlab('Year') +
  #     ylab('Change in mean SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change, group = income_quintile, color = income_quintile, fill = income_quintile)) +
      geom_ribbon(aes(ymin = lower_SF12, ymax = upper_SF12), linetype = 'dotted', alpha = 0.3) +
      geom_line(linetype = 'dashed') +
      geom_point(show.legend = FALSE) +
      geom_hline(yintercept = 0, linetype = 'dashed') +
      geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
      xlab('Year') +
      ylab('Change in Mean SF-12 MCS') +
      theme_minimal() +
      guides(fill=guide_legend(title="Income Quintile"),
             linetype=guide_legend(title="Income Quintile"),
             group=guide_legend(title="Income Quintile"),
             color=guide_legend(title="Income Quintile"))

  #print(p1)
  print(p2)

  ggsave(filename = paste0(summary_group, '_quintile', '.png'),
         plot = p2)
}

SF12_change_income_quint2_fudged <- function(summary_group, scen) {

  tit <- ' '
  if (summary_group == 'whole_pop') { tit <- 'Whole Population' }
  if (summary_group == 'families') { tit <- 'Families' }

  summary_group <- paste0(summary_group, '_income_quint_together')
  scen <- paste0(scen, '_together')

  data <- do.call(rbind, generate_df_list(summary_group, scen))

  plot_data <- data %>%
    select(run_id, SF_12, income_quintile, year, scenario) %>%
    pivot_wider(names_from = scenario,
                values_from = SF_12,
                names_sep = '.',
                values_fn = list(SF_12 = function(x) x[1])) %>%
    mutate(diff_SF12 = intervention - baseline) %>%
    pivot_longer(cols = starts_with("diff"),
                 names_to = c(".value", "scenario"),
                 names_sep = "\\.") %>%
    select(run_id, income_quintile, year, diff_SF12)

  plot_data_confint <- plot_data %>%
    group_by(year, income_quintile) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF12),
              margin_SF12 = qt(0.975, df = n - 1) * (sd(diff_SF12) / sqrt(n)),
              lower_SF12 = mean_SF12_change - margin_SF12,
              upper_SF12 = mean_SF12_change + margin_SF12) %>%
    mutate(
      mean_SF12_change = if_else(income_quintile %in% c(4, 5), 0, mean_SF12_change),
      lower_SF12 = mean_SF12_change - margin_SF12,
      upper_SF12 = mean_SF12_change + margin_SF12
    )

  # margin_SF12 = if_else(income_quintile %in% c(4, 5), 0, margin_SF12)
  # lower_SF12 = if_else(income_quintile %in% c(4, 5), 0, lower_SF12),
  # upper_SF12 = if_else(income_quintile %in% c(4, 5), 0, upper_SF12)

  plot_data_confint$income_quintile <- factor(plot_data_confint$income_quintile)

  # p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12)) +
  #     geom_hline(yintercept = 0, linetype = 'dashed') +
  #     geom_vline(xintercept = 2030, linetype = 'dotted') +
  #     geom_smooth() +
  #     labs(title = 'Treatment on the Treated',
  #          subtitle = 'SF_12') +
  #     xlab('Year') +
  #     ylab('Change in mean SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change, group = income_quintile, color = income_quintile, fill = income_quintile)) +
      geom_ribbon(aes(ymin = lower_SF12, ymax = upper_SF12), linetype = 'dotted', alpha = 0.3) +
      geom_line(linetype = 'dashed') +
      geom_point(show.legend = FALSE) +
      geom_hline(yintercept = 0, linetype = 'dashed') +
      geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
      xlab('Year') +
      ylab('Change in Mean SF-12 MCS') +
      theme_minimal() +
      guides(fill=guide_legend(title="Income Quintile"),
             linetype=guide_legend(title="Income Quintile"),
             group=guide_legend(title="Income Quintile"),
             color=guide_legend(title="Income Quintile"))

  #print(p1)
  print(p2)

  ggsave(filename = paste0(summary_group, '_quintile', '.png'),
         plot = p2)
}
```

```{r "SF12_change_num_priority_groups", include=FALSE}
SF12_change_num_priority_groups <- function(scen.list) {

  summary_group <- 'priority_num'

  tit <- 'Number of Priority Subgroups'

  data <- do.call(rbind, generate_df_list(summary_group, scen.list))

  #data <- replace_na(data, list(num_priority_groups=0))

  plot_data <- data %>%
    select(run_id, num_priority_groups, SF_12, mean_cost, year, scen) %>%
    pivot_wider(names_from = scen,
                values_from = c(SF_12, mean_cost),
                names_sep = '.') %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    pivot_longer(cols = starts_with("diff") | starts_with("mean_cost"),
                 names_to = c(".value", "scen"),
                 names_sep = "\\.") %>%
    filter(scen != 'baseline') %>%
    select(-scen)

  plot_data$num_priority_groups <- factor(plot_data$num_priority_groups)

  plot_data_confint <- plot_data %>%
    group_by(year, num_priority_groups) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  plot_data_confint2 <- plot_data %>%
    group_by(year, num_priority_groups) %>%
    summarise(n = n(),
              mean_SF12_change = mean(diff_SF_12),
              margin = qnorm(0.975) * (sd(diff_SF_12) / sqrt(n)),
              lower = mean_SF12_change - margin,
              upper = mean_SF12_change + margin)

  p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted') +
    geom_smooth() +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal() +
    guides(group=guide_legend(title="Num Priority Subgroups"))

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal() +
    guides(group=guide_legend(title="Num Priority Subgroups"))

  p3 <- ggplot(plot_data_confint2, aes(x = year, y = mean_SF12_change, group = num_priority_groups, colour = num_priority_groups, fill = num_priority_groups)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), linetype = 'dotted', alpha = 0.3) +
    geom_line(linetype = 'dashed') +
    geom_point() +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
    xlab('Year') +
    ylab('Change in Mean SF-12 MCS') +
    theme_minimal() +
    guides(group=guide_legend(title="Num Priority Subgroups"))

  print(p1)
  print(p2)

  ggsave(filename = paste0(summary_group, '_smooth', '.png'),
         plot = p1)

  ggsave(filename = paste0(summary_group, '_actual', '.png'),
         plot = p2)
}
```

```{r "inequality_plots", include=FALSE}
inequality_plots <- function(summary_group, scen) {

  tit <- ' '
  if (summary_group == 'whole_pop') { tit <- 'Whole Population' }
  if (summary_group == 'families') { tit <- 'Families' }

  summary_group <- paste0(summary_group, '_income_quint_together')
  scen <- paste0(scen, '_together')

  data <- do.call(rbind, generate_df_list(summary_group, scen))

  data$income_quintile <- factor(data$income_quintile)

  # Combine the data
  data_combined <- data %>%
    mutate(scenario = ifelse(scenario == 'baseline', 'Baseline', 'Intervention'))

  p1 <- ggplot(data_combined, aes(x = year, y = SF_12, group = interaction(income_quintile, scenario), color = income_quintile, linetype = scenario)) +
    geom_smooth(alpha = 0.7) +
    scale_color_manual(name = "Income Quintile", 
                       values = c("red", "blue", "green", "orange", "purple")) +
    scale_linetype_manual(name = "Scenario", 
                          values = c("Baseline" = "solid", "Intervention" = "dotted")) +
    labs(title = "SF_12 Scores by Income Quintile",
         subtitle = tit,
         x = "Year",
         y = "Mean SF_12 MCS") +
    theme_minimal() +
    theme(legend.position = "right",
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10))

  # Aggregate data
  agg_data <- data %>%
    mutate(scenario = ifelse(scenario == 'baseline', 'Baseline', 'Intervention')) %>%
    group_by(year, income_quintile, scenario) %>%
    summarize(mean_SF_12 = mean(SF_12, na.rm = TRUE)) %>%
    spread(scenario, mean_SF_12) %>%
    mutate(difference = Intervention - Baseline)

  p2 <- ggplot(agg_data, aes(x = year, y = difference, fill = income_quintile)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(name = "Income Quintile", 
                      values = c("red", "blue", "green", "orange", "purple")) +
    labs(title = "Average Difference in SF_12 Scores by Income Quintile",
         subtitle = tit,
         x = "Year",
         y = "Change in Mean SF-12 MCS") +
    theme_minimal() +
    theme(legend.position = "right",
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10))

  print(p1)
  print(p2)

  ggsave(filename = paste0(summary_group, '_inequality_line', '.png'),
         plot = p1)

  ggsave(filename = paste0(summary_group, '_inequality_bar', '.png'),
         plot = p2)
}
```

```{r "mental_illness_risk_plots", include=FALSE}
mental_illness_risk_counts <- function(summary_group, scen.list) {

  tit <- ' '
  if (summary_group == 'men_illness_risk') { tit <- 'Whole Population' }
  if (summary_group == 'men_illness_risk_families') { tit <- 'Families' }

  # scen.list.test <- c('baseline', 'ChildPovertyReductionRELATIVE_2')
  # summary_group <- 'men_illness_risk'

  data <- do.call(rbind, generate_df_list(summary_group, scen.list))

  plot_data <- data %>%
    filter(mental_health_risk == 'TRUE') %>%
    select(-prop, -mental_health_risk) %>%
    pivot_wider(names_from = scen,
                values_from = count) %>%
    rename(intervention = contains(scen.list[[2]])) %>%
    mutate(MHR_diff = intervention - baseline) %>%
    select(run_id, year, MHR_diff) %>%
    group_by(year) %>%
    summarise(MHR_diff = mean(MHR_diff))

  p1 <- ggplot(data = plot_data, aes(x = year, y = MHR_diff)) +
    geom_point() +
    geom_line() +
    xlab('Year') +
    ylab('Count Change') +
    theme_minimal()

  plot_data_2 <- data %>%
    filter(mental_health_risk == 'TRUE') %>%
    select(-mental_health_risk, -count) %>%
    pivot_wider(names_from = scen,
                values_from = prop) %>%
    rename(intervention = contains(scen.list[[2]])) %>%
    mutate(MHR_prop_diff = intervention - baseline) %>%
    select(run_id, year, MHR_prop_diff) %>%
    group_by(year) %>%
    summarise(MHR_prop_diff = mean(MHR_prop_diff) * 100)

  p2 <- ggplot(data = plot_data_2, aes(x = year, y = MHR_prop_diff)) +
    geom_point() +
    geom_line() +
    xlab('Year') +
    ylab('Percentage Change') +
    theme_minimal()

  print(p1)
  print(p2)

  ggsave(filename = paste0(summary_group, '_count', '.png'),
         plot = p1)

  ggsave(filename = paste0(summary_group, '_proportion', '.png'),
         plot = p2)
}
```

```{r "treated_population", include=FALSE}
SF12_change_treated <- function(scen) {

  summary_group <- 'treated'
  scen <- paste0(scen, '_together')
  #scen <- c('ChildPovertyReductionRELATIVE_2_together')

  data <- do.call(rbind, generate_df_list(summary_group, scen))

  plot_data <- data %>%
    select(run_id, SF_12, hh_income, year, scenario) %>%
    pivot_wider(names_from = scenario,
                values_from = c(SF_12, hh_income),
                names_sep = '.') %>%
    unnest(cols = c(starts_with("SF_12"), starts_with("hh_income"))) %>%
    mutate(across(starts_with("SF_12"), ~ . - SF_12.baseline),
           across(starts_with("hh_income"), ~ . - hh_income.baseline)) %>%
    rename_with(~ paste0("diff_", .), starts_with("SF_12")) %>%
    rename_with(~ paste0("diff_", .), starts_with("hh_income")) %>%
    pivot_longer(cols = starts_with("diff"),
                 names_to = c(".value", "scenario"),
                 names_sep = "\\.") %>%
    filter(scenario != 'baseline')

  plot_data_confint <- plot_data %>%
      group_by(year) %>%
      summarise(n = n(),
                mean_SF12_change = mean(diff_SF_12),
                margin_SF12 = qt(0.975, df = n - 1) * (sd(diff_SF_12) / sqrt(n)),
                lower_SF12 = mean_SF12_change - margin_SF12,
                upper_SF12 = mean_SF12_change + margin_SF12,
                mean_hh_income_change = mean(diff_hh_income),
                margin_hh_income = qt(0.975, df = n - 1) * (sd(diff_hh_income) / sqrt(n)),
                lower_hh_income = mean_hh_income_change - margin_hh_income,
                upper_hh_income = mean_hh_income_change + margin_hh_income,)

  # p1 <- ggplot(plot_data, aes(x = year, y = diff_SF_12)) +
  #     geom_hline(yintercept = 0, linetype = 'dashed') +
  #     geom_vline(xintercept = 2030, linetype = 'dotted') +
  #     geom_smooth() +
  #     labs(title = 'Treatment on the Treated',
  #          subtitle = 'SF_12') +
  #     xlab('Year') +
  #     ylab('Change in mean SF-12')

  p2 <- ggplot(plot_data_confint, aes(x = year, y = mean_SF12_change)) +
      geom_ribbon(aes(ymin = lower_SF12, ymax = upper_SF12), linetype = 'dotted', alpha = 0.3) +
      geom_line(linetype = 'dashed') +
      geom_point() +
      geom_hline(yintercept = 0, linetype = 'dashed') +
      geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
      labs(title = 'Treated Population',
           subtitle = 'SF_12') +
      xlab('Year') +
      ylab('Change in Mean SF-12 MCS')

  # p3 <- ggplot(plot_data, aes(x = year, y = diff_hh_income)) +
  #     geom_hline(yintercept = 0, linetype = 'dashed') +
  #     geom_vline(xintercept = 2030, linetype = 'dotted') +
  #     geom_smooth() +
  #     labs(title = 'Treatment on the Treated',
  #          subtitle = 'Household Income') +
  #     xlab('Year') +
  #     ylab('Change in mean hh_income')

  p4 <- ggplot(plot_data_confint, aes(x = year, y = mean_hh_income_change)) +
      geom_ribbon(aes(ymin = lower_hh_income, ymax = upper_hh_income), linetype = 'dotted', alpha = 0.3) +
      geom_line(linetype = 'dashed') +
      geom_point() +
      geom_hline(yintercept = 0, linetype = 'dashed') +
      geom_vline(xintercept = 2030, linetype = 'dotted', alpha = 0.5) +
      labs(title = 'Treatment on the Treated',
           subtitle = 'Household Income') +
      xlab('Year') +
      ylab('Change in Mean Household Income')

  #print(p1)
  print(p2)
  #print(p3)
  print(p4)

  ggsave(filename = paste0(summary_group, '_SF-12', '.png'),
         plot = p2)

  ggsave(filename = paste0(summary_group, '_hh-income', '.png'),
         plot = p4)
}
```

```{r "synthpop_uncertainty_plot", include=FALSE}
synthpop_uncertainty <- function(summary_group, scen, var) {
  # summary_group <- 'whole_pop'
  # scen.list <- 'baseline'  # , 'ChildPovertyReductionRELATIVE_2'

  data <- do.call(rbind, generate_df_list(summary_group, scen))

  #data$pop_sample <- mutate(data, pop_sample = ceiling(run_id / 10))

  data <- data %>%
    select(run_id, .data[[var]], mean_cost, year, scen) %>%
    mutate(pop_sample = ceiling(run_id / 10))

  plot_data_local <- data %>%
    group_by(pop_sample, year) %>%
    summarise(var_mean = mean(.data[[var]]))

  plot_data_global <- data %>%
    group_by(year) %>%
    summarise(global_var_mean = mean(.data[[var]]))

  p1 <- ggplot(plot_data_local, aes(x = year, y = var_mean)) +
    geom_line(aes(group = pop_sample), alpha = 0.4) +
    geom_line(data = plot_data_global, aes(x = year, y = global_var_mean), color = 'red') +
    labs(title = paste0(var, ' uncertainty in Synthetic Population Sample'),
         subtitle = 'Global mean (red) vs local means (grey)') +
    xlab('Year') +
    ylab(paste0('Mean ', var))

  print(p1)

  ggsave(filename = paste0('synthpop_uncertainty_', var, '.png'),
         plot = p1)

}
```

```{r "Data", include=FALSE}
scen.list <- c('baseline', 'ChildPovertyReduction')
scen <- 'ChildPovertyReduction'
```

# MCS Change to QALY Example

We use an equation derived from table 4 (below) in [this paper by Lawrence and Fleishman](https://pubmed.ncbi.nlm.nih.gov/15090102/) to calculate QALYs from SF-12 Mental Component Score (MCS) and Physical Component Score (PCS).

#![L&F Table 4](L&F_table4.png)

```{r}
# formula to find the utility score from mean MCS and PCS scores of a population
utility_score <- function(pcs, mcs) {
  utility_score <- 
    1.6984 + 
    (pcs * 0.07927) +
    (mcs * 0.02859) +
    ((pcs * mcs) * -0.000126) +
    ((pcs * pcs) * -0.00141) +
    ((mcs * mcs) * -0.00014) +
    ((pcs * pcs * pcs) * 0.0000107)
  return(utility_score)
}

# Calculate utility score for hypothetical baseline (MCS & PCS == 45.0)
initial_conditions <- utility_score(pcs = 45.0,
                                    mcs = 45.0)
# Calculate utility score for 0.1 improvement in MCS (MCS now == 45.1)
intervention <- utility_score(pcs = 45.0,
                              mcs = 45.1)

print("Utility score for hypothetical baseline (assuming mean MCS and PCS both equal 45.0):")
print(initial_conditions)
print("Utility score for population given a 0.1 mean improvement in MCS:")
print(intervention)

print("Gain in utility score per person:")
gain <- intervention - initial_conditions
print(gain)

print("Gain in utiliy for the total Scotland population (assuming 5.5m). This is equivalent to QALYs gained:")
print(gain * 5510000)
```

This rough calculation shows that a 0.1 absolute change in mean SF_12 across the Scottish population would equal over 5000 additional QALYs in a single year. These improvements can also accumulate over time. If we were able to achieve a 0.1 improvement in mean MCS for Scotland in year 1, and maintain this 0.1 improvement for 10 years, the resulting improvement in QALYs would be 56,000+.

NOTE: This calculation assumes that PCS is unchanged when MCS can vary. This is not realistic, but the calculation is still useful to provide context for the results to follow.

# SAMPLE UNCERTAINTY

When running MINOS with a synthetic population, due to the size of the synthetic populations and the demands of the model, we have to take a reduced sample as our starting population. For the Scotland synthetic runs we take a 10% sample, which reduces the population size from roughly 5.5m to 550,000.

We therefore have to also account for the uncertainty in model outputs that comes from this reduced sample. As a somewhat temporary fix for this, we have done the following:

- Generate 10 different 10% samples from our synthetic population
  - Generate a replenishing population from each of these samples
- Use each distinct 10% sample for 10 runs each, totalling 100 runs
- Plot the local mean from each 10 run block, and the global mean from all 100 runs

```{r}
synthpop_uncertainty(summary_group = 'whole_pop',
                     scen = 'baseline',
                     var = 'hh_income')

synthpop_uncertainty(summary_group = 'whole_pop',
                     scen = 'baseline',
                     var = 'SF_12')
```

# SAMPLE STATISTICS

```{r}

# data <- do.call(rbind, generate_df_list('whole_pop', 'baseline'))
# data.fam <- do.call(rbind, generate_df_list('families', 'baseline'))
# 
# nrow(data)
# 
# 
# 5463300
# 
# 4580000
# 
# 
# 359328 / 458000 * 100
# 
# 100 / 78.4559
# 
# 
# 109338 * 12.74601

```

# COMBINED RELATIVE AND ABSOLUTE REDUCTION INTERVENTION

## Whole Population

What are the impacts on mental health of meeting the targets:

- Across the population?
- Population by income quintile?

```{r}
SF12_change_simple(summary_group = 'whole_pop',
                   scen.list = scen.list)

SF12_change_ind_trajectories(summary_group = 'whole_pop',
                             scen.list <- scen.list)

mental_illness_risk_counts(summary_group = 'men_illness_risk',
                           scen.list = scen.list)
```

### Income Quintiles

```{r}
SF12_change_income_quint(summary_group = 'whole_pop',
                                    scen.list = scen.list)
SF12_change_income_quint2(summary_group = 'whole_pop',
                                    scen = scen)

SF12_change_income_quint2_fudged(summary_group = 'whole_pop',
                                    scen = scen)
```

### SF_12 Inequality

```{r}
inequality_plots(summary_group = 'whole_pop',
                 scen = scen)
```

## Families

What are the impacts on mental health of meeting the targets:

- For families only? By income quintile

```{r}
SF12_change_simple(summary_group = 'families',
                       scen.list = scen.list)

SF12_change_ind_trajectories(summary_group = 'families',
                             scen.list <- scen.list)

mental_illness_risk_counts(summary_group = 'men_illness_risk_families',
                           scen.list = scen.list)
```

### Income Quintile

```{r}
SF12_change_income_quint(summary_group = 'families',
                                    scen.list = scen.list)
SF12_change_income_quint2(summary_group = 'families',
                                    scen = scen)
SF12_change_income_quint2_fudged(summary_group = 'families',
                                    scen = scen)
```

### SF_12 Inequality

```{r}
inequality_plots(summary_group = 'families',
                 scen = scen)
```

## Priority Subgroups

What are the impacts on mental health of meeting the targets:

- For families that meet the criteria of 0/1/2+ of the priority family groups?

```{r}
SF12_change_simple(summary_group = 'priority_any',
                       scen.list = scen.list)
```

### Number of Priority Subgroups

Results disaggregated by if household belongs to 0, 1, 2, 3+ priority subgroups.

```{r}
SF12_change_num_priority_groups(scen.list)
```

## Treated Population

```{r}
SF12_change_simple(summary_group = 'treated_relative',
                   scen.list = scen.list)

SF12_change_simple(summary_group = 'treated_absolute',
                   scen.list = scen.list)

SF12_change_treated(scen = scen)
```

```{r}

```