### 2024-10-02 minos/outcomes/spatial_analysis_gbis.Rmd

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

spatial analysis notebook for chapter.

msoa/lsoa/la level maps for GBIS MCS/PCS/QALY.

regressions
ICER plots
disaggregation of good performing LSOAs/MSOAs.

```{r}
library(here)
library(ggplot2)
library(xtable)
library(dplyr)

# load 2020 data.
full_pop_data <- read.csv(here::here("data", "scaled_manchester_aligned_US", "2020_US_cohort.csv"))

full_pop_data$FP10 <- ((full_pop_data$yearly_energy * 10) > (full_pop_data$net_hh_income * 12))
full_pop_data$is_green <- "Total Population"

full_pop_data_copy <- full_pop_data
full_pop_data_copy2 <- full_pop_data

# load green csvs for gbis/gbis.
gbis_green_lsoas <- read.csv(here::here("plots", "cost_vs_gain_by_lsoa_QALYs_manchester_synthetic_baseline_GBIS_2025_diffmap.pdfgreen_lsoas.csv"))
#gbis_green_lsoas <- read.csv(here::here("plots", "cost_vs_gain_by_lsoa_QALYs_manchester_synthetic_baseline_GBIS_2030_diffmap.pdfgreen_lsoas.csv"))
gbis_expensive_non_green_lsoas <- read.csv(here::here("plots", "cost_vs_gain_by_lsoa_QALYs_manchester_synthetic_baseline_GBIS_2025_diffmap.pdfexpensive_non_green_lsoas.csv"))

#distributional differences for heating and income between full pop and green pop.
#green_subpopulation <- full_pop_data[which(full_pop_data$ZoneID %in% gbis_green_lsoas$x), ]
full_pop_data[which(full_pop_data$ZoneID %in% gbis_green_lsoas$x), "is_green"] <- "Cost-Effective LSOAs"
full_pop_data <- full_pop_data[which(full_pop_data$is_green != "Total Population"), ] 
#full_pop_data[which(!(full_pop_data$ZoneID %in% gbis_green_lsoas$x)), "is_green"] <- "Not Green LSOAs"

full_pop_data_copy2[which(full_pop_data_copy2$ZoneID %in% gbis_expensive_non_green_lsoas$x), "is_green"] <-  "Expensive and not Cost-Effective LSOAs"
#full_pop_data_copy2 <- full_pop_data_copy2[which(full_pop_data$is_green == "All Not Expensive and Not Green LSOAs"), ] 
full_pop_data_copy2 <- full_pop_data_copy2[which(full_pop_data_copy2$is_green != "Total Population"), ]

full_pop_data <- rbind(full_pop_data, full_pop_data_copy, full_pop_data_copy2)
#full_pop_data <- full_pop_data[which(full_pop_data$is_green != "F"), ] 
#print(table(full_pop_data$is_green))
full_pop_data$housing_quality <- factor(full_pop_data$housing_quality, levels=c("Low", "Medium", "High"), ordered=T)
```

```{r}

aggregate_by_group_continuous <- function(data, v) {
  out_table <- full_pop_data %>%
    group_by(is_green) %>%
     summarise("GBIS Variable Mean" = mean(!!sym(v)))  %>%
        mutate(Variable = "", .before=is_green) %>%
          rename("LSOA Type" = is_green)
    out_table$Variable[1] <- v
  return (out_table)
}

aggregate_by_group_discrete <- function(data, v) {
  out_table <- full_pop_data %>%
    group_by(is_green, !!sym(v)) %>%
     summarise(count = n())  %>%
      ungroup() %>%
        mutate(freq = sum(count), .by=is_green) %>%
          mutate("GBIS Percentages" = 100*count / freq) %>%
            mutate(Variable = "", .before=is_green) %>%
              rename("Variable Levels" = !!sym(v), "LSOA Type" = "is_green") %>%
                select(-one_of(c("count", "freq"))) %>%
                  arrange("Variable Levels")

  out_table$`Variable Levels` <- as.character(out_table$`Variable Levels`)
  out_table$Variable[1] <- v
  return (out_table)
}

discrete_var_table <- data.frame()

for (var in c('housing_quality' ,'FP10', 'ethnicity', 'simd_decile')){
  discrete_var_table <- rbind(discrete_var_table, aggregate_by_group_discrete(full_pop_data, var))
}

continuous_var_table <- data.frame()

for (var in c('hh_income', "yearly_energy", "hh_rent","hh_mortgage", "nutrition_quality", "ncars", "council_tax", "yearly_electric", "yearly_gas", "yearly_oil", "yearly_gas_electric", "yearly_other_fuel", "age")){
  continuous_var_table <- rbind(continuous_var_table, aggregate_by_group_continuous(full_pop_data, var))
}

epcg_discrete_table <- read.csv(here::here("plots", "EPCG_discrete_var_table.csv"))
epcg_continuous_table <- read.csv(here::here("plots", "EPCG_continuous_var_table.csv"))
colnames(epcg_discrete_table)[3] <- "Variable Levels"
colnames(epcg_discrete_table)[2] <- "LSOA Type"
colnames(epcg_continuous_table)[2] <- "LSOA Type"

#colnames(epcg_continuous_table)[3] <- "Variable Levels"

discrete_var_table <- discrete_var_table %>% left_join(epcg_discrete_table, by=c("Variable", "LSOA Type", "Variable Levels"))
colnames(discrete_var_table)[5] <- "EPCG Percentages"

continuous_var_table$`EPCG Variable Mean` <- epcg_continuous_table$Variable.Mean
continuous_var_table <- continuous_var_table[, c("Variable", "LSOA Type", "EPCG Variable Mean", "GBIS Variable Mean")]

print(xtable(discrete_var_table, type = "latex"), file = here::here("plots", "GBIS_discrete_var_table.tex"))
print(xtable(continuous_var_table, type = "latex"), file = here::here("plots", "GBIS_continuous_var_table.tex"))

```

```{r}

ggplot(full_pop_data, aes(x = hh_income, colour = is_green)) +
  geom_density() +labs(x="Household Disposable Income", y="Density", colour="LSOA Type")#+
  #labs(x="Intervention Cost", y="QALY Cumulative Gain", color="IMD Decile")
ggsave(here::here("plots","EPCG_hh_income.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = SF_12, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_SF_12.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = SF_12_PCS, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_SF_12_PCS.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = yearly_energy, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_yearly_energy.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = yearly_gas, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_yearly_gas.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = yearly_electric, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_yearly_electric.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = yearly_gas_electric, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_yearly_gas_electric.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = age, colour = is_green)) +
  geom_density()+labs(x="Age", y="Density", colour="LSOA Type")
ggsave(here::here("plots","EPCG_age.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = yearly_oil, colour = is_green)) +
  geom_density()
ggsave(here::here("plots","EPCG_yearly_oil.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x = nutrition_quality, colour = is_green)) +
  geom_density()

ggplot(full_pop_data, aes(x = hh_mortgage, colour = is_green)) +
  geom_density()

```

```{r}
ggplot(full_pop_data, aes(x=as.factor(heating), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )
ggsave(here::here("plots","EPCG_heating.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(housing_quality), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )
ggsave(here::here("plots","EPCG_housing_quality.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(neighbourhood_safety), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )
ggsave(here::here("plots","EPCG_neighbourhood_safety.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(FP10), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" ) +labs(x="FP10", y="Density", fill="LSOA Type")
ggsave(here::here("plots","EPCG_FP10.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(ethnicity), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" ) +labs(x="Ethncity", y="Density", fill="LSOA Type")
ggsave(here::here("plots","EPCG_ethnicity.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(ncars), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )
ggsave(here::here("plots","EPCG_ncars.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(sex), fill=as.factor(is_green)))+
  geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" ) +labs(x="Sex", y="Density", fill="LSOA Type")
ggsave(here::here("plots","EPCG_sex.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(simd_decile), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )+labs(x="IMD Decile", y="Density", fill="LSOA Type")
ggsave(here::here("plots","EPCG_simd_decile.pdf"), plot=last_plot())

ggplot(full_pop_data, aes(x=as.factor(financial_situation), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

ggplot(full_pop_data, aes(x=as.factor(simd_decile), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

ggplot(full_pop_data, aes(x=as.factor(is_overcrowded), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

ggplot(full_pop_data, aes(x=as.factor(S7_labour_state), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

ggplot(full_pop_data, aes(x=as.factor(education_state), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

ggplot(full_pop_data, aes(x=as.factor(housing_tenure), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

ggplot(full_pop_data, aes(x=as.factor(number_of_rooms), fill=as.factor(is_green)))+
    geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" )

```

<!-- # do the same again for baseline and intervened green pops? -->

<!-- ```{r} -->

<!-- # baseline and gbis green lsoa populations in 2025.   -->

<!-- baseline_2025 <- read.csv(here::here("output", "energy_manchester_scaled", "baseline", "2024_07_25_15_08_03", "0001_run_id_2025.csv")) -->
<!-- gbis_2025 <- read.csv(here::here("output", "energy_manchester_scaled", "GBIS", "2024_07_26_18_23_14", "0001_run_id_2025.csv")) -->

<!-- baseline_2025 <-  baseline_2025[which(baseline_2025$ZoneID %in% gbis_green_lsoas$x), ] -->
<!-- gbis_2025 <-  gbis_2025[which(gbis_2025$ZoneID %in% gbis_green_lsoas$x), ] -->

<!-- baseline_2025 <- baseline_2025[, c("hh_income", "SF_12", "SF_12_PCS", "yearly_energy", "heating", "housing_quality", "neighbourhood_safety")] -->
<!-- gbis_2025 <- gbis_2025[, c("hh_income", "SF_12", "SF_12_PCS", "yearly_energy", "heating", "housing_quality", "neighbourhood_safety")] -->

<!-- baseline_2025$group <- "baseline" -->
<!-- gbis_2025$group <- "gbis" -->

<!-- plot_data <- rbind(baseline_2025, gbis_2025) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ggplot(plot_data, aes(x = hh_income, colour = group)) + -->
<!--   geom_density() -->

<!-- ggplot(plot_data, aes(x = SF_12, colour = group)) + -->
<!--   geom_density() -->

<!-- ggplot(plot_data, aes(x = SF_12_PCS, colour = group)) + -->
<!--   geom_density() -->

<!-- ggplot(plot_data, aes(x = yearly_energy, colour = group)) + -->
<!--   geom_density() -->
<!-- ``` -->
<!-- ```{r} -->
<!-- ggplot(plot_data, aes(x=as.factor(heating), fill=as.factor(group)))+ -->
<!--   geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" ) -->

<!-- ggplot(plot_data, aes(x=as.factor(housing_quality), fill=as.factor(group)))+ -->
<!--   geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" ) -->

<!-- ggplot(plot_data, aes(x=as.factor(neighbourhood_safety), fill=as.factor(group)))+ -->
<!--   geom_bar(aes( y=..count../tapply(..count.., ..fill.. ,sum)[..fill..]), position="dodge" ) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- baseline_2025 <- read.csv(here::here("output", "energy_manchester_scaled", "baseline", "2024_07_25_15_08_03", "0001_run_id_2025.csv")) -->
<!-- gbis_2025 <- read.csv(here::here("output", "energy_manchester_scaled", "GBIS", "2024_07_26_18_23_14", "0001_run_id_2025.csv")) -->

<!-- baseline_2025 <-  baseline_2025[which(!(baseline_2025$ZoneID %in% gbis_green_lsoas$x)), ] -->
<!-- gbis_2025 <-  gbis_2025[which(!(gbis_2025$ZoneID %in% gbis_green_lsoas$x)), ] -->

<!-- baseline_2025 <- baseline_2025[, c("hh_income", "SF_12", "SF_12_PCS", "yearly_energy", "heating", "housing_quality", "neighbourhood_safety")] -->
<!-- gbis_2025 <- gbis_2025[, c("hh_income", "SF_12", "SF_12_PCS", "yearly_energy", "heating", "housing_quality", "neighbourhood_safety")] -->

<!-- baseline_2025$group <- "baseline" -->
<!-- gbis_2025$group <- "gbis" -->

<!-- plot_data <- rbind(baseline_2025, gbis_2025) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ggplot(plot_data, aes(x = hh_income, colour = group)) + -->
<!--   geom_density() -->

<!-- ggplot(plot_data, aes(x = SF_12, colour = group)) + -->
<!--   geom_density() -->

<!-- ggplot(plot_data, aes(x = SF_12_PCS, colour = group)) + -->
<!--   geom_density() -->

<!-- ggplot(plot_data, aes(x = yearly_energy, colour = group)) + -->
<!--   geom_density() -->
<!-- ``` -->