### 2024-10-02 minos/outcomes/spatial_energy_validation.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r}

library(here)
source(here::here("minos", "outcomes", "minos_SF12_maps.R"))

electricity_consumption_data <- read.csv(here::here("persistent_data", "spatial_data", "UK_electricity_by_lsoa.csv"))
electricity_consumption_data <- electricity_consumption_data[, c("LSOA.code", "Median..consumption..kWh.per.meter.")]
colnames(electricity_consumption_data) <- c("ZoneID", "median_electricity")
electricity_consumption_data$median_electricity <- as.numeric(gsub(",", "", electricity_consumption_data$median_electricity))

gas_consumption_data <- read.csv(here::here("persistent_data", "spatial_data", "UK_gas_by_lsoa.csv"))
gas_consumption_data <- gas_consumption_data[, c("LSOA.code", "Median..consumption..kWh.per.meter.")]
colnames(gas_consumption_data) <- c("ZoneID", "median_gas")
gas_consumption_data$median_gas <- as.numeric(gsub(",", "", gas_consumption_data$median_gas))

# get geojson data. convert to tibble.
lsoa_data <- st_read(here::here("persistent_data", "spatial_data", "UK_super_outputs.geojson"))
gmca_la_names <- c("Bolton", "Bury", "Rochdale", "Oldham", "Tameside", "Stockport", "Manchester", "Trafford", "Salford", "Wigan")
lsoa_data <- lsoa_data[which(lsoa_data$LAName %in% gmca_la_names), ]

# get 2020 data. 
# aggregate four yearly energy consumptions median by lsoa. 
full_pop_data <- read.csv(here::here("data", "scaled_manchester_aligned_US", "2020_US_cohort.csv"))

elec_aggregate <- aggregate(full_pop_data$yearly_electric, list(full_pop_data$ZoneID), median)
elec_gas_aggregate <- aggregate(full_pop_data$yearly_gas_electric, list(full_pop_data$ZoneID), median)
gas_aggregate <- aggregate(full_pop_data$yearly_gas, list(full_pop_data$ZoneID), median)
oil_aggregate <- aggregate(full_pop_data$yearly_oil, list(full_pop_data$ZoneID), median)
other_aggregate <- aggregate(full_pop_data$yearly_other_fuel, list(full_pop_data$ZoneID), median)

colnames(elec_aggregate) <- c("ZoneID", "yearly_electric")
elec_aggregate$yearly_gas_electric <- elec_gas_aggregate$x
elec_aggregate$yearly_gas <- gas_aggregate$x
elec_aggregate$yearly_oil <- oil_aggregate$x
elec_aggregate$yearly_other_fuel <- other_aggregate$x

#attach geojson data.
elec_aggregate2 <- merge(elec_aggregate, lsoa_data, by="ZoneID")

# plot it.

v <- 'yearly_electric'
elec_map <- ggplot(data = elec_aggregate2, aes(geometry = geometry, fill=.data[[v]])) +
    #geom_sf() + # black outlines on borders for clarity.
    geom_sf(color=NA) + # black outlines on borders for clarity.
    scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
    labs(fill="Yearly Electricity Consumption (£s)")  + # colour bar label.
    theme(legend.position="bottom") + # change plot to widescreen dimensions.
    xlab("Longitude") + # axis labels
    ylab("Latitude")
print(elec_map)
ggsave(here::here("plots","electric_map.pdf"), plot=last_plot())

v <- 'yearly_gas_electric'
gas_elec_map <- ggplot(data = elec_aggregate2, aes(geometry = geometry, fill=.data[[v]])) +
    #geom_sf() + # black outlines on borders for clarity.
    geom_sf(color=NA) + # black outlines on borders for clarity.
    scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
    labs(fill="Yearly Gas and Electricity Consumption (£s)")  + # colour bar label.
    theme(legend.position="bottom") + # change plot to widescreen dimensions.
    xlab("Longitude") + # axis labels
    ylab("Latitude")
print(gas_elec_map)
ggsave(here::here("plots","gas_electric_map.pdf"), plot=last_plot())

v <- 'yearly_gas'
gas_map <- ggplot(data = elec_aggregate2, aes(geometry = geometry, fill=.data[[v]])) +
    #geom_sf() + # black outlines on borders for clarity.
    geom_sf(color=NA) + # black outlines on borders for clarity.
    scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
    labs(fill="Yearly Gas Consumption (£s)")  + # colour bar label.
    theme(legend.position="bottom") + # change plot to widescreen dimensions.
    xlab("Longitude") + # axis labels
    ylab("Latitude")
print(gas_map)
ggsave(here::here("plots","gas_map.pdf"), plot=last_plot())

v <- 'yearly_oil'
oil_map <- ggplot(data = elec_aggregate2, aes(geometry = geometry, fill=.data[[v]])) +
    #geom_sf() + # black outlines on borders for clarity.
    geom_sf(color=NA) + # black outlines on borders for clarity.
    scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
    labs(fill="Yearly Fuel Oil Consumption (£s)") + # colour bar label.
    theme(legend.position="bottom") + # change plot to widescreen dimensions.
    xlab("Longitude") + # axis labels
    ylab("Latitude")
print(oil_map)
ggsave(here::here("plots","oil_map.pdf"), plot=last_plot())

v <- 'yearly_other_fuel'
other_map <- ggplot(data = elec_aggregate2, aes(geometry = geometry, fill=.data[[v]])) +
    #geom_sf() + # black outlines on borders for clarity.
    geom_sf(color=NA) + # black outlines on borders for clarity.
    scale_fill_viridis_c(alpha = 1.0, direction=-1) + # use viridis colour scale and reverse it  so brighter is better.
    labs(fill="Yearly Other Fuel Consumption (£s)")  + # colour bar label.
    theme(legend.position="bottom") + # change plot to widescreen dimensions.
    xlab("Longitude") + # axis labels
    ylab("Latitude")
print(other_map)
ggsave(here::here("plots","other_fuel_map.pdf"), plot=last_plot())

# quantile plots as well.

pdf(here::here("plots", "electricty_quantile_plot.pdf"))
hist(elec_aggregate2$yearly_electric, xlab="Yearly Electricity Cost", main="")
#plot(sort(elec_aggregate2$yearly_electric), xlab="Observation", ylab = "Monthly Electricity Cost")
dev.off()

pdf(here::here("plots", "gas_electricty_quantile_plot.pdf"))
hist(elec_aggregate2$yearly_gas_electric, xlab="YearlyGas and Electricity Cost", main="")
#plot(sort(elec_aggregate2$yearly_gas_electric), xlab="Observation", ylab = "Monthly Gas and Electricity Cost")
dev.off()

pdf(here::here("plots", "gas_quantile_plot.pdf"))
hist(elec_aggregate2$yearly_gas, xlab="Yearly Gas Cost", main="")
#plot(sort(elec_aggregate2$yearly_gas), xlab="Observation", ylab = "Monthly Gas Cost")
dev.off()

pdf(here::here("plots", "oil_quantile_plot.pdf"))
hist(elec_aggregate2$yearly_oil, xlab="Yearly Fuel Oil Cost", main="")
#plot(sort(elec_aggregate2$yearly_oil), xlab="Observation", ylab = "Monthly Fuel Oil Cost")
dev.off()

pdf(here::here("plots", "other_fuel_quantile_plot.pdf"))
hist(elec_aggregate2$yearly_other_fuel, xlab="Yearly Other Fuel Cost", main="")
#plot(sort(elec_aggregate2$yearly_other_fuel), xlab="Observation", ylab = "Monthly Other Fuel Cost")
dev.off()

#qqplot(elec_aggregate2$yearly_electric)

# rolling in DESNZ real median KWH usage for validation. 
elec_aggregate3 <- merge(elec_aggregate2, electricity_consumption_data, by="ZoneID")
elec_aggregate4 <- merge(elec_aggregate3, gas_consumption_data, by="ZoneID")

# simple linear regression between KWH use and energy bill. 
print(summary(lm(median_electricity ~ yearly_electric-1, data=elec_aggregate4)))
print(summary(lm(median_gas ~ yearly_gas-1, data=elec_aggregate4)))

print(summary(lm(yearly_electric ~ median_electricity -1, data=elec_aggregate4)))
print(summary(lm(yearly_gas~median_gas -1, data=elec_aggregate4)))

plot(elec_aggregate4$yearly_gas, elec_aggregate4$median_gas)
plot(elec_aggregate4$yearly_electric, elec_aggregate4$median_electric)

```