### 2024-10-23 docsrc/documentation/notebooks/mental_well_being.Rmd

---
title: "Mental Well-Being"
output: html_document
bibliography: 
  - "`r here::here('docsrc', 'refs.bib')`"
---

## Mental Well-Being

Introductory fluff. Why do we need this module? test reference [@1987:nelson].

```{r load_data, include=FALSE}
library(here)
source(here::here("docsrc", "documentation", "notebooks", "sphinx_notebook_utils.R"))
model_path <- here::here("data", "transitions", "SF_12", "lmm", "SF_12_LMM.rds")
model <- readRDS(model_path)

obs <- model@frame$SF_12
preds <- predict(model, type='response')
v <- "SF_12"
```

### Methods

What methods are used? Justification due to output data type. explanation of model output.

### Data

What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

### Results

What are the results. Coefficients tables. diagnostic plots. measures of goodness of fit.

```{r SF12_Output, echo=FALSE}
options(width=120)
#ols_output("data/transitions/SF_12/OLS/SF_12_2018_2019.rds")
#ols_histogram("data/transitions/SF_12/OLS/SF_12_2018_2019.rds", "next_SF_12")

#glmm_output(model_path)
continuous_densities(obs, preds, v)
print(summary(model))
plot(model)
```

### References