### 2024-10-23 docsrc/documentation/notebooks/tobacco.Rmd

---
title: "Tobacco"
output:
  pdf_document: default
  html_document: default
bibliography: 
  - "`r here::here('docsrc', 'refs.bib')`"
---

```{r setup, include=FALSE}
library(here)
library(glmnet)
library(plyr)

source(here::here("docsrc", "documentation", "notebooks", "sphinx_notebook_utils.R"))
model_path <- here::here("data", "transitions", "ncigs", "rfo", "ncigs_RFO.rds")
model <- readRDS(model_path)

data <- read.csv(here::here("data", "final_US", "2020_US_cohort.csv"))
obs <- data$ncigs
data$ncigs_last <- data$ncigs

preds <- ranger::predictions(stats::predict(model, data))
column_names <- colnames(preds)
preds <-glmnet::rmult(preds) - 1
preds <- 5 * preds

v <- "ncigs"

```

## Tobacco

Number of cigarettes consumed is an indicator of several mental illnesses including anxiety [@lawrence2010anxiety].

### Data

  What variables are included? Why is this output chosen. What explanatory variables are used and why are they chosen

```{r tobacco_data}
discrete_barplot(obs, "ncigs")
```

### Methods

The number of zero inflated values is higher than expected for a count distribution such as a poisson distribution. This inflation occurs naturally as a large proportion (over 50\%) of the population do not smoke. There are two sources of cigarette consumption that can be modelled using zero inflated models. In this case a zero-inflated poisson (ZIP) is used. Two models are fitted simulatenously. One is a logistic regression that estimates whether a person smokes cigarettes or not. This provides a simple probability of smoking or not. The second is a poisson counts model estimating the number of cigarettes consumed.

### Data

Two set of variables are needed for the logistic and poisson parts of the ZIP model respectively.

Variables that predict how much a person smokes.

age. persons age. generally older people and very young smoke. 
SF_12. wellbeing estimates number of cigarettes smoked.
labour_state. whether a person is employed or not.
ethnicity. certain ethnicities more likely to smoke cigarettes. 
education_state. highest qualification.
job_sec job quality
hh_income household income
ncigs previous number consumed.

Variables that predict whether a person smokes

ethnicity. certain ethnicities more likely to smoke cigarettes. 
labour_state. whether a person is employed or not.
age
SF_12. wellbeing estimates number of cigarettes smoked.
ncigs previous number consumed.

### Results

Almost all coefficients significant. Particularly prevous consumption of cigarettes. Good estimation of the number of non-smokers in the population at around 55\%. Counts of smoking are underdispersed and fail to estimate consumption over 20 cigarettes.

```{r tobacco_output, echo=FALSE}
options(width=120)
discrete_barplot2(obs, preds, "housing_quality") 
rfo_output(model)
```

### References